### 2023-02-25 src/tex/ms.tex
initial commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@kelvin33550336}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-02-25 src/tex/ms.tex
paste paper

@@
 \usepackage{showyourwork}
+
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{verbatim}
+\usepackage{mathtools}
+\usepackage{needspace}
+\usepackage{enumitem}
+\usepackage{etoolbox}
+\usepackage{graphicx}
+\usepackage{microtype}
+\usepackage{afterpage}
+\usepackage{bm}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage[dvipsnames, usenames]{xcolor}
+\usepackage{tabularx,booktabs,multirow}
+\usepackage{float}
+\usepackage[italicdiff]{physics}
+
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
+\newcommand{\ripple}{\texttt{ripple}}
+\newcommand{\python}{\texttt{python}}
+\newcommand{\jax}{\texttt{jax}}
+\newcommand{\zdethp}{\texttt{zdethp}}
+
+\newcommand{\cuhk}{\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, N.T., Hong Kong}}
+\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}}
+\newcommand{\JHU}{\affiliation{William H. Miller III Department of Physics and Astronomy, Johns Hopkins University, Baltimore, Maryland 21218, USA}}
 
@@
 % Title
-\title{An open source scientific article}
+\title{Recalibrating Gravitational Wave Phenomenological Waveform Model}
 
 % Author list
-\author{@kelvin33550336}
+\author{Kelvin K.~H.~Lam} 
+\email{kelvin33550336@gmail.com}
+\cuhk
+\author{Kaze W.~K.~Wong} 
+\flatiron
+\author{Thomas D.~P.~Edwards}
+\JHU
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    We present a simple and general method of recalibrating gravitational wave (GW) phenomenological waveform models jointly.
+By using {\jax} and {\ripple}, we can perform automatic differentiation to functions, which allows us to use gradient-based optimization methods to recalibrate waveform coefficients in IMRPhenomD model.
+This method reduces systematic bias previously introduced to the model and generally can improve waveform accuracy.
+With recalibrated coefficients, we found that the typical \textit{mismatch} has a $50\%$ decrease.
+Furthermore, we analyze the accuracy base on the waveform's intrinsic parameters.
+We found that waveform accuracy has significant dependence on black hole spin.
+Reduced spin approximation introduces degeneracy in spin, which prevented further improvement.
+We isolated regions in the parameter space that does not fit the waveform ansatz.
+These results allow us to understand more about how to develop newer phenomenological models.
 \end{abstract}
@@
 % Main body with filler text
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction} \label{sec:intro}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results.
+This new O4 run is expected to double the rate of current binary black hole (BBH) observations \cite{abbott2020prospects}.
+Additionally, the sensitivity of interferometers will be increased to capture more details of GW.
+Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information.
+Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses.
+Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \cite{purrer2020gravitational}.
+Hence, it is necessary for us to develop and improve GW models.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+% Waveform families and slight motivation of why we take Phenom model
+Currently, three families of GW models are commonly used.
+They are the effective-one-body (EOB) \cite{taracchini2014effective}, Numerical Relativity (NR) surrogate \cite{varma2019surrogate} and phenomenological (Phenom) models \cite{husa2016frequency,khan2016frequency}.
+EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations.
+While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations.
+This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run.
+This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.
+
+Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision.
+In traditional numerical calculations, derivatives are usually obtained through numerical derivatives.
+Symbolic derivatives were available but it was less efficient.
+Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations.
+In {\python}, packages including \texttt{pytorch} \cite{pytorch}, \texttt{tensorflow} \cite{tensorflow2015-whitepaper}, etc.
+utilizes AD to train machine learning models.
+AD's algorithm is intuitive in nature.
+Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations.
+Since these operations are fundamental, they were saved as pairs internally.
+Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative.
+Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure.
+{\ripple} \cite{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
+It was first implemented in \texttt{lalsuite} using \texttt{C}.
+In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD.
+Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses.
+% Need to cite references for AD
+
+In this paper, we investigate the possibility of further improving the accuracy of IMRPhenomD by jointly optimizing all the fitting coefficients given NR waveforms, and what constraints one may face when trying to further improve.
+We find that simply by applying gradient descent algorithm, one can obtain a better set of waveform coefficients, thus improving the accuracy of the model.
+Furthermore, by comparing the accuracy of optimized and original waveforms, we find that model-generated waveforms are very sensitive to their intrinsic parameters.
+Specifically, IMRPhenomD favors certain parts of the parameter space.
+This means IMRPhenomD introduces systematic bias to other GW analysis tasks.
+This showcases the flaws of the ansatz and allows us to have a deeper understanding of Phenom models.
+% \kw{Briefly describe the high level finding in this paper}
+
+The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model.
+We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms.
+Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}.
+In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms.
+We also show how the optimization result differs with waveforms of different intrinsic parameters.
+In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \cite{khan2016frequency}.
+We also explain how reduced spin parameterization affects the accuracy of the model.
+
+\section{Method} \label{sec:method}
+
+\subsection{Waveform Model} \label{subsec:waveform_model}
+
+IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events.
+It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \cite{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses.
+% Using {\ripple} \cite{ripple} and {\jax} \cite{jax2018github}, we are able to easily recalibrate the model. 
+IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
+\begin{align}
+	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
+\end{align}
+In each segment, the amplitude and phase are made using simple functions of frequency such as polynomials or lorentzians.
+These simple analytic functions consists of parameters $\Lambda^i$, which are defined as follows.
+\begin{align} \label{eq:Lambda}
+	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
+	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
+	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
+	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
+\end{align}
+$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters including symmetric mass ratio $\eta$ and post-newtonian spin parameter $\chi_{\mathrm{PN}}$, 
+\begin{align}
+	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
+\end{align}
+Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller black hole respectively.
+Finally, the individual segments are connected in a way that the final waveform is continuous in its first derivative.
+
+From Eq.
+\ref{eq:Lambda}, we can see that the values of $\lambda$ are fundamental to waveform generation, which can significantly affect the shape of final waveforms.
+Thus, having a set of accurate waveform coefficients is important.
+Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations.
+In the case of \cite{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms.
+For each NR simulation, they can obtain one set of $\Lambda$'s through optimization.
+With many sets of $\Lambda$, $\lambda$ are subsequently found by fitting against Eq.
+\ref{eq:Lambda}.
+However, this piece-wise fitting deployed by \cite{khan2016frequency} causes calibrated coefficients to have systematic biases.
+Since coefficients were not fitted jointly, correlations between segments are omitted.
+Hence, the model introduces systematic biases to $\lambda$ and will be accumulated and passed along to the data analysis pipeline.
+Also, their calibration procedure does not guarantee to have the optimal set of $\lambda$, since the process of connecting segments alters the previously fitted waveform.
+The model-generated waveforms do not take this effect into account, and thus introduces additional inaccuracies.
+
+% they did not use mismatch as the loss function
+
+Instead, we recalibrate coefficients jointly, which we can remove inaccuracies and biases discussed above, and can improve model accuracy.
+In the past, due to the complex nature of GW strains and piece-wise formalism of IMRPhenomD, non-linear fitting was difficult to be performed in optimizing coefficients.
+Hence, piece-wise optimization was done to obtain coefficients.
+However, with {\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus allowing the use of gradient-based algorithms for us to recalibrate the model.
+
+\subsection{Mismatch} \label{subsec:mismatch}
+
+To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \cite{husa2016frequency}. It is defined as 
+\begin{align} \label{eq:mismatch}
+	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
+\end{align}
+where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
+\begin{align}\label{eq:inner_prod}
+	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
+\end{align}
+where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.
+
+In the original calibration \cite{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
+The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \cite{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
+We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown.
+This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length.
+When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.
+
+We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned high-power (\zdethp) noise spectrum and a constant spectrum.
+Since in any data analysis tasks, model-generated waveforms will ultimately be used to compare with GW observations, they are subjected to the sensitivity of detectors.
+Hence, it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than a non-weighted spectrum.
+However, with sensitivity curve involved, we have to choose a corresponding mass scale.
+Here, we simply choose $M=50M_{\odot}$ for all waveforms, since this mass scale is common in LIGO observations.
+On the other hand, we also recalibrated coefficients with the constant spectrum for a more general purpose.
+
+% GW strain treatment from time-series to frequency domain
+In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms.
+To compute such mismatches, NR waveforms from the SXS catalog are taken \cite{boyle2019sxs}.
+These NR waveforms are calculated using the \texttt{SpEC} code.
+Initially, NR waveforms are in the form of time-series data.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters.
+We taper the time-series using Tukey window.
+\footnote{
+	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
+} 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
+
+\subsection{Optimization Scheme} \label{subsec:optimization}
+
+Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus, we can utilize simple gradient descent to recalibrate the model.
+
+% \kw{Mention the main advantage in this study is we can optimize everything at once.}
+
+% To recalibrate the model, we have to define loss functions that quantify the difference between the model-generated waveforms and NR waveforms.
+In the original calibration, they took the mean square error as the loss function for optimization.
+Instead, we directly use mismatch as our metric for loss functions.
+However, the mismatch of individual waveforms is not useful in optimization, since it does not represent all waveforms with various intrinsic parameters.
+We ought to calibrate coefficients that suit waveforms with different intrinsic parameters simultaneously, hence waveforms in different regions of parameter space should be used in the optimization scheme.
+Here, we chose two different loss functions to examine their ability to optimize the model, e.g.
+mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and fractional loss $\mathcal{L}_{\mathrm{fl}}$.
+
+% Do something much more direct:
+To recalibrate the model, we need to define a loss function that quantifies the difference between the model-generated waveforms and NR waveforms.
+We use mismatch over the set of all training waveforms as the loss function.
+While in \cite{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
+The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest.
+One of which takes a simple average and the other takes a normalized average based on initial mismatch.
+
+% To recalibrate the model, we need to define a loss function that quantify ...
+We use mismatch over the set of all training waveforms as the loss function.
+We choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
+The mismatch is averaged over the training dataset to cover the parameter space of interest.
+(used)
+\begin{align}
+	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
+	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+\end{align}	
+$\mathcal{M}_i$ represents the mismatch of an individual training waveform, $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual training waveform, and $N$ is the total number of individual training waveforms.
+For the first choice, it takes the mean of the mismatches, which serves as a basic loss function.
+However, it is prone to be dominated by a single waveform with a large mismatch.
+Other waveforms with smaller mismatches would be insignificant comparatively, and might not be able to improve under such optimization.
+Alternatively, the second choice, the mean of normalized mismatches, eliminates the aforementioned issue.
+Nevertheless, it excludes the information on initial mismatches.
+$\mathcal{L}_{\mathrm{fl}}$ restricts every training waveform to decrease at similar rates, hence it is hard to obtain optimized waveforms with mismatches in the same order of magnitude.
+Instead, their ratios in mismatches would remain approximately the same.
+Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust and final individual mismatches would be in a similar order of magnitude.
+In this paper, we display the results of using both loss functions and examine the difference between them.
+
+We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process.
+Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \cite{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation.
+As BAM waveforms are not publicly available, we cannot take the identical training set as them.
+Instead, we take the available waveforms from the SXS catalog to construct our loss function.
+They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+Notice the maximum mass ratio in our training set is lower than that in \cite{khan2016frequency}.
+Instead of having $q\leq18$, the maximum mass ratio is set as 8.
+This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+In fact, the SXS catalog only has NR waveforms with $q\leq10$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed have $q\leq8$.
+Hence, we set the maximum mass ratio to 8.
+
+With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
+\begin{algorithm}[H]
+	\hspace*{\algorithmicindent} \textbf{Input} $\mathcal{L}_{\mathrm{ini}}$
+	\begin{algorithmic}[1]
+		\State{$\mathcal{L}(\lambda) \gets \mathcal{L}_{\mathrm{ini}}$}
+		\State{$\alpha \gets 10^{-6}$}
+		\State{$i \gets 0$}
+		\While{$i<30000$}
+		\State{$\mathcal{L}'(\lambda) \gets \texttt{jax.grad}(\mathcal{L})$}
+		\State{$\lambda \gets \lambda-\alpha\mathcal{L}'(\lambda)$}
+		\EndWhile
+	\end{algorithmic}
+	\hspace*{\algorithmicindent} \textbf{Output} $\lambda$
+	\caption{Gradient Descent}
+\end{algorithm}
+Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \cite{khan2016frequency}.
+We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
+With such an optimization procedure, we obtain the optimized coefficients.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
+	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
+	\label{fig:intrin_space}
+\end{figure}
+
+\begin{table}[t]
+	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
+		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
+		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
+		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
+		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
+		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
+		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
+		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
+		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
+		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
+		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
+	\end{tabularx}
+	\caption{List of waveforms used to recalibrate the model.
+The mass ratio $q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$.
+Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration.
+\cite{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do have access.}
+	\label{tab:q148}
+\end{table}
+\begin{table}[t]
+	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
+		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
+		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
+		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
+		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
+	\end{tabularx}
+	\caption{Additional waveforms used in further recalibration.}
+	\label{tab:q1248}
+\end{table}
+
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=\textwidth]{figures/0154_waveform_comparison.pdf}
+	\caption{Comparison between original and optimized IMRPhenomD waveforms.
+Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and $\chi_1=\chi_2=-0.8$.
+The original mismatch is around $2.8\times10^{-4}$ and the optimized mismatch is around $5.3\times10^{-5}$.
+Top: It shows the amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD waveform.
+Bottom: It shows the relative error of amplitudes between NR and IMRPhenomD waveforms, and the absolute error of phases between NR and IMRPhenomD waveforms}
+	\label{fig:0154}
+\end{figure*}
+
+\section{Result and Comparison with Original Model} \label{sec:result}
+
+First, to judge the performance of the optimization scheme, we take $\sim530$ NR waveforms from the SXS catalog.
+We choose testing waveforms with negligible eccentricity (${e<2\times10^{-3}}$) and precession (${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model.
+We see that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters that are within the parameter space spanned by training waveforms.
+Hence, we can take these testing waveforms to compare with the original model.
+
+We first examine the effect of joint optimization on a single waveform.
+In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error.
+We have chosen one of the testing waveforms listed in \cite{khan2016frequency} for a fair comparison.
+% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.
+
+Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig.
+\ref{fig:q148}.
+We find the distribution's peak has shifted to the end with a lower mismatch.
+Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
+Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement.
+We see that in Fig.~\ref{fig:q148_fl}, the median of distribution has a $22.9\%$ decrease.
+However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
+This is due to the problem mentioned in section \ref{subsec:optimization}.
+Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.
+
+% \kw{We then compute the mismatch for all testing waveforms and plot the distribution of the mismatch in figure X. We find the distribution...}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/q148.pdf}
+	\caption{Distributions of waveform mismatches.
+We used $\mathcal{L}_{\mathrm{mean}}$ in optimization with training waveforms listed in Tab.~\ref{tab:q148}.
+Mismatches are weighted with the constant noise spectrum.
+Comparing the distributions, the median has decreased by $50.0\%$.}
+	\label{fig:q148}
+\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/q148_fl.pdf}
+	\caption{Distributions of waveform mismatches.
+We used $\mathcal{L}_{\mathrm{fl}}$ in optimization with training waveforms listed in Tab.~\ref{tab:q148}.
+Mismatches are weighted with the constant noise spectrum.
+The median has decreased by $22.9\%$.}
+	\label{fig:q148_fl}
+\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
+	\caption{Distributions of $\log_{10}$ difference in mismatch.
+The distribution labeled \texttt{q148} uses training waveforms listed in Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+Mismatches are calculated using the constant noise spectrum with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
+	\label{fig:q148_q1248_compare}
+\end{figure}
+
+Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch.
+The shape of the distribution is similar to Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}.
+This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch.
+Thus, it should be a model that fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum instead of the constant spectrum.
+
+With the success of improving waveforms, we increase the number of training waveforms for optimization.
+Taking additional waveforms listed in Tab.~\ref{tab:q1248}, we obtain a new set of coefficients.
+We see in Fig.
+\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small improvement.
+The high mismatch tail of the optimized distribution remains similar in length and endpoint as the original distribution, meaning that they cannot be improved using our procedure.
+Likewise, using additional waveforms to optimize loss function with {\zdethp} spectrum shows the same result, where the new distribution barely shows any improvement.
+
+% We believe that either the ansatz does not suit waveforms within certain regions of parameter space or the set of optimized coefficients falls to a minimum where they do not fit some testing waveforms.
+If in some region of parameter space, the ansatz does not fit well with NR waveforms, it would never be able to have consistent improvement under optimization.
+Instead, it would surf around the loss manifold with random fluctuations in mismatches.
+If the set of coefficients does not fit, then dividing the parameter space into regions and fitting separate sets of coefficients should give better results.
+
+Given the ansatz used in the waveform model is unlikely to be fully compatible with NR, and the optimization procedure is done over a distribution of waveforms with different source parameters, it is conceivable that there are some trade-off in accuracy of the waveform models between different part of the parameter space.
+If this is the reason why the high mismatch tail is not improving during the joint-optimization, separating the parameter space into smaller subspace should help alleviate this issue.
+On the other hand, if the ansatz does not have the right parameterized form to capture the behavior of the NR waveforms as a function of the intrinsic parameters, the result should be always biased, and we should not expect any improvement even if we separate the parameter space into smaller space during training.
+
+Since we know intrinsic parameters play an important role in the ansatz, we would like to investigate how intrinsic parameters affect the recalibration process.
+First, we plot the parameter space of $q$ vs.
+$\chi_{\mathrm{PN}}$ in Fig.
+\ref{fig:ps_q148_qchi}.
+In the low mass ratio region, waveforms with both positive and negative log differences are mixed up.
+On the other hand, along the horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve under recalibration.
+
+Furthermore, we plot the parameter space of $\chi_1$ vs.
+$\chi_2$ in Fig.~\ref{fig:ps_q148}.
+Waveforms along the diagonal axis, i.e.
+$\chi_1\approx\chi_2$, show good mismatch improvements.
+We expect to see this feature since the original coefficients were fitted using NR waveforms with equal or similar spin, hence the model prefers waveforms with similar spin.
+One interesting feature is how the second and fourth quadrants respond to optimization.
+In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms generally improve with along optimization.
+However, mismatches in the fourth quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization.
+Most waveforms even turned worse after optimization.
+These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}.
+
+% Waveform mismatches in the fourth quadrant, i.e.
+$\chi_1>0$ and $\chi_2<0$, performed worse after optimized.
+IMRPhenomD is constructed using single spin parameter, where there are degeneracy between $\chi_1$ and $\chi_2$.
+In regions with opposite extreme spin components, degeneracy significantly alters intrinsic parameters passed into waveform generation, hence the fourth quadrant shows such feature.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
+	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$.
+We use the same mismatch data as Fig.~\ref{fig:q148}.
+Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.}
+	\label{fig:ps_q148_qchi}
+\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/ps_q148.pdf}
+	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. This shows the same result as Fig.~\ref{fig:ps_q148_qchi}, except the axes are different.}
+	\label{fig:ps_q148}
+\end{figure}
+
+\begin{table}[t]
+	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0172 & 1.0 & 0.98     & 0.98     \\
+		SXS:BBH:0152 & 1.0 & 0.60     & 0.60     \\
+		SXS:BBH:0001 & 1.0 & 0.00     & 0.00     \\
+		SXS:BBH:1417 & 4.0 & 0.40     & 0.50     \\
+		SXS:BBH:0167 & 4.0 & 0.00     & 0.00     \\
+		SXS:BBH:1426 & 8.0 & 0.48     & 0.75     \\
+		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule
+		SXS:BBH:0370 & 1.0 & -0.20    & 0.40     \\
+		SXS:BBH:2092 & 1.0 & -0.50    & 0.50     \\
+		SXS:BBH:0330 & 1.0 & -0.80    & 0.80     \\
+		SXS:BBH:2116 & 2.0 & -0.30    & 0.30     \\
+		SXS:BBH:2111 & 2.0 & -0.60    & 0.60     \\
+		SXS:BBH:0335 & 2.0 & -0.80    & 0.80     \\
+		SXS:BBH:0263 & 3.0 & -0.60    & 0.60     \\
+		SXS:BBH:2133 & 3.0 & -0.73    & 0.85     \\
+		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule
+		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
+		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
+		SXS:BBH:0001 & 1.0 & 0.00     & 0.00     \\
+		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
+		SXS:BBH:0167 & 4.0 & 0.00     & 0.00     \\
+		SXS:BBH:1419 & 8.0 & -0.80    & -0.80    \\
+		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule
+		SXS:BBH:0304 & 1.0 & 0.50     & -0.50    \\
+		SXS:BBH:0327 & 1.0 & 0.80     & -0.80    \\
+		SXS:BBH:2123 & 2.0 & 0.30     & -0.30    \\
+		SXS:BBH:2128 & 2.0 & 0.60     & -0.60    \\
+		SXS:BBH:2132 & 2.0 & 0.87     & -0.85    \\
+		SXS:BBH:2153 & 3.0 & 0.30     & -0.30    \\
+		SXS:BBH:0045 & 3.0 & 0.50     & -0.50    \\
+		SXS:BBH:0292 & 3.0 & 0.73     & -0.85    \\ \midrule\bottomrule
+	\end{tabularx}
+	\caption{List of waveforms used in recalibrating coefficients in 4 quadrants.
+From top to down are the first, second, third and fourth quadrants.
+Note that for the first and third quadrants, waveforms are chosen to have equal or similar spins, while the training waveforms for the second and fourth quadrants are chosen to have opposite spins.}
+	\label{tab:quadrants}
+\end{table}
+
+As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is significantly different in different regions in the parameter space, we split the parameter space into 4 quadrants and perform separate fitting with training waveforms listed in Table \ref{tab:quadrants}.
+Note that in the second and fourth quadrants, there are not enough waveforms with $q>4$, hence the result are only valid up to $q\leq4$.
+From Fig.~\ref{fig:ps_q148_quadrant}, we see that all waveforms, except those in the fourth quadrant, show improvement in mismatch.
+Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
+First, we focus on waveforms with the same spin direction, i.e.
+$\chi_1\chi_2>0$.
+Waveforms lying in the neighborhood of the diagonal axis have noticeable improvements in mismatch due to most training waveforms being equal-spin waveforms.
+For the second quadrant, waveforms improved significantly with only a few defects due to some testing waveforms having $q>4$.
+In the fourth quadrant, most optimized waveforms have a higher mismatch than the original waveforms, as indicated by the positive log difference of mismatches.
+As performing optimization in a smaller subspace does not show any improvement compared to the original result, the ansatz mostly does not fit waveforms in the fourth quadrant.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
+	\caption{Distributions of mismatches after optimizing in separate quadrants.
+We use a constant noise spectrum to calculate mismatch and $\mathcal{L}_{\mathrm{mean}}$ for the loss function.
+Generally, most waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant.
+}
+	\label{fig:all_quadrants}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrant.pdf}
+	\caption{Parameter space of testing waveforms. Each quadrant is fitted independently. Colorbar represents log difference of mismatches before and after optimization.}
+	\label{fig:ps_q148_quadrant}
+\end{figure}
+
+\section{Discussion} \label{sec:discussion}
+
+%\begin{enumerate}
+%	\item Comparison with the original paper is not the same. (Different waveforms, different ways to modify the waveform, different range of frequencies)
+%	\item More waveforms are taken so it should be more robust, since the parameter space coverage is better.
+%    \item Problems with IMRPhenomD model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat)
+%    \item Can in principle split parameter space into regions that can be separately optimized.
+If not, then its must be the model's problem.
+Would the region classification be made automatic.
+Are there any algorithms to do so?
+%    \item Future work (Repeat on other GW models. Mapping between NR surrogate to see any systematics.)
+%\end{enumerate}
+
+We have shown the result of recalibrating waveform coefficients.
+One thing to note is that our recalibration procedure is not exactly the same as the original calibration.
+For instance, we use a different set of NR waveforms, frequency range, etc.
+Nonetheless, as the decrease in mismatch is rather significant, this optimization procedure should be able to improve the accuracy of IMRPhenomD on a similar scale regardless of the differences.
+Here, this result serves as a demonstration of the general method used.
+% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for both amplitude and phase.
+Similarly, for the merger-ringdown region, the optimized waveforms also show great improvement.
+
+We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with more training waveforms only has a small increase in the accuracy of the waveform model.
+We believe that by increasing the number of waveforms, the accuracy will not have a significant change as the waveform coefficients are already over-determined.
+Using more calibration NR waveforms will not further improve the model significantly.
+This suggests the form of the parameterized ansatz is not suitable for certain regions in the parameter space, thus mismatches of only a few waveforms decreased while other waveforms remain at the high mismatch tail with little changes.
+This implies that the model is ultimately restricted by the flexibility of the ansatz.
+
+One of the major problems causing inaccuracy in the ansatz is the reduced spin approximation.
+In IMRPhenomD, it is modeled using a single spin parameter, namely $\chi_{\mathrm{PN}}$ as outlined in Sec.
+\ref{sec:method}.
+Parameterizing a BBH merger with one spin parameter introduces degeneracy within the parameter space.
+Events with distinct black hole spins could result in equal $\chi_{\mathrm{PN}}$, thus generating the same waveform.
+Especially with high unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as events with small spin, thus giving an inaccurate result.
+Generally, the approximation gives straight lines of degeneracy in the parameter space, with its slope (always negative) dependent on the mass ratio.
+From Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves better in the top left region than the bottom right.
+To try to accommodate this issue, we split the parameter space into 4 quadrants as described in Sec.
+\ref{sec:result}.
+However, even with separate optimizations, we see in Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar mismatches as before while the second quadrant further improved.
+This suggests the ansatz is region-specific, with a higher preference for BBH events with $\chi_1<0$ and $\chi_2>0$.
+
+Separating regions into 4 quadrants is done purely out of simplicity.
+To give a more comprehensive analysis, one should be systematic about region selection.
+One can use level set estimation algorithms to obtain systematic regions of interest.
+This general algorithm reveals further degeneracies or issues within the ansatz.
+Then, recalibrating such individual regions might give better results.
+Alternatively, one can select regions according to the lines of degeneracy.
+However, with limited NR waveforms, such a selection scheme is not viable for us.
+In the future, with more NR waveforms spanning the entire parameter space, one can perform optimization with fewer restrictions.
+
+While our work focused mainly on the IMRPhenomD model, this simple yet general method can be utilized in other differentiable GW models.
+For instance, within the same family, IMRPhenomX \cite{pratten2020setting} or IMRPhenomP \cite{hannam2014simple} models.
+By jointly fitting a new set of coefficients, it is expected that both models can be improved since they are constructed in a similar way as the IMRPhenomD model.
+For example, they also use PN approximant as part of the ansatz in the inspiral segment.
+One interesting result might arise while recalibrating IMRPhenomX model \cite{pratten2020setting}.
+Since it is parameterized by an additional anti-symmetric spin parameter, it is expected to not show the same degeneracy as described above.
+Further analysis might give insights into the systematics of Phenom models.
+Moreover, this method can be applied to other GW model families, such as NR surrogate models \cite{varma2019surrogate} or EOB models \cite{taracchini2014effective}.
+NR waveform calibration procedures could be made easier and are likely to improve current models.
+
+% Can we further check systematic bias by injection? But how can we know the real mass or spin in the first place if we disregard GW models.
+
+\section{Conclusion} \label{sec:conclusion}
+
+In this paper, we have presented a systematic method to recalibrate GW models.
+This method utilizes {\jax}'s automatic differentiation to apply derivative-based optimization to recalibrate GW models jointly.
+Using the new implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in conjunction with NR waveforms from the SXS catalog, we recalibrate waveform coefficients of the IMRPhenomD model.
+In general, the waveform accuracy can be improved by 50\%.
+Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement.
+In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148} and \ref{fig:q148_fl}.
+By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
+
+Furthermore, we investigated how the intrinsic parameters affect the improvement.
+Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a certain preference for waveforms lying in the second quadrant while the fourth quadrant cannot be improved.
+To further test this result, we recalibrate waveforms in separate regions in parameter space.
+As shown in Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further improvement to the second quadrant while the fourth quadrant shows similar result.
+This indicates that the model's ansatz does not fit waveforms in the fourth quadrant.
+This phenomenon is due to the reduced spin approximation used in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
+
+While we naively separate the optimization process into 4 quadrants, one can perform systematic region-selection.
+In principle, we can apply this general method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP models.
+Then, we can perform all the above analyses to understand how to construct better GW Phenom models in the future.
+
+% From analyzing mismatches in parameter space, we investigated the accuracy of generated waveforms based on their mass ratio and spin.
+Similar studies can be performed on other GW models to give better criteria to
+
+% Improved the IMRPhenomD model. Found regions that can be improved. Checked ansatz.
+
+\section{ACKNOWLEDGMENTS}
 

### 2023-02-25 src/scripts/intrin_space.py
test script

@@ -1,0 +1,21 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+train_data = np.loadtxt(
+    paths.data / "ps_train.txt", unpack=True
+)
+
+test_data = np.loadtxt(
+    paths.data / "q148.txt", unpack=True
+)
+
+# train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
+# test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
+plt.scatter(test_data[:,0], test_data[:,1], s=30, label='testing')
+plt.scatter(train_data[:,0], train_data[:,1], s=100, label='training')
+plt.legend()
+plt.xlabel('$q$')
+plt.ylabel('$\chi_{\mathrm{PN}}$')
+plt.savefig(paths.figures / "intrin_space.pdf", bbox_inches="tight")

### 2023-02-25 src/scripts/intrin_space.py
script test 2

@@ -3,13 +3,8 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
-train_data = np.loadtxt(
-    paths.data / "ps_train.txt", unpack=True
-)
-
-test_data = np.loadtxt(
-    paths.data / "q148.txt", unpack=True
-)
+train_data = np.loadtxt(paths.data / "ps_train.dat")
+test_data = np.loadtxt(paths.data / "q148.dat")
 
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]

### 2023-02-26 src/tex/ms.tex
minor change
test
test script
minor fix

@@
 \usepackage{algpseudocode}
-\usepackage[dvipsnames, usenames]{xcolor}
 \usepackage{tabularx,booktabs,multirow}
@@
 
-\definecolor{rb4}{HTML}{27408B}
-\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
-\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
+%\definecolor{rb4}{HTML}{27408B}
+%\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+%\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
 \newcommand{\ripple}{\texttt{ripple}}
@@
 In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results.
-This new O4 run is expected to double the rate of current binary black hole (BBH) observations \cite{abbott2020prospects}.
+This new O4 run is expected to double the rate of current binary black hole (BBH) observations \citep{abbott2020prospects}.
 Additionally, the sensitivity of interferometers will be increased to capture more details of GW.
@@
 Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses.
-Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \cite{purrer2020gravitational}.
+Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \citep{purrer2020gravitational}.
 Hence, it is necessary for us to develop and improve GW models.
@@
 Currently, three families of GW models are commonly used.
-They are the effective-one-body (EOB) \cite{taracchini2014effective}, Numerical Relativity (NR) surrogate \cite{varma2019surrogate} and phenomenological (Phenom) models \cite{husa2016frequency,khan2016frequency}.
+They are the effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models \citep{husa2016frequency,khan2016frequency}.
 EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations.
@@
 Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations.
-In {\python}, packages including \texttt{pytorch} \cite{pytorch}, \texttt{tensorflow} \cite{tensorflow2015-whitepaper}, etc.
+In {\python}, packages including \texttt{pytorch} \citep{pytorch}, \texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc.
 utilizes AD to train machine learning models.
@@
 Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure.
-{\ripple} \cite{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
+{\ripple} \citep{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
 It was first implemented in \texttt{lalsuite} using \texttt{C}.
@@
 We also show how the optimization result differs with waveforms of different intrinsic parameters.
-In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \cite{khan2016frequency}.
+In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \citep{khan2016frequency}.
 We also explain how reduced spin parameterization affects the accuracy of the model.
@@
 IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events.
-It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \cite{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses.
-% Using {\ripple} \cite{ripple} and {\jax} \cite{jax2018github}, we are able to easily recalibrate the model. 
+It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses.
+% Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
 IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
@@
 Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations.
-In the case of \cite{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms.
+In the case of \citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms.
 For each NR simulation, they can obtain one set of $\Lambda$'s through optimization.
@@
 \ref{eq:Lambda}.
-However, this piece-wise fitting deployed by \cite{khan2016frequency} causes calibrated coefficients to have systematic biases.
+However, this piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated coefficients to have systematic biases.
 Since coefficients were not fitted jointly, correlations between segments are omitted.
@@
 
-To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \cite{husa2016frequency}. It is defined as 
+To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
 \begin{align} \label{eq:mismatch}
@@
 
-In the original calibration \cite{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
+In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
 The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \cite{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
+Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
 We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown.
@@
 In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms.
-To compute such mismatches, NR waveforms from the SXS catalog are taken \cite{boyle2019sxs}.
+To compute such mismatches, NR waveforms from the SXS catalog are taken \citep{boyle2019sxs}.
 These NR waveforms are calculated using the \texttt{SpEC} code.
@@
 We use mismatch over the set of all training waveforms as the loss function.
-While in \cite{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
+While in \citep{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
 The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest.
@@
 We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process.
-Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \cite{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation.
+Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation.
 As BAM waveforms are not publicly available, we cannot take the identical training set as them.
@@
 They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
-Notice the maximum mass ratio in our training set is lower than that in \cite{khan2016frequency}.
+Notice the maximum mass ratio in our training set is lower than that in \citep{khan2016frequency}.
 Instead of having $q\leq18$, the maximum mass ratio is set as 8.
@@
 \end{algorithm}
-Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \cite{khan2016frequency}.
+Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \citep{khan2016frequency}.
 We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
@@
 \begin{figure}[t]
+	\script{intrin_space.py}
 	\centering
@@
 \begin{table}[t]
+	\centering
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
@@
 Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration.
-\cite{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do have access.}
+\citep{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do not have access.}
 	\label{tab:q148}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/0154_waveform_comparison.pdf}
+	\includegraphics[width=\textwidth]{../static/0154_waveform_comparison.pdf}
 	\caption{Comparison between original and optimized IMRPhenomD waveforms.
@@
 In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error.
-We have chosen one of the testing waveforms listed in \cite{khan2016frequency} for a fair comparison.
+We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison.
 % In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/q148.pdf}
+	\includegraphics[width=\columnwidth]{../static/q148.pdf}
 	\caption{Distributions of waveform mismatches.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/q148_fl.pdf}
+	\includegraphics[width=\columnwidth]{../static/q148_fl.pdf}
 	\caption{Distributions of waveform mismatches.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
+	\includegraphics[width=\columnwidth]{../static/q148_q1248_compare.pdf}
 	\caption{Distributions of $\log_{10}$ difference in mismatch.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
+	\includegraphics[width=\columnwidth]{../static/ps_q148_qchi.pdf}
 	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/ps_q148.pdf}
+	\includegraphics[width=\columnwidth]{../static/ps_q148.pdf}
 	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. This shows the same result as Fig.~\ref{fig:ps_q148_qchi}, except the axes are different.}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
+	\includegraphics[width=\columnwidth]{../static/all_quadrants.pdf}
 	\caption{Distributions of mismatches after optimizing in separate quadrants.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrant.pdf}
+	\includegraphics[width=\columnwidth]{../static/ps_q148_quadrant.pdf}
 	\caption{Parameter space of testing waveforms. Each quadrant is fitted independently. Colorbar represents log difference of mismatches before and after optimization.}
@@
 While our work focused mainly on the IMRPhenomD model, this simple yet general method can be utilized in other differentiable GW models.
-For instance, within the same family, IMRPhenomX \cite{pratten2020setting} or IMRPhenomP \cite{hannam2014simple} models.
+For instance, within the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP \citep{hannam2014simple} models.
 By jointly fitting a new set of coefficients, it is expected that both models can be improved since they are constructed in a similar way as the IMRPhenomD model.
 For example, they also use PN approximant as part of the ansatz in the inspiral segment.
-One interesting result might arise while recalibrating IMRPhenomX model \cite{pratten2020setting}.
+One interesting result might arise while recalibrating IMRPhenomX model \citep{pratten2020setting}.
 Since it is parameterized by an additional anti-symmetric spin parameter, it is expected to not show the same degeneracy as described above.
 Further analysis might give insights into the systematics of Phenom models.
-Moreover, this method can be applied to other GW model families, such as NR surrogate models \cite{varma2019surrogate} or EOB models \cite{taracchini2014effective}.
+Moreover, this method can be applied to other GW model families, such as NR surrogate models \citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}.
 NR waveform calibration procedures could be made easier and are likely to improve current models.

### 2023-02-26 src/scripts/0154.py
add more scripts

@@ -1,0 +1,45 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "0154_compare.dat")
+
+f_uniform = data[:,0]
+NR_amp = data[:,1]
+NR_angle = data[:,2]
+IMR_amp = data[:,3]
+IMR_angle = data[:,4]
+IMR_opt_amp = data[:,5]
+IMR_opt_angle = data[:,6]
+
+with plt.style.context(['seaborn-paper']):
+    fig, ax = plt.subplots(2, 2, figsize=(15, 10))
+
+    ax[0, 0].loglog(f_uniform, NR_amp, label='NR')
+    ax[0, 0].loglog(f_uniform, IMR_amp, label='IMR original')
+    ax[0, 0].loglog(f_uniform, IMR_opt_amp, label='IMR optimized')
+    ax[0, 0].set_xlabel(r'$Mf$')
+    ax[0, 0].set_ylabel(r'$|\tilde{h}(f)|$')
+    ax[0, 0].legend()
+
+    ax[0, 1].plot(f_uniform, NR_angle, label='NR')
+    ax[0, 1].plot(f_uniform, IMR_angle, label='IMR original')
+    ax[0, 1].plot(f_uniform, IMR_opt_angle, label='IMR optimized')
+    ax[0, 1].set_xlabel(r'$Mf$')
+    ax[0, 1].set_ylabel(r'$\phi$')
+    ax[0, 1].legend()
+
+    ax[1, 0].plot(f_uniform, np.abs(NR_amp - IMR_amp) / NR_amp, label='IMR original')
+    ax[1, 0].plot(f_uniform, np.abs(NR_amp - IMR_opt_amp) / NR_amp, label='IMR optimized')
+    ax[1, 0].set_xlabel(r'$Mf$')
+    ax[1, 0].set_ylabel(r'$\Delta|\tilde{h}(f)|$')
+    ax[1, 0].legend()
+
+    ax[1, 1].plot(f_uniform, np.abs(NR_angle - IMR_angle), label='IMR original')
+    ax[1, 1].plot(f_uniform, np.abs(NR_angle - IMR_opt_angle), label='IMR optimized')
+    ax[1, 1].set_xlabel(r'$Mf$')
+    ax[1, 1].set_ylabel(r'$\Delta\phi$')
+    ax[1, 1].legend()
+
+plt.savefig(paths.figures / "0154.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/all_quadrants.py

@@ -1,0 +1,15 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "all_quadrants.dat")
+
+bins = np.arange(-5, -1, 0.25)
+plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.legend()
+plt.ylabel('Count')
+plt.xlabel('$\log(\mathrm{mismatch})$')
+
+plt.savefig(paths.figures / "all_quadrants.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/ps_q148_chi1chi2.py

@@ -1,0 +1,16 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "q148.dat")
+
+logdiff = np.log10(data[:,5] / data[:,4])
+cm = plt.cm.get_cmap('RdYlBu')
+colors = logdiff
+sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.xlabel('$\chi_1$')
+plt.ylabel('$\chi_2$')
+
+plt.savefig(paths.figures / "ps_q148_qchi.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/ps_q148_qchi.py

@@ -1,0 +1,16 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "q148.dat")
+
+logdiff = np.log10(data[:,5] / data[:,4])
+cm = plt.cm.get_cmap('RdYlBu')
+colors = logdiff
+sc = plt.scatter(data[:,0], data[:,1], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.xlabel('$q$')
+plt.ylabel('$\chi_{\mathrm{PN}}$')
+
+plt.savefig(paths.figures / "ps_q148_qchi.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/ps_q148_quadrants.py

@@ -1,0 +1,16 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "q148_quadrants.dat")
+
+logdiff = np.log10(data[:,5] / data[:,4])
+cm = plt.cm.get_cmap('RdYlBu')
+colors = logdiff
+sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.xlabel('$\chi_1$')
+plt.ylabel('$\chi_2$')
+
+plt.savefig(paths.figures / "ps_q148_quadrants.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/q148.py

@@ -1,0 +1,14 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "q148.dat")
+
+bins = np.arange(-5, -1, 0.25)
+plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.legend()
+plt.ylabel('Count')
+plt.xlabel('$\log(\mathrm{mismatch})$')
+plt.savefig(paths.figures / "q148.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/q148_fl.py

@@ -1,0 +1,14 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+data = np.loadtxt(paths.data / "q148_fl.dat")
+
+bins = np.arange(-5, -1, 0.25)
+plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.legend()
+plt.ylabel('Count')
+plt.xlabel('$\log(\mathrm{mismatch})$')
+plt.savefig(paths.figures / "q148_fl.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/q148_q1248_compare.py

@@ -1,0 +1,19 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+q148_data = np.loadtxt(paths.data / "q148.dat")
+q1248_data = np.loadtxt(paths.data / "q1248.dat")
+
+bins = np.arange(-2, 2, 0.125)
+q148_logdiff = np.log10(q148_data[5]/q148_data[4])
+q1248_logdiff = np.log10(q1248_data[5]/q1248_data[4])
+
+plt.hist(q148_logdiff, bins=bins, alpha=0.6, label='q148')
+plt.hist(q1248_logdiff, bins=bins, alpha=0.6, label='q1248')
+
+plt.legend()
+plt.ylabel('Count')
+plt.xlabel('$\log(\mathrm{mismatch})$')
+plt.savefig(paths.figures / "q148_q1248_compare.pdf", bbox_inches="tight")

### 2023-02-26 src/scripts/q148_q1248_compare.py
fix q1248

@@ -7,8 +7,8 @@ q148_data = np.loadtxt(paths.data / "q148.dat")
 q1248_data = np.loadtxt(paths.data / "q1248.dat")
 
 bins = np.arange(-2, 2, 0.125)
-q148_logdiff = np.log10(q148_data[5]/q148_data[4])
-q1248_logdiff = np.log10(q1248_data[5]/q1248_data[4])
+q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
+q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:.4])
 
 plt.hist(q148_logdiff, bins=bins, alpha=0.6, label='q148')
 plt.hist(q1248_logdiff, bins=bins, alpha=0.6, label='q1248')

### 2023-02-26 src/scripts/q148_q1248_compare.py
fix q1248

@@ -8,7 +8,7 @@ q1248_data = np.loadtxt(paths.data / "q1248.dat")
 
 bins = np.arange(-2, 2, 0.125)
 q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
-q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:.4])
+q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
 
 plt.hist(q148_logdiff, bins=bins, alpha=0.6, label='q148')
 plt.hist(q1248_logdiff, bins=bins, alpha=0.6, label='q1248')

### 2023-02-27 src/scripts/ps_q148_chi1chi2.py
fix q1248

@@ -13,4 +13,4 @@ plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', ro
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 
-plt.savefig(paths.figures / "ps_q148_qchi.pdf", bbox_inches="tight")
+plt.savefig(paths.figures / "ps_q148_chi1chi2.pdf", bbox_inches="tight")

### 2023-02-27 src/scripts/all_quadrants.py
fix

@@ -6,8 +6,8 @@ import paths
 data = np.loadtxt(paths.data / "all_quadrants.dat")
 
 bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.hist(np.log10(data[4]), bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[5]), bins=bins, alpha=0.6, label='optimized')
 plt.legend()
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')

### 2023-02-27 src/scripts/all_quadrants.py
fix

@@ -6,8 +6,8 @@ import paths
 data = np.loadtxt(paths.data / "all_quadrants.dat")
 
 bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[4]), bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[5]), bins=bins, alpha=0.6, label='optimized')
+plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
 plt.legend()
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')

### 2023-02-27 src/scripts/q148.py
fix all

@@ -6,8 +6,8 @@ import paths
 data = np.loadtxt(paths.data / "q148.dat")
 
 bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
 plt.legend()
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')

### 2023-02-27 src/scripts/q148_fl.py

@@ -6,8 +6,8 @@ import paths
 data = np.loadtxt(paths.data / "q148_fl.dat")
 
 bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[4]).values, bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[5].values), bins=bins, alpha=0.6, label='optimized')
+plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
+plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
 plt.legend()
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')

### 2023-02-27 src/tex/ms.tex
add more scripts
fix figure
fix name
figure style

@@
 \begin{figure*}[t]
+	\script{0154.py}
 	\centering
-	\includegraphics[width=\textwidth]{../static/0154_waveform_comparison.pdf}
+	\includegraphics[width=\textwidth]{figures/0154.pdf}
 	\caption{Comparison between original and optimized IMRPhenomD waveforms.
@@
 \begin{figure}[t]
+	\script{q148.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/q148.pdf}
-	\caption{Distributions of waveform mismatches.
-We used $\mathcal{L}_{\mathrm{mean}}$ in optimization with training waveforms listed in Tab.~\ref{tab:q148}.
-Mismatches are weighted with the constant noise spectrum.
-Comparing the distributions, the median has decreased by $50.0\%$.}
+	\includegraphics[width=\columnwidth]{figures/q148.pdf}
+	\caption{Distributions of waveform mismatches calculated using both $\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in recalibration.
+We use training waveforms listed in Tab.~\ref{tab:q148} and mismatches are weighted with the constant noise spectrum.
+For the $\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\% while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by 22.9\%.}
 	\label{fig:q148}
@@
 \begin{figure}[t]
+	\script{q148_q1248_compare.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/q148_fl.pdf}
-	\caption{Distributions of waveform mismatches.
-We used $\mathcal{L}_{\mathrm{fl}}$ in optimization with training waveforms listed in Tab.~\ref{tab:q148}.
-Mismatches are weighted with the constant noise spectrum.
-The median has decreased by $22.9\%$.}
-	\label{fig:q148_fl}
-\end{figure}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\columnwidth]{../static/q148_q1248_compare.pdf}
+	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
 	\caption{Distributions of $\log_{10}$ difference in mismatch.
@@
 \begin{figure}[t]
+	\script{ps_q148_qchi.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/ps_q148_qchi.pdf}
+	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
 	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$.
-We use the same mismatch data as Fig.~\ref{fig:q148}.
+We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.
 Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.}
@@
 \begin{figure}[t]
+	\script{ps_q148_chi1chi2.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/ps_q148.pdf}
-	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. This shows the same result as Fig.~\ref{fig:ps_q148_qchi}, except the axes are different.}
+	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
+	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$.
+We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
 	\label{fig:ps_q148}
@@
 \begin{figure}[t]
+	\script{all_quadrants.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/all_quadrants.pdf}
+	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
 	\caption{Distributions of mismatches after optimizing in separate quadrants.
@@
 \begin{figure}[t]
+	\script{ps_q148_quadrants.py}
 	\centering
-	\includegraphics[width=\columnwidth]{../static/ps_q148_quadrant.pdf}
+	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
 	\caption{Parameter space of testing waveforms. Each quadrant is fitted independently. Colorbar represents log difference of mismatches before and after optimization.}

### 2023-02-27 src/scripts/0154.py

@@ -3,6 +3,38 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 15,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+
+mpl.rcParams.update(params)
+
 data = np.loadtxt(paths.data / "0154_compare.dat")
 
 f_uniform = data[:,0]
@@ -13,33 +45,32 @@ IMR_angle = data[:,4]
 IMR_opt_amp = data[:,5]
 IMR_opt_angle = data[:,6]
 
-with plt.style.context(['seaborn-paper']):
-    fig, ax = plt.subplots(2, 2, figsize=(15, 10))
+fig, ax = plt.subplots(2, 2, figsize=(15, 10))
 
-    ax[0, 0].loglog(f_uniform, NR_amp, label='NR')
-    ax[0, 0].loglog(f_uniform, IMR_amp, label='IMR original')
-    ax[0, 0].loglog(f_uniform, IMR_opt_amp, label='IMR optimized')
-    ax[0, 0].set_xlabel(r'$Mf$')
-    ax[0, 0].set_ylabel(r'$|\tilde{h}(f)|$')
-    ax[0, 0].legend()
+ax[0, 0].loglog(f_uniform, NR_amp, label='NR')
[…]

### 2023-02-27 src/scripts/all_quadrants.py

@@ -3,11 +3,44 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
-data = np.loadtxt(paths.data / "all_quadrants.dat")
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
+q148 = np.loadtxt(paths.data / "all_quadrants.dat")
 
 bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
+plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
+plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0")
+plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.3, color="C1")
+plt.hist(np.log10(q148[:,5]), bins=bins, label="Optimized", histtype="step", color="C1")
 plt.legend()
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')

### 2023-02-27 src/scripts/intrin_space.py

@@ -3,6 +3,38 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 18,
+    "legend.fontsize": 15,
+    "legend.frameon": True,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+
+mpl.rcParams.update(params)
+
 train_data = np.loadtxt(paths.data / "ps_train.dat")
 test_data = np.loadtxt(paths.data / "q148.dat")
 

### 2023-02-27 src/scripts/ps_q148_chi1chi2.py

@@ -3,13 +3,44 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
 data = np.loadtxt(paths.data / "q148.dat")
 
 logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 

### 2023-02-27 src/scripts/ps_q148_qchi.py

@@ -3,6 +3,37 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
 data = np.loadtxt(paths.data / "q148.dat")
 
 logdiff = np.log10(data[:,5] / data[:,4])

### 2023-02-27 src/scripts/ps_q148_quadrants.py

@@ -3,13 +3,44 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
 data = np.loadtxt(paths.data / "q148_quadrants.dat")
 
 logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 

### 2023-02-27 src/scripts/q148.py

@@ -3,11 +3,47 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
-data = np.loadtxt(paths.data / "q148.dat")
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
 
-bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
+mpl.rcParams.update(params)
+
+q148 = np.loadtxt(paths.data / "q148.dat")
+q148_fl = np.loadtxt(paths.data / "q148_fl.dat")
+bins = np.arange(-5, -1, 0.2)
+
+plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
+plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0")
+plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.05, color="C1")
+plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1")
+plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C3")
[…]

### 2023-02-27 src/scripts/q148_fl.py

@@ -1,14 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib as mpl
-import paths
-
-data = np.loadtxt(paths.data / "q148_fl.dat")
-
-bins = np.arange(-5, -1, 0.25)
-plt.hist(np.log10(data[:,4]), bins=bins, alpha=0.6, label='original')
-plt.hist(np.log10(data[:,5]), bins=bins, alpha=0.6, label='optimized')
-plt.legend()
-plt.ylabel('Count')
-plt.xlabel('$\log(\mathrm{mismatch})$')
-plt.savefig(paths.figures / "q148_fl.pdf", bbox_inches="tight")

### 2023-02-27 src/scripts/q148_q1248_compare.py

@@ -6,14 +6,16 @@ import paths
 q148_data = np.loadtxt(paths.data / "q148.dat")
 q1248_data = np.loadtxt(paths.data / "q1248.dat")
 
-bins = np.arange(-2, 2, 0.125)
+bins = np.arange(-2, 2, 0.2)
 q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
 q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
 
-plt.hist(q148_logdiff, bins=bins, alpha=0.6, label='q148')
-plt.hist(q1248_logdiff, bins=bins, alpha=0.6, label='q1248')
+plt.hist(q148_logdiff, bins=bins, alpha=0.3, color="C0")
+plt.hist(q148_logdiff, bins=bins, label="q148", histtype="step", color="C0")
+plt.hist(q1248_logdiff, bins=bins, alpha=0.3, color="C1")
+plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1")
 
-plt.legend()
+plt.legend(prop={"family":"Courier New"})
 plt.ylabel('Count')
 plt.xlabel('$\log(\mathrm{mismatch})$')
 plt.savefig(paths.figures / "q148_q1248_compare.pdf", bbox_inches="tight")

### 2023-03-06 src/tex/ms.tex
fix algorithm

@@
 
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{verbatim}
-\usepackage{mathtools}
-\usepackage{needspace}
-\usepackage{enumitem}
-\usepackage{etoolbox}
 \usepackage{graphicx}
-\usepackage{microtype}
-\usepackage{afterpage}
-\usepackage{bm}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+\usepackage{algorithmic}
+
+\SetKwInput{Parameters}{Parameters}
+\SetKwInput{Variables}{Variables}
+
 \usepackage{tabularx,booktabs,multirow}
@@
 With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
-\begin{algorithm}[H]
-	\hspace*{\algorithmicindent} \textbf{Input} $\mathcal{L}_{\mathrm{ini}}$
-	\begin{algorithmic}[1]
-		\State{$\mathcal{L}(\lambda) \gets \mathcal{L}_{\mathrm{ini}}$}
-		\State{$\alpha \gets 10^{-6}$}
-		\State{$i \gets 0$}
-		\While{$i<30000$}
-		\State{$\mathcal{L}'(\lambda) \gets \texttt{jax.grad}(\mathcal{L})$}
-		\State{$\lambda \gets \lambda-\alpha\mathcal{L}'(\lambda)$}
-		\EndWhile
-	\end{algorithmic}
-	\hspace*{\algorithmicindent} \textbf{Output} $\lambda$
-	\caption{Gradient Descent}
+\begin{algorithm}
+	\caption{Gradient descent pseudocode}
+	\KwIn{initial coefficients $\lambda_i$}
+	\Parameters{number of iterations $N$, learning rate $\alpha$}
+	\Variables{current coefficients $\lambda$, mismatch gradient $\nabla\mathcal{L}$}
+	\KwResult{output coefficients $\lambda$}
+	$\lambda\leftarrow\lambda_i$\\
+	\tcc{Gradient Descent}
+	\For{$i<N$}{
+		$\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
+		$\nabla\mathcal{L}\leftarrow AutoDiff(\mathcal{L})$\\
+		$\lambda\leftarrow\lambda-\alpha\nabla\mathcal{L}$\\
+	}
+	\Return{$\lambda$}
 \end{algorithm}
-Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \citep{khan2016frequency}.
+Here, $\lambda_i$ are the original coefficients given in \citep{khan2016frequency}.
 We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
@@
 Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement.
-We see that in Fig.~\ref{fig:q148_fl}, the median of distribution has a $22.9\%$ decrease.
+We see that the median of distribution has a $22.9\%$ decrease.
 However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
@@
 Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch.
-The shape of the distribution is similar to Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}.
+The shape of the distribution is similar to Fig.~\ref{fig:q148}.
 This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch.
@@
 Most waveforms even turned worse after optimization.
-These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}.
+These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
 
@@
 Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement.
-In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148} and \ref{fig:q148_fl}.
+In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148}.
 By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.

### 2023-03-06 src/scripts/all_quadrants.py

@@ -36,7 +36,7 @@ mpl.rcParams.update(params)
 
 q148 = np.loadtxt(paths.data / "all_quadrants.dat")
 
-bins = np.arange(-5, -1, 0.25)
+bins = np.arange(-5, -1, 0.2)
 plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
 plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0")
 plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.3, color="C1")

### 2023-03-06 src/scripts/ps_q148_qchi.py

@@ -40,7 +40,7 @@ logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,0], data[:,1], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=15)
+plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
 plt.xlabel('$q$')
 plt.ylabel('$\chi_{\mathrm{PN}}$')
 

### 2023-03-13 src/tex/ms.tex
Wrap text

@@
 \begin{abstract}
-    We present a simple and general method of recalibrating gravitational wave (GW) phenomenological waveform models jointly.
-By using {\jax} and {\ripple}, we can perform automatic differentiation to functions, which allows us to use gradient-based optimization methods to recalibrate waveform coefficients in IMRPhenomD model.
-This method reduces systematic bias previously introduced to the model and generally can improve waveform accuracy.
-With recalibrated coefficients, we found that the typical \textit{mismatch} has a $50\%$ decrease.
-Furthermore, we analyze the accuracy base on the waveform's intrinsic parameters.
-We found that waveform accuracy has significant dependence on black hole spin.
-Reduced spin approximation introduces degeneracy in spin, which prevented further improvement.
-We isolated regions in the parameter space that does not fit the waveform ansatz.
-These results allow us to understand more about how to develop newer phenomenological models.
+    We present a simple and general method of recalibrating gravitational wave
+    (GW) phenomenological waveform models jointly. By using {\jax} and
+    {\ripple}, we can perform automatic differentiation to functions, which
+    allows us to use gradient-based optimization methods to recalibrate waveform
+    coefficients in IMRPhenomD model. This method reduces systematic bias
+    previously introduced to the model and generally can improve waveform
+    accuracy. With recalibrated coefficients, we found that the typical
+    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
+    accuracy base on the waveform's intrinsic parameters. We found that waveform
+    accuracy has significant dependence on black hole spin. Reduced spin
+    approximation introduces degeneracy in spin, which prevented further
+    improvement. We isolated regions in the parameter space that does not fit
+    the waveform ansatz. These results allow us to understand more about how to
+    develop newer phenomenological models. 
 \end{abstract}
@@
 
-In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results.
-This new O4 run is expected to double the rate of current binary black hole (BBH) observations \citep{abbott2020prospects}.
-Additionally, the sensitivity of interferometers will be increased to capture more details of GW.
-Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information.
-Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses.
-Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \citep{purrer2020gravitational}.
-Hence, it is necessary for us to develop and improve GW models.
+In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
+will finish its maintenance and start observing new gravitational wave (GW)
+results. This new O4 run is expected to double the rate of current binary black
+hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
+sensitivity of interferometers will be increased to capture more details of GW.
+Having instruments with higher sensitivity, GW models of equal or higher
+accuracy than observations should be used to extract GW information. Otherwise,
+the extracted information would be affected more by GW models instead of
+interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
+GW models are accurate enough for current analyses, the accuracy of current
+models will no longer suffice for future data analyses
+\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
+improve GW models.
 
 % Waveform families and slight motivation of why we take Phenom model
-Currently, three families of GW models are commonly used.
-They are the effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models \citep{husa2016frequency,khan2016frequency}.
-EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations.
-While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations.
-This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run.
-This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.
+Currently, three families of GW models are commonly used. They are the
+effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
+(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
+\citep{husa2016frequency,khan2016frequency}. EOB models are constructed by
+mapping two masses onto an effective body under an effective metric; NR
+surrogate models construct waveforms using combinations of NR waveforms; Phenom
+models are formulated using specific ansatz and inspiral approximations. While
+EOB and NR surrogate models give better waveform approximants, Phenom waveforms
+can be produced much faster, hence it is used mostly in data analysis tasks that
+requires many waveform generations. This advantage scales up in data analysis
+tasks such as matched filtering and parameter estimation, where many waveforms
+are required in each run. This motivates us to improve upon the current
+framework of Phenom models, thus can retain the advantage of fast waveform
+generation while improving the model's accuracy.
 
-Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision.
-In traditional numerical calculations, derivatives are usually obtained through numerical derivatives.
-Symbolic derivatives were available but it was less efficient.
-Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations.
-In {\python}, packages including \texttt{pytorch} \citep{pytorch}, \texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc.
-utilizes AD to train machine learning models.
-AD's algorithm is intuitive in nature.
-Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations.
-Since these operations are fundamental, they were saved as pairs internally.
-Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative.
-Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure.
-{\ripple} \citep{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
-It was first implemented in \texttt{lalsuite} using \texttt{C}.
-In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD.
-Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses.
+Automatic differentiation (AD) is a method to calculate derivatives of functions
+up to machine precision. In traditional numerical calculations, derivatives are
+usually obtained through numerical derivatives. Symbolic derivatives were
+available but it was less efficient. Both methods were not viable in machine
+learning, where back-propagation requires precise and rapid derivative
+calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch},
+\texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train
+machine learning models. AD's algorithm is intuitive in nature. Functions
+defined are decomposed into tree structures of primitives, such as addition or
+function evaluations. Since these operations are fundamental, they were saved as
+pairs internally. Differentiation proceeds forward following the tree structure,
+with the application of the chain rule in each step to evaluate its derivative.
+Analytic derivatives of such operations are applied in each step and the desired
+derivative can then be obtained by composing back the original function
+according to the original structure. {\ripple} \citep{ripple} was a new
+implementation of IMRPhenomD, one of the Phenom models. It was first implemented
+in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
+rewritten using \jax, another {\python} package that supports AD. Using
+{\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
+allowing one to freely use derivative-based algorithms to perform data analyses. 
 % Need to cite references for AD
 
-In this paper, we investigate the possibility of further improving the accuracy of IMRPhenomD by jointly optimizing all the fitting coefficients given NR waveforms, and what constraints one may face when trying to further improve.
-We find that simply by applying gradient descent algorithm, one can obtain a better set of waveform coefficients, thus improving the accuracy of the model.
-Furthermore, by comparing the accuracy of optimized and original waveforms, we find that model-generated waveforms are very sensitive to their intrinsic parameters.
-Specifically, IMRPhenomD favors certain parts of the parameter space.
-This means IMRPhenomD introduces systematic bias to other GW analysis tasks.
-This showcases the flaws of the ansatz and allows us to have a deeper understanding of Phenom models.
+In this paper, we investigate the possibility of further improving the accuracy
+of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
+waveforms, and what constraints one may face when trying to further improve. We
+find that simply by applying gradient descent algorithm, one can obtain a better
+set of waveform coefficients, thus improving the accuracy of the model.
+Furthermore, by comparing the accuracy of optimized and original waveforms, we
+find that model-generated waveforms are very sensitive to their intrinsic
+parameters. Specifically, IMRPhenomD favors certain parts of the parameter
+space. This means IMRPhenomD introduces systematic bias to other GW analysis
+tasks. This showcases the flaws of the ansatz and allows us to have a deeper
+understanding of Phenom models.  
 % \kw{Briefly describe the high level finding in this paper}
 
-The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model.
-We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms.
-Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}.
-In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms.
-We also show how the optimization result differs with waveforms of different intrinsic parameters.
-In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \citep{khan2016frequency}.
-We also explain how reduced spin parameterization affects the accuracy of the model.
+The rest of the paper is structured as follows: In
+Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and
+have a detailed explanation of how waveform coefficients are introduced to the
+model. We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained
+how we calculate it in conjunction with NR waveforms. Then, we describe the
+optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}. In
+Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
+of optimized waveforms with original waveforms. We also show how the
+optimization result differs with waveforms of different intrinsic parameters. In
+Sec.~\ref{sec:discussion}, we address the difference between our calibrating
+procedure with \citep{khan2016frequency}. We also explain how reduced spin
+parameterization affects the accuracy of the model.
 
@@
 
-IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events.
-It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses.
+IMRPhenomD is a frequency-domain phenomenological model for modeling
+spin-aligned BBH coalescence events. It was developed as an overhaul of previous
+Phenom models, such as  IMRPhenomB/C
+\citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW
+data analyses. 
 % Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
@@
 \end{align}
-In each segment, the amplitude and phase are made using simple functions of frequency such as polynomials or lorentzians.
-These simple analytic functions consists of parameters $\Lambda^i$, which are defined as follows.
+In each segment, the amplitude and phase are made using simple functions of
+frequency such as polynomials or lorentzians. These simple analytic functions
+consists of parameters $\Lambda^i$, which are defined as follows.
 \begin{align} \label{eq:Lambda}
@@
 \end{align}
-$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters including symmetric mass ratio $\eta$ and post-newtonian spin parameter $\chi_{\mathrm{PN}}$, 
+$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters
+including symmetric mass ratio $\eta$ and post-newtonian spin parameter
+$\chi_{\mathrm{PN}}$, 
 \begin{align}
@@
 \end{align}
-Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller black hole respectively.
-Finally, the individual segments are connected in a way that the final waveform is continuous in its first derivative.
+Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller
+black hole respectively. Finally, the individual segments are connected in a way
+that the final waveform is continuous in its first derivative.
 
-From Eq.
-\ref{eq:Lambda}, we can see that the values of $\lambda$ are fundamental to waveform generation, which can significantly affect the shape of final waveforms.
-Thus, having a set of accurate waveform coefficients is important.
-Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations.
-In the case of \citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms.
-For each NR simulation, they can obtain one set of $\Lambda$'s through optimization.
-With many sets of $\Lambda$, $\lambda$ are subsequently found by fitting against Eq.
-\ref{eq:Lambda}.
-However, this piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated coefficients to have systematic biases.
-Since coefficients were not fitted jointly, correlations between segments are omitted.
-Hence, the model introduces systematic biases to $\lambda$ and will be accumulated and passed along to the data analysis pipeline.
-Also, their calibration procedure does not guarantee to have the optimal set of $\lambda$, since the process of connecting segments alters the previously fitted waveform.
-The model-generated waveforms do not take this effect into account, and thus introduces additional inaccuracies.
+From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are
+fundamental to waveform generation, which can significantly affect the shape of
+final waveforms. Thus, having a set of accurate waveform coefficients is
+important. Generally, waveform coefficients are obtained by calibrating with NR
+waveforms, which are waveforms computed using NR simulations. In the case of
+\citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated
+waveforms to NR waveforms. For each NR simulation, they can obtain one set of
+$\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are
+subsequently found by fitting against Eq. \ref{eq:Lambda}. However, this
+piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated
+coefficients to have systematic biases. Since coefficients were not fitted
+jointly, correlations between segments are omitted. Hence, the model introduces
+systematic biases to $\lambda$ and will be accumulated and passed along to the
+data analysis pipeline. Also, their calibration procedure does not guarantee to
+have the optimal set of $\lambda$, since the process of connecting segments
+alters the previously fitted waveform. The model-generated waveforms do not take
+this effect into account, and thus introduces additional inaccuracies.
 
@@
 
-Instead, we recalibrate coefficients jointly, which we can remove inaccuracies and biases discussed above, and can improve model accuracy.
-In the past, due to the complex nature of GW strains and piece-wise formalism of IMRPhenomD, non-linear fitting was difficult to be performed in optimizing coefficients.
-Hence, piece-wise optimization was done to obtain coefficients.
-However, with {\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus allowing the use of gradient-based algorithms for us to recalibrate the model.
+Instead, we recalibrate coefficients jointly, which we can remove inaccuracies
+and biases discussed above, and can improve model accuracy. In the past, due to
+the complex nature of GW strains and piece-wise formalism of IMRPhenomD,
+non-linear fitting was difficult to be performed in optimizing coefficients.
+Hence, piece-wise optimization was done to obtain coefficients. However, with
+{\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus
+allowing the use of gradient-based algorithms for us to recalibrate the model.
 
@@
 
-To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
+To quantify the difference between waveforms, we use the standard
+\textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is
+defined as 
 \begin{align} \label{eq:mismatch}
@@
 \end{align}
-where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
+where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are
+time shift and phase shift respectively. The inner product is defined as 
 \begin{align}\label{eq:inner_prod}
@@
 \end{align}
-where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.
+where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
+$S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
+the maximum and minimum frequencies in GW frequency spectra.
 
-In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
-The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
-We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown.
-This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length.
-When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.
+In the original calibration \citep{khan2016frequency,husa2016frequency},
+training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
+(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
+waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since most NR waveforms used
+have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
+which they are long enough to contain part of the inspiral segment and all
+merger and ringdown frequency information. We take the frequency limits as
+$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,
+where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
+the IMRPhenomD's frequency range, except the minimum frequency is set higher
+than that in the original calibration due to NR length. When compared with
+IMRPhenomC, the frequency range is slightly extended to have a higher maximum
+frequency.
 
-We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned high-power (\zdethp) noise spectrum and a constant spectrum.
-Since in any data analysis tasks, model-generated waveforms will ultimately be used to compare with GW observations, they are subjected to the sensitivity of detectors.
-Hence, it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than a non-weighted spectrum.
-However, with sensitivity curve involved, we have to choose a corresponding mass scale.
-Here, we simply choose $M=50M_{\odot}$ for all waveforms, since this mass scale is common in LIGO observations.
-On the other hand, we also recalibrated coefficients with the constant spectrum for a more general purpose.
+We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned
+high-power (\zdethp) noise spectrum and a constant spectrum. Since in any data
+analysis tasks, model-generated waveforms will ultimately be used to compare
+with GW observations, they are subjected to the sensitivity of detectors. Hence,
+it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than
+a non-weighted spectrum. However, with sensitivity curve involved, we have to
+choose a corresponding mass scale. Here, we simply choose $M=50M_{\odot}$ for
+all waveforms, since this mass scale is common in LIGO observations. On the
+other hand, we also recalibrated coefficients with the constant spectrum for a
+more general purpose.
 
 % GW strain treatment from time-series to frequency domain
-In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms.
-To compute such mismatches, NR waveforms from the SXS catalog are taken \citep{boyle2019sxs}.
-These NR waveforms are calculated using the \texttt{SpEC} code.
-Initially, NR waveforms are in the form of time-series data.
-Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
-Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters.
-We taper the time-series using Tukey window.
-\footnote{
-	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
-} 
-Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
+In our work, mismatches are used as a measure of discrepancies between
+IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms
+from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are
+calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form
+of time-series data. Since time-series data is oscillatory, performing
+optimization in the time-domain is not ideal. Hence, we had chosen to transform
+NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
+same intrinsic parameters. We taper the time-series using Tukey window.
+\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
+$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
+entire BBH event. } Then, the frequency spectra can be obtained by taking the
+Fourier transform of the time-series. Then, mismatches can be calculated through
+Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
 
@@
 
-Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus, we can utilize simple gradient descent to recalibrate the model.
+Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus,
+we can utilize simple gradient descent to recalibrate the model.
 
-% \kw{Mention the main advantage in this study is we can optimize everything at once.}
+% \kw{Mention the main advantage in this study is we can optimize everything at
+% once.}
 
-% To recalibrate the model, we have to define loss functions that quantify the difference between the model-generated waveforms and NR waveforms.
-In the original calibration, they took the mean square error as the loss function for optimization.
-Instead, we directly use mismatch as our metric for loss functions.
-However, the mismatch of individual waveforms is not useful in optimization, since it does not represent all waveforms with various intrinsic parameters.
-We ought to calibrate coefficients that suit waveforms with different intrinsic parameters simultaneously, hence waveforms in different regions of parameter space should be used in the optimization scheme.
-Here, we chose two different loss functions to examine their ability to optimize the model, e.g.
-mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and fractional loss $\mathcal{L}_{\mathrm{fl}}$.
+% To recalibrate the model, we have to define loss functions that quantify the
+% difference between the model-generated waveforms and NR waveforms. In the
+% original calibration, they took the mean square error as the loss function for
+% optimization. Instead, we directly use mismatch as our metric for loss
+% functions. However, the mismatch of individual waveforms is not useful in
+% optimization, since it does not represent all waveforms with various intrinsic
+% parameters. We ought to calibrate coefficients that suit waveforms with
+% different intrinsic parameters simultaneously, hence waveforms in different
+% regions of parameter space should be used in the optimization scheme. Here, we
+% chose two different loss functions to examine their ability to optimize the
+% model, e.g. mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and
+% fractional loss $\mathcal{L}_{\mathrm{fl}}$.
 
 % Do something much more direct:
-To recalibrate the model, we need to define a loss function that quantifies the difference between the model-generated waveforms and NR waveforms.
-We use mismatch over the set of all training waveforms as the loss function.
-While in \citep{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
-The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest.
-One of which takes a simple average and the other takes a normalized average based on initial mismatch.
+To recalibrate the model, we need to define a loss function that quantifies the
+difference between the model-generated waveforms and NR waveforms. We use
+mismatch over the set of all training waveforms as the loss function. While in
+\citep{khan2016frequency}, they chose to use the mean square error as the loss
+function, we choose mismatch as the loss function because it is closely related
+to downstream objectives such as the likelihood in parameter estimation. The
+mismatch is averaged over the training dataset in two ways to cover the
+parameter space of interest. One of which takes a simple average and the other
+takes a normalized average based on initial mismatch.
 
 % To recalibrate the model, we need to define a loss function that quantify ...
-We use mismatch over the set of all training waveforms as the loss function.
-We choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
-The mismatch is averaged over the training dataset to cover the parameter space of interest.
-(used)
+% We use mismatch over the set of all training waveforms as the loss function.
+% We choose mismatch as the loss function because it is closely related to
+% downstream objectives such as the likelihood in parameter estimation. The
+% mismatch is averaged over the training dataset to cover the parameter space of
+% interest. (used)
 \begin{align}
@@
 \end{align}	
-$\mathcal{M}_i$ represents the mismatch of an individual training waveform, $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual training waveform, and $N$ is the total number of individual training waveforms.
-For the first choice, it takes the mean of the mismatches, which serves as a basic loss function.
-However, it is prone to be dominated by a single waveform with a large mismatch.
-Other waveforms with smaller mismatches would be insignificant comparatively, and might not be able to improve under such optimization.
-Alternatively, the second choice, the mean of normalized mismatches, eliminates the aforementioned issue.
-Nevertheless, it excludes the information on initial mismatches.
-$\mathcal{L}_{\mathrm{fl}}$ restricts every training waveform to decrease at similar rates, hence it is hard to obtain optimized waveforms with mismatches in the same order of magnitude.
-Instead, their ratios in mismatches would remain approximately the same.
-Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust and final individual mismatches would be in a similar order of magnitude.
-In this paper, we display the results of using both loss functions and examine the difference between them.
+$\mathcal{M}_i$ represents the mismatch of an individual training waveform,
+$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
+training waveform, and $N$ is the total number of individual training waveforms.
+For the first choice, it takes the mean of the mismatches, which serves as a
+basic loss function. However, it is prone to be dominated by a single waveform
+with a large mismatch. Other waveforms with smaller mismatches would be
+insignificant comparatively, and might not be able to improve under such
+optimization. Alternatively, the second choice, the mean of normalized
+mismatches, eliminates the aforementioned issue. Nevertheless, it excludes the
+information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
+training waveform to decrease at similar rates, hence it is hard to obtain
+optimized waveforms with mismatches in the same order of magnitude. Instead,
+their ratios in mismatches would remain approximately the same. Conversely,
+$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+and final individual mismatches would be in a similar order of magnitude. In
+this paper, we display the results of using both loss functions and examine the
+difference between them.
 
-We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process.
-Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation.
-As BAM waveforms are not publicly available, we cannot take the identical training set as them.
-Instead, we take the available waveforms from the SXS catalog to construct our loss function.
-They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
-Notice the maximum mass ratio in our training set is lower than that in \citep{khan2016frequency}.
-Instead of having $q\leq18$, the maximum mass ratio is set as 8.
-This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
-In fact, the SXS catalog only has NR waveforms with $q\leq10$.
-Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed have $q\leq8$.
-Hence, we set the maximum mass ratio to 8.
+We have chosen 11-16 NR waveforms from the set of waveforms used in the original
+calibration process. Originally, 19 waveforms are taken from NR simulations for
+calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they
+are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not
+publicly available, we cannot take the identical training set as them. Instead,
+we take the available waveforms from the SXS catalog to construct our loss
+function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the
+maximum mass ratio in our training set is lower than that in
+\citep{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio
+is set as 8. This is because SXS catalog does not have NR waveforms with
+extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with
+$q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model
+with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the
+maximum mass ratio to 8.
 
-With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
+With both mismatches, we apply gradient descent to optimize the tunable
+coefficients as follows: 
 \begin{algorithm}
 	\caption{Gradient descent pseudocode}
-	\KwIn{initial coefficients $\lambda_i$}
-	\Parameters{number of iterations $N$, learning rate $\alpha$}
-	\Variables{current coefficients $\lambda$, mismatch gradient $\nabla\mathcal{L}$}
-	\KwResult{output coefficients $\lambda$}
-	$\lambda\leftarrow\lambda_i$\\
+	\KwIn{initial coefficients $\lambda_i$} \Parameters{number of iterations
+	$N$, learning rate $\alpha$} \Variables{current coefficients $\lambda$,
+	mismatch gradient $\nabla\mathcal{L}$} \KwResult{output coefficients
+	$\lambda$} $\lambda\leftarrow\lambda_i$\\
 	\tcc{Gradient Descent}
-	\For{$i<N$}{
-		$\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
+	\For{$i<N$}{ $\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
 		$\nabla\mathcal{L}\leftarrow AutoDiff(\mathcal{L})$\\
 		$\lambda\leftarrow\lambda-\alpha\nabla\mathcal{L}$\\
-	}
-	\Return{$\lambda$}
+	} \Return{$\lambda$}
 \end{algorithm}
-Here, $\lambda_i$ are the original coefficients given in \citep{khan2016frequency}.
-We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
-With such an optimization procedure, we obtain the optimized coefficients.
+Here, $\lambda_i$ are the original coefficients given in
+\citep{khan2016frequency}. We take them as the initial waveform coefficients
+because they lie in the neighborhood of the minimum that we wish to find. With
+such an optimization procedure, we obtain the optimized coefficients.
 
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule
-		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
-		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
+		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
+		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
 		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
@@
 	\end{tabularx}
-	\caption{List of waveforms used to recalibrate the model.
-The mass ratio $q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$.
-Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration.
-\citep{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do not have access.}
+	\caption{List of waveforms used to recalibrate the model. The mass ratio
+	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
+	here, 9 of them are also used in the original IMRPhenomD calibration.
+	\citep{khan2016frequency} The two remaining waveforms were from BAM
+	simulation, to which we do not have access.}
 	\label{tab:q148}
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule
-		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
-		SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
+		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
+		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
 		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
@@
 	\caption{Comparison between original and optimized IMRPhenomD waveforms.
-Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and $\chi_1=\chi_2=-0.8$.
-The original mismatch is around $2.8\times10^{-4}$ and the optimized mismatch is around $5.3\times10^{-5}$.
-Top: It shows the amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD waveform.
-Bottom: It shows the relative error of amplitudes between NR and IMRPhenomD waveforms, and the absolute error of phases between NR and IMRPhenomD waveforms}
+	Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and
+	$\chi_1=\chi_2=-0.8$. The original mismatch is around $2.8\times10^{-4}$ and
+	the optimized mismatch is around $5.3\times10^{-5}$. Top: It shows the
+	amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD
+	waveform. Bottom: It shows the relative error of amplitudes between NR and
+	IMRPhenomD waveforms, and the absolute error of phases between NR and
+	IMRPhenomD waveforms}
 	\label{fig:0154}
@@
 
-First, to judge the performance of the optimization scheme, we take $\sim530$ NR waveforms from the SXS catalog.
-We choose testing waveforms with negligible eccentricity (${e<2\times10^{-3}}$) and precession (${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model.
-We see that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters that are within the parameter space spanned by training waveforms.
-Hence, we can take these testing waveforms to compare with the original model.
+First, to judge the performance of the optimization scheme, we take $\sim530$ NR
+waveforms from the SXS catalog. We choose testing waveforms with negligible
+eccentricity (${e<2\times10^{-3}}$) and precession
+(${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model. We see
+that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters
+that are within the parameter space spanned by training waveforms. Hence, we can
+take these testing waveforms to compare with the original model.
 
-We first examine the effect of joint optimization on a single waveform.
-In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error.
-We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison.
-% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.
+We first examine the effect of joint optimization on a single waveform. In
+Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy
+than the original waveform, especially in the inspiral region, where the
+amplitude has a $50\%$ decrease in error. We have chosen one of the testing
+waveforms listed in \citep{khan2016frequency} for a fair comparison. 
+% In general, most waveforms within the neighborhood of the training waveforms
+% show improvement in mismatches.
 
-Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig.
-\ref{fig:q148}.
-We find the distribution's peak has shifted to the end with a lower mismatch.
-Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
-Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement.
-We see that the median of distribution has a $22.9\%$ decrease.
+Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute
+the mismatch for all testing waveforms and plot the distribution of the mismatch
+in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end
+with a lower mismatch. Quantitatively, the peak has lowered by almost an
+order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
+Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar
+improvement. We see that the median of distribution has a $22.9\%$ decrease.
 However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
 This is due to the problem mentioned in section \ref{subsec:optimization}.
-Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.
+Nevertheless, both distributions show improvement and the potential problem did
+not affect the optimization scheme significantly.
 
-% \kw{We then compute the mismatch for all testing waveforms and plot the distribution of the mismatch in figure X. We find the distribution...}
+% \kw{We then compute the mismatch for all testing waveforms and plot the
+% distribution of the mismatch in figure X. We find the distribution...}
 
@@
 	\includegraphics[width=\columnwidth]{figures/q148.pdf}
-	\caption{Distributions of waveform mismatches calculated using both $\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in recalibration.
-We use training waveforms listed in Tab.~\ref{tab:q148} and mismatches are weighted with the constant noise spectrum.
-For the $\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\% while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by 22.9\%.}
+	\caption{Distributions of waveform mismatches calculated using both
+	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in
+	recalibration. We use training waveforms listed in Tab.~\ref{tab:q148} and
+	mismatches are weighted with the constant noise spectrum. For the
+	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\%
+	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
+	22.9\%.}
 	\label{fig:q148}
@@
 	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
-	\caption{Distributions of $\log_{10}$ difference in mismatch.
-The distribution labeled \texttt{q148} uses training waveforms listed in Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
-Mismatches are calculated using the constant noise spectrum with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
+	\caption{Distributions of $\log_{10}$ difference in mismatch. The
+	distribution labeled \texttt{q148} uses training waveforms listed in
+	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
+	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
+	using the constant noise spectrum with the loss function
+	$\mathcal{L}_{\mathrm{mean}}$.}
 	\label{fig:q148_q1248_compare}
@@
 
-Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch.
-The shape of the distribution is similar to Fig.~\ref{fig:q148}.
-This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch.
-Thus, it should be a model that fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum instead of the constant spectrum.
+Similarly, using the same procedures, distributions of mismatches calculated
+using {\zdethp} noise spectrum show better improvement than the weighted
+mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}. This
+effect is expected since the IMRPhenomD model was originally constructed and
+fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that
+fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum
+instead of the constant spectrum.
 
-With the success of improving waveforms, we increase the number of training waveforms for optimization.
-Taking additional waveforms listed in Tab.~\ref{tab:q1248}, we obtain a new set of coefficients.
-We see in Fig.
-\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small improvement.
-The high mismatch tail of the optimized distribution remains similar in length and endpoint as the original distribution, meaning that they cannot be improved using our procedure.
-Likewise, using additional waveforms to optimize loss function with {\zdethp} spectrum shows the same result, where the new distribution barely shows any improvement.
+With the success of improving waveforms, we increase the number of training
+waveforms for optimization. Taking additional waveforms listed in
+Tab.~\ref{tab:q1248}, we obtain a new set of coefficients. We see in Fig.
+\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small
+improvement. The high mismatch tail of the optimized distribution remains
+similar in length and endpoint as the original distribution, meaning that they
+cannot be improved using our procedure. Likewise, using additional waveforms to
+optimize loss function with {\zdethp} spectrum shows the same result, where the
+new distribution barely shows any improvement.
 
-% We believe that either the ansatz does not suit waveforms within certain regions of parameter space or the set of optimized coefficients falls to a minimum where they do not fit some testing waveforms.
-If in some region of parameter space, the ansatz does not fit well with NR waveforms, it would never be able to have consistent improvement under optimization.
-Instead, it would surf around the loss manifold with random fluctuations in mismatches.
-If the set of coefficients does not fit, then dividing the parameter space into regions and fitting separate sets of coefficients should give better results.
+% We believe that either the ansatz does not suit waveforms within certain
+% regions of parameter space or the set of optimized coefficients falls to a
+% minimum where they do not fit some testing waveforms. If in some region of
+% parameter space, the ansatz does not fit well with NR waveforms, it would
+% never be able to have consistent improvement under optimization. Instead, it
+% would surf around the loss manifold with random fluctuations in mismatches. If
+% the set of coefficients does not fit, then dividing the parameter space into
+% regions and fitting separate sets of coefficients should give better results.
 
-Given the ansatz used in the waveform model is unlikely to be fully compatible with NR, and the optimization procedure is done over a distribution of waveforms with different source parameters, it is conceivable that there are some trade-off in accuracy of the waveform models between different part of the parameter space.
-If this is the reason why the high mismatch tail is not improving during the joint-optimization, separating the parameter space into smaller subspace should help alleviate this issue.
-On the other hand, if the ansatz does not have the right parameterized form to capture the behavior of the NR waveforms as a function of the intrinsic parameters, the result should be always biased, and we should not expect any improvement even if we separate the parameter space into smaller space during training.
+Given the ansatz used in the waveform model is unlikely to be fully compatible
+with NR, and the optimization procedure is done over a distribution of waveforms
+with different source parameters, it is conceivable that there are some
+trade-off in accuracy of the waveform models between different part of the
+parameter space. If this is the reason why the high mismatch tail is not
+improving during the joint-optimization, separating the parameter space into
+smaller subspace should help alleviate this issue. On the other hand, if the
+ansatz does not have the right parameterized form to capture the behavior of the
+NR waveforms as a function of the intrinsic parameters, the result should be
+always biased, and we should not expect any improvement even if we separate the
+parameter space into smaller space during training.
 
-Since we know intrinsic parameters play an important role in the ansatz, we would like to investigate how intrinsic parameters affect the recalibration process.
-First, we plot the parameter space of $q$ vs.
-$\chi_{\mathrm{PN}}$ in Fig.
-\ref{fig:ps_q148_qchi}.
-In the low mass ratio region, waveforms with both positive and negative log differences are mixed up.
-On the other hand, along the horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve under recalibration.
+Since we know intrinsic parameters play an important role in the ansatz, we
+would like to investigate how intrinsic parameters affect the recalibration
+process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
+Fig. \ref{fig:ps_q148_qchi}. In the low mass ratio region, waveforms with both
+positive and negative log differences are mixed up. On the other hand, along the
+horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve
+under recalibration.
 
-Furthermore, we plot the parameter space of $\chi_1$ vs.
-$\chi_2$ in Fig.~\ref{fig:ps_q148}.
-Waveforms along the diagonal axis, i.e.
-$\chi_1\approx\chi_2$, show good mismatch improvements.
-We expect to see this feature since the original coefficients were fitted using NR waveforms with equal or similar spin, hence the model prefers waveforms with similar spin.
-One interesting feature is how the second and fourth quadrants respond to optimization.
-In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms generally improve with along optimization.
-However, mismatches in the fourth quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization.
-Most waveforms even turned worse after optimization.
-These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
+Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
+Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e.
+$\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this
+feature since the original coefficients were fitted using NR waveforms with
+equal or similar spin, hence the model prefers waveforms with similar spin. One
+interesting feature is how the second and fourth quadrants respond to
+optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
+generally improve with along optimization. However, mismatches in the fourth
+quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most
+waveforms even turned worse after optimization. These waveforms correspond to
+the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
 
-% Waveform mismatches in the fourth quadrant, i.e.
-$\chi_1>0$ and $\chi_2<0$, performed worse after optimized.
-IMRPhenomD is constructed using single spin parameter, where there are degeneracy between $\chi_1$ and $\chi_2$.
-In regions with opposite extreme spin components, degeneracy significantly alters intrinsic parameters passed into waveform generation, hence the fourth quadrant shows such feature.
+% Waveform mismatches in the fourth quadrant, i.e. $\chi_1>0$ and $\chi_2<0$,
+% performed worse after optimized. IMRPhenomD is constructed using single spin
+% parameter, where there are degeneracy between $\chi_1$ and $\chi_2$. In
+% regions with opposite extreme spin components, degeneracy significantly alters
+% intrinsic parameters passed into waveform generation, hence the fourth
+% quadrant shows such feature.
 
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
-	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$.
-We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.
-Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.}
+	\caption{Parameter space of testing waveforms of
+	$q$~vs.~$\chi_{\mathrm{PN}}$. We use the recalibrated result from
+	$\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training
+	waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the
+	$\log_{10}$ difference between optimized and original unweighted
+	mismatches.}
 	\label{fig:ps_q148_qchi}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
-	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$.
-We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
+	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. We
+	use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the
+	constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
 	\label{fig:ps_q148}
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule
-		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
-		SXS:BBH:0172 & 1.0 & 0.98     & 0.98     \\
+		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
+		\midrule\midrule SXS:BBH:0172 & 1.0 & 0.98     & 0.98     \\
 		SXS:BBH:0152 & 1.0 & 0.60     & 0.60     \\
@@
 		SXS:BBH:1426 & 8.0 & 0.48     & 0.75     \\
-		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule
-		SXS:BBH:0370 & 1.0 & -0.20    & 0.40     \\
+		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule SXS:BBH:0370 & 1.0
+		& -0.20    & 0.40     \\
 		SXS:BBH:2092 & 1.0 & -0.50    & 0.50     \\
@@
 		SXS:BBH:2133 & 3.0 & -0.73    & 0.85     \\
-		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule
-		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
+		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule SXS:BBH:0156 & 1.0
+		& -0.95    & -0.95    \\
 		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
@@
 		SXS:BBH:1419 & 8.0 & -0.80    & -0.80    \\
-		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule
-		SXS:BBH:0304 & 1.0 & 0.50     & -0.50    \\
+		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule SXS:BBH:0304 & 1.0
+		& 0.50     & -0.50    \\
 		SXS:BBH:0327 & 1.0 & 0.80     & -0.80    \\
@@
 
-As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is significantly different in different regions in the parameter space, we split the parameter space into 4 quadrants and perform separate fitting with training waveforms listed in Table \ref{tab:quadrants}.
-Note that in the second and fourth quadrants, there are not enough waveforms with $q>4$, hence the result are only valid up to $q\leq4$.
-From Fig.~\ref{fig:ps_q148_quadrant}, we see that all waveforms, except those in the fourth quadrant, show improvement in mismatch.
-Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
+As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is
+significantly different in different regions in the parameter space, we split
+the parameter space into 4 quadrants and perform separate fitting with training
+waveforms listed in Table \ref{tab:quadrants}. Note that in the second and
+fourth quadrants, there are not enough waveforms with $q>4$, hence the result
+are only valid up to $q\leq4$. From Fig.~\ref{fig:ps_q148_quadrant}, we see that
+all waveforms, except those in the fourth quadrant, show improvement in
+mismatch. Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
 First, we focus on waveforms with the same spin direction, i.e.
-$\chi_1\chi_2>0$.
-Waveforms lying in the neighborhood of the diagonal axis have noticeable improvements in mismatch due to most training waveforms being equal-spin waveforms.
-For the second quadrant, waveforms improved significantly with only a few defects due to some testing waveforms having $q>4$.
-In the fourth quadrant, most optimized waveforms have a higher mismatch than the original waveforms, as indicated by the positive log difference of mismatches.
-As performing optimization in a smaller subspace does not show any improvement compared to the original result, the ansatz mostly does not fit waveforms in the fourth quadrant.
+$\chi_1\chi_2>0$. Waveforms lying in the neighborhood of the diagonal axis have
+noticeable improvements in mismatch due to most training waveforms being
+equal-spin waveforms. For the second quadrant, waveforms improved significantly
+with only a few defects due to some testing waveforms having $q>4$. In the
+fourth quadrant, most optimized waveforms have a higher mismatch than the
+original waveforms, as indicated by the positive log difference of mismatches.
+As performing optimization in a smaller subspace does not show any improvement
+compared to the original result, the ansatz mostly does not fit waveforms in the
+fourth quadrant.
 
@@
 	\caption{Distributions of mismatches after optimizing in separate quadrants.
-We use a constant noise spectrum to calculate mismatch and $\mathcal{L}_{\mathrm{mean}}$ for the loss function.
-Generally, most waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant.
-}
+	We use a constant noise spectrum to calculate mismatch and
+	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. Generally, most
+	waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant. }
 	\label{fig:all_quadrants}
@@
 
-%\begin{enumerate}
-%	\item Comparison with the original paper is not the same. (Different waveforms, different ways to modify the waveform, different range of frequencies)
-%	\item More waveforms are taken so it should be more robust, since the parameter space coverage is better.
-%    \item Problems with IMRPhenomD model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat)
-%    \item Can in principle split parameter space into regions that can be separately optimized.
-If not, then its must be the model's problem.
-Would the region classification be made automatic.
-Are there any algorithms to do so?
-%    \item Future work (Repeat on other GW models. Mapping between NR surrogate to see any systematics.)
-%\end{enumerate}
+%\begin{enumerate} \item Comparison with the original paper is not the same.
+%   (Different waveforms, different ways to modify the waveform, different range
+%   of frequencies) \item More waveforms are taken so it should be more robust,
+%   since the parameter space coverage is better. \item Problems with IMRPhenomD
+%   model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space
+%   $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat) \item
+%   Can in principle split parameter space into regions that can be separately
+%   optimized. If not, then its must be the model's problem. Would the region
+%   classification be made automatic. Are there any algorithms to do so? \item
+%   Future work (Repeat on other GW models. Mapping between NR surrogate to see
+%   any systematics.) \end{enumerate}
 
-We have shown the result of recalibrating waveform coefficients.
-One thing to note is that our recalibration procedure is not exactly the same as the original calibration.
-For instance, we use a different set of NR waveforms, frequency range, etc.
-Nonetheless, as the decrease in mismatch is rather significant, this optimization procedure should be able to improve the accuracy of IMRPhenomD on a similar scale regardless of the differences.
-Here, this result serves as a demonstration of the general method used.
-% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for both amplitude and phase.
-Similarly, for the merger-ringdown region, the optimized waveforms also show great improvement.
+We have shown the result of recalibrating waveform coefficients. One thing to
+note is that our recalibration procedure is not exactly the same as the original
+calibration. For instance, we use a different set of NR waveforms, frequency
+range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
+optimization procedure should be able to improve the accuracy of IMRPhenomD on a
+similar scale regardless of the differences. Here, this result serves as a
+demonstration of the general method used.  
+% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for
+% both amplitude and phase. Similarly, for the merger-ringdown region, the
+% optimized waveforms also show great improvement.
 
-We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with more training waveforms only has a small increase in the accuracy of the waveform model.
-We believe that by increasing the number of waveforms, the accuracy will not have a significant change as the waveform coefficients are already over-determined.
-Using more calibration NR waveforms will not further improve the model significantly.
-This suggests the form of the parameterized ansatz is not suitable for certain regions in the parameter space, thus mismatches of only a few waveforms decreased while other waveforms remain at the high mismatch tail with little changes.
-This implies that the model is ultimately restricted by the flexibility of the ansatz.
+We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with
+more training waveforms only has a small increase in the accuracy of the
+waveform model. We believe that by increasing the number of waveforms, the
+accuracy will not have a significant change as the waveform coefficients are
+already over-determined. Using more calibration NR waveforms will not further
+improve the model significantly. This suggests the form of the parameterized
+ansatz is not suitable for certain regions in the parameter space, thus
+mismatches of only a few waveforms decreased while other waveforms remain at the
+high mismatch tail with little changes. This implies that the model is
+ultimately restricted by the flexibility of the ansatz.
 
-One of the major problems causing inaccuracy in the ansatz is the reduced spin approximation.
-In IMRPhenomD, it is modeled using a single spin parameter, namely $\chi_{\mathrm{PN}}$ as outlined in Sec.
-\ref{sec:method}.
-Parameterizing a BBH merger with one spin parameter introduces degeneracy within the parameter space.
-Events with distinct black hole spins could result in equal $\chi_{\mathrm{PN}}$, thus generating the same waveform.
-Especially with high unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as events with small spin, thus giving an inaccurate result.
-Generally, the approximation gives straight lines of degeneracy in the parameter space, with its slope (always negative) dependent on the mass ratio.
-From Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves better in the top left region than the bottom right.
-To try to accommodate this issue, we split the parameter space into 4 quadrants as described in Sec.
-\ref{sec:result}.
-However, even with separate optimizations, we see in Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar mismatches as before while the second quadrant further improved.
-This suggests the ansatz is region-specific, with a higher preference for BBH events with $\chi_1<0$ and $\chi_2>0$.
+One of the major problems causing inaccuracy in the ansatz is the reduced spin
+approximation. In IMRPhenomD, it is modeled using a single spin parameter,
+namely $\chi_{\mathrm{PN}}$ as outlined in Sec. \ref{sec:method}. Parameterizing
+a BBH merger with one spin parameter introduces degeneracy within the parameter
+space. Events with distinct black hole spins could result in equal
+$\chi_{\mathrm{PN}}$, thus generating the same waveform. Especially with high
+unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as
+events with small spin, thus giving an inaccurate result. Generally, the
+approximation gives straight lines of degeneracy in the parameter space, with
+its slope (always negative) dependent on the mass ratio. From
+Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves
+better in the top left region than the bottom right. To try to accommodate this
+issue, we split the parameter space into 4 quadrants as described in Sec.
+\ref{sec:result}. However, even with separate optimizations, we see in
+Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar
+mismatches as before while the second quadrant further improved. This suggests
+the ansatz is region-specific, with a higher preference for BBH events with
+$\chi_1<0$ and $\chi_2>0$.
 
-Separating regions into 4 quadrants is done purely out of simplicity.
-To give a more comprehensive analysis, one should be systematic about region selection.
-One can use level set estimation algorithms to obtain systematic regions of interest.
-This general algorithm reveals further degeneracies or issues within the ansatz.
-Then, recalibrating such individual regions might give better results.
-Alternatively, one can select regions according to the lines of degeneracy.
-However, with limited NR waveforms, such a selection scheme is not viable for us.
-In the future, with more NR waveforms spanning the entire parameter space, one can perform optimization with fewer restrictions.
+Separating regions into 4 quadrants is done purely out of simplicity. To give a
+more comprehensive analysis, one should be systematic about region selection.
+One can use level set estimation algorithms to obtain systematic regions of
+interest. This general algorithm reveals further degeneracies or issues within
+the ansatz. Then, recalibrating such individual regions might give better
+results. Alternatively, one can select regions according to the lines of
+degeneracy. However, with limited NR waveforms, such a selection scheme is not
+viable for us. In the future, with more NR waveforms spanning the entire
+parameter space, one can perform optimization with fewer restrictions.
 
-While our work focused mainly on the IMRPhenomD model, this simple yet general method can be utilized in other differentiable GW models.
-For instance, within the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP \citep{hannam2014simple} models.
-By jointly fitting a new set of coefficients, it is expected that both models can be improved since they are constructed in a similar way as the IMRPhenomD model.
-For example, they also use PN approximant as part of the ansatz in the inspiral segment.
-One interesting result might arise while recalibrating IMRPhenomX model \citep{pratten2020setting}.
-Since it is parameterized by an additional anti-symmetric spin parameter, it is expected to not show the same degeneracy as described above.
-Further analysis might give insights into the systematics of Phenom models.
-Moreover, this method can be applied to other GW model families, such as NR surrogate models \citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}.
-NR waveform calibration procedures could be made easier and are likely to improve current models.
+While our work focused mainly on the IMRPhenomD model, this simple yet general
+method can be utilized in other differentiable GW models. For instance, within
+the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP
+\citep{hannam2014simple} models. By jointly fitting a new set of coefficients,
+it is expected that both models can be improved since they are constructed in a
+similar way as the IMRPhenomD model. For example, they also use PN approximant
+as part of the ansatz in the inspiral segment. One interesting result might
+arise while recalibrating IMRPhenomX model \citep{pratten2020setting}. Since it
+is parameterized by an additional anti-symmetric spin parameter, it is expected
+to not show the same degeneracy as described above. Further analysis might give
+insights into the systematics of Phenom models. Moreover, this method can be
+applied to other GW model families, such as NR surrogate models
+\citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}. NR
+waveform calibration procedures could be made easier and are likely to improve
+current models.
 
-% Can we further check systematic bias by injection? But how can we know the real mass or spin in the first place if we disregard GW models.
+% Can we further check systematic bias by injection? But how can we know the
+% real mass or spin in the first place if we disregard GW models.
 
@@
 In this paper, we have presented a systematic method to recalibrate GW models.
-This method utilizes {\jax}'s automatic differentiation to apply derivative-based optimization to recalibrate GW models jointly.
-Using the new implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in conjunction with NR waveforms from the SXS catalog, we recalibrate waveform coefficients of the IMRPhenomD model.
-In general, the waveform accuracy can be improved by 50\%.
-Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement.
-In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148}.
-By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
+This method utilizes {\jax}'s automatic differentiation to apply
+derivative-based optimization to recalibrate GW models jointly. Using the new
+implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
+conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
+coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
+improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
+mismatches have a slightly better improvement. In contrast, different types of
+loss function result in significantly different final mismatch distributions,
+where the result can be seen in Fig.~\ref{fig:q148}. By increasing the number of
+training waveforms, we see a slight improvement increase in
+Fig.~\ref{fig:q148_q1248_compare}.
 
-Furthermore, we investigated how the intrinsic parameters affect the improvement.
-Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a certain preference for waveforms lying in the second quadrant while the fourth quadrant cannot be improved.
-To further test this result, we recalibrate waveforms in separate regions in parameter space.
-As shown in Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further improvement to the second quadrant while the fourth quadrant shows similar result.
-This indicates that the model's ansatz does not fit waveforms in the fourth quadrant.
-This phenomenon is due to the reduced spin approximation used in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
+Furthermore, we investigated how the intrinsic parameters affect the
+improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
+certain preference for waveforms lying in the second quadrant while the fourth
+quadrant cannot be improved. To further test this result, we recalibrate
+waveforms in separate regions in parameter space. As shown in
+Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further
+improvement to the second quadrant while the fourth quadrant shows similar
+result. This indicates that the model's ansatz does not fit waveforms in the
+fourth quadrant. This phenomenon is due to the reduced spin approximation used
+in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$
+are introduced.
 
-While we naively separate the optimization process into 4 quadrants, one can perform systematic region-selection.
-In principle, we can apply this general method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP models.
-Then, we can perform all the above analyses to understand how to construct better GW Phenom models in the future.
+While we naively separate the optimization process into 4 quadrants, one can
+perform systematic region-selection. In principle, we can apply this general
+method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
+models. Then, we can perform all the above analyses to understand how to
+construct better GW Phenom models in the future.
 
-% From analyzing mismatches in parameter space, we investigated the accuracy of generated waveforms based on their mass ratio and spin.
-Similar studies can be performed on other GW models to give better criteria to
+% From analyzing mismatches in parameter space, we investigated the accuracy of
+% generated waveforms based on their mass ratio and spin. Similar studies can be
+% performed on other GW models to give better criteria to
 
-% Improved the IMRPhenomD model. Found regions that can be improved. Checked ansatz.
+% Improved the IMRPhenomD model. Found regions that can be improved. Checked
+% ansatz.
 

### 2023-03-13 src/tex/ms.tex
update

@@
 (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
-\citep{husa2016frequency,khan2016frequency}. EOB models are constructed by
-mapping two masses onto an effective body under an effective metric; NR
-surrogate models construct waveforms using combinations of NR waveforms; Phenom
-models are formulated using specific ansatz and inspiral approximations. While
-EOB and NR surrogate models give better waveform approximants, Phenom waveforms
-can be produced much faster, hence it is used mostly in data analysis tasks that
-requires many waveform generations. This advantage scales up in data analysis
-tasks such as matched filtering and parameter estimation, where many waveforms
-are required in each run. This motivates us to improve upon the current
-framework of Phenom models, thus can retain the advantage of fast waveform
-generation while improving the model's accuracy.
+\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
+the newest models}. EOB models are constructed by mapping two masses onto an
+effective body under an effective metric; NR surrogate models construct
+waveforms using combinations of NR waveforms; Phenom models are formulated using
+specific ansatz and inspiral approximations. While EOB and NR surrogate models
+give better waveform approximants, Phenom waveforms can be produced much faster,
+hence it is used mostly in data analysis tasks that requires many waveform
+generations. This advantage scales up in data analysis tasks such as matched
+filtering and parameter estimation, where many waveforms are required in each
+run. This motivates us to improve upon the current framework of Phenom models,
+thus can retain the advantage of fast waveform generation while improving the
+model's accuracy.
 
@@
 
-The rest of the paper is structured as follows: In
-Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and
-have a detailed explanation of how waveform coefficients are introduced to the
-model. We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained
-how we calculate it in conjunction with NR waveforms. Then, we describe the
-optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}. In
+The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
+review the parameterization of the IMRPhenomD model and the mismatch function
+that is used as an objective function for the calibration, followed by outlining the
+specific optimization scheme used for recalibration. In
 Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
@@
 
-IMRPhenomD is a frequency-domain phenomenological model for modeling
-spin-aligned BBH coalescence events. It was developed as an overhaul of previous
-Phenom models, such as  IMRPhenomB/C
-\citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW
-data analyses. 
-% Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
-IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
+In order to recalibrate the model, we have to understand what parameters the model has.
+Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
+For interested readers, please refer to \cite{khan2016frequency} for more details on construction of the model.
+
+The IMRPhenomD model is constructed by combining three individually fitted parts
+into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
+\kw{Check if the following equation is correct. I think it is off by some smoothing factors.}
 \begin{align}
-	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
+	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta).
 \end{align}
-In each segment, the amplitude and phase are made using simple functions of
-frequency such as polynomials or lorentzians. These simple analytic functions
-consists of parameters $\Lambda^i$, which are defined as follows.
+
+Instead of fitting the strain, which is a highly oscillatory function that is
+difficult to fit, the amplitude and phase are fitted since they are smoother
+functions. In each part, the amplitude and phase are made using simple functions
+of frequency such as polynomials or lorentzians. These simple analytic functions
+consists of parameters $\Lambda^i$, which are defined as follows. \kw{Can you
+change equation one such that it consists of $\Lambda$? Also be explicit about
+the form of the function used for the fitting. Like write down the polynomials
+and lorentizans for the specific parts.}
 \begin{align} \label{eq:Lambda}
@@
 	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
-	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
+	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2),
 \end{align}
-$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters
-including symmetric mass ratio $\eta$ and post-newtonian spin parameter
-$\chi_{\mathrm{PN}}$, 
+where $\lambda$ are fitting coefficients obtained during calibration, $\eta$ is
+the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
+parameter, which is defined as 
 \begin{align}
@@
 \end{align}
-Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller
-black hole respectively. Finally, the individual segments are connected in a way
-that the final waveform is continuous in its first derivative.
+Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
+respectively. Finally, the individual segments are connected in a way that the
+final waveform is continuous in its first derivative \kw{expand a bit}.
 

### 2023-03-13 src/tex/ms.tex
update

@@
 
+\kw{Need to work on this paragraph}
 From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are
@@
 $\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are
-subsequently found by fitting against Eq. \ref{eq:Lambda}. However, this
-piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated
-coefficients to have systematic biases. Since coefficients were not fitted
-jointly, correlations between segments are omitted. Hence, the model introduces
-systematic biases to $\lambda$ and will be accumulated and passed along to the
-data analysis pipeline. Also, their calibration procedure does not guarantee to
-have the optimal set of $\lambda$, since the process of connecting segments
-alters the previously fitted waveform. The model-generated waveforms do not take
-this effect into account, and thus introduces additional inaccuracies.
+subsequently found by fitting against Eq. \ref{eq:Lambda}. Since the fitting
+procedure is done in a piece-wise manner, the correlations between different
+segments are omitted, which could limit the accuracy of the model. Also,
+their calibration procedure does not guarantee to have the optimal set of
+$\lambda$, since the process of connecting segments alters the previously fitted
+waveform. The model-generated waveforms do not take this effect into account,
+and thus introduces additional inaccuracies.
 

### 2023-03-16 src/tex/ms.tex
Update mismatch

@@
 
-To quantify the difference between waveforms, we use the standard
-\textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is
-defined as 
+In order to recalibrate the model, we need to define an objective function that
+quantify the performance of the model. A common choice for such metric is the
+\textit{mismatch} function \citep{husa2016frequency}. It is defined as
 \begin{align} \label{eq:mismatch}
@@
 \end{align}
-where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are
-time shift and phase shift respectively. The inner product is defined as 
+where $h_{1,2}$ are the two GW waveform we are comparing, and $t_0$ and $\phi_0$
+are time shift and phase shift respectively. The $\langle h_1, h_2 \rangle$ is
+commonly referred as the inner product, which is defined as 
 \begin{align}\label{eq:inner_prod}
-	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
+	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
 \end{align}
@@
 $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the maximum and minimum frequencies in GW frequency spectra.
+the relevant maximum and minimum frequencies for the integration. The mismatch
+is a quantity that is closely related to the mean square error (MSE) between the
+two waveforms, as it is often considered a weighted version of the MSE.
 

### 2023-03-16 src/tex/ms.tex
Reword PSD paragraph

@@
 where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
+$S_n(f)$ is the power spectral density(PSD) of noise from the instrument, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
 the relevant maximum and minimum frequencies for the integration. The mismatch
@@
 
+\kw{Remind me why don't we use the hybrid waveforms again? Cause we couldn't do it?}
 In the original calibration \citep{khan2016frequency,husa2016frequency},
@@
 
-We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned
-high-power (\zdethp) noise spectrum and a constant spectrum. Since in any data
-analysis tasks, model-generated waveforms will ultimately be used to compare
-with GW observations, they are subjected to the sensitivity of detectors. Hence,
-it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than
-a non-weighted spectrum. However, with sensitivity curve involved, we have to
-choose a corresponding mass scale. Here, we simply choose $M=50M_{\odot}$ for
-all waveforms, since this mass scale is common in LIGO observations. On the
-other hand, we also recalibrated coefficients with the constant spectrum for a
-more general purpose.
+To compute the integral shown in Eq. \ref{eq:inner_prod}, one need to choose a
+relevant noise PSD. Since a common goal for creating waveform approximent is to
+deploy the waveform model in downstream data analysis tasks, a reasonable choice
+for the noise PSD should be closely related to the instrument of interest. We
+choose the zero-detuned high-power (\zdethp) noise PSD \kw{add citation} as our
+first choice of noise PSD. Because the total mass of the system simple scales
+the frequency of the waveform, we need to choose a corresponding mass scale to
+match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
+all waveforms, since this mass scale is common in LIGO observations. In addition
+to a detector-dependent PSD, we are also interested in a more
+instrument-agnostic case such that we can evaluate the effectiveness of the
+recalibration without considering detector effect. In this case, we set the
+noise PSD to be a constant value.
 

### 2023-03-20 src/tex/ms.tex
First comments added

--- 2023-03-13 src/tex/ms.tex
@@
 
-%\definecolor{rb4}{HTML}{27408B}
+\definecolor{rr}{RGB}{173, 37, 37}
 %\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 %\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
+\newcommand{\te}[1]{{\color{rr}[TE: #1 ]}}
 \newcommand{\ripple}{\texttt{ripple}}
@@
 \begin{abstract}
-    We present a simple and general method of recalibrating gravitational wave
-    (GW) phenomenological waveform models jointly. By using {\jax} and
-    {\ripple}, we can perform automatic differentiation to functions, which
-    allows us to use gradient-based optimization methods to recalibrate waveform
-    coefficients in IMRPhenomD model. This method reduces systematic bias
-    previously introduced to the model and generally can improve waveform
-    accuracy. With recalibrated coefficients, we found that the typical
-    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
-    accuracy base on the waveform's intrinsic parameters. We found that waveform
-    accuracy has significant dependence on black hole spin. Reduced spin
-    approximation introduces degeneracy in spin, which prevented further
-    improvement. We isolated regions in the parameter space that does not fit
-    the waveform ansatz. These results allow us to understand more about how to
-    develop newer phenomenological models. 
+    We present a simple and general method of recalibrating gravitational wave (GW) phenomenological waveform models jointly.
+By using {\jax} and {\ripple}, we can perform automatic differentiation to functions, which allows us to use gradient-based optimization methods to recalibrate waveform coefficients in IMRPhenomD model.
+This method reduces systematic bias previously introduced to the model and generally can improve waveform accuracy.
+With recalibrated coefficients, we found that the typical \textit{mismatch} has a $50\%$ decrease.
+Furthermore, we analyze the accuracy base on the waveform's intrinsic parameters.
+We found that waveform accuracy has significant dependence on black hole spin.
+Reduced spin approximation introduces degeneracy in spin, which prevented further improvement.
+We isolated regions in the parameter space that does not fit the waveform ansatz.
+These results allow us to understand more about how to develop newer phenomenological models.
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
+\te{General points:}
 
-In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
-will finish its maintenance and start observing new gravitational wave (GW)
-results. This new O4 run is expected to double the rate of current binary black
-hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
-sensitivity of interferometers will be increased to capture more details of GW.
-Having instruments with higher sensitivity, GW models of equal or higher
-accuracy than observations should be used to extract GW information. Otherwise,
-the extracted information would be affected more by GW models instead of
-interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
-GW models are accurate enough for current analyses, the accuracy of current
-models will no longer suffice for future data analyses
-\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
-improve GW models.
+In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results.
+This new O4 run is expected to double the rate of current binary black hole (BBH) observations \citep{abbott2020prospects}.
+Additionally, the sensitivity of interferometers will be increased to capture more details of GW.
+Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information.
+Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses.
+Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \citep{purrer2020gravitational}.
+Hence, it is necessary for us to develop and improve GW models.
 
 % Waveform families and slight motivation of why we take Phenom model
-Currently, three families of GW models are commonly used. They are the
-effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
-(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
-\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
-the newest models}. EOB models are constructed by mapping two masses onto an
-effective body under an effective metric; NR surrogate models construct
-waveforms using combinations of NR waveforms; Phenom models are formulated using
-specific ansatz and inspiral approximations. While EOB and NR surrogate models
-give better waveform approximants, Phenom waveforms can be produced much faster,
-hence it is used mostly in data analysis tasks that requires many waveform
-generations. This advantage scales up in data analysis tasks such as matched
-filtering and parameter estimation, where many waveforms are required in each
-run. This motivates us to improve upon the current framework of Phenom models,
-thus can retain the advantage of fast waveform generation while improving the
-model's accuracy.
+Currently, three families of GW models are commonly used.
+They are the effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models \citep{husa2016frequency,khan2016frequency}.
+EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations.
+While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations.
+This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run.
+This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.
 
-Automatic differentiation (AD) is a method to calculate derivatives of functions
-up to machine precision. In traditional numerical calculations, derivatives are
-usually obtained through numerical derivatives. Symbolic derivatives were
-available but it was less efficient. Both methods were not viable in machine
-learning, where back-propagation requires precise and rapid derivative
-calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch},
-\texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train
-machine learning models. AD's algorithm is intuitive in nature. Functions
-defined are decomposed into tree structures of primitives, such as addition or
-function evaluations. Since these operations are fundamental, they were saved as
-pairs internally. Differentiation proceeds forward following the tree structure,
-with the application of the chain rule in each step to evaluate its derivative.
-Analytic derivatives of such operations are applied in each step and the desired
-derivative can then be obtained by composing back the original function
-according to the original structure. {\ripple} \citep{ripple} was a new
-implementation of IMRPhenomD, one of the Phenom models. It was first implemented
-in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
-rewritten using \jax, another {\python} package that supports AD. Using
-{\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
-allowing one to freely use derivative-based algorithms to perform data analyses. 
+Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision.
+In traditional numerical calculations, derivatives are usually obtained through numerical derivatives.
+Symbolic derivatives were available but it was less efficient.
+Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations.
+In {\python}, packages including \texttt{pytorch} \citep{pytorch}, \texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc.
+utilizes AD to train machine learning models.
+AD's algorithm is intuitive in nature.
+Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations.
+Since these operations are fundamental, they were saved as pairs internally.
+Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative.
+Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure.
+{\ripple} \citep{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
+It was first implemented in \texttt{lalsuite} using \texttt{C}.
+In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD.
+Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses.
 % Need to cite references for AD
 
-In this paper, we investigate the possibility of further improving the accuracy
-of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
-waveforms, and what constraints one may face when trying to further improve. We
-find that simply by applying gradient descent algorithm, one can obtain a better
-set of waveform coefficients, thus improving the accuracy of the model.
-Furthermore, by comparing the accuracy of optimized and original waveforms, we
-find that model-generated waveforms are very sensitive to their intrinsic
-parameters. Specifically, IMRPhenomD favors certain parts of the parameter
-space. This means IMRPhenomD introduces systematic bias to other GW analysis
-tasks. This showcases the flaws of the ansatz and allows us to have a deeper
-understanding of Phenom models.  
+In this paper, we investigate the possibility of further improving the accuracy of IMRPhenomD by jointly optimizing all the fitting coefficients given NR waveforms, and what constraints one may face when trying to further improve.
+We find that simply by applying gradient descent algorithm, one can obtain a better set of waveform coefficients, thus improving the accuracy of the model.
+Furthermore, by comparing the accuracy of optimized and original waveforms, we find that model-generated waveforms are very sensitive to their intrinsic parameters.
+Specifically, IMRPhenomD favors certain parts of the parameter space.
+This means IMRPhenomD introduces systematic bias to other GW analysis tasks.
+This showcases the flaws of the ansatz and allows us to have a deeper understanding of Phenom models.
 % \kw{Briefly describe the high level finding in this paper}
 
-The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
-review the parameterization of the IMRPhenomD model and the mismatch function
-that is used as an objective function for the calibration, followed by outlining the
-specific optimization scheme used for recalibration. In
-Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
-of optimized waveforms with original waveforms. We also show how the
-optimization result differs with waveforms of different intrinsic parameters. In
-Sec.~\ref{sec:discussion}, we address the difference between our calibrating
-procedure with \citep{khan2016frequency}. We also explain how reduced spin
-parameterization affects the accuracy of the model.
+The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model.
+We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms.
+Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}.
+In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms.
+We also show how the optimization result differs with waveforms of different intrinsic parameters.
+In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \citep{khan2016frequency}.
+We also explain how reduced spin parameterization affects the accuracy of the model.
 
@@
 
-In order to recalibrate the model, we have to understand what parameters the model has.
-Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
-For interested readers, please refer to \cite{khan2016frequency} for more details on construction of the model.
-
-The IMRPhenomD model is constructed by combining three individually fitted parts
-into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
-\kw{Check if the following equation is correct. I think it is off by some smoothing factors.}
+IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events.
+It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses.
+% Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
+IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
 \begin{align}
-	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta).
+	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
 \end{align}
-
-Instead of fitting the strain, which is a highly oscillatory function that is
-difficult to fit, the amplitude and phase are fitted since they are smoother
-functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. These simple analytic functions
-consists of parameters $\Lambda^i$, which are defined as follows. \kw{Can you
-change equation one such that it consists of $\Lambda$? Also be explicit about
-the form of the function used for the fitting. Like write down the polynomials
-and lorentizans for the specific parts.}
+In each segment, the amplitude and phase are made using simple functions of frequency such as polynomials or lorentzians.
+These simple analytic functions consists of parameters $\Lambda^i$, which are defined as follows.
 \begin{align} \label{eq:Lambda}
@@
 	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
-	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2),
+	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
 \end{align}
-where $\lambda$ are fitting coefficients obtained during calibration, $\eta$ is
-the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
-parameter, which is defined as 
+$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters including symmetric mass ratio $\eta$ and post-newtonian spin parameter $\chi_{\mathrm{PN}}$, 
 \begin{align}
@@
 \end{align}
-Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
-respectively. Finally, the individual segments are connected in a way that the
-final waveform is continuous in its first derivative \kw{expand a bit}.
+Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller black hole respectively.
+Finally, the individual segments are connected in a way that the final waveform is continuous in its first derivative.
 
-\kw{Need to work on this paragraph}
-From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are
-fundamental to waveform generation, which can significantly affect the shape of
-final waveforms. Thus, having a set of accurate waveform coefficients is
-important. Generally, waveform coefficients are obtained by calibrating with NR
-waveforms, which are waveforms computed using NR simulations. In the case of
-\citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated
-waveforms to NR waveforms. For each NR simulation, they can obtain one set of
-$\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are
-subsequently found by fitting against Eq. \ref{eq:Lambda}. Since the fitting
-procedure is done in a piece-wise manner, the correlations between different
-segments are omitted, which could limit the accuracy of the model. Also,
-their calibration procedure does not guarantee to have the optimal set of
-$\lambda$, since the process of connecting segments alters the previously fitted
-waveform. The model-generated waveforms do not take this effect into account,
-and thus introduces additional inaccuracies.
+From Eq.
+\ref{eq:Lambda}, we can see that the values of $\lambda$ are fundamental to waveform generation, which can significantly affect the shape of final waveforms.
+Thus, having a set of accurate waveform coefficients is important.
+Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations.
+In the case of \citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms.
+For each NR simulation, they can obtain one set of $\Lambda$'s through optimization.
+With many sets of $\Lambda$, $\lambda$ are subsequently found by fitting against Eq.
+\ref{eq:Lambda}.
+However, this piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated coefficients to have systematic biases.
+Since coefficients were not fitted jointly, correlations between segments are omitted.
+Hence, the model introduces systematic biases to $\lambda$ and will be accumulated and passed along to the data analysis pipeline.
+Also, their calibration procedure does not guarantee to have the optimal set of $\lambda$, since the process of connecting segments alters the previously fitted waveform.
+The model-generated waveforms do not take this effect into account, and thus introduces additional inaccuracies.
 
@@
 
-Instead, we recalibrate coefficients jointly, which we can remove inaccuracies
-and biases discussed above, and can improve model accuracy. In the past, due to
-the complex nature of GW strains and piece-wise formalism of IMRPhenomD,
-non-linear fitting was difficult to be performed in optimizing coefficients.
-Hence, piece-wise optimization was done to obtain coefficients. However, with
-{\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus
-allowing the use of gradient-based algorithms for us to recalibrate the model.
+Instead, we recalibrate coefficients jointly, which we can remove inaccuracies and biases discussed above, and can improve model accuracy.
+In the past, due to the complex nature of GW strains and piece-wise formalism of IMRPhenomD, non-linear fitting was difficult to be performed in optimizing coefficients.
+Hence, piece-wise optimization was done to obtain coefficients.
+However, with {\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus allowing the use of gradient-based algorithms for us to recalibrate the model.
 
@@
 
-To quantify the difference between waveforms, we use the standard
-\textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is
-defined as 
+To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
 \begin{align} \label{eq:mismatch}
@@
 \end{align}
-where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are
-time shift and phase shift respectively. The inner product is defined as 
+where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
 \begin{align}\label{eq:inner_prod}
@@
 \end{align}
-where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the maximum and minimum frequencies in GW frequency spectra.
+where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.
 
-In the original calibration \citep{khan2016frequency,husa2016frequency},
-training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
-(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
-waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used
-have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
-which they are long enough to contain part of the inspiral segment and all
-merger and ringdown frequency information. We take the frequency limits as
-$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,
-where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set higher
-than that in the original calibration due to NR length. When compared with
-IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency.
+In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
+The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
+We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown.
+This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length.
+When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.
+\te{What is the frequency spacing?}
 
-We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned
-high-power (\zdethp) noise spectrum and a constant spectrum. Since in any data
-analysis tasks, model-generated waveforms will ultimately be used to compare
-with GW observations, they are subjected to the sensitivity of detectors. Hence,
-it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than
-a non-weighted spectrum. However, with sensitivity curve involved, we have to
-choose a corresponding mass scale. Here, we simply choose $M=50M_{\odot}$ for
-all waveforms, since this mass scale is common in LIGO observations. On the
-other hand, we also recalibrated coefficients with the constant spectrum for a
-more general purpose.
+\te{I think this discussion needs to be much more precise.
+Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
+So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
+We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned high-power (\zdethp) noise spectrum and a constant spectrum.
+Since in any data analysis tasks, model-generated waveforms will ultimately be used to compare with GW observations, they are subjected to the sensitivity of detectors.
+Hence, it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than a non-weighted spectrum.
+However, with sensitivity curve involved, we have to choose a corresponding mass scale.
+Here, we simply choose $M=50M_{\odot}$ for all waveforms, since this mass scale is common in LIGO observations.
+On the other hand, we also recalibrated coefficients with the constant spectrum for a more general purpose.
 
 % GW strain treatment from time-series to frequency domain
-In our work, mismatches are used as a measure of discrepancies between
-IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms
-from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are
-calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form
-of time-series data. Since time-series data is oscillatory, performing
-optimization in the time-domain is not ideal. Hence, we had chosen to transform
-NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
-same intrinsic parameters. We taper the time-series using Tukey window.
-\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-entire BBH event. } Then, the frequency spectra can be obtained by taking the
-Fourier transform of the time-series. Then, mismatches can be calculated through
-Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
+In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms.
+To compute such mismatches, NR waveforms from the SXS catalog are taken \citep{boyle2019sxs}.
+These NR waveforms are calculated using the \texttt{SpEC} code.
+Initially, NR waveforms are in the form of time-series data.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters.
+We taper the time-series using Tukey window.
+\footnote{
+	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
+} 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
 
@@
 
-Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus,
-we can utilize simple gradient descent to recalibrate the model.
+Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus, we can utilize simple gradient descent to recalibrate the model.
 
-% \kw{Mention the main advantage in this study is we can optimize everything at
-% once.}
+% \kw{Mention the main advantage in this study is we can optimize everything at once.}
 
-% To recalibrate the model, we have to define loss functions that quantify the
-% difference between the model-generated waveforms and NR waveforms. In the
-% original calibration, they took the mean square error as the loss function for
-% optimization. Instead, we directly use mismatch as our metric for loss
-% functions. However, the mismatch of individual waveforms is not useful in
-% optimization, since it does not represent all waveforms with various intrinsic
-% parameters. We ought to calibrate coefficients that suit waveforms with
-% different intrinsic parameters simultaneously, hence waveforms in different
-% regions of parameter space should be used in the optimization scheme. Here, we
-% chose two different loss functions to examine their ability to optimize the
-% model, e.g. mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and
-% fractional loss $\mathcal{L}_{\mathrm{fl}}$.
+% To recalibrate the model, we have to define loss functions that quantify the difference between the model-generated waveforms and NR waveforms.
+In the original calibration, they took the mean square error as the loss function for optimization.
+Instead, we directly use mismatch as our metric for loss functions.
+However, the mismatch of individual waveforms is not useful in optimization, since it does not represent all waveforms with various intrinsic parameters.
+We ought to calibrate coefficients that suit waveforms with different intrinsic parameters simultaneously, hence waveforms in different regions of parameter space should be used in the optimization scheme.
+Here, we chose two different loss functions to examine their ability to optimize the model, e.g.
+mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and fractional loss $\mathcal{L}_{\mathrm{fl}}$.
 
 % Do something much more direct:
-To recalibrate the model, we need to define a loss function that quantifies the
-difference between the model-generated waveforms and NR waveforms. We use
-mismatch over the set of all training waveforms as the loss function. While in
-\citep{khan2016frequency}, they chose to use the mean square error as the loss
-function, we choose mismatch as the loss function because it is closely related
-to downstream objectives such as the likelihood in parameter estimation. The
-mismatch is averaged over the training dataset in two ways to cover the
-parameter space of interest. One of which takes a simple average and the other
-takes a normalized average based on initial mismatch.
+To recalibrate the model, we need to define a loss function that quantifies the difference between the model-generated waveforms and NR waveforms.
+We use mismatch over the set of all training waveforms as the loss function.
+While in \citep{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
+The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest.
+One of which takes a simple average and the other takes a normalized average based on initial mismatch.
 
 % To recalibrate the model, we need to define a loss function that quantify ...
-% We use mismatch over the set of all training waveforms as the loss function.
-% We choose mismatch as the loss function because it is closely related to
-% downstream objectives such as the likelihood in parameter estimation. The
-% mismatch is averaged over the training dataset to cover the parameter space of
-% interest. (used)
+We use mismatch over the set of all training waveforms as the loss function.
+We choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation.
+The mismatch is averaged over the training dataset to cover the parameter space of interest.
+(used)
 \begin{align}
@@
 \end{align}	
-$\mathcal{M}_i$ represents the mismatch of an individual training waveform,
-$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of individual training waveforms.
-For the first choice, it takes the mean of the mismatches, which serves as a
-basic loss function. However, it is prone to be dominated by a single waveform
-with a large mismatch. Other waveforms with smaller mismatches would be
-insignificant comparatively, and might not be able to improve under such
-optimization. Alternatively, the second choice, the mean of normalized
-mismatches, eliminates the aforementioned issue. Nevertheless, it excludes the
-information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
-training waveform to decrease at similar rates, hence it is hard to obtain
-optimized waveforms with mismatches in the same order of magnitude. Instead,
-their ratios in mismatches would remain approximately the same. Conversely,
-$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and final individual mismatches would be in a similar order of magnitude. In
-this paper, we display the results of using both loss functions and examine the
-difference between them.
+$\mathcal{M}_i$ represents the mismatch of an individual training waveform, $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual training waveform, and $N$ is the total number of individual training waveforms.
+For the first choice, it takes the mean of the mismatches, which serves as a basic loss function.
+However, it is prone to be dominated by a single waveform with a large mismatch.
+Other waveforms with smaller mismatches would be insignificant comparatively, and might not be able to improve under such optimization.
+Alternatively, the second choice, the mean of normalized mismatches, eliminates the aforementioned issue.
+Nevertheless, it excludes the information on initial mismatches.
+$\mathcal{L}_{\mathrm{fl}}$ restricts every training waveform to decrease at similar rates, hence it is hard to obtain optimized waveforms with mismatches in the same order of magnitude.
+Instead, their ratios in mismatches would remain approximately the same.
+Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust and final individual mismatches would be in a similar order of magnitude.
+In this paper, we display the results of using both loss functions and examine the difference between them.
 
-We have chosen 11-16 NR waveforms from the set of waveforms used in the original
-calibration process. Originally, 19 waveforms are taken from NR simulations for
-calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they
-are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not
-publicly available, we cannot take the identical training set as them. Instead,
-we take the available waveforms from the SXS catalog to construct our loss
-function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the
-maximum mass ratio in our training set is lower than that in
-\citep{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio
-is set as 8. This is because SXS catalog does not have NR waveforms with
-extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with
-$q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model
-with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the
-maximum mass ratio to 8.
+We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process.
+Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation.
+As BAM waveforms are not publicly available, we cannot take the identical training set as them.
+Instead, we take the available waveforms from the SXS catalog to construct our loss function.
+They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+Notice the maximum mass ratio in our training set is lower than that in \citep{khan2016frequency}.
+Instead of having $q\leq18$, the maximum mass ratio is set as 8.
+This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+In fact, the SXS catalog only has NR waveforms with $q\leq10$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed have $q\leq8$.
+Hence, we set the maximum mass ratio to 8.
 
-With both mismatches, we apply gradient descent to optimize the tunable
-coefficients as follows: 
+\te{I think we need a bit more discussion about the stopping criterion here.
+What is N, and why did we choose it to be this number?
+Ideally we would actually have a plot of the loss function during training.
+I don't have a good intuition for if its noisy.
+Does it plateau?
+If we wanted to ensure that people don't complain about overfitting we could also plot the loss of the validation set to show its not going up.}
+
+With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
 \begin{algorithm}
 	\caption{Gradient descent pseudocode}
-	\KwIn{initial coefficients $\lambda_i$} \Parameters{number of iterations
-	$N$, learning rate $\alpha$} \Variables{current coefficients $\lambda$,
-	mismatch gradient $\nabla\mathcal{L}$} \KwResult{output coefficients
-	$\lambda$} $\lambda\leftarrow\lambda_i$\\
+	\KwIn{initial coefficients $\lambda_i$}
+	\Parameters{number of iterations $N$, learning rate $\alpha$}
+	\Variables{current coefficients $\lambda$, mismatch gradient $\nabla\mathcal{L}$}
+	\KwResult{output coefficients $\lambda$}
+	$\lambda\leftarrow\lambda_i$\\
 	\tcc{Gradient Descent}
-	\For{$i<N$}{ $\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
+	\For{$i<N$}{
+		$\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
 		$\nabla\mathcal{L}\leftarrow AutoDiff(\mathcal{L})$\\
 		$\lambda\leftarrow\lambda-\alpha\nabla\mathcal{L}$\\
-	} \Return{$\lambda$}
+	}
+	\Return{$\lambda$}
 \end{algorithm}
-Here, $\lambda_i$ are the original coefficients given in
-\citep{khan2016frequency}. We take them as the initial waveform coefficients
-because they lie in the neighborhood of the minimum that we wish to find. With
-such an optimization procedure, we obtain the optimized coefficients.
+Here, $\lambda_i$ are the original coefficients given in \citep{khan2016frequency}.
+We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
+With such an optimization procedure, we obtain the optimized coefficients.
 
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
-		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
 		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
@@
 	\end{tabularx}
-	\caption{List of waveforms used to recalibrate the model. The mass ratio
-	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
-	here, 9 of them are also used in the original IMRPhenomD calibration.
-	\citep{khan2016frequency} The two remaining waveforms were from BAM
-	simulation, to which we do not have access.}
+	\caption{List of waveforms used to recalibrate the model.
+The mass ratio $q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$.
+Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration.
+\citep{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do not have access.}
 	\label{tab:q148}
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
-		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
 		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
@@
 	\caption{Comparison between original and optimized IMRPhenomD waveforms.
-	Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and
-	$\chi_1=\chi_2=-0.8$. The original mismatch is around $2.8\times10^{-4}$ and
-	the optimized mismatch is around $5.3\times10^{-5}$. Top: It shows the
-	amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD
-	waveform. Bottom: It shows the relative error of amplitudes between NR and
-	IMRPhenomD waveforms, and the absolute error of phases between NR and
-	IMRPhenomD waveforms}
+Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and $\chi_1=\chi_2=-0.8$.
+The original mismatch is around $2.8\times10^{-4}$ and the optimized mismatch is around $5.3\times10^{-5}$.
+Top: It shows the amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD waveform.
+Bottom: It shows the relative error of amplitudes between NR and IMRPhenomD waveforms, and the absolute error of phases between NR and IMRPhenomD waveforms \te{I realize that this is also an error in the ripple paper, but the bottom panels should have abs round the whole y axis, otherwise why are they strictly positive?
+I actually think just plotting the error itself i.e.
+allowing for positive and negative values would potentially look better.
+Its also useful to add the vertical lines corresponding the different waveform regions.
+Otherwise one might think that the oscillations will line up with this and its some weird artifact of the splitting.
+The top and bottom panels should probably share x axes, or at least have the same scales so that they can be compared easily.
+}}
 	\label{fig:0154}
@@
 
-First, to judge the performance of the optimization scheme, we take $\sim530$ NR
-waveforms from the SXS catalog. We choose testing waveforms with negligible
-eccentricity (${e<2\times10^{-3}}$) and precession
-(${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model. We see
-that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters
-that are within the parameter space spanned by training waveforms. Hence, we can
-take these testing waveforms to compare with the original model.
+First, to judge the performance of the optimization scheme, we take $\sim530$ NR waveforms from the SXS catalog.
+We choose testing waveforms with negligible eccentricity (${e<2\times10^{-3}}$) and precession (${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model.
+We see that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters that are within the parameter space spanned by training waveforms.
+Hence, we can take these testing waveforms to compare with the original model.
 
-We first examine the effect of joint optimization on a single waveform. In
-Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy
-than the original waveform, especially in the inspiral region, where the
-amplitude has a $50\%$ decrease in error. We have chosen one of the testing
-waveforms listed in \citep{khan2016frequency} for a fair comparison. 
-% In general, most waveforms within the neighborhood of the training waveforms
-% show improvement in mismatches.
+We first examine the effect of joint optimization on a single waveform.
+In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error.
+We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison.
+% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.
 
-Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute
-the mismatch for all testing waveforms and plot the distribution of the mismatch
-in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end
-with a lower mismatch. Quantitatively, the peak has lowered by almost an
-order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
-Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar
-improvement. We see that the median of distribution has a $22.9\%$ decrease.
+Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig.
+\ref{fig:q148}.
+We find the distribution's peak has shifted to the end with a lower mismatch.
+Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
+Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement.
+We see that the median of distribution has a $22.9\%$ decrease.
 However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
 This is due to the problem mentioned in section \ref{subsec:optimization}.
-Nevertheless, both distributions show improvement and the potential problem did
-not affect the optimization scheme significantly.
+Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.
 
-% \kw{We then compute the mismatch for all testing waveforms and plot the
-% distribution of the mismatch in figure X. We find the distribution...}
+% \kw{We then compute the mismatch for all testing waveforms and plot the distribution of the mismatch in figure X. We find the distribution...}
 
@@
 	\includegraphics[width=\columnwidth]{figures/q148.pdf}
-	\caption{Distributions of waveform mismatches calculated using both
-	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in
-	recalibration. We use training waveforms listed in Tab.~\ref{tab:q148} and
-	mismatches are weighted with the constant noise spectrum. For the
-	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\%
-	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
-	22.9\%.}
+	\caption{Distributions of waveform mismatches calculated using both $\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in recalibration.
+We use training waveforms listed in Tab.~\ref{tab:q148} and mismatches are weighted with the constant noise spectrum.
+For the $\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\% while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by 22.9\%.}
 	\label{fig:q148}
@@
 	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
-	\caption{Distributions of $\log_{10}$ difference in mismatch. The
-	distribution labeled \texttt{q148} uses training waveforms listed in
-	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
-	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant noise spectrum with the loss function
-	$\mathcal{L}_{\mathrm{mean}}$.}
+	\caption{Distributions of $\log_{10}$ difference in mismatch.
+The distribution labeled \texttt{q148} uses training waveforms listed in Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+Mismatches are calculated using the constant noise spectrum with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
 	\label{fig:q148_q1248_compare}
@@
 
-Similarly, using the same procedures, distributions of mismatches calculated
-using {\zdethp} noise spectrum show better improvement than the weighted
-mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}. This
-effect is expected since the IMRPhenomD model was originally constructed and
-fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that
-fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum
-instead of the constant spectrum.
+Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch.
+The shape of the distribution is similar to Fig.~\ref{fig:q148}.
+This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch.
+Thus, it should be a model that fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum instead of the constant spectrum.
 
-With the success of improving waveforms, we increase the number of training
-waveforms for optimization. Taking additional waveforms listed in
-Tab.~\ref{tab:q1248}, we obtain a new set of coefficients. We see in Fig.
-\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small
-improvement. The high mismatch tail of the optimized distribution remains
-similar in length and endpoint as the original distribution, meaning that they
-cannot be improved using our procedure. Likewise, using additional waveforms to
-optimize loss function with {\zdethp} spectrum shows the same result, where the
-new distribution barely shows any improvement.
+With the success of improving waveforms, we increase the number of training waveforms for optimization.
+Taking additional waveforms listed in Tab.~\ref{tab:q1248}, we obtain a new set of coefficients.
+We see in Fig.
+\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small improvement.
+The high mismatch tail of the optimized distribution remains similar in length and endpoint as the original distribution, meaning that they cannot be improved using our procedure.
+Likewise, using additional waveforms to optimize loss function with {\zdethp} spectrum shows the same result, where the new distribution barely shows any improvement.
 
-% We believe that either the ansatz does not suit waveforms within certain
-% regions of parameter space or the set of optimized coefficients falls to a
-% minimum where they do not fit some testing waveforms. If in some region of
-% parameter space, the ansatz does not fit well with NR waveforms, it would
-% never be able to have consistent improvement under optimization. Instead, it
-% would surf around the loss manifold with random fluctuations in mismatches. If
-% the set of coefficients does not fit, then dividing the parameter space into
-% regions and fitting separate sets of coefficients should give better results.
+% We believe that either the ansatz does not suit waveforms within certain regions of parameter space or the set of optimized coefficients falls to a minimum where they do not fit some testing waveforms.
+If in some region of parameter space, the ansatz does not fit well with NR waveforms, it would never be able to have consistent improvement under optimization.
+Instead, it would surf around the loss manifold with random fluctuations in mismatches.
+If the set of coefficients does not fit, then dividing the parameter space into regions and fitting separate sets of coefficients should give better results.
 
-Given the ansatz used in the waveform model is unlikely to be fully compatible
-with NR, and the optimization procedure is done over a distribution of waveforms
-with different source parameters, it is conceivable that there are some
-trade-off in accuracy of the waveform models between different part of the
-parameter space. If this is the reason why the high mismatch tail is not
-improving during the joint-optimization, separating the parameter space into
-smaller subspace should help alleviate this issue. On the other hand, if the
-ansatz does not have the right parameterized form to capture the behavior of the
-NR waveforms as a function of the intrinsic parameters, the result should be
-always biased, and we should not expect any improvement even if we separate the
-parameter space into smaller space during training.
+Given the ansatz used in the waveform model is unlikely to be fully compatible with NR, and the optimization procedure is done over a distribution of waveforms with different source parameters, it is conceivable that there are some trade-off in accuracy of the waveform models between different part of the parameter space.
+If this is the reason why the high mismatch tail is not improving during the joint-optimization, separating the parameter space into smaller subspace should help alleviate this issue.
+On the other hand, if the ansatz does not have the right parameterized form to capture the behavior of the NR waveforms as a function of the intrinsic parameters, the result should be always biased, and we should not expect any improvement even if we separate the parameter space into smaller space during training.
 
-Since we know intrinsic parameters play an important role in the ansatz, we
-would like to investigate how intrinsic parameters affect the recalibration
-process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
-Fig. \ref{fig:ps_q148_qchi}. In the low mass ratio region, waveforms with both
-positive and negative log differences are mixed up. On the other hand, along the
-horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve
-under recalibration.
+Since we know intrinsic parameters play an important role in the ansatz, we would like to investigate how intrinsic parameters affect the recalibration process.
+First, we plot the parameter space of $q$ vs.
+$\chi_{\mathrm{PN}}$ in Fig.
+\ref{fig:ps_q148_qchi}.
+In the low mass ratio region, waveforms with both positive and negative log differences are mixed up.
+On the other hand, along the horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve under recalibration.
 
-Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
-Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e.
-$\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this
-feature since the original coefficients were fitted using NR waveforms with
-equal or similar spin, hence the model prefers waveforms with similar spin. One
-interesting feature is how the second and fourth quadrants respond to
-optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
-generally improve with along optimization. However, mismatches in the fourth
-quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most
-waveforms even turned worse after optimization. These waveforms correspond to
-the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
+Furthermore, we plot the parameter space of $\chi_1$ vs.
+$\chi_2$ in Fig.~\ref{fig:ps_q148}.
+Waveforms along the diagonal axis, i.e.
+$\chi_1\approx\chi_2$, show good mismatch improvements.
+We expect to see this feature since the original coefficients were fitted using NR waveforms with equal or similar spin, hence the model prefers waveforms with similar spin.
+One interesting feature is how the second and fourth quadrants respond to optimization.
+In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms generally improve with along optimization.
+However, mismatches in the fourth quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization.
+Most waveforms even turned worse after optimization.
+These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
+\te{I think this kind of behaviour needs to be matched with the discussion about our stopping criteria.
+If it gets worse, why aren't we stopping earlier?}
 
-% Waveform mismatches in the fourth quadrant, i.e. $\chi_1>0$ and $\chi_2<0$,
-% performed worse after optimized. IMRPhenomD is constructed using single spin
-% parameter, where there are degeneracy between $\chi_1$ and $\chi_2$. In
-% regions with opposite extreme spin components, degeneracy significantly alters
-% intrinsic parameters passed into waveform generation, hence the fourth
-% quadrant shows such feature.
+% Waveform mismatches in the fourth quadrant, i.e.
+$\chi_1>0$ and $\chi_2<0$, performed worse after optimized.
+IMRPhenomD is constructed using single spin parameter, where there are degeneracy between $\chi_1$ and $\chi_2$.
+In regions with opposite extreme spin components, degeneracy significantly alters intrinsic parameters passed into waveform generation, hence the fourth quadrant shows such feature.
 
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
-	\caption{Parameter space of testing waveforms of
-	$q$~vs.~$\chi_{\mathrm{PN}}$. We use the recalibrated result from
-	$\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training
-	waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the
-	$\log_{10}$ difference between optimized and original unweighted
-	mismatches.}
+	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$.
+We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.
+Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.
+\te{I find this plot a little odd.
+The bounding of between -1 and 1 gives the impression that the colourbar has strict limits between these regions.
+Are we truncating the colors?And if so, why?}}
 	\label{fig:ps_q148_qchi}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
-	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. We
-	use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the
-	constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
+	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$.
+We use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
 	\label{fig:ps_q148}
@@
 	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
-		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
-		\midrule\midrule SXS:BBH:0172 & 1.0 & 0.98     & 0.98     \\
+		\toprule\midrule
+		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
+		SXS:BBH:0172 & 1.0 & 0.98     & 0.98     \\
 		SXS:BBH:0152 & 1.0 & 0.60     & 0.60     \\
@@
 		SXS:BBH:1426 & 8.0 & 0.48     & 0.75     \\
-		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule SXS:BBH:0370 & 1.0
-		& -0.20    & 0.40     \\
+		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule
+		SXS:BBH:0370 & 1.0 & -0.20    & 0.40     \\
 		SXS:BBH:2092 & 1.0 & -0.50    & 0.50     \\
@@
 		SXS:BBH:2133 & 3.0 & -0.73    & 0.85     \\
-		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule SXS:BBH:0156 & 1.0
-		& -0.95    & -0.95    \\
+		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule
+		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
 		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
@@
 		SXS:BBH:1419 & 8.0 & -0.80    & -0.80    \\
-		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule SXS:BBH:0304 & 1.0
-		& 0.50     & -0.50    \\
+		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule
+		SXS:BBH:0304 & 1.0 & 0.50     & -0.50    \\
 		SXS:BBH:0327 & 1.0 & 0.80     & -0.80    \\
@@
 
-As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is
-significantly different in different regions in the parameter space, we split
-the parameter space into 4 quadrants and perform separate fitting with training
-waveforms listed in Table \ref{tab:quadrants}. Note that in the second and
-fourth quadrants, there are not enough waveforms with $q>4$, hence the result
-are only valid up to $q\leq4$. From Fig.~\ref{fig:ps_q148_quadrant}, we see that
-all waveforms, except those in the fourth quadrant, show improvement in
-mismatch. Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
+As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is significantly different in different regions in the parameter space, we split the parameter space into 4 quadrants and perform separate fitting with training waveforms listed in Table \ref{tab:quadrants}.
+Note that in the second and fourth quadrants, there are not enough waveforms with $q>4$, hence the result are only valid up to $q\leq4$.
+From Fig.~\ref{fig:ps_q148_quadrant}, we see that all waveforms, except those in the fourth quadrant, show improvement in mismatch.
+Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
 First, we focus on waveforms with the same spin direction, i.e.
-$\chi_1\chi_2>0$. Waveforms lying in the neighborhood of the diagonal axis have
-noticeable improvements in mismatch due to most training waveforms being
-equal-spin waveforms. For the second quadrant, waveforms improved significantly
-with only a few defects due to some testing waveforms having $q>4$. In the
-fourth quadrant, most optimized waveforms have a higher mismatch than the
-original waveforms, as indicated by the positive log difference of mismatches.
-As performing optimization in a smaller subspace does not show any improvement
-compared to the original result, the ansatz mostly does not fit waveforms in the
-fourth quadrant.
+$\chi_1\chi_2>0$.
+Waveforms lying in the neighborhood of the diagonal axis have noticeable improvements in mismatch due to most training waveforms being equal-spin waveforms.
+For the second quadrant, waveforms improved significantly with only a few defects due to some testing waveforms having $q>4$.
+In the fourth quadrant, most optimized waveforms have a higher mismatch than the original waveforms, as indicated by the positive log difference of mismatches.
+As performing optimization in a smaller subspace does not show any improvement compared to the original result, the ansatz mostly does not fit waveforms in the fourth quadrant.
 
@@
 	\caption{Distributions of mismatches after optimizing in separate quadrants.
-	We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. Generally, most
-	waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant. }
+We use a constant noise spectrum to calculate mismatch and $\mathcal{L}_{\mathrm{mean}}$ for the loss function.
+Generally, most waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant.
+\te{It would be nice to have four histograms for the four quadrants here.
+Ideally for both the original and optimized waveform but this might get a little cluttered.
+Maybe just plot the best and worst quadrant?}\te{This figure also doesn't appear to be referenced in the text?}}
 	\label{fig:all_quadrants}
@@
 
-%\begin{enumerate} \item Comparison with the original paper is not the same.
-%   (Different waveforms, different ways to modify the waveform, different range
-%   of frequencies) \item More waveforms are taken so it should be more robust,
-%   since the parameter space coverage is better. \item Problems with IMRPhenomD
-%   model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space
-%   $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat) \item
-%   Can in principle split parameter space into regions that can be separately
-%   optimized. If not, then its must be the model's problem. Would the region
-%   classification be made automatic. Are there any algorithms to do so? \item
-%   Future work (Repeat on other GW models. Mapping between NR surrogate to see
-%   any systematics.) \end{enumerate}
+%\begin{enumerate}
+%	\item Comparison with the original paper is not the same. (Different waveforms, different ways to modify the waveform, different range of frequencies)
+%	\item More waveforms are taken so it should be more robust, since the parameter space coverage is better.
+%    \item Problems with IMRPhenomD model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat)
+%    \item Can in principle split parameter space into regions that can be separately optimized.
+If not, then its must be the model's problem.
+Would the region classification be made automatic.
+Are there any algorithms to do so?
+%    \item Future work (Repeat on other GW models. Mapping between NR surrogate to see any systematics.)
+%\end{enumerate}
 
-We have shown the result of recalibrating waveform coefficients. One thing to
-note is that our recalibration procedure is not exactly the same as the original
-calibration. For instance, we use a different set of NR waveforms, frequency
-range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
-optimization procedure should be able to improve the accuracy of IMRPhenomD on a
-similar scale regardless of the differences. Here, this result serves as a
-demonstration of the general method used.  
-% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for
-% both amplitude and phase. Similarly, for the merger-ringdown region, the
-% optimized waveforms also show great improvement.
+We have shown the result of recalibrating waveform coefficients.
+One thing to note is that our recalibration procedure is not exactly the same as the original calibration.
+For instance, we use a different set of NR waveforms, frequency range, etc.
+Nonetheless, as the decrease in mismatch is rather significant, this optimization procedure should be able to improve the accuracy of IMRPhenomD on a similar scale regardless of the differences.
+Here, this result serves as a demonstration of the general method used.
+% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for both amplitude and phase.
+Similarly, for the merger-ringdown region, the optimized waveforms also show great improvement.
 
-We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with
-more training waveforms only has a small increase in the accuracy of the
-waveform model. We believe that by increasing the number of waveforms, the
-accuracy will not have a significant change as the waveform coefficients are
-already over-determined. Using more calibration NR waveforms will not further
-improve the model significantly. This suggests the form of the parameterized
-ansatz is not suitable for certain regions in the parameter space, thus
-mismatches of only a few waveforms decreased while other waveforms remain at the
-high mismatch tail with little changes. This implies that the model is
-ultimately restricted by the flexibility of the ansatz.
+We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with more training waveforms only has a small increase in the accuracy of the waveform model.
+We believe that by increasing the number of waveforms, the accuracy will not have a significant change as the waveform coefficients are already over-determined.
+Using more calibration NR waveforms will not further improve the model significantly.
+This suggests the form of the parameterized ansatz is not suitable for certain regions in the parameter space, thus mismatches of only a few waveforms decreased while other waveforms remain at the high mismatch tail with little changes.
+This implies that the model is ultimately restricted by the flexibility of the ansatz.
 
-One of the major problems causing inaccuracy in the ansatz is the reduced spin
-approximation. In IMRPhenomD, it is modeled using a single spin parameter,
-namely $\chi_{\mathrm{PN}}$ as outlined in Sec. \ref{sec:method}. Parameterizing
-a BBH merger with one spin parameter introduces degeneracy within the parameter
-space. Events with distinct black hole spins could result in equal
-$\chi_{\mathrm{PN}}$, thus generating the same waveform. Especially with high
-unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as
-events with small spin, thus giving an inaccurate result. Generally, the
-approximation gives straight lines of degeneracy in the parameter space, with
-its slope (always negative) dependent on the mass ratio. From
-Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves
-better in the top left region than the bottom right. To try to accommodate this
-issue, we split the parameter space into 4 quadrants as described in Sec.
-\ref{sec:result}. However, even with separate optimizations, we see in
-Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar
-mismatches as before while the second quadrant further improved. This suggests
-the ansatz is region-specific, with a higher preference for BBH events with
-$\chi_1<0$ and $\chi_2>0$.
+One of the major problems causing inaccuracy in the ansatz is the reduced spin approximation.
+In IMRPhenomD, it is modeled using a single spin parameter, namely $\chi_{\mathrm{PN}}$ as outlined in Sec.
+\ref{sec:method}.
+Parameterizing a BBH merger with one spin parameter introduces degeneracy within the parameter space.
+Events with distinct black hole spins could result in equal $\chi_{\mathrm{PN}}$, thus generating the same waveform.
+Especially with high unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as events with small spin, thus giving an inaccurate result.
+Generally, the approximation gives straight lines of degeneracy in the parameter space, with its slope (always negative) dependent on the mass ratio.
+From Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves better in the top left region than the bottom right.
+To try to accommodate this issue, we split the parameter space into 4 quadrants as described in Sec.
+\ref{sec:result}.
+However, even with separate optimizations, we see in Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar mismatches as before while the second quadrant further improved.
+This suggests the ansatz is region-specific, with a higher preference for BBH events with $\chi_1<0$ and $\chi_2>0$.
 
-Separating regions into 4 quadrants is done purely out of simplicity. To give a
-more comprehensive analysis, one should be systematic about region selection.
-One can use level set estimation algorithms to obtain systematic regions of
-interest. This general algorithm reveals further degeneracies or issues within
-the ansatz. Then, recalibrating such individual regions might give better
-results. Alternatively, one can select regions according to the lines of
-degeneracy. However, with limited NR waveforms, such a selection scheme is not
-viable for us. In the future, with more NR waveforms spanning the entire
-parameter space, one can perform optimization with fewer restrictions.
+Separating regions into 4 quadrants is done purely out of simplicity.
+To give a more comprehensive analysis, one should be systematic about region selection.
+One can use level set estimation algorithms to obtain systematic regions of interest.
+This general algorithm reveals further degeneracies or issues within the ansatz.
+Then, recalibrating such individual regions might give better results.
+Alternatively, one can select regions according to the lines of degeneracy.
+However, with limited NR waveforms, such a selection scheme is not viable for us.
+In the future, with more NR waveforms spanning the entire parameter space, one can perform optimization with fewer restrictions.
 
-While our work focused mainly on the IMRPhenomD model, this simple yet general
-method can be utilized in other differentiable GW models. For instance, within
-the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP
-\citep{hannam2014simple} models. By jointly fitting a new set of coefficients,
-it is expected that both models can be improved since they are constructed in a
-similar way as the IMRPhenomD model. For example, they also use PN approximant
-as part of the ansatz in the inspiral segment. One interesting result might
-arise while recalibrating IMRPhenomX model \citep{pratten2020setting}. Since it
-is parameterized by an additional anti-symmetric spin parameter, it is expected
-to not show the same degeneracy as described above. Further analysis might give
-insights into the systematics of Phenom models. Moreover, this method can be
-applied to other GW model families, such as NR surrogate models
-\citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}. NR
-waveform calibration procedures could be made easier and are likely to improve
-current models.
+While our work focused mainly on the IMRPhenomD model, this simple yet general method can be utilized in other differentiable GW models.
+For instance, within the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP \citep{hannam2014simple} models.
+By jointly fitting a new set of coefficients, it is expected that both models can be improved since they are constructed in a similar way as the IMRPhenomD model.
+For example, they also use PN approximant as part of the ansatz in the inspiral segment.
+One interesting result might arise while recalibrating IMRPhenomX model \citep{pratten2020setting}.
+Since it is parameterized by an additional anti-symmetric spin parameter, it is expected to not show the same degeneracy as described above.
+Further analysis might give insights into the systematics of Phenom models.
+Moreover, this method can be applied to other GW model families, such as NR surrogate models \citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}.
+NR waveform calibration procedures could be made easier and are likely to improve current models.
 
-% Can we further check systematic bias by injection? But how can we know the
-% real mass or spin in the first place if we disregard GW models.
+% Can we further check systematic bias by injection? But how can we know the real mass or spin in the first place if we disregard GW models.
 
@@
 In this paper, we have presented a systematic method to recalibrate GW models.
-This method utilizes {\jax}'s automatic differentiation to apply
-derivative-based optimization to recalibrate GW models jointly. Using the new
-implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
-conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
-coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
-improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
-mismatches have a slightly better improvement. In contrast, different types of
-loss function result in significantly different final mismatch distributions,
-where the result can be seen in Fig.~\ref{fig:q148}. By increasing the number of
-training waveforms, we see a slight improvement increase in
-Fig.~\ref{fig:q148_q1248_compare}.
+This method utilizes {\jax}'s automatic differentiation to apply derivative-based optimization to recalibrate GW models jointly.
+Using the new implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in conjunction with NR waveforms from the SXS catalog, we recalibrate waveform coefficients of the IMRPhenomD model.
+In general, the waveform accuracy can be improved by 50\%.
+Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement.
+In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148}.
+By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
 
-Furthermore, we investigated how the intrinsic parameters affect the
-improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-certain preference for waveforms lying in the second quadrant while the fourth
-quadrant cannot be improved. To further test this result, we recalibrate
-waveforms in separate regions in parameter space. As shown in
-Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further
-improvement to the second quadrant while the fourth quadrant shows similar
-result. This indicates that the model's ansatz does not fit waveforms in the
-fourth quadrant. This phenomenon is due to the reduced spin approximation used
-in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$
-are introduced.
+Furthermore, we investigated how the intrinsic parameters affect the improvement.
+Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a certain preference for waveforms lying in the second quadrant while the fourth quadrant cannot be improved.
+To further test this result, we recalibrate waveforms in separate regions in parameter space.
+As shown in Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further improvement to the second quadrant while the fourth quadrant shows similar result.
+This indicates that the model's ansatz does not fit waveforms in the fourth quadrant.
+This phenomenon is due to the reduced spin approximation used in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
 
-While we naively separate the optimization process into 4 quadrants, one can
-perform systematic region-selection. In principle, we can apply this general
-method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
-models. Then, we can perform all the above analyses to understand how to
-construct better GW Phenom models in the future.
+While we naively separate the optimization process into 4 quadrants, one can perform systematic region-selection.
+In principle, we can apply this general method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP models.
+Then, we can perform all the above analyses to understand how to construct better GW Phenom models in the future.
 
-% From analyzing mismatches in parameter space, we investigated the accuracy of
-% generated waveforms based on their mass ratio and spin. Similar studies can be
-% performed on other GW models to give better criteria to
+% From analyzing mismatches in parameter space, we investigated the accuracy of generated waveforms based on their mass ratio and spin.
+Similar studies can be performed on other GW models to give better criteria to
 
-% Improved the IMRPhenomD model. Found regions that can be improved. Checked
-% ansatz.
+% Improved the IMRPhenomD model. Found regions that can be improved. Checked ansatz.
 

### 2023-03-27 src/tex/ms.tex
Add kw comment
update

--- 2023-03-16 src/tex/ms.tex
@@
 
-\usepackage{tabularx,booktabs,multirow}
+\usepackage{tabularx,booktabs,multirow,amsmath}
 \usepackage{float}
@@
 \newcommand{\zdethp}{\texttt{zdethp}}
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\definecolor{cyan}{HTML}{0097A7}
+\newcommand{\kl}[1]{{\color{cyan}[KL: #1 ]}}
 
@@
 Currently, three families of GW models are commonly used. They are the
-effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
-(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
-\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
+effective-one-body (EOB) \citep{ossokine2020multipolar, cotesta2020frequency, taracchini2014effective}, Numerical Relativity
+(NR) surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2} and phenomenological (Phenom) models
+\citep{husa2016frequency, khan2016frequency, garcia2020multimode, pratten2021computationally} \kw{Add more citations related to
 the newest models}. EOB models are constructed by mapping two masses onto an
@@
 Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
-For interested readers, please refer to \cite{khan2016frequency} for more details on construction of the model.
+For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
 
@@
 \kw{Check if the following equation is correct. I think it is off by some smoothing factors.} 
-\begin{align}
-	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta).
+\kl{I don't remember there is any smoothing factors.
+They are just connected using step functions, then the entire waveform is made to have continuous first derivative by fixing some of the coefficients for the two connections.}
+\begin{align}\label{eq:joint_waveform}
+	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
 \end{align}
@@
 functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. These simple analytic functions
-consists of parameters $\Lambda^i$, which are defined as follows. \kw{Can you
+of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
+\begin{equation}\label{eq:amplitude}
+\begin{aligned}
+	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
+	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
+	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
+	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
+\end{aligned}
+\end{equation}
+where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
+The ansatzes for phase can be found in \citep{khan2016frequency}.
+These simple analytic functions
+consists of parameters $\Lambda^i=\{\rho_i,\delta_i,\gamma_i\}$, which are defined as follows. \kw{Can you
 change equation one such that it consists of $\Lambda$? Also be explicit about
@@
 Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
-respectively. Finally, the individual segments are connected in a way that the
-final waveform is continuous in its first derivative \kw{expand a bit}.
+respectively. Finally, the individual segments are first connected directly using 
+step functions. Then, by fixing coefficients in the intermediate waveform, one can make 
+the final waveform is continuous in its first derivative \kw{expand a bit}.
 
 \kw{Need to work on this paragraph}
-From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are
-fundamental to waveform generation, which can significantly affect the shape of
-final waveforms. Thus, having a set of accurate waveform coefficients is
-important. Generally, waveform coefficients are obtained by calibrating with NR
-waveforms, which are waveforms computed using NR simulations. In the case of
-\citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated
-waveforms to NR waveforms. For each NR simulation, they can obtain one set of
-$\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are
-subsequently found by fitting against Eq. \ref{eq:Lambda}. Since the fitting
-procedure is done in a piece-wise manner, the correlations between different
-segments are omitted, which could limit the accuracy of the model. Also,
-their calibration procedure does not guarantee to have the optimal set of
-$\lambda$, since the process of connecting segments alters the previously fitted
-waveform. The model-generated waveforms do not take this effect into account,
-and thus introduces additional inaccuracies.
+Combining Eq.~\ref{eq:joint_waveform},~\ref{eq:amplitude}~and~\ref{eq:Lambda}, 
+we can see that the entire waveform is non-linear in $\lambda$. A slightly 
+inaccurate set of $\lambda$ can significantly affect the shape of the generated 
+waveforms. Thus, having a set of accurate waveform coefficients is important 
+and fundamental to having accurate GW models. Generally, waveform coefficients 
+are obtained by calibrating with NR waveforms, which are waveforms computed 
+using NR simulations. In the case of \citep{khan2016frequency}, they first 
+obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+the correlations between different segments are omitted, which could limit the 
+accuracy of the model. Also, since fitting was performed before connecting 
+individual segments, the final waveform does not guarantee to achieve the optimal 
+waveform. The connecting procedure can alter the previously fitted waveform. 
+Hence, the model generated waveforms contains additional inaccuracies.
 
-% they did not use mismatch as the loss function
+%their calibration procedure does not guarantee to have the optimal set of
+%$\lambda$, since the process of connecting segments alters the previously fitted
+%waveform. The model-generated waveforms do not take this effect into account,
+%and thus introduces additional inaccuracies.
 
@@
 
-\subsection{Mismatch} \label{subsec:mismatch}
+\subsection{Loss Function} \label{subsec:loss}
 
@@
 
+Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
+However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
+To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
+We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
+\begin{align}
+	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
+	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+\end{align}	
+where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
+$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
+training waveform, and $N$ is the total number of individual training waveforms.
+Note that we choose to use two different averages, since they have different 
+preferences in optimization base on waveform mismatches.
+For the first choice, simple average serves as the simplest choice of loss function, but is prone to be dominated by a single waveform with a large mismatch.
+Other waveforms with smaller mismatches would be
+insignificant comparatively, and might not be able to improve under such
+optimization. Alternatively, the second choice, normalized average eliminates the aforementioned issue. Nevertheless, it excludes the
+information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
+training waveform to decrease at similar rates, hence it is hard to obtain
+optimized waveforms with mismatches in the same order of magnitude. Instead,
+their ratios in mismatches would remain approximately the same. Conversely,
+$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+and individual mismatches would be in a similar order of magnitude after 
+optimization. In this paper, we showcase the results of using both loss functions 
+and examine the differences between them.
+
+\subsection{Optimization Scheme} \label{subsec:optimization}
+
+To compute the loss functions, we have to take NR waveforms for calculating the mismatch. We choose 11-16 NR waveforms from the set of waveforms used in the original
+calibration process as training waveforms. Originally, 19 waveforms are taken from 
+NR simulations for calibrating IMRPhenomD \citep{khan2016frequency}, 
+which are waveforms from the SXS catalog \citep{boyle2019sxs} or BAM simulation. 
+As BAM waveforms are not publicly available, we cannot take the identical training 
+set as them.
+Instead, we take the available waveforms from the SXS catalog to construct our loss function.
+Training waveforms used are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+The training waveforms chosen has maximum mass ratio to be 8.
+This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+In fact, the SXS catalog only has NR waveforms with $q\leq10$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed from LIGO have $q\leq8$.
+Hence, we calibrate IMRPhenomD with waveforms of $q\leq8$.
+
+In the SXS catalog, NR waveforms are in the form of time-series strain.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
+same intrinsic parameters. We taper the time-series using Tukey window.
+\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
+	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
+	entire GW strain. } 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series.
+
+Other than NR waveforms, one need to choose a relevant noise PSD for 
+mismatch. Since a common goal for creating waveform approximant is to
+deploy the waveform model in downstream data analysis tasks, a reasonable choice
+for the noise PSD should be closely related to the instrument of interest. We
+choose the zero-detuned high-power (\zdethp) noise PSD as our
+first choice of noise PSD \citep{aasi2015advanced}. Because the total mass of the system simple scales
+the frequency of the waveform, we need to choose a corresponding mass scale to
+match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
+all waveforms, since this mass scale is common in LIGO observations. In addition
+to a detector-dependent PSD, we are also interested in a more
+instrument-agnostic case such that we can evaluate the effectiveness of the
+recalibration without considering detector effect. In this case, we set the
+noise PSD to be a constant value.
+
 \kw{Remind me why don't we use the hybrid waveforms again? Cause we couldn't do it?}
-In the original calibration \citep{khan2016frequency,husa2016frequency},
-training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
+We shall notify readers that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
+In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
 (SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
 waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used
-have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
+Instead, we solely use NR waveforms for comparison, since we only exploring the possibility of optimizing waveform models.
+Thus, for simplicity sake, we ignore this procedure.
+On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
 which they are long enough to contain part of the inspiral segment and all
@@
 
-To compute the integral shown in Eq. \ref{eq:inner_prod}, one need to choose a
-relevant noise PSD. Since a common goal for creating waveform approximent is to
-deploy the waveform model in downstream data analysis tasks, a reasonable choice
-for the noise PSD should be closely related to the instrument of interest. We
-choose the zero-detuned high-power (\zdethp) noise PSD \kw{add citation} as our
-first choice of noise PSD. Because the total mass of the system simple scales
-the frequency of the waveform, we need to choose a corresponding mass scale to
-match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
-all waveforms, since this mass scale is common in LIGO observations. In addition
-to a detector-dependent PSD, we are also interested in a more
-instrument-agnostic case such that we can evaluate the effectiveness of the
-recalibration without considering detector effect. In this case, we set the
-noise PSD to be a constant value.
+With the loss function evaluated, we apply gradient descent to optimize the tunable
+coefficients as shown in Algorithm~\ref{alg:gradient}.
+We take $\lambda_i$ to be the original coefficients given in \citep{khan2016frequency}.
+We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find.
+With
+such an optimization procedure, we can systematically obtain the optimized coefficients.
 
-% GW strain treatment from time-series to frequency domain
-In our work, mismatches are used as a measure of discrepancies between
-IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms
-from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are
-calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form
-of time-series data. Since time-series data is oscillatory, performing
-optimization in the time-domain is not ideal. Hence, we had chosen to transform
-NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
-same intrinsic parameters. We taper the time-series using Tukey window.
-\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-entire BBH event. } Then, the frequency spectra can be obtained by taking the
-Fourier transform of the time-series. Then, mismatches can be calculated through
-Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
-
-\subsection{Optimization Scheme} \label{subsec:optimization}
-
-Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus,
-we can utilize simple gradient descent to recalibrate the model.
-
-% \kw{Mention the main advantage in this study is we can optimize everything at
-% once.}
-
-% To recalibrate the model, we have to define loss functions that quantify the
-% difference between the model-generated waveforms and NR waveforms. In the
-% original calibration, they took the mean square error as the loss function for
-% optimization. Instead, we directly use mismatch as our metric for loss
-% functions. However, the mismatch of individual waveforms is not useful in
-% optimization, since it does not represent all waveforms with various intrinsic
-% parameters. We ought to calibrate coefficients that suit waveforms with
-% different intrinsic parameters simultaneously, hence waveforms in different
-% regions of parameter space should be used in the optimization scheme. Here, we
-% chose two different loss functions to examine their ability to optimize the
-% model, e.g. mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and
-% fractional loss $\mathcal{L}_{\mathrm{fl}}$.
-
-% Do something much more direct:
-To recalibrate the model, we need to define a loss function that quantifies the
-difference between the model-generated waveforms and NR waveforms. We use
-mismatch over the set of all training waveforms as the loss function. While in
-\citep{khan2016frequency}, they chose to use the mean square error as the loss
-function, we choose mismatch as the loss function because it is closely related
-to downstream objectives such as the likelihood in parameter estimation. The
-mismatch is averaged over the training dataset in two ways to cover the
-parameter space of interest. One of which takes a simple average and the other
-takes a normalized average based on initial mismatch.
-
-% To recalibrate the model, we need to define a loss function that quantify ...
-% We use mismatch over the set of all training waveforms as the loss function.
-% We choose mismatch as the loss function because it is closely related to
-% downstream objectives such as the likelihood in parameter estimation. The
-% mismatch is averaged over the training dataset to cover the parameter space of
-% interest. (used)
-\begin{align}
-	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
-\end{align}	
-$\mathcal{M}_i$ represents the mismatch of an individual training waveform,
-$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of individual training waveforms.
-For the first choice, it takes the mean of the mismatches, which serves as a
-basic loss function. However, it is prone to be dominated by a single waveform
-with a large mismatch. Other waveforms with smaller mismatches would be
-insignificant comparatively, and might not be able to improve under such
-optimization. Alternatively, the second choice, the mean of normalized
-mismatches, eliminates the aforementioned issue. Nevertheless, it excludes the
-information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
-training waveform to decrease at similar rates, hence it is hard to obtain
-optimized waveforms with mismatches in the same order of magnitude. Instead,
-their ratios in mismatches would remain approximately the same. Conversely,
-$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and final individual mismatches would be in a similar order of magnitude. In
-this paper, we display the results of using both loss functions and examine the
-difference between them.
-
-We have chosen 11-16 NR waveforms from the set of waveforms used in the original
-calibration process. Originally, 19 waveforms are taken from NR simulations for
-calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they
-are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not
-publicly available, we cannot take the identical training set as them. Instead,
-we take the available waveforms from the SXS catalog to construct our loss
-function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the
-maximum mass ratio in our training set is lower than that in
-\citep{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio
-is set as 8. This is because SXS catalog does not have NR waveforms with
-extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with
-$q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model
-with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the
-maximum mass ratio to 8.
-
-With both mismatches, we apply gradient descent to optimize the tunable
-coefficients as follows: 
-\begin{algorithm}
+\begin{algorithm}[t]
 	\caption{Gradient descent pseudocode}
+	\label{alg:gradient}
 	\KwIn{initial coefficients $\lambda_i$} \Parameters{number of iterations
@@
 \end{algorithm}
-Here, $\lambda_i$ are the original coefficients given in
-\citep{khan2016frequency}. We take them as the initial waveform coefficients
-because they lie in the neighborhood of the minimum that we wish to find. With
-such an optimization procedure, we obtain the optimized coefficients.
 

### 2023-03-31 src/tex/ms.tex
GPT pass of Result

--- 2023-03-16 src/tex/ms.tex
@@
 \newcommand{\zdethp}{\texttt{zdethp}}
+\definecolor{rb4}{HTML}{27408B}
+\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 
@@
 
-First, to judge the performance of the optimization scheme, we take $\sim530$ NR
-waveforms from the SXS catalog. We choose testing waveforms with negligible
-eccentricity (${e<2\times10^{-3}}$) and precession
-(${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model. We see
-that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters
-that are within the parameter space spanned by training waveforms. Hence, we can
-take these testing waveforms to compare with the original model.
+To evaluate the effectiveness of our optimization approach, we analyze
+approximately 530 NR waveforms from the SXS catalog in this study. We
+specifically select waveforms with negligible eccentricity
+(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
+consistent with the constraints of the waveform model. As shown in
+Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
+waveforms fall within the parameter space defined by the training waveforms.
+This suggests that these test waveforms can be used for direct comparison with
+the original model.
 
-We first examine the effect of joint optimization on a single waveform. In
-Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy
-than the original waveform, especially in the inspiral region, where the
-amplitude has a $50\%$ decrease in error. We have chosen one of the testing
-waveforms listed in \citep{khan2016frequency} for a fair comparison. 
+We begin by investigating the impact of joint optimization on an individual
+waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
+outperforms the original waveform, particularly in the inspiral region, where
+the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
+selected one of the testing waveforms from the catalog presented in
+\citep{khan2016frequency}.
 % In general, most waveforms within the neighborhood of the training waveforms
@@
 
-Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute
-the mismatch for all testing waveforms and plot the distribution of the mismatch
-in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end
-with a lower mismatch. Quantitatively, the peak has lowered by almost an
-order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
-Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar
-improvement. We see that the median of distribution has a $22.9\%$ decrease.
-However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
-This is due to the problem mentioned in section \ref{subsec:optimization}.
-Nevertheless, both distributions show improvement and the potential problem did
-not affect the optimization scheme significantly.
+We use a constant noise spectrum, $\mathcal{L}{\mathrm{mean}}$, to evaluate the
+mismatch of all testing waveforms, and present the resulting distribution in
+Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
+mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
+in the median. When using $\mathcal{L}{\mathrm{fl}}$, we observe a similar
+improvement with a 22.9\% decrease in the median. However, the distribution lacks
+a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
+Despite this, both distributions exhibit significant improvement, and the
+optimization scheme remains unaffected.
 

### 2023-03-31 src/tex/ms.tex
GPT pass

@@
 
-Similarly, using the same procedures, distributions of mismatches calculated
-using {\zdethp} noise spectrum show better improvement than the weighted
-mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}. This
-effect is expected since the IMRPhenomD model was originally constructed and
-fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that
-fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum
-instead of the constant spectrum.
+Applying the same methods, we observe that the distributions of mismatches
+calculated using the {\zdethp} noise spectrum exhibit superior improvement
+compared to the weighted mismatch. The shape of the distribution is similar to
+that shown in Fig. \ref{fig:q148}. This result is expected, as the IMRPhenomD
+model was initially constructed and fitted using the {\zdethp} weighted
+mismatch. Consequently, the model is anticipated to closely fit the NR waveforms
+while incorporating the influence of the {\zdethp} noise spectrum, rather than a
+constant spectrum.
 
-With the success of improving waveforms, we increase the number of training
-waveforms for optimization. Taking additional waveforms listed in
-Tab.~\ref{tab:q1248}, we obtain a new set of coefficients. We see in Fig.
-\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small
-improvement. The high mismatch tail of the optimized distribution remains
-similar in length and endpoint as the original distribution, meaning that they
-cannot be improved using our procedure. Likewise, using additional waveforms to
-optimize loss function with {\zdethp} spectrum shows the same result, where the
-new distribution barely shows any improvement.
+Motivated by the successful improvement of the waveforms, we expand the training
+dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. The new
+set of coefficients generated from this optimization process yields only
+marginal improvements in the newly produced waveforms, as seen in Fig.
+\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
+distribution remains comparable in length and endpoint to the original
+distribution, indicating that our procedure is incapable of improving these
+waveforms. Similarly, utilizing the {\zdethp} spectrum to optimize the loss
+function with additional waveforms leads to minimal improvement in the resulting
+distribution.
 
@@
 
-Given the ansatz used in the waveform model is unlikely to be fully compatible
-with NR, and the optimization procedure is done over a distribution of waveforms
-with different source parameters, it is conceivable that there are some
-trade-off in accuracy of the waveform models between different part of the
-parameter space. If this is the reason why the high mismatch tail is not
-improving during the joint-optimization, separating the parameter space into
-smaller subspace should help alleviate this issue. On the other hand, if the
-ansatz does not have the right parameterized form to capture the behavior of the
-NR waveforms as a function of the intrinsic parameters, the result should be
-always biased, and we should not expect any improvement even if we separate the
-parameter space into smaller space during training.
+Given that the waveform model's ansatz may not be entirely compatible with NR,
+and the optimization procedure is carried out over a distribution of waveforms
+with varying source parameters, it is conceivable that some trade-offs in
+accuracy exist between different parts of the parameter space. If this is the
+cause of the high mismatch tail's failure to improve during joint-optimization,
+segmenting the parameter space into smaller subspaces should alleviate this
+problem. On the other hand, if the ansatz lacks the correct parameterized form
+to capture the NR waveforms' behavior as a function of the intrinsic parameters,
+the results will always be biased, and we should not expect any improvement,
+even if we segment the parameter space during training.
 

### 2023-03-31 src/tex/ms.tex
Chat

@@
 process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
-Fig. \ref{fig:ps_q148_qchi}. In the low mass ratio region, waveforms with both
-positive and negative log differences are mixed up. On the other hand, along the
-horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve
-under recalibration.
+Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
+more consistent improvement, whereas spinning waveforms can either improve or
+worsen in terms of mismatch. \kw{What does this means?}
 
@@
 interesting feature is how the second and fourth quadrants respond to
-optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
+optimization. \kw{How did you count quadrant?} In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
 generally improve with along optimization. However, mismatches in the fourth
@@
 waveforms even turned worse after optimization. These waveforms correspond to
-the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.
-
-% Waveform mismatches in the fourth quadrant, i.e. $\chi_1>0$ and $\chi_2<0$,
-% performed worse after optimized. IMRPhenomD is constructed using single spin
-% parameter, where there are degeneracy between $\chi_1$ and $\chi_2$. In
-% regions with opposite extreme spin components, degeneracy significantly alters
-% intrinsic parameters passed into waveform generation, hence the fourth
-% quadrant shows such feature.
+the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}. \kw{Give justification.}
 
@@
 
-As we see in Fig.~\ref{fig:ps_q148} that the recalibration procedure is
-significantly different in different regions in the parameter space, we split
-the parameter space into 4 quadrants and perform separate fitting with training
-waveforms listed in Table \ref{tab:quadrants}. Note that in the second and
-fourth quadrants, there are not enough waveforms with $q>4$, hence the result
-are only valid up to $q\leq4$. From Fig.~\ref{fig:ps_q148_quadrant}, we see that
-all waveforms, except those in the fourth quadrant, show improvement in
-mismatch. Many features seen in Fig.~\ref{fig:ps_q148} can be found here again.
-First, we focus on waveforms with the same spin direction, i.e.
-$\chi_1\chi_2>0$. Waveforms lying in the neighborhood of the diagonal axis have
-noticeable improvements in mismatch due to most training waveforms being
-equal-spin waveforms. For the second quadrant, waveforms improved significantly
-with only a few defects due to some testing waveforms having $q>4$. In the
-fourth quadrant, most optimized waveforms have a higher mismatch than the
-original waveforms, as indicated by the positive log difference of mismatches.
-As performing optimization in a smaller subspace does not show any improvement
-compared to the original result, the ansatz mostly does not fit waveforms in the
-fourth quadrant.
+We divided the parameter space into four quadrants to analyze the effect of the
+recalibration procedure on each region separately (Fig.\ref{fig:ps_q148}). The
+training waveforms used for fitting are listed in Table \ref{tab:quadrants}. The
+second and fourth quadrants have limited data for $q>4$, hence the results are
+only valid up to $q\leq4$. From Fig.\ref{fig:ps_q148_quadrant}, we observe
+improvements in waveform mismatch for all waveforms, except those in the fourth
+quadrant. For equal-spin waveforms lying in the proximity of the diagonal axis,
+the mismatch improves significantly due to most training waveforms being
+equal-spin. In the second quadrant, the improvement is noticeable, except for a
+few defects caused by some testing waveforms having $q>4$. The optimized
+waveforms in the fourth quadrant have a higher mismatch than the original ones,
+indicating poor fitting. Further optimization in a smaller subspace does not
+improve the results.
 
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Parameter space of testing waveforms. Each quadrant is fitted independently. Colorbar represents log difference of mismatches before and after optimization.}
+	\caption{\kw{The color scheme makes the plot a bit hard to read.}Parameter
+	space of testing waveforms. Each quadrant is fitted independently. Colorbar
+	represents log difference of mismatches before and after optimization.}
 	\label{fig:ps_q148_quadrant}
@@
 
-We see that in Fig.~\ref{fig:q148_q1248_compare}, performing optimization with
-more training waveforms only has a small increase in the accuracy of the
-waveform model. We believe that by increasing the number of waveforms, the
-accuracy will not have a significant change as the waveform coefficients are
-already over-determined. Using more calibration NR waveforms will not further
-improve the model significantly. This suggests the form of the parameterized
-ansatz is not suitable for certain regions in the parameter space, thus
-mismatches of only a few waveforms decreased while other waveforms remain at the
-high mismatch tail with little changes. This implies that the model is
-ultimately restricted by the flexibility of the ansatz.
+The results presented in Fig.~\ref{fig:q148_q1248_compare} demonstrate that
+increasing the number of training waveforms used in waveform optimization yields
+only a marginal increase in accuracy. Our analysis suggests that this marginal
+improvement is a consequence of over-determination of the waveform coefficients.
+Consequently, increasing the number of calibration NR waveforms is unlikely to
+result in any significant improvement of the model's accuracy. These
+observations suggest that the parameterized ansatz employed in our study may not
+be suitable for certain regions in the parameter space, leading to mismatches
+for some waveforms while other waveforms remain at the high mismatch tail with
+negligible changes. This highlights the constraints of the model's flexibility
+that ultimately limit its performance.
 
-One of the major problems causing inaccuracy in the ansatz is the reduced spin
-approximation. In IMRPhenomD, it is modeled using a single spin parameter,
-namely $\chi_{\mathrm{PN}}$ as outlined in Sec. \ref{sec:method}. Parameterizing
-a BBH merger with one spin parameter introduces degeneracy within the parameter
-space. Events with distinct black hole spins could result in equal
-$\chi_{\mathrm{PN}}$, thus generating the same waveform. Especially with high
-unequal spin events, $\chi_{\mathrm{PN}}$ would identify them as the same as
-events with small spin, thus giving an inaccurate result. Generally, the
-approximation gives straight lines of degeneracy in the parameter space, with
-its slope (always negative) dependent on the mass ratio. From
-Fig.~\ref{fig:ps_q148}, we see that along a degeneracy line, the ansatz behaves
-better in the top left region than the bottom right. To try to accommodate this
-issue, we split the parameter space into 4 quadrants as described in Sec.
-\ref{sec:result}. However, even with separate optimizations, we see in
-Fig.~\ref{fig:ps_q148_quadrant} that the fourth quadrant still shows similar
-mismatches as before while the second quadrant further improved. This suggests
-the ansatz is region-specific, with a higher preference for BBH events with
+The reduced spin approximation is a major contributor to the inaccuracies
+observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
+parameter, $\chi_{\mathrm{PN}}$, as described in Sec. \ref{sec:method}. The
+parameterization of BBH mergers using a single spin parameter results in a
+degeneracy within the parameter space. Specifically, black hole events with
+different spins may generate the same waveform due to identical values of
+$\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
+unequal spin events. This degeneracy produces straight lines in the parameter
+space with negative slopes that depend on the mass ratio, as shown in
+Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top left
+region than the bottom right along a degeneracy line. In an attempt to address
+this issue, we partitioned the parameter space into four quadrants, as described
+in Sec. \ref{sec:result}. However, even with separate optimizations for each
+quadrant, Fig.\ref{fig:ps_q148_quadrant} indicates that the fourth quadrant
+exhibits similar mismatches as before, while the second quadrant's performance
+has improved. This observation suggests that the ansatz is specific to certain
+regions of the parameter space, with a preference for BBH events featuring
 $\chi_1<0$ and $\chi_2>0$.
 
-Separating regions into 4 quadrants is done purely out of simplicity. To give a
-more comprehensive analysis, one should be systematic about region selection.
-One can use level set estimation algorithms to obtain systematic regions of
-interest. This general algorithm reveals further degeneracies or issues within
-the ansatz. Then, recalibrating such individual regions might give better
-results. Alternatively, one can select regions according to the lines of
-degeneracy. However, with limited NR waveforms, such a selection scheme is not
-viable for us. In the future, with more NR waveforms spanning the entire
-parameter space, one can perform optimization with fewer restrictions.
+The division of the parameter space into four quadrants was a simple approach
+taken for practical reasons. A more systematic approach would involve the use of
+level set estimation algorithms to identify regions of interest within the
+parameter space. Such an algorithm can reveal additional degeneracies or issues
+that may exist within the ansatz. One possible strategy is to recalibrate
+individual regions of interest to achieve better results. An alternative
+approach is to select regions based on degeneracy lines. However, due to the
+limited number of NR waveforms available, we were unable to implement this
+approach. With more NR waveforms available in the future that cover the entire
+parameter space, we can perform optimization with fewer restrictions and select
+regions more systematically.
 
-While our work focused mainly on the IMRPhenomD model, this simple yet general
-method can be utilized in other differentiable GW models. For instance, within
-the same family, IMRPhenomX \citep{pratten2020setting} or IMRPhenomP
-\citep{hannam2014simple} models. By jointly fitting a new set of coefficients,
-it is expected that both models can be improved since they are constructed in a
-similar way as the IMRPhenomD model. For example, they also use PN approximant
-as part of the ansatz in the inspiral segment. One interesting result might
-arise while recalibrating IMRPhenomX model \citep{pratten2020setting}. Since it
-is parameterized by an additional anti-symmetric spin parameter, it is expected
-to not show the same degeneracy as described above. Further analysis might give
-insights into the systematics of Phenom models. Moreover, this method can be
-applied to other GW model families, such as NR surrogate models
-\citep{varma2019surrogate} or EOB models \citep{taracchini2014effective}. NR
-waveform calibration procedures could be made easier and are likely to improve
-current models.
+Although our study primarily focused on the IMRPhenomD model, this simple yet
+versatile approach can be applied to other differentiable GW models, such as the
+IMRPhenomP \citep{hannam2014simple} or IMRPhenomX \citep{pratten2020setting} 
+models within the same family. By jointly optimizing a new set of coefficients,
+it is expected that both models can be enhanced since they share similar
+construction principles to the IMRPhenomD model. For instance, they also use PN
+approximants as part of the ansatz in the inspiral segment. Interestingly, while
+recalibrating the IMRPhenomX model \citep{pratten2020setting}, which is
+parameterized by an additional anti-symmetric spin parameter, it is expected not
+to exhibit the degeneracy previously described. A more detailed investigation
+may provide valuable insights into the systematics of the Phenom models.
+Furthermore, this approach is applicable to other GW model families, such as NR
+surrogate models \citep{varma2019surrogate} or EOB models
+\citep{taracchini2014effective}. Such an approach could simplify NR waveform
+calibration procedures and lead to the improvement of existing models.
 

### 2023-04-03 src/tex/ms.tex
Adding more comments

--- 2023-03-31 src/tex/ms.tex
@@
 \newcommand{\zdethp}{\texttt{zdethp}}
-\definecolor{rb4}{HTML}{27408B}
-\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
 
@@
 \section{Introduction} \label{sec:intro}
+\te{General points:
+\begin{itemize}
+	\item Overall the text needs to be refined
+	\item Plot styles should be made more uniform
+\end{itemize}
+}
 
-In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
-will finish its maintenance and start observing new gravitational wave (GW)
-results. This new O4 run is expected to double the rate of current binary black
-hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
-sensitivity of interferometers will be increased to capture more details of GW.
-Having instruments with higher sensitivity, GW models of equal or higher
-accuracy than observations should be used to extract GW information. Otherwise,
-the extracted information would be affected more by GW models instead of
-interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
-GW models are accurate enough for current analyses, the accuracy of current
-models will no longer suffice for future data analyses
-\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
-improve GW models.
+In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results.
+This new O4 run is expected to double the rate of current binary black hole (BBH) observations \citep{abbott2020prospects}.
+Additionally, the sensitivity of interferometers will be increased to capture more details of GW.
+Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information.
+Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses.
+Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \citep{purrer2020gravitational}.
+Hence, it is necessary for us to develop and improve GW models.
 
 % Waveform families and slight motivation of why we take Phenom model
-Currently, three families of GW models are commonly used. They are the
-effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
-(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
-\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
-the newest models}. EOB models are constructed by mapping two masses onto an
-effective body under an effective metric; NR surrogate models construct
-waveforms using combinations of NR waveforms; Phenom models are formulated using
-specific ansatz and inspiral approximations. While EOB and NR surrogate models
-give better waveform approximants, Phenom waveforms can be produced much faster,
-hence it is used mostly in data analysis tasks that requires many waveform
-generations. This advantage scales up in data analysis tasks such as matched
-filtering and parameter estimation, where many waveforms are required in each
-run. This motivates us to improve upon the current framework of Phenom models,
-thus can retain the advantage of fast waveform generation while improving the
-model's accuracy.
+Currently, three families of GW models are commonly used.
+They are the effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models \citep{husa2016frequency,khan2016frequency}.
+EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations.
+While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations.
+This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run.
+This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.
 
-Automatic differentiation (AD) is a method to calculate derivatives of functions
-up to machine precision. In traditional numerical calculations, derivatives are
-usually obtained through numerical derivatives. Symbolic derivatives were
-available but it was less efficient. Both methods were not viable in machine
-learning, where back-propagation requires precise and rapid derivative
-calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch},
-\texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train
-machine learning models. AD's algorithm is intuitive in nature. Functions
-defined are decomposed into tree structures of primitives, such as addition or
-function evaluations. Since these operations are fundamental, they were saved as
-pairs internally. Differentiation proceeds forward following the tree structure,
-with the application of the chain rule in each step to evaluate its derivative.
-Analytic derivatives of such operations are applied in each step and the desired
-derivative can then be obtained by composing back the original function
-according to the original structure. {\ripple} \citep{ripple} was a new
-implementation of IMRPhenomD, one of the Phenom models. It was first implemented
-in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
-rewritten using \jax, another {\python} package that supports AD. Using
-{\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
-allowing one to freely use derivative-based algorithms to perform data analyses. 
+Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision.
+In traditional numerical calculations, derivatives are usually obtained through numerical derivatives.
+Symbolic derivatives were available but it was less efficient.
+Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations.
+In {\python}, packages including \texttt{pytorch} \citep{pytorch}, \texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc.
+utilizes AD to train machine learning models.
+AD's algorithm is intuitive in nature.
+Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations.
+Since these operations are fundamental, they were saved as pairs internally.
+Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative.
+Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure.
+{\ripple} \citep{ripple} was a new implementation of IMRPhenomD, one of the Phenom models.
+It was first implemented in \texttt{lalsuite} using \texttt{C}.
+In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD.
+Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses.
 % Need to cite references for AD
 
-In this paper, we investigate the possibility of further improving the accuracy
-of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
-waveforms, and what constraints one may face when trying to further improve. We
-find that simply by applying gradient descent algorithm, one can obtain a better
-set of waveform coefficients, thus improving the accuracy of the model.
-Furthermore, by comparing the accuracy of optimized and original waveforms, we
-find that model-generated waveforms are very sensitive to their intrinsic
-parameters. Specifically, IMRPhenomD favors certain parts of the parameter
-space. This means IMRPhenomD introduces systematic bias to other GW analysis
-tasks. This showcases the flaws of the ansatz and allows us to have a deeper
-understanding of Phenom models.  
+In this paper, we investigate the possibility of further improving the accuracy of IMRPhenomD by jointly optimizing all the fitting coefficients given NR waveforms, and what constraints one may face when trying to further improve.
+We find that simply by applying gradient descent algorithm, one can obtain a better set of waveform coefficients, thus improving the accuracy of the model.
+Furthermore, by comparing the accuracy of optimized and original waveforms, we find that model-generated waveforms are very sensitive to their intrinsic parameters.
+Specifically, IMRPhenomD favors certain parts of the parameter space.
+This means IMRPhenomD introduces systematic bias to other GW analysis tasks.
+This showcases the flaws of the ansatz and allows us to have a deeper understanding of Phenom models.
 % \kw{Briefly describe the high level finding in this paper}
 
-The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
-review the parameterization of the IMRPhenomD model and the mismatch function
-that is used as an objective function for the calibration, followed by outlining the
-specific optimization scheme used for recalibration. In
-Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
-of optimized waveforms with original waveforms. We also show how the
-optimization result differs with waveforms of different intrinsic parameters. In
-Sec.~\ref{sec:discussion}, we address the difference between our calibrating
-procedure with \citep{khan2016frequency}. We also explain how reduced spin
-parameterization affects the accuracy of the model.
+The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model.
+We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms.
+Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}.
+In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms.
+We also show how the optimization result differs with waveforms of different intrinsic parameters.
+In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \citep{khan2016frequency}.
+We also explain how reduced spin parameterization affects the accuracy of the model.
 
@@
 
-In order to recalibrate the model, we need to define an objective function that
-quantify the performance of the model. A common choice for such metric is the
-\textit{mismatch} function \citep{husa2016frequency}. It is defined as
+To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
 \begin{align} \label{eq:mismatch}
@@
 \end{align}
-where $h_{1,2}$ are the two GW waveform we are comparing, and $t_0$ and $\phi_0$
-are time shift and phase shift respectively. The $\langle h_1, h_2 \rangle$ is
-commonly referred as the inner product, which is defined as 
+where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
 \begin{align}\label{eq:inner_prod}
-	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
+	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
 \end{align}
-where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the power spectral density(PSD) of noise from the instrument, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the relevant maximum and minimum frequencies for the integration. The mismatch
-is a quantity that is closely related to the mean square error (MSE) between the
-two waveforms, as it is often considered a weighted version of the MSE.
+where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.
 
-\kw{Remind me why don't we use the hybrid waveforms again? Cause we couldn't do it?}
-In the original calibration \citep{khan2016frequency,husa2016frequency},
-training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
-(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
-waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used
-have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
-which they are long enough to contain part of the inspiral segment and all
-merger and ringdown frequency information. We take the frequency limits as
-$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,
-where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set higher
-than that in the original calibration due to NR length. When compared with
-IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency.
+In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms.
+The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information.
+We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown.
+This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length.
+When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.
+\te{What is the frequency spacing?} \te{Why do we not just go to the cutoff of 0.2, or do the NR waveforms end earlier?
+Think these choices need more motivation.}
 
-To compute the integral shown in Eq. \ref{eq:inner_prod}, one need to choose a
-relevant noise PSD. Since a common goal for creating waveform approximent is to
-deploy the waveform model in downstream data analysis tasks, a reasonable choice
-for the noise PSD should be closely related to the instrument of interest. We
-choose the zero-detuned high-power (\zdethp) noise PSD \kw{add citation} as our
-first choice of noise PSD. Because the total mass of the system simple scales
-the frequency of the waveform, we need to choose a corresponding mass scale to
-match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
-all waveforms, since this mass scale is common in LIGO observations. In addition
-to a detector-dependent PSD, we are also interested in a more
-instrument-agnostic case such that we can evaluate the effectiveness of the
-recalibration without considering detector effect. In this case, we set the
-noise PSD to be a constant value.
+\te{I think this discussion needs to be much more precise.
+Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
+So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
+We choose to use two different noise spectra $S_n(f)$, namely the zero-detuned high-power (\zdethp) noise spectrum and a constant spectrum.
+Since in any data analysis tasks, model-generated waveforms will ultimately be used to compare with GW observations, they are subjected to the sensitivity of detectors.
+Hence, it is sensible to have waveforms calibrated with {\zdethp} spectrum rather than a non-weighted spectrum.
+However, with sensitivity curve involved, we have to choose a corresponding mass scale.
+Here, we simply choose $M=50M_{\odot}$ for all waveforms, since this mass scale is common in LIGO observations.
+\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}
+On the other hand, we also recalibrated coefficients with the constant spectrum for a more general purpose.
 
 % GW strain treatment from time-series to frequency domain
-In our work, mismatches are used as a measure of discrepancies between
-IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms
-from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are
-calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form
-of time-series data. Since time-series data is oscillatory, performing
-optimization in the time-domain is not ideal. Hence, we had chosen to transform
-NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
-same intrinsic parameters. We taper the time-series using Tukey window.
-\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-entire BBH event. } Then, the frequency spectra can be obtained by taking the
-Fourier transform of the time-series. Then, mismatches can be calculated through
-Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
+In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms.
+To compute such mismatches, NR waveforms from the SXS catalog are taken \citep{boyle2019sxs}.
+These NR waveforms are calculated using the \texttt{SpEC} code.
+Initially, NR waveforms are in the form of time-series data.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters.
+We taper the time-series using Tukey window.
+\footnote{
+	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
+} 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
 
@@
 
-To evaluate the effectiveness of our optimization approach, we analyze
-approximately 530 NR waveforms from the SXS catalog in this study. We
-specifically select waveforms with negligible eccentricity
-(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
-consistent with the constraints of the waveform model. As shown in
-Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
-waveforms fall within the parameter space defined by the training waveforms.
-This suggests that these test waveforms can be used for direct comparison with
-the original model.
+First, to judge the performance of the optimization scheme, we take $\sim530$ NR waveforms from the SXS catalog.
+We choose testing waveforms with negligible eccentricity (${e<2\times10^{-3}}$) and precession (${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model.
+We see that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters that are within the parameter space spanned by training waveforms.
+Hence, we can take these testing waveforms to compare with the original model.
 
-We begin by investigating the impact of joint optimization on an individual
-waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
-outperforms the original waveform, particularly in the inspiral region, where
-the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
-selected one of the testing waveforms from the catalog presented in
-\citep{khan2016frequency}.
-% In general, most waveforms within the neighborhood of the training waveforms
-% show improvement in mismatches.
+We first examine the effect of joint optimization on a single waveform.
+In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error.
+We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison.
+% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.
 
-We use a constant noise spectrum, $\mathcal{L}{\mathrm{mean}}$, to evaluate the
-mismatch of all testing waveforms, and present the resulting distribution in
-Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
-mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
-in the median. When using $\mathcal{L}{\mathrm{fl}}$, we observe a similar
-improvement with a 22.9\% decrease in the median. However, the distribution lacks
-a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
-Despite this, both distributions exhibit significant improvement, and the
-optimization scheme remains unaffected.
+Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig.
+\ref{fig:q148}.
+We find the distribution's peak has shifted to the end with a lower mismatch.
+Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
+Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement.
+We see that the median of distribution has a $22.9\%$ decrease.
+However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
+This is due to the problem mentioned in section \ref{subsec:optimization}.
+Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.
 
-% \kw{We then compute the mismatch for all testing waveforms and plot the
-% distribution of the mismatch in figure X. We find the distribution...}
+% \kw{We then compute the mismatch for all testing waveforms and plot the distribution of the mismatch in figure X. We find the distribution...}
 

### 2023-04-12 src/tex/ms.tex
update

--- 2023-03-31 src/tex/ms.tex
@@
 \newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
+\definecolor{cyan}{HTML}{0097A7}
+\newcommand{\kl}[1]{{\color{cyan}[KL: #1 ]}}
+\definecolor{rr}{RGB}{173, 37, 37}
+\newcommand{\te}[1]{{\color{rr}[TE: #1 ]}}
 
@@
 Currently, three families of GW models are commonly used. They are the
-effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
-(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
-\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
-the newest models}. EOB models are constructed by mapping two masses onto an
+effective-one-body (EOB) \citep{ossokine2020multipolar, cotesta2020frequency, taracchini2014effective}, Numerical Relativity
+(NR) surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2} and phenomenological (Phenom) models
+\citep{husa2016frequency, khan2016frequency, garcia2020multimode, pratten2021computationally}. 
+EOB models are constructed by mapping two masses onto an
 effective body under an effective metric; NR surrogate models construct
@@
 according to the original structure. {\ripple} \citep{ripple} was a new
-implementation of IMRPhenomD, one of the Phenom models. It was first implemented
+implementation of IMRPhenomD, one of the models within the Phenom family. It was first implemented
 in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
-rewritten using \jax, another {\python} package that supports AD. Using
+rewritten using \jax, a {\python} package that supports AD. Using
 {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
@@
 functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. These simple analytic functions
-consists of parameters $\Lambda^i$, which are defined as follows. \kw{Can you
-change equation one such that it consists of $\Lambda$? Also be explicit about
-the form of the function used for the fitting. Like write down the polynomials
-and lorentizans for the specific parts.}
+of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
+\begin{equation}\label{eq:amplitude}
+\begin{aligned}
+	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
+	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
+	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
+	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
+\end{aligned}
+\end{equation}
+where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
+The ansatzes for phase can be found in \citep{khan2016frequency}.
+These simple analytic functions
+consists of parameters $\Lambda^i=\{\rho_i,\delta_i,\gamma_i\}$, which are defined as follows. 
 \begin{align} \label{eq:Lambda}
@@
 Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
-respectively. Finally, the individual segments are connected in a way that the
-final waveform is continuous in its first derivative \kw{expand a bit}.
+respectively. Finally, the individual segments are first connected directly using 
+step functions. Then, by fixing coefficients in the intermediate segment, one can make 
+the final waveform is continuous in its first derivative.
 
-\kw{Need to work on this paragraph}
-From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are
-fundamental to waveform generation, which can significantly affect the shape of
-final waveforms. Thus, having a set of accurate waveform coefficients is
-important. Generally, waveform coefficients are obtained by calibrating with NR
-waveforms, which are waveforms computed using NR simulations. In the case of
-\citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated
-waveforms to NR waveforms. For each NR simulation, they can obtain one set of
-$\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are
-subsequently found by fitting against Eq. \ref{eq:Lambda}. Since the fitting
-procedure is done in a piece-wise manner, the correlations between different
-segments are omitted, which could limit the accuracy of the model. Also,
-their calibration procedure does not guarantee to have the optimal set of
-$\lambda$, since the process of connecting segments alters the previously fitted
-waveform. The model-generated waveforms do not take this effect into account,
-and thus introduces additional inaccuracies.
-
-% they did not use mismatch as the loss function
+%\kw{Need to work on this paragraph}
+Combining Eq.~\ref{eq:joint_waveform},~\ref{eq:amplitude}~and~\ref{eq:Lambda}, 
+we can see that the entire waveform is non-linear in $\lambda$. A slightly 
+inaccurate set of $\lambda$ can significantly affect the shape of the generated 
+waveforms. Thus, having a set of accurate waveform coefficients is important 
+and fundamental to having accurate GW models. Generally, waveform coefficients 
+are obtained by calibrating with NR waveforms, which are waveforms computed 
+using NR simulations. In the case of \citep{khan2016frequency}, they first 
+obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+the correlations between different segments are omitted, which could limit the 
+accuracy of the model. Also, since fitting was performed before connecting 
+individual segments, the final waveform does not guarantee to achieve the optimal 
+waveform. The connecting procedure alters the previously fitted waveform. 
+Hence, the model generated waveforms contains additional inaccuracies.
 
@@
 
-\kw{Remind me why don't we use the hybrid waveforms again? Cause we couldn't do it?}
-In the original calibration \citep{khan2016frequency,husa2016frequency},
-training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB
-(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
-waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since most NR waveforms used
-have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
-which they are long enough to contain part of the inspiral segment and all
-merger and ringdown frequency information. We take the frequency limits as
-$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,
-where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set higher
-than that in the original calibration due to NR length. When compared with
-IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency.
+\kl{The maximum is calculated by comparing the phase of IMRPhenomD and NR waveforms. We know that they are off by a linear relation of $f$, 
+\begin{align*}
+	\phi_{NR}-\phi_{IMR}=(2\pi t_0)f+\phi_0.
+\end{align*}
+Since I have $\phi_{NR}$, $\phi_{IMR}$ and the frequency array, I did a linear regression to get $2\pi t_0$ and $\phi_0$.}
 
-To compute the integral shown in Eq. \ref{eq:inner_prod}, one need to choose a
-relevant noise PSD. Since a common goal for creating waveform approximent is to
-deploy the waveform model in downstream data analysis tasks, a reasonable choice
-for the noise PSD should be closely related to the instrument of interest. We
-choose the zero-detuned high-power (\zdethp) noise PSD \kw{add citation} as our
-first choice of noise PSD. Because the total mass of the system simple scales
-the frequency of the waveform, we need to choose a corresponding mass scale to
-match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
-all waveforms, since this mass scale is common in LIGO observations. In addition
-to a detector-dependent PSD, we are also interested in a more
-instrument-agnostic case such that we can evaluate the effectiveness of the
-recalibration without considering detector effect. In this case, we set the
-noise PSD to be a constant value.
-
-% GW strain treatment from time-series to frequency domain
-In our work, mismatches are used as a measure of discrepancies between
-IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms
-from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are
-calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form
-of time-series data. Since time-series data is oscillatory, performing
-optimization in the time-domain is not ideal. Hence, we had chosen to transform
-NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
-same intrinsic parameters. We taper the time-series using Tukey window.
-\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-entire BBH event. } Then, the frequency spectra can be obtained by taking the
-Fourier transform of the time-series. Then, mismatches can be calculated through
-Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.
-
-\subsection{Optimization Scheme} \label{subsec:optimization}
-
-Using \ripple, we can make use of {\jax} to evaluate gradients accurately. Thus,
-we can utilize simple gradient descent to recalibrate the model.
-
-% \kw{Mention the main advantage in this study is we can optimize everything at
-% once.}
-
-% To recalibrate the model, we have to define loss functions that quantify the
-% difference between the model-generated waveforms and NR waveforms. In the
-% original calibration, they took the mean square error as the loss function for
-% optimization. Instead, we directly use mismatch as our metric for loss
-% functions. However, the mismatch of individual waveforms is not useful in
-% optimization, since it does not represent all waveforms with various intrinsic
-% parameters. We ought to calibrate coefficients that suit waveforms with
-% different intrinsic parameters simultaneously, hence waveforms in different
-% regions of parameter space should be used in the optimization scheme. Here, we
-% chose two different loss functions to examine their ability to optimize the
-% model, e.g. mean of individual mismatches $\mathcal{L}_{\mathrm{mean}}$ and
-% fractional loss $\mathcal{L}_{\mathrm{fl}}$.
-
-% Do something much more direct:
-To recalibrate the model, we need to define a loss function that quantifies the
-difference between the model-generated waveforms and NR waveforms. We use
-mismatch over the set of all training waveforms as the loss function. While in
-\citep{khan2016frequency}, they chose to use the mean square error as the loss
-function, we choose mismatch as the loss function because it is closely related
-to downstream objectives such as the likelihood in parameter estimation. The
-mismatch is averaged over the training dataset in two ways to cover the
-parameter space of interest. One of which takes a simple average and the other
-takes a normalized average based on initial mismatch.
-
-% To recalibrate the model, we need to define a loss function that quantify ...
-% We use mismatch over the set of all training waveforms as the loss function.
-% We choose mismatch as the loss function because it is closely related to
-% downstream objectives such as the likelihood in parameter estimation. The
-% mismatch is averaged over the training dataset to cover the parameter space of
-% interest. (used)
+Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
+However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
+To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
+We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
 \begin{align}
@@
 \end{align}	
-$\mathcal{M}_i$ represents the mismatch of an individual training waveform,
+where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
 $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
 training waveform, and $N$ is the total number of individual training waveforms.
-For the first choice, it takes the mean of the mismatches, which serves as a
-basic loss function. However, it is prone to be dominated by a single waveform
-with a large mismatch. Other waveforms with smaller mismatches would be
+Note that we choose to use two different averages, since they have different 
+preferences in optimization base on waveform mismatches.
+For the first choice, simple average serves as the simplest choice of loss function, but is prone to be dominated by a single waveform with a large mismatch.
+Other waveforms with smaller mismatches would be
 insignificant comparatively, and might not be able to improve under such
-optimization. Alternatively, the second choice, the mean of normalized
-mismatches, eliminates the aforementioned issue. Nevertheless, it excludes the
+optimization. Alternatively, the second choice, normalized average eliminates the aforementioned issue. Nevertheless, it excludes the
 information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
@@
 $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and final individual mismatches would be in a similar order of magnitude. In
-this paper, we display the results of using both loss functions and examine the
-difference between them.
+and individual mismatches would be in a similar order of magnitude after 
+optimization. In this paper, we showcase the results of using both loss functions 
+and examine the differences between them.
 
-We have chosen 11-16 NR waveforms from the set of waveforms used in the original
-calibration process. Originally, 19 waveforms are taken from NR simulations for
-calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they
-are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not
-publicly available, we cannot take the identical training set as them. Instead,
-we take the available waveforms from the SXS catalog to construct our loss
-function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the
-maximum mass ratio in our training set is lower than that in
-\citep{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio
-is set as 8. This is because SXS catalog does not have NR waveforms with
-extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with
-$q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model
-with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the
-maximum mass ratio to 8.
+\subsection{Optimization Scheme} \label{subsec:optimization}
 
-With both mismatches, we apply gradient descent to optimize the tunable
-coefficients as follows: 
-\begin{algorithm}
+To compute the loss functions, we have to take NR waveforms for calculating the mismatch. We choose 11-16 NR waveforms from the set of waveforms used in the original
+calibration process as training waveforms. Originally, 19 waveforms are taken from 
+NR simulations for calibrating IMRPhenomD \citep{khan2016frequency}, 
+which are waveforms from the SXS catalog \citep{boyle2019sxs} or BAM simulation. 
+As BAM waveforms are not publicly available, we cannot take the identical training 
+set as them.
+Instead, we take the available waveforms from the SXS catalog to construct our loss function.
+Training waveforms used are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+The training waveforms chosen has maximum mass ratio to be 8.
+This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+In fact, the SXS catalog only has NR waveforms with $q\leq10$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed from LIGO have $q\leq8$.
+Hence, we calibrate IMRPhenomD with waveforms of $q\leq8$.
+
+In the SXS catalog, NR waveforms are in the form of time-series strain.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
+same intrinsic parameters. We taper the time-series using Tukey window.
+\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
+	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
+	entire GW strain. } 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series.
+
+\te{I think this discussion needs to be much more precise.
+Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
+So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
+\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}
+
+% Need to change the focus to flat PSD instead of ALIGO PSD
+Other than NR waveforms, one need to choose a relevant noise PSD for 
+mismatch.
+We have opted to use a flat PSD for this purpose, as it provides results that are independent of the detector sensitivity and mass scale.
+The use of a flat PSD ensures that the improvement in accuracy is due solely to the difference in high-dimensional fitting and piece-wise fitting.
+Furthermore, we are interested in examining the effect of introducing a detector PSD on the optimization process.
+For this, we have chosen the zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
+Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD.
+Therefore, we have selected a mass scale of $M=50M_{\odot}$ for all waveforms, as this is a commonly observed mass scale in LIGO observations.
+We will then examine the differences in optimization.
+
+%Other than NR waveforms, one need to choose a relevant noise PSD for 
+%mismatch. Since a common goal for creating waveform approximant is to
+%deploy the waveform model in downstream data analysis tasks, a reasonable choice
+%for the noise PSD should be closely related to the instrument of interest. We
+%choose the zero-detuned high-power (\zdethp) noise PSD as our
+%first choice of noise PSD \citep{aasi2015advanced}. Because the total mass of the system simple scales
+%the frequency of the waveform, we need to choose a corresponding mass scale to
+%match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
+%all waveforms, since this mass scale is common in LIGO observations. In addition
+%to a detector-dependent PSD, we are also interested in a more
+%instrument-agnostic case such that we can evaluate the effectiveness of the
+%recalibration without considering detector effect. In this case, we set the
+%noise PSD to be a constant value.
+
+We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
+In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
+(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
+waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models.
+Thus, for simplicity sake, we ignore this procedure.
+On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
+which they are long enough to contain part of the inspiral segment and all
+merger and ringdown frequency information. We take the frequency limits as
+$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	% 0.001 Hz?
+where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
+the IMRPhenomD's frequency range, except the minimum frequency is set higher
+than that in the original calibration due to NR length. When compared with
+IMRPhenomC, the frequency range is slightly extended to have a higher maximum
+frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
+which is sufficient to capture all features of GW strain.
+
+With the loss function evaluated, we apply gradient descent to optimize the tunable
+coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
+original coefficients given in \citep{khan2016frequency}. We take them as the initial 
+waveform coefficients because they lie in the neighborhood of the minimum that we wish 
+to find. Then, by taking $\alpha=10^{-6}$ and $N=30000$, we can optimize the 
+coefficients systematically. 
+\te{I think we need a bit more discussion about the stopping criterion here.
+What is N, and why did we choose it to be this number?
+Ideally we would actually have a plot of the loss function during training.
+I don't have a good intuition for if its noisy.
+Does it plateau?
+If we wanted to ensure that people don't complain about overfitting we could also plot the loss of the validation set to show its not going up.}
+
+\begin{figure}
+	\script{loss.py}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/loss.pdf}
+%	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
+	\label{fig:loss}
+\end{figure}
+
+\begin{algorithm}[t]
 	\caption{Gradient descent pseudocode}
+	\label{alg:gradient}
 	\KwIn{initial coefficients $\lambda_i$} \Parameters{number of iterations
@@
 
-To evaluate the effectiveness of our optimization approach, we analyze
-approximately 530 NR waveforms from the SXS catalog in this study. We
-specifically select waveforms with negligible eccentricity
-(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
-consistent with the constraints of the waveform model. As shown in
-Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
-waveforms fall within the parameter space defined by the training waveforms.
-This suggests that these test waveforms can be used for direct comparison with
-the original model.
+First, to judge the performance of the optimization scheme, we take $\sim530$ NR
+waveforms from the SXS catalog. We choose testing waveforms with negligible
+eccentricity (${e<2\times10^{-3}}$) and precession
+(${\chi_{x,y}<5\times10^{-3}}$) to fit with the limitation of the model. We see
+that in Fig.~\ref{fig:intrin_space}, testing waveforms have intrinsic parameters
+that are within the parameter space spanned by training waveforms. Hence, we can
+take these testing waveforms to compare with the original model.
 
-We begin by investigating the impact of joint optimization on an individual
-waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
-outperforms the original waveform, particularly in the inspiral region, where
-the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
-selected one of the testing waveforms from the catalog presented in
-\citep{khan2016frequency}.
+We first examine the effect of joint optimization on a single waveform. In
+Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy
+than the original waveform, especially in the inspiral region, where the
+amplitude has a $50\%$ decrease in error. We have chosen one of the testing
+waveforms listed in \citep{khan2016frequency} for a fair comparison. 
 % In general, most waveforms within the neighborhood of the training waveforms
@@
 
-We use a constant noise spectrum, $\mathcal{L}{\mathrm{mean}}$, to evaluate the
-mismatch of all testing waveforms, and present the resulting distribution in
-Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
-mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
-in the median. When using $\mathcal{L}{\mathrm{fl}}$, we observe a similar
-improvement with a 22.9\% decrease in the median. However, the distribution lacks
-a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
-Despite this, both distributions exhibit significant improvement, and the
-optimization scheme remains unaffected.
+Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute
+the mismatch for all testing waveforms and plot the distribution of the mismatch
+in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end
+with a lower mismatch. Quantitatively, the peak has lowered by almost an
+order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease.
+Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar
+improvement. We see that the median of distribution has a $22.9\%$ decrease.
+However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}.
+This is due to the problem mentioned in section \ref{subsec:optimization}.
+Nevertheless, both distributions show improvement and the potential problem did
+not affect the optimization scheme significantly.
 
@@
 
-Applying the same methods, we observe that the distributions of mismatches
-calculated using the {\zdethp} noise spectrum exhibit superior improvement
-compared to the weighted mismatch. The shape of the distribution is similar to
-that shown in Fig. \ref{fig:q148}. This result is expected, as the IMRPhenomD
-model was initially constructed and fitted using the {\zdethp} weighted
-mismatch. Consequently, the model is anticipated to closely fit the NR waveforms
-while incorporating the influence of the {\zdethp} noise spectrum, rather than a
-constant spectrum.
+Similarly, using the same procedures, distributions of mismatches calculated
+using {\zdethp} noise spectrum show better improvement than the weighted
+mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}. This
+effect is expected since the IMRPhenomD model was originally constructed and
+fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that
+fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum
+instead of the constant spectrum.
 
-Motivated by the successful improvement of the waveforms, we expand the training
-dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. The new
-set of coefficients generated from this optimization process yields only
-marginal improvements in the newly produced waveforms, as seen in Fig.
-\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
-distribution remains comparable in length and endpoint to the original
-distribution, indicating that our procedure is incapable of improving these
-waveforms. Similarly, utilizing the {\zdethp} spectrum to optimize the loss
-function with additional waveforms leads to minimal improvement in the resulting
-distribution.
+With the success of improving waveforms, we increase the number of training
+waveforms for optimization. Taking additional waveforms listed in
+Tab.~\ref{tab:q1248}, we obtain a new set of coefficients. We see in Fig.
+\ref{fig:q148_q1248_compare}, new waveforms produced only have a very small
+improvement. The high mismatch tail of the optimized distribution remains
+similar in length and endpoint as the original distribution, meaning that they
+cannot be improved using our procedure. Likewise, using additional waveforms to
+optimize loss function with {\zdethp} spectrum shows the same result, where the
+new distribution barely shows any improvement.
 
@@
 
-Given that the waveform model's ansatz may not be entirely compatible with NR,
-and the optimization procedure is carried out over a distribution of waveforms
-with varying source parameters, it is conceivable that some trade-offs in
-accuracy exist between different parts of the parameter space. If this is the
-cause of the high mismatch tail's failure to improve during joint-optimization,
-segmenting the parameter space into smaller subspaces should alleviate this
-problem. On the other hand, if the ansatz lacks the correct parameterized form
-to capture the NR waveforms' behavior as a function of the intrinsic parameters,
-the results will always be biased, and we should not expect any improvement,
-even if we segment the parameter space during training.
+Given the ansatz used in the waveform model is unlikely to be fully compatible
+with NR, and the optimization procedure is done over a distribution of waveforms
+with different source parameters, it is conceivable that there are some
+trade-off in accuracy of the waveform models between different part of the
+parameter space. If this is the reason why the high mismatch tail is not
+improving during the joint-optimization, separating the parameter space into
+smaller subspace should help alleviate this issue. On the other hand, if the
+ansatz does not have the right parameterized form to capture the behavior of the
+NR waveforms as a function of the intrinsic parameters, the result should be
+always biased, and we should not expect any improvement even if we separate the
+parameter space into smaller space during training.
 

### 2023-04-12 src/scripts/loss.py

@@ -1,0 +1,42 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 16,
+    "axes.titlesize": 16,
+    "xtick.labelsize": 14,
+    "ytick.labelsize": 14,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
+data = np.loadtxt(paths.data / "q148_loss.dat")
+
+plt.plot(data[:,0], data[:,1])
+plt.ylabel('Loss')
+plt.xlabel('Epoch')
+plt.savefig(paths.figures / "loss.pdf", bbox_inches="tight")

### 2023-04-15 src/tex/ms.tex
added loss figure and other updates

@@
 
+% Put more focus on the joint optimization instead of having a set of more accurate 
+% coefficients, since the method is the main point
 In this paper, we investigate the possibility of further improving the accuracy
@@
 understanding of Phenom models.
-% \kw{Briefly describe the high level finding in this paper}
 
@@
 functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
+of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
+amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
 \begin{equation}\label{eq:amplitude}
@@
 Note that we choose to use two different averages, since they have different 
-preferences in optimization base on waveform mismatches.
-For the first choice, simple average serves as the simplest choice of loss function, but is prone to be dominated by a single waveform with a large mismatch.
-Other waveforms with smaller mismatches would be
-insignificant comparatively, and might not be able to improve under such
-optimization. Alternatively, the second choice, normalized average eliminates the aforementioned issue. Nevertheless, it excludes the
+preferences in optimization base on waveform mismatches. For the first choice, 
+simple average serves as the simplest choice of loss function, but is prone to 
+be dominated by a single waveform with a large mismatch. Other waveforms with 
+smaller mismatches would be insignificant comparatively, and might not be able 
+to improve under such optimization. Alternatively, the second choice, normalized 
+average eliminates the aforementioned issue. Nevertheless, it excludes the
 information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
@@
 Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD.
-Therefore, we have selected a mass scale of $M=50M_{\odot}$ for all waveforms, as this is a commonly observed mass scale in LIGO observations.
+We selected a mass scale of $M=50M_{\odot}$ for all waveforms, as this is a commonly observed mass scale in LIGO observations.
 We will then examine the differences in optimization.
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-%	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
+	\caption{The value of loss function against number of iterations. The code is terminated after 30000 steps }
 	\label{fig:loss}
@@
 more consistent improvement, whereas spinning waveforms can either improve or
-worsen in terms of mismatch. \kw{What does this means?}
+worsen in terms of mismatch base on their intrinsic parameters. 
+%\kw{What does this means?}
 
@@
 equal or similar spin, hence the model prefers waveforms with similar spin. One
-interesting feature is how the second and fourth quadrants respond to
-optimization. \kw{How did you count quadrant?} In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
-generally improve with along optimization. However, mismatches in the fourth
-quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most
+interesting feature is how the top-left ($\chi_1<0$ and $\chi_2>0$) and bottom-right ($\chi_1>0$ and $\chi_2<0$) regions respond to
+optimization. In the top-left region, waveforms
+generally improve with along optimization. However, mismatches in the bottom-right region do not improve after optimization. Most
 waveforms even turned worse after optimization. These waveforms correspond to
@@
 	\end{tabularx}
-	\caption{List of waveforms used in recalibrating coefficients in 4 quadrants.
-From top to down are the first, second, third and fourth quadrants.
-Note that for the first and third quadrants, waveforms are chosen to have equal or similar spins, while the training waveforms for the second and fourth quadrants are chosen to have opposite spins.}
+	\caption{List of waveforms used in recalibrating coefficients in 4 regions.
+From top to down are the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$), bottom-left ($\chi_1,\chi_2<0$) and bottom-right ($\chi_1>0>\chi_2$) regions.
+Note that for the top-right and bottom-left regions, waveforms are chosen to have equal or similar spins, while the training waveforms for the other two regions are chosen to have opposite spins.}
 	\label{tab:quadrants}
@@
 
-We divided the parameter space into four quadrants to analyze the effect of the
+We divided the parameter space into four regions to analyze the effect of the
 recalibration procedure on each region separately (Fig.\ref{fig:ps_q148}). The
 training waveforms used for fitting are listed in Table \ref{tab:quadrants}. The
-second and fourth quadrants have limited data for $q>4$, hence the results are
+top-left and bottom-right regions have limited data for $q>4$, hence the results are
 only valid up to $q\leq4$. From Fig.\ref{fig:ps_q148_quadrant}, we observe
-improvements in waveform mismatch for all waveforms, except those in the fourth
-quadrant. For equal-spin waveforms lying in the proximity of the diagonal axis,
+improvements in waveform mismatch for all waveforms, except those in the bottom-right region. For equal-spin waveforms lying in the proximity of the diagonal axis,
 the mismatch improves significantly due to most training waveforms being
-equal-spin. In the second quadrant, the improvement is noticeable, except for a
+equal-spin. In the top-left region, the improvement is noticeable, except for a
 few defects caused by some testing waveforms having $q>4$. The optimized
-waveforms in the fourth quadrant have a higher mismatch than the original ones,
+waveforms in the bottom-right region have a higher mismatch than the original ones,
 indicating poor fitting. Further optimization in a smaller subspace does not
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches after optimizing in separate quadrants.
+	\caption{Distributions of mismatches after optimizing in separate regions.
 	We use a constant noise spectrum to calculate mismatch and
 	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. Generally, most
-	waveforms with mismatch $>10^{-2}$ lies in the fourth quadrant. }
+	waveforms with mismatch $>10^{-2}$ lies in the bottom-right region. }
 	\label{fig:all_quadrants}
@@
 	\caption{\kw{The color scheme makes the plot a bit hard to read.}Parameter
-	space of testing waveforms. Each quadrant is fitted independently. Colorbar
+	space of testing waveforms. Each region is fitted independently. Colorbar
 	represents log difference of mismatches before and after optimization.}
@@
 region than the bottom right along a degeneracy line. In an attempt to address
-this issue, we partitioned the parameter space into four quadrants, as described
+this issue, we partitioned the parameter space into four regions, as described
 in Sec. \ref{sec:result}. However, even with separate optimizations for each
-quadrant, Fig.\ref{fig:ps_q148_quadrant} indicates that the fourth quadrant
-exhibits similar mismatches as before, while the second quadrant's performance
+regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the bottom-right region
+exhibits similar mismatches as before, while the top-left region's performance
 has improved. This observation suggests that the ansatz is specific to certain
@@
 
-The division of the parameter space into four quadrants was a simple approach
+The division of the parameter space into four regions was a simple approach
 taken for practical reasons. A more systematic approach would involve the use of
@@
 improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-certain preference for waveforms lying in the second quadrant while the fourth
+certain preference for waveforms lying in the top-left region while the fourth
 quadrant cannot be improved. To further test this result, we recalibrate
@@
 Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further
-improvement to the second quadrant while the fourth quadrant shows similar
-result. This indicates that the model's ansatz does not fit waveforms in the
-fourth quadrant. This phenomenon is due to the reduced spin approximation used
+improvement to the top-left region while the bottom-right region shows similar
+result. This indicates that the model's ansatz does not fit waveforms in this region. This phenomenon is due to the reduced spin approximation used
 in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$
@@
 
-While we naively separate the optimization process into 4 quadrants, one can
+While we naively separate the optimization process into 4 regions, one can
 perform systematic region-selection. In principle, we can apply this general

### 2023-04-15 src/scripts/all_quadrants.py

@@ -7,10 +7,10 @@ params = {
     "font.size": 18,
     "legend.fontsize": 12,
     "legend.frameon": True,
-    "axes.labelsize": 18,
-    "axes.titlesize": 18,
-    "xtick.labelsize": 18,
-    "ytick.labelsize": 18,
+    "axes.labelsize": 16,
+    "axes.titlesize": 16,
+    "xtick.labelsize": 14,
+    "ytick.labelsize": 14,
     "axes.unicode_minus": False,
     "figure.figsize": (7, 5),
     "xtick.top": True,
@@ -35,14 +35,32 @@ params = {
 mpl.rcParams.update(params)
 
 q148 = np.loadtxt(paths.data / "all_quadrants.dat")
+data2 = np.array([i for i in q148 if (i[2] < 0 and i[3] > 0)])
+data4 = np.array([i for i in q148 if (i[2] > 0 and i[3] < 0)])
 
-bins = np.arange(-5, -1, 0.2)
-plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
-plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0")
-plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.3, color="C1")
-plt.hist(np.log10(q148[:,5]), bins=bins, label="Optimized", histtype="step", color="C1")
-plt.legend()
-plt.ylabel('Count')
-plt.xlabel('$\log(\mathrm{mismatch})$')
+bins = np.arange(-5, -1, 0.25)
+
+fig, ax = plt.subplots(2, figsize=(6, 10))
+
+ax[0].hist(np.log10(data2[:,4]), bins=bins, label="top-left original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
+ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=1.3, alpha=0.8)
+ax[0].set_ylabel('Count')
+ax[0].legend(loc='upper left')
+
+ax[1].hist(np.log10(data4[:,4]), bins=bins, label="bottom-right original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
+ax[1].hist(np.log10(data4[:,5]), bins=bins, label="bottom-right optimized", histtype="step", color="C1", linewidth=1.3, alpha=0.8)
+ax[1].set_xlabel('log(mismatch)')
[…]

### 2023-04-15 src/scripts/ps_q148_qchi.py

@@ -40,7 +40,9 @@ logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,0], data[:,1], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar = plt.colorbar(sc)
+cbar.ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
 plt.xlabel('$q$')
 plt.ylabel('$\chi_{\mathrm{PN}}$')
 

### 2023-04-15 src/scripts/ps_q148_quadrants.py

@@ -40,7 +40,9 @@ logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-plt.colorbar(sc).ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar = plt.colorbar(sc)
+cbar.ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 

### 2023-04-15 src/tex/ms.tex
minor caption fix

@@
 So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
-This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
-\te{This choice seems pretty arbitrary. Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70. This choice is important and should be more informed.}
-
-% Need to change the focus to flat PSD instead of ALIGO PSD
-Other than NR waveforms, one need to choose a relevant noise PSD for 
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}Other than NR waveforms, one need to choose a relevant noise PSD for
 mismatch.
@@
 We selected a mass scale of $M=50M_{\odot}$ for all waveforms, as this is a commonly observed mass scale in LIGO observations.
-We will then examine the differences in optimization.
+\te{This choice seems pretty arbitrary.
+Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70.
+This choice is important and should be more informed.}We will then examine the differences in optimization.
 
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches after optimizing in separate regions.
+	\caption{Distributions of mismatches in the top-left and bottom-right regions.
 	We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ for the loss function. Generally, most
-	waveforms with mismatch $>10^{-2}$ lies in the bottom-right region. }
+	$\mathcal{L}_{\mathrm{mean}}$ for the loss function.
+We see waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened.
+\kl{I tried to overlap these two histograms together but it makes them very hard to read.
+}}
 	\label{fig:all_quadrants}

### 2023-04-21 src/tex/ms.tex
figure fix

@@
 
-\te{I think this discussion needs to be much more precise.
+%\te{I think this discussion needs to be much more precise.
 Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
 So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
-This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}Other than NR waveforms, one need to choose a relevant noise PSD for
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
+Other than NR waveforms, one need to choose a relevant noise PSD for 
 mismatch.
 We have opted to use a flat PSD for this purpose, as it provides results that are independent of the detector sensitivity and mass scale.
-The use of a flat PSD ensures that the improvement in accuracy is due solely to the difference in high-dimensional fitting and piece-wise fitting.
+The use of a flat PSD ensures that the improvement in accuracy is due mainly to the difference in high-dimensional fitting and piece-wise fitting, but not due to the use of a different mass scale.
 Furthermore, we are interested in examining the effect of introducing a detector PSD on the optimization process.
@@
 Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD.
-We selected a mass scale of $M=50M_{\odot}$ for all waveforms, as this is a commonly observed mass scale in LIGO observations.
-\te{This choice seems pretty arbitrary.
+We selected an arbitrary mass scale of $M=50M_{\odot}$ for all waveforms for demonstration, as this is a commonly observed mass scale in LIGO observations.
+%\te{This choice seems pretty arbitrary.
 Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70.
@@
 merger and ringdown frequency information. We take the frequency limits as
-$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	% 0.001 Hz?
+$f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
 where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
@@
 waveform coefficients because they lie in the neighborhood of the minimum that we wish 
-to find. Then, by taking $\alpha=10^{-6}$ and $N=30000$, we can optimize the 
-coefficients systematically. 
-\te{I think we need a bit more discussion about the stopping criterion here.
+to find. Then, by taking $\alpha=10^{-6}$, we see in Fig. \ref{fig:loss} plateau at 
+around $N=30000$, hence we end the optimization here.  
+%\te{I think we need a bit more discussion about the stopping criterion here.
 What is N, and why did we choose it to be this number?
@@
 \citep{khan2016frequency}.
-% In general, most waveforms within the neighborhood of the training waveforms
-% show improvement in mismatches.
 
-We use a constant noise spectrum, $\mathcal{L}{\mathrm{mean}}$, to evaluate the
+We use the constant PSD weighted loss function, $\mathcal{L}_{\mathrm{mean}}$, to evaluate the
 mismatch of all testing waveforms, and present the resulting distribution in
@@
 mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
-in the median. When using $\mathcal{L}{\mathrm{fl}}$, we observe a similar
+in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
 improvement with a 22.9\% decrease in the median. However, the distribution lacks
@@
 optimization scheme remains unaffected.
-
-% \kw{We then compute the mismatch for all testing waveforms and plot the
-% distribution of the mismatch in figure X. We find the distribution...}
 
@@
 	\caption{Distributions of waveform mismatches calculated using both
-	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$ in
-	recalibration. We use training waveforms listed in Tab.~\ref{tab:q148} and
-	mismatches are weighted with the constant noise spectrum. For the
+	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$. We use 
+	training waveforms listed in Tab.~\ref{tab:q148} and
+	mismatches are weighted with the constant PSD. For the
 	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\%
@@
 	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant noise spectrum with the loss function
-	$\mathcal{L}_{\mathrm{mean}}$.}
+	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
 	\label{fig:q148_q1248_compare}
@@
 Applying the same methods, we observe that the distributions of mismatches
-calculated using the {\zdethp} noise spectrum exhibit superior improvement
+calculated using the {\zdethp} PSD exhibit superior improvement
 compared to the weighted mismatch. The shape of the distribution is similar to
@@
 mismatch. Consequently, the model is anticipated to closely fit the NR waveforms
-while incorporating the influence of the {\zdethp} noise spectrum, rather than a
-constant spectrum.
+while incorporating the influence of the {\zdethp} PSD, rather than a
+constant PSD.
 
@@
 distribution, indicating that our procedure is incapable of improving these
-waveforms. Similarly, utilizing the {\zdethp} spectrum to optimize the loss
+waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
 function with additional waveforms leads to minimal improvement in the resulting
@@
 and the optimization procedure is carried out over a distribution of waveforms
-with varying source parameters, it is conceivable that some trade-offs in
+with varying intrinsic parameters, it is conceivable that some trade-offs in
 accuracy exist between different parts of the parameter space. If this is the
@@
 Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
-more consistent improvement, whereas spinning waveforms can either improve or
-worsen in terms of mismatch base on their intrinsic parameters. 
-%\kw{What does this means?}
-
-Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
+more consistent improvement, since the ansatz are developed base on non-spinning 
+waveforms. Also, the original coefficients were fitted using NR waveforms with
+equal or similar spin, hence the model prefers waveforms with similar spin. On 
+the other hand, spinning waveforms can either improve or worsen in terms of 
+mismatch base on their intrinsic parameters. To further discuss the behavior of 
+non-spinning waveforms, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
 Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. 
-$\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this
-feature since the original coefficients were fitted using NR waveforms with
-equal or similar spin, hence the model prefers waveforms with similar spin. One
-interesting feature is how the top-left ($\chi_1<0$ and $\chi_2>0$) and bottom-right ($\chi_1>0$ and $\chi_2<0$) regions respond to
-optimization. In the top-left region, waveforms
-generally improve with along optimization. However, mismatches in the bottom-right region do not improve after optimization. Most
-waveforms even turned worse after optimization. These waveforms correspond to
-the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}. \kw{Give justification.}
+$\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
+Meanwhile, the top-left ($\chi_1<0$ and $\chi_2>0$) and bottom-right ($\chi_1>0$ 
+and $\chi_2<0$) regions respond to optimization differently. In the top-left region, waveforms
+generally improve with along optimization. However, mismatches in the bottom-right 
+region do not improve after optimization. Most waveforms even turned worse after 
+optimization.
 
@@
 We divided the parameter space into four regions to analyze the effect of the
-recalibration procedure on each region separately (Fig.\ref{fig:ps_q148}). The
-training waveforms used for fitting are listed in Table \ref{tab:quadrants}. The
+recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148}). The
+training waveforms used for fitting are listed in Table~\ref{tab:quadrants}. The
 top-left and bottom-right regions have limited data for $q>4$, hence the results are
-only valid up to $q\leq4$. From Fig.\ref{fig:ps_q148_quadrant}, we observe
-improvements in waveform mismatch for all waveforms, except those in the bottom-right region. For equal-spin waveforms lying in the proximity of the diagonal axis,
-the mismatch improves significantly due to most training waveforms being
-equal-spin. In the top-left region, the improvement is noticeable, except for a
-few defects caused by some testing waveforms having $q>4$. The optimized
-waveforms in the bottom-right region have a higher mismatch than the original ones,
-indicating poor fitting. Further optimization in a smaller subspace does not
-improve the results.
+only valid up to $q\leq4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe
+improvements in waveform mismatch for all waveforms, except those in the bottom-right 
+region. For equal-spin waveforms lying in the proximity of the diagonal axis,
+the mismatch has additional improvement due to most training waveforms being
+equal-spin. In the top-left region, the improvement is significant, except for a
+few defects caused by some testing waveforms having $q>4$. This can be seen clearer in 
+Fig.~\ref{fig:all_quadrants}, which we see that the optimized histogram shifts downward uniformly. 
+The optimized waveforms in the bottom-right region have a higher mismatch than the original ones,
+indicating poor fitting. These waveforms correspond to the waveforms in the high mismatch tail 
+in Fig.~\ref{fig:q148}, which we can see in Fig.~\ref{fig:all_quadrants}. Thus, we show that 
+further optimization in a smaller subspace does not improve the results.
 
@@
 construction principles to the IMRPhenomD model. For instance, they also use PN
-approximants as part of the ansatz in the inspiral segment. Interestingly, while
-recalibrating the IMRPhenomX model \citep{pratten2020setting}, which is
+approximants as part of the ansatz in the inspiral segment. It will be interesting 
+to recalibratethe IMRPhenomX model \citep{pratten2020setting}. Because it is
 parameterized by an additional anti-symmetric spin parameter, it is expected not
@@
 Furthermore, this approach is applicable to other GW model families, such as NR
-surrogate models \citep{varma2019surrogate} or EOB models
-\citep{taracchini2014effective}. Such an approach could simplify NR waveform
-calibration procedures and lead to the improvement of existing models.
-
-% Can we further check systematic bias by injection? But how can we know the
-% real mass or spin in the first place if we disregard GW models.
+surrogate models or EOB models introduced previously. Such an approach could 
+simplify NR waveform calibration procedures and lead to the improvement of 
+existing models.
 
@@
 mismatches have a slightly better improvement. In contrast, different types of
-loss function result in significantly different final mismatch distributions,
-where the result can be seen in Fig.~\ref{fig:q148}. By increasing the number of
-training waveforms, we see a slight improvement increase in
-Fig.~\ref{fig:q148_q1248_compare}.
+loss function result in significantly different final mismatch distributions. As 
+seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
+$\mathcal{L}_{fl}$. By increasing the number of training waveforms, we see a 
+slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
 
@@
 improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-certain preference for waveforms lying in the top-left region while the fourth
-quadrant cannot be improved. To further test this result, we recalibrate
-waveforms in separate regions in parameter space. As shown in
-Fig.~\ref{fig:ps_q148_quadrant}, this recalibration process gives further
+certain preference for waveforms lying in the top-left region while the bottom-right 
+region cannot be improved. To further test this result, we recalibrate
+waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
+we can see that this recalibration process gives further
 improvement to the top-left region while the bottom-right region shows similar
-result. This indicates that the model's ansatz does not fit waveforms in this region. This phenomenon is due to the reduced spin approximation used
-in parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$
-are introduced.
+result. This indicates that the ansatz does not fit waveforms in this region. 
+This phenomenon is due to the reduced spin approximation used in parameterizing the 
+ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
 

### 2023-04-21 src/scripts/all_quadrants.py

@@ -7,10 +7,10 @@ params = {
     "font.size": 18,
     "legend.fontsize": 12,
     "legend.frameon": True,
-    "axes.labelsize": 16,
-    "axes.titlesize": 16,
-    "xtick.labelsize": 14,
-    "ytick.labelsize": 14,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
     "axes.unicode_minus": False,
     "figure.figsize": (7, 5),
     "xtick.top": True,
@@ -38,18 +38,22 @@ q148 = np.loadtxt(paths.data / "all_quadrants.dat")
 data2 = np.array([i for i in q148 if (i[2] < 0 and i[3] > 0)])
 data4 = np.array([i for i in q148 if (i[2] > 0 and i[3] < 0)])
 
-bins = np.arange(-5, -1, 0.25)
+bins = np.arange(-5, -1, 0.2)
 
 fig, ax = plt.subplots(2, figsize=(6, 10))
 
-ax[0].hist(np.log10(data2[:,4]), bins=bins, label="top-left original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
-ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=1.3, alpha=0.8)
+ax[0].hist(np.log10(data2[:,4]), bins=bins, alpha=0.3, color="C0")
+ax[0].hist(np.log10(data2[:,4]), bins=bins, label="top-left original", histtype="step", color="C0", linewidth=1.3)
+ax[0].hist(np.log10(data2[:,5]), bins=bins, alpha=0.3, color="C1")
+ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=1.3)
 ax[0].set_ylabel('Count')
 ax[0].legend(loc='upper left')
 
+ax[1].hist(np.log10(data4[:,4]), bins=bins, alpha=0.3, color="C0")
 ax[1].hist(np.log10(data4[:,4]), bins=bins, label="bottom-right original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
+ax[1].hist(np.log10(data2[:,5]), bins=bins, alpha=0.3, color="C1")
[…]

### 2023-04-21 src/scripts/ps_q148_qchi.py

@@ -41,7 +41,7 @@ cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,0], data[:,1], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
 cbar = plt.colorbar(sc)
-cbar.ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
 cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
 plt.xlabel('$q$')
 plt.ylabel('$\chi_{\mathrm{PN}}$')

### 2023-04-21 src/scripts/ps_q148_quadrants.py

@@ -41,7 +41,7 @@ cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
 cbar = plt.colorbar(sc)
-cbar.ax.set_ylabel('$\log(MM_{\mathrm{opt}}/MM_{\mathrm{ori}})$', rotation=270, labelpad=20)
+cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
 cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')

### 2023-04-21 src/scripts/q148.py

@@ -39,12 +39,17 @@ q148_fl = np.loadtxt(paths.data / "q148_fl.dat")
 bins = np.arange(-5, -1, 0.2)
 
 plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
-plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0")
+plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0", linewidth=1.3)
+
 plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.05, color="C1")
-plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1")
+plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=1.3)
+plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", ls="--")
+
 plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C3")
-plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C3")
+plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C3", linewidth=1.3)
+plt.axvline(x=np.log10(np.median(q148_fl[:,5])), color="C2", ls="--")
+
 plt.legend()
 plt.ylabel('Count')
-plt.xlabel('$\log(\mathrm{mismatch})$')
+plt.xlabel('$\log(\mathcal{M})$')
 plt.savefig(paths.figures / "q148.pdf", bbox_inches="tight")

### 2023-04-21 src/scripts/q148_q1248_compare.py

@@ -3,6 +3,37 @@ import matplotlib.pyplot as plt
 import matplotlib as mpl
 import paths
 
+params = {
+    "font.size": 18,
+    "legend.fontsize": 12,
+    "legend.frameon": True,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
+    "axes.unicode_minus": False,
+    "figure.figsize": (7, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
 q148_data = np.loadtxt(paths.data / "q148.dat")
 q1248_data = np.loadtxt(paths.data / "q1248.dat")
 
@@ -11,11 +42,11 @@ q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
 q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
 
 plt.hist(q148_logdiff, bins=bins, alpha=0.3, color="C0")
-plt.hist(q148_logdiff, bins=bins, label="q148", histtype="step", color="C0")
+plt.hist(q148_logdiff, bins=bins, label="q148", histtype="step", color="C0", linewidth=1.3)
 plt.hist(q1248_logdiff, bins=bins, alpha=0.3, color="C1")
-plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1")
+plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1", linewidth=1.3)
 
-plt.legend(prop={"family":"Courier New"})
+plt.legend()
 plt.ylabel('Count')
-plt.xlabel('$\log(\mathrm{mismatch})$')
[…]

### 2023-04-25 src/tex/ms.tex
fix figures again

@@
 
-% Abstract with filler text
 \begin{abstract}
@@
 
-% Main body with filler text
 \section{Introduction} \label{sec:intro}
@@
 	mismatches are weighted with the constant PSD. For the
-	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median decreased by 50.0\%
+	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median (dashed lines) decreased by 50.0\%
 	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
@@
 	$\mathcal{L}_{\mathrm{mean}}$ for the loss function.
-We see waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened.
+Waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened, as indicated by the median (dashed lines).
 \kl{I tried to overlap these two histograms together but it makes them very hard to read.
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{\kw{The color scheme makes the plot a bit hard to read.}Parameter
-	space of testing waveforms. Each region is fitted independently. Colorbar
+	\caption{Parameter space of testing waveforms. Each region is fitted independently. Colorbar
 	represents log difference of mismatches before and after optimization.}

### 2023-04-25 src/scripts/all_quadrants.py

@@ -44,16 +44,20 @@ fig, ax = plt.subplots(2, figsize=(6, 10))
 
 ax[0].hist(np.log10(data2[:,4]), bins=bins, alpha=0.3, color="C0")
 ax[0].hist(np.log10(data2[:,4]), bins=bins, label="top-left original", histtype="step", color="C0", linewidth=1.3)
+ax[0].axvline(x=np.log10(np.median(data2[:,4])), color="C0", ls="--")
 ax[0].hist(np.log10(data2[:,5]), bins=bins, alpha=0.3, color="C1")
 ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=1.3)
+ax[0].axvline(x=np.log10(np.median(data2[:,5])), color="C1", ls="--")
 ax[0].set_ylabel('Count')
 ax[0].legend(loc='upper left')
 
 ax[1].hist(np.log10(data4[:,4]), bins=bins, alpha=0.3, color="C0")
 ax[1].hist(np.log10(data4[:,4]), bins=bins, label="bottom-right original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
-ax[1].hist(np.log10(data2[:,5]), bins=bins, alpha=0.3, color="C1")
+ax[1].axvline(x=np.log10(np.median(data4[:,4])), color="C0", ls="--")
+ax[1].hist(np.log10(data4[:,5]), bins=bins, alpha=0.3, color="C1")
 ax[1].hist(np.log10(data4[:,5]), bins=bins, label="bottom-right optimized", histtype="step", color="C1", linewidth=1.3, alpha=0.8)
-ax[1].set_xlabel('log(\mathcal{M})')
+ax[1].axvline(x=np.log10(np.median(data4[:,5])), color="C1", ls="--")
+ax[1].set_xlabel('$\log(\mathcal{M})$')
 ax[1].set_ylabel('Count')
 ax[1].legend(loc='upper left')
 

### 2023-04-25 src/scripts/loss.py

@@ -37,6 +37,6 @@ mpl.rcParams.update(params)
 data = np.loadtxt(paths.data / "q148_loss.dat")
 
 plt.plot(data[:,0], data[:,1])
-plt.ylabel('Loss')
-plt.xlabel('Epoch')
+plt.ylabel('$\mathcal{L}_{\mathrm{mean}}$')
+plt.xlabel('N')
 plt.savefig(paths.figures / "loss.pdf", bbox_inches="tight")

### 2023-04-25 src/scripts/q148.py

@@ -40,12 +40,13 @@ bins = np.arange(-5, -1, 0.2)
 
 plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
 plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0", linewidth=1.3)
+plt.axvline(x=np.log10(np.median(q148[:,4])), color="C0", ls="--")
 
 plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.05, color="C1")
 plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=1.3)
 plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", ls="--")
 
-plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C3")
+plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C2")
 plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C3", linewidth=1.3)
 plt.axvline(x=np.log10(np.median(q148_fl[:,5])), color="C2", ls="--")
 

### 2023-05-01 src/tex/ms.tex

@@
 in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
+
 improvement with a 22.9\% decrease in the median. However, the distribution lacks
@@
 with varying intrinsic parameters, it is conceivable that some trade-offs in
+
 accuracy exist between different parts of the parameter space. If this is the
@@
 Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
+
 more consistent improvement, since the ansatz are developed base on non-spinning 
@@
 	\caption{Parameter space of testing waveforms. Each region is fitted independently. Colorbar
+
 	represents log difference of mismatches before and after optimization.}
@@
 region than the bottom right along a degeneracy line. In an attempt to address
+
 this issue, we partitioned the parameter space into four regions, as described
@@
 exhibits similar mismatches as before, while the top-left region's performance
+
 has improved. This observation suggests that the ansatz is specific to certain
@@
 The division of the parameter space into four regions was a simple approach
+
 taken for practical reasons. A more systematic approach would involve the use of
@@
 to recalibratethe IMRPhenomX model \citep{pratten2020setting}. Because it is
+
 parameterized by an additional anti-symmetric spin parameter, it is expected not

### 2023-05-03 src/tex/ms.tex
Add Acknowlegment

@@
 Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
-
 more consistent improvement, since the ansatz are developed base on non-spinning 
@@
 	\caption{Parameter space of testing waveforms. Each region is fitted independently. Colorbar
-
 	represents log difference of mismatches before and after optimization.}
@@
 \section{Discussion} \label{sec:discussion}
-
-%\begin{enumerate} \item Comparison with the original paper is not the same.
-%   (Different waveforms, different ways to modify the waveform, different range
-%   of frequencies) \item More waveforms are taken so it should be more robust,
-%   since the parameter space coverage is better. \item Problems with IMRPhenomD
-%   model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space
-%   $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat) \item
-%   Can in principle split parameter space into regions that can be separately
-%   optimized. If not, then its must be the model's problem. Would the region
-%   classification be made automatic. Are there any algorithms to do so? \item
-%   Future work (Repeat on other GW models. Mapping between NR surrogate to see
-%   any systematics.) \end{enumerate}
 
@@
 
+We thank Will M.~Farr, Max Isi, and Mark Hannam for helpful discussions; we also
+thank Carl-Johan Haster, Neil J.~Cornish and Thomas Dent for comments on the
+draft. The Flatiron Institute is a division of the Simons Foundation. T.E.\ is
+supported by the Horizon Postdoctoral Fellowship.
 \bibliography{bib}

### 2023-05-03 src/tex/ms.tex
Working through result

--- 2023-04-25 src/tex/ms.tex
@@
 
-To evaluate the effectiveness of our optimization approach, we analyze
-approximately 530 NR waveforms from the SXS catalog in this study. We
-specifically select waveforms with negligible eccentricity
+To evaluate the effectiveness of the optimization, we analyze approximately 530
+NR waveforms \kw{Just say how many you have} from the SXS catalog in this study.
+We specifically select waveforms with negligible eccentricity
 (${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
@@
 
-We begin by investigating the impact of joint optimization on an individual
-waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
-outperforms the original waveform, particularly in the inspiral region, where
-the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
-selected one of the testing waveforms from the catalog presented in
-\citep{khan2016frequency}.
+To illustrate the improvement on an individual waveform level, we compare a
+particular waveform before and after optimization in Fig. \ref{fig:0154}.
+Compared to the original IMRPhenomD waveform, we can see the optimized waveform
+has smaller residual from NR waveform both in amplitude and phase, particularly
+in the inspiral region, where the amplitude displays a $50\%$ reduction in
+error. For a fair comparison, we selected one of the testing waveforms from the
+catalog presented in \citep{khan2016frequency}.
 
-We use the constant PSD weighted loss function, $\mathcal{L}_{\mathrm{mean}}$, to evaluate the
-mismatch of all testing waveforms, and present the resulting distribution in
-Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
+To quantify the effect of optimization over the entire dataset, We evaluate the
+mismatch of all testing waveforms using a constant PSD weighted loss function,
+$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
+\ref{fig:q148}. The peak of the distribution has shifted towards a lower
 mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
 in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
-improvement with a 22.9\% decrease in the median. However, the distribution lacks
-a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
-Despite this, both distributions exhibit significant improvement, and the
-optimization scheme remains unaffected.
+improvement with a 22.9\% decrease in the median. Even though the distribution
+lacks a clear peak due to the problem discussed in Section
+\ref{subsec:optimization}, both distributions show observable improvement, and
+the optimization scheme remains unaffected \kw{What about the optimization scheme?}.
 
@@
 
+\kw{Which plot is this paragraph referring to?}
 Applying the same methods, we observe that the distributions of mismatches
-calculated using the {\zdethp} PSD exhibit superior improvement
-compared to the weighted mismatch. The shape of the distribution is similar to
-that shown in Fig. \ref{fig:q148}. This result is expected, as the IMRPhenomD
-model was initially constructed and fitted using the {\zdethp} weighted
-mismatch. Consequently, the model is anticipated to closely fit the NR waveforms
-while incorporating the influence of the {\zdethp} PSD, rather than a
-constant PSD.
+calculated using the {\zdethp} PSD exhibit superior improvement compared to the
+weighted mismatch \kw{What do you mean by weighted mismatch? Like flat PSD?}.
+The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
+This result is expected, as the IMRPhenomD model was initially constructed and
+fitted using the {\zdethp} weighted mismatch. Consequently, the model is
+anticipated to closely fit the NR waveforms while incorporating the influence of
+the {\zdethp} PSD, rather than a constant PSD.
 
 Motivated by the successful improvement of the waveforms, we expand the training
-dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. The new
-set of coefficients generated from this optimization process yields only
-marginal improvements in the newly produced waveforms, as seen in Fig.
-\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
-distribution remains comparable in length and endpoint to the original
-distribution, indicating that our procedure is incapable of improving these
-waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
-function with additional waveforms leads to minimal improvement in the resulting
+dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248} \kw{What
+special about the q=2 events?}. The new set of coefficients generated from this
+optimization process yields only marginal improvements in the newly produced
+waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
+of the optimized distribution remains comparable in length and endpoint to the
+original distribution, indicating that our procedure is incapable of improving
+these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
+function with additional waveforms leads to similar improvement in the resulting
 distribution.
@@
 
-%\begin{enumerate} \item Comparison with the original paper is not the same.
-%   (Different waveforms, different ways to modify the waveform, different range
-%   of frequencies) \item More waveforms are taken so it should be more robust,
-%   since the parameter space coverage is better. \item Problems with IMRPhenomD
-%   model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space
-%   $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat) \item
-%   Can in principle split parameter space into regions that can be separately
-%   optimized. If not, then its must be the model's problem. Would the region
-%   classification be made automatic. Are there any algorithms to do so? \item
-%   Future work (Repeat on other GW models. Mapping between NR surrogate to see
-%   any systematics.) \end{enumerate}
-
 We have shown the result of recalibrating waveform coefficients. One thing to

### 2023-05-03 src/tex/ms.tex
Update

@@
 % regions and fitting separate sets of coefficients should give better results.
-
-Given that the waveform model's ansatz may not be entirely compatible with NR,
-and the optimization procedure is carried out over a distribution of waveforms
-with varying intrinsic parameters, it is conceivable that some trade-offs in
-accuracy exist between different parts of the parameter space. If this is the
-cause of the high mismatch tail's failure to improve during joint-optimization,
+Although most of the waveforms show improvement in figure \ref{fig:q148}, the
+high mismatch tail of the distribution remains unaffected. Given that the
+waveform model's ansatz may not be entirely compatible with NR, and the
+optimization procedure is carried out over a distribution of waveforms with
+varying intrinsic parameters, it is conceivable that some trade-offs in accuracy
+exist between different parts of the parameter space. If this is the cause of
+the lack of improvement in the high mismatch tail of the distribution,
 segmenting the parameter space into smaller subspaces should alleviate this

### 2023-05-03 src/scripts/q148.py
plot update

@@ -38,16 +38,16 @@ q148 = np.loadtxt(paths.data / "q148.dat")
 q148_fl = np.loadtxt(paths.data / "q148_fl.dat")
 bins = np.arange(-5, -1, 0.2)
 
-plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.3, color="C0")
+plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.1, color="C0")
 plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0", linewidth=1.3)
 plt.axvline(x=np.log10(np.median(q148[:,4])), color="C0", ls="--")
 
 plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.05, color="C1")
-plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=1.3)
+plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=3)
 plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", ls="--")
 
 plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C2")
-plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C3", linewidth=1.3)
+plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C2", linewidth=3)
 plt.axvline(x=np.log10(np.median(q148_fl[:,5])), color="C2", ls="--")
 
 plt.legend()

### 2023-05-03 src/tex/ms.tex

--- 2023-05-03 src/tex/ms.tex
@@
 
+%\begin{enumerate} \item Comparison with the original paper is not the same.
+%   (Different waveforms, different ways to modify the waveform, different range
+%   of frequencies) \item More waveforms are taken so it should be more robust,
+%   since the parameter space coverage is better. \item Problems with IMRPhenomD
+%   model, i.e. $\chi_{\mathrm{eff}}$ reduces the dimension of spin space
+%   $\chi_1, \chi_2$, so the model is flawed. Comparison issues. (Caveat) \item
+%   Can in principle split parameter space into regions that can be separately
+%   optimized. If not, then its must be the model's problem. Would the region
+%   classification be made automatic. Are there any algorithms to do so? \item
+%   Future work (Repeat on other GW models. Mapping between NR surrogate to see
+%   any systematics.) \end{enumerate}
+
 We have shown the result of recalibrating waveform coefficients. One thing to
@@
 
-We thank Will M.~Farr, Max Isi, and Mark Hannam for helpful discussions; we also
-thank Carl-Johan Haster, Neil J.~Cornish and Thomas Dent for comments on the
-draft. The Flatiron Institute is a division of the Simons Foundation. T.E.\ is
-supported by the Horizon Postdoctoral Fellowship.
 \bibliography{bib}

### 2023-05-11 src/tex/ms.tex
update

--- 2023-05-03 src/tex/ms.tex
@@
 into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
-\kw{Check if the following equation is correct. I think it is off by some smoothing factors.} 
-\kl{I don't remember there is any smoothing factors.
-They are just connected using step functions, then the entire waveform is made to have continuous first derivative by fixing some of the coefficients for the two connections.}
 \begin{align}\label{eq:joint_waveform}
@@
 \end{align}
-
 Instead of fitting the strain, which is a highly oscillatory function that is
@@
 $S_n(f)$ is the power spectral density(PSD) of noise from the instrument, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the relevant maximum and minimum frequencies for the integration. The mismatch
+the relevant maximum and minimum frequencies for the integration. $t_0$ and $\phi_0$ are maximized by performing regression between $h_1$ and $h_2$. The mismatch
 is a quantity that is closely related to the mean square error (MSE) between the
 two waveforms, as it is often considered a weighted version of the MSE.
-
-\kl{The maximum is calculated by comparing the phase of IMRPhenomD and NR waveforms. We know that they are off by a linear relation of $f$, 
-\begin{align*}
-	\phi_{NR}-\phi_{IMR}=(2\pi t_0)f+\phi_0.
-\end{align*}
-Since I have $\phi_{NR}$, $\phi_{IMR}$ and the frequency array, I did a linear regression to get $2\pi t_0$ and $\phi_0$.}
 
@@
 in Fig.~\ref{fig:q148}, which we can see in Fig.~\ref{fig:all_quadrants}. Thus, we show that 
-further optimization in a smaller subspace does not improve the results.
+further optimization in a smaller subspace does not improve the results. \kw{I am a bit confused about why you can get worse result.}
 

### 2023-05-11 src/scripts/all_quadrants.py

@@ -38,37 +38,24 @@ q148 = np.loadtxt(paths.data / "all_quadrants.dat")
 data2 = np.array([i for i in q148 if (i[2] < 0 and i[3] > 0)])
 data4 = np.array([i for i in q148 if (i[2] > 0 and i[3] < 0)])
 
-bins = np.arange(-5, -1, 0.2)
+bins = np.linspace(-5, -1, 15)
 
-fig, ax = plt.subplots(2, figsize=(6, 10))
+fig, ax = plt.subplots(2, figsize=(6, 10), sharex=True)
+plt.subplots_adjust(hspace=0.05)
 
 ax[0].hist(np.log10(data2[:,4]), bins=bins, alpha=0.3, color="C0")
-ax[0].hist(np.log10(data2[:,4]), bins=bins, label="top-left original", histtype="step", color="C0", linewidth=1.3)
-ax[0].axvline(x=np.log10(np.median(data2[:,4])), color="C0", ls="--")
-ax[0].hist(np.log10(data2[:,5]), bins=bins, alpha=0.3, color="C1")
-ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=1.3)
-ax[0].axvline(x=np.log10(np.median(data2[:,5])), color="C1", ls="--")
+ax[0].axvline(x=np.log10(np.median(data2[:,4])), color="C0", ls="--", alpha=0.4)
+ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=3)
+ax[0].axvline(x=np.log10(np.median(data2[:,5])), color="C1", ls="--", alpha=0.4)
 ax[0].set_ylabel('Count')
 ax[0].legend(loc='upper left')
 
 ax[1].hist(np.log10(data4[:,4]), bins=bins, alpha=0.3, color="C0")
-ax[1].hist(np.log10(data4[:,4]), bins=bins, label="bottom-right original", histtype="step", color="C0", linewidth=1.3, alpha=0.8)
-ax[1].axvline(x=np.log10(np.median(data4[:,4])), color="C0", ls="--")
-ax[1].hist(np.log10(data4[:,5]), bins=bins, alpha=0.3, color="C1")
-ax[1].hist(np.log10(data4[:,5]), bins=bins, label="bottom-right optimized", histtype="step", color="C1", linewidth=1.3, alpha=0.8)
[…]

### 2023-05-11 src/scripts/loss.py

@@ -34,9 +34,17 @@ params = {
 
 mpl.rcParams.update(params)
 
-data = np.loadtxt(paths.data / "q148_loss.dat")
+train_data = np.loadtxt(paths.data / "train_waveform_loss.txt")
+validation_data = np.loadtxt(paths.data / "validation_waveform_loss.txt")
 
-plt.plot(data[:,0], data[:,1])
-plt.ylabel('$\mathcal{L}_{\mathrm{mean}}$')
-plt.xlabel('N')
+fig, ax1 = plt.subplots()
+
+ax2 = ax1.twinx()
+ax1.plot(validation_data[:,0], validation_data[:,1], color="C1", linewidth=2, label='validation')
+ax2.plot(train_data[:,0], train_data[:,1], color="C0", linewidth=2, label='train')
+
+ax1.set_xlabel('N')
+ax1.set_ylabel('validation loss (dashed)')
+ax2.set_ylabel('test loss')
+
 plt.savefig(paths.figures / "loss.pdf", bbox_inches="tight")

### 2023-05-11 src/scripts/q148.py

@@ -35,20 +35,13 @@ params = {
 mpl.rcParams.update(params)
 
 q148 = np.loadtxt(paths.data / "q148.dat")
-q148_fl = np.loadtxt(paths.data / "q148_fl.dat")
-bins = np.arange(-5, -1, 0.2)
+bins = np.linspace(-5, -1, 15)
 
-plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.1, color="C0")
-plt.hist(np.log10(q148[:,4]), bins=bins, label="Original", histtype="step", color="C0", linewidth=1.3)
-plt.axvline(x=np.log10(np.median(q148[:,4])), color="C0", ls="--")
+plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.2, color="C0", label="Original")
+plt.axvline(x=np.log10(np.median(q148[:,4])), color="C0", zorder=-20, alpha=0.4, ls=":")
 
-plt.hist(np.log10(q148[:,5]), bins=bins, alpha=0.05, color="C1")
 plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=3)
-plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", ls="--")
-
-plt.hist(np.log10(q148_fl[:,5]), bins=bins, alpha=0.05, color="C2")
-plt.hist(np.log10(q148_fl[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{fl}}$", histtype="step", color="C2", linewidth=3)
-plt.axvline(x=np.log10(np.median(q148_fl[:,5])), color="C2", ls="--")
+plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", zorder=-20, alpha=0.4, ls=":")
 
 plt.legend()
 plt.ylabel('Count')

### 2023-05-11 src/scripts/q148_q1248_compare.py

@@ -42,9 +42,7 @@ q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
 q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
 
 plt.hist(q148_logdiff, bins=bins, alpha=0.3, color="C0")
-plt.hist(q148_logdiff, bins=bins, label="q148", histtype="step", color="C0", linewidth=1.3)
-plt.hist(q1248_logdiff, bins=bins, alpha=0.3, color="C1")
-plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1", linewidth=1.3)
+plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1", linewidth=3)
 
 plt.legend()
 plt.ylabel('Count')

### 2023-05-15 src/tex/ms.tex
update figures

@@
 
-%\definecolor{rb4}{HTML}{27408B}
-%\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
-%\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
 \newcommand{\ripple}{\texttt{ripple}}
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{The value of loss function against number of iterations. The code is terminated after 30000 steps }
+	\caption{Loss functions against number of iterations. We take the set of coefficients at the minimum of the validation loss.}
 	\label{fig:loss}
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
+	\caption{Parameter space with mass ratio $q$ against post-newtonian spin parameter $\chi_{\mathrm{PN}}$. Orange: Training waveforms; Blue: Testing waveforms}
 	\label{fig:intrin_space}
@@
 lacks a clear peak due to the problem discussed in Section
-\ref{subsec:optimization}, both distributions show observable improvement, and
-the optimization scheme remains unaffected \kw{What about the optimization scheme?}.
+\ref{subsec:optimization}, both distributions show observable improvement.
 
@@
 	\includegraphics[width=\columnwidth]{figures/q148.pdf}
-	\caption{Distributions of waveform mismatches calculated using both
-	$\mathcal{L}_{\mathrm{mean}}$ and $\mathcal{L}_{\mathrm{fl}}$. We use 
-	training waveforms listed in Tab.~\ref{tab:q148} and
-	mismatches are weighted with the constant PSD. For the
-	$\mathcal{L}_{\mathrm{mean}}$ distributions, the median (dashed lines) decreased by 50.0\%
-	while the median of $\mathcal{L}_{\mathrm{fl}}$ distribution decreased by
-	22.9\%.}
+	\caption{Distributions of waveform mismatches calculated using 
+	$\mathcal{L}_{\mathrm{mean}}$. We use training waveforms listed in 
+	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
+	Dashed lines represent the median of the distributions, which has decreased by 45.3\%.}
 	\label{fig:q148}
@@
 	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$.}
+	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$. 
+	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
 	\label{fig:q148_q1248_compare}
@@
 
-\kw{Which plot is this paragraph referring to?}
-Applying the same methods, we observe that the distributions of mismatches
+%\kw{Which plot is this paragraph referring to?}
+Applying the same methods, we will see that the distributions of mismatches
 calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
-weighted mismatch \kw{What do you mean by weighted mismatch? Like flat PSD?}.
+unweighted mismatch.
 The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
@@
 Motivated by the successful improvement of the waveforms, we expand the training
-dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248} \kw{What
-special about the q=2 events?}. The new set of coefficients generated from this
+dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. We 
+specifically choose to use $q=2$ events since we have abundant $q=2$ NR waveforms 
+to validate the final result. The new set of coefficients generated from this
 optimization process yields only marginal improvements in the newly produced
@@
 $\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
-Meanwhile, the top-left ($\chi_1<0$ and $\chi_2>0$) and bottom-right ($\chi_1>0$ 
-and $\chi_2<0$) regions respond to optimization differently. In the top-left region, waveforms
-generally improve with along optimization. However, mismatches in the bottom-right 
-region do not improve after optimization. Most waveforms even turned worse after 
-optimization.
+Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right ($\chi_1>0>\chi_2$) 
+regions respond to optimization differently. In the top-left 
+region, waveforms generally improve with along optimization. However, mismatches 
+in the bottom-right region do not improve after optimization. Most waveforms even 
+turned worse after optimization.
 
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
-	\caption{Parameter space of testing waveforms of
-	$q$~vs.~$\chi_{\mathrm{PN}}$. We use the recalibrated result from
-	$\mathcal{L}_{\mathrm{mean}}$ with the constant noise spectrum and training
-	waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the
-	$\log_{10}$ difference between optimized and original unweighted
-	mismatches.}
+	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
+	We show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with constant PSD 
+	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
+	$\log_{10}$ difference between optimized and original mismatches.}
 	\label{fig:ps_q148_qchi}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
-	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. We
-	use the recalibrated result from $\mathcal{L}_{\mathrm{mean}}$ with the
-	constant noise spectrum and training waveforms in Tab.~\ref{tab:q148}.}
+	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
+	show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with the
+	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
 	\label{fig:ps_q148}
@@
 		SXS:BBH:1426 & 8.0 & 0.48     & 0.75     \\
-		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule SXS:BBH:0370 & 1.0
-		& -0.20    & 0.40     \\
+		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule 
+		SXS:BBH:0370 & 1.0 & -0.20    & 0.40     \\
 		SXS:BBH:2092 & 1.0 & -0.50    & 0.50     \\
@@
 		SXS:BBH:2133 & 3.0 & -0.73    & 0.85     \\
-		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule SXS:BBH:0156 & 1.0
-		& -0.95    & -0.95    \\
+		SXS:BBH:0263 & 4.0 & -0.80    & 0.80     \\ \midrule 
+		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
 		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
@@
 		SXS:BBH:1419 & 8.0 & -0.80    & -0.80    \\
-		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule SXS:BBH:0304 & 1.0
-		& 0.50     & -0.50    \\
+		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\ \midrule 
+		SXS:BBH:0304 & 1.0 & 0.50     & -0.50    \\
 		SXS:BBH:0327 & 1.0 & 0.80     & -0.80    \\
@@
 From top to down are the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$), bottom-left ($\chi_1,\chi_2<0$) and bottom-right ($\chi_1>0>\chi_2$) regions.
-Note that for the top-right and bottom-left regions, waveforms are chosen to have equal or similar spins, while the training waveforms for the other two regions are chosen to have opposite spins.}
+Note that for the top-right and bottom-left regions, waveforms are chosen to have $\chi_1\approx\chi_2$, while the training waveforms for the other two regions are chosen to have $\chi_1\approx-\chi_2$.}
 	\label{tab:quadrants}
@@
 	We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ for the loss function.
-Waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened, as indicated by the median (dashed lines).
-\kl{I tried to overlap these two histograms together but it makes them very hard to read.
-}}
+	$\mathcal{L}_{\mathrm{mean}}$ as the loss function.
+Waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened, as indicated by the median (dashed lines).}
 	\label{fig:all_quadrants}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Parameter space of testing waveforms. Each region is fitted independently. Colorbar
-	represents log difference of mismatches before and after optimization.}
+	\caption{Parameter space of testing waveforms.
+Each region is fitted independently with waveforms listed in Tab.~\ref{tab:quadrants}.
+Colorbar represents the $\log_{10}$ difference of mismatches before and after optimization.}
 	\label{fig:ps_q148_quadrant}
@@
 seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
-$\mathcal{L}_{fl}$. By increasing the number of training waveforms, we see a 
+$\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
 slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.

### 2023-05-15 src/scripts/all_quadrants.py

@@ -43,16 +43,16 @@ bins = np.linspace(-5, -1, 15)
 fig, ax = plt.subplots(2, figsize=(6, 10), sharex=True)
 plt.subplots_adjust(hspace=0.05)
 
-ax[0].hist(np.log10(data2[:,4]), bins=bins, alpha=0.3, color="C0")
+ax[0].hist(np.log10(data2[:,4]), bins=bins, label='Original', alpha=0.3, color="C0")
 ax[0].axvline(x=np.log10(np.median(data2[:,4])), color="C0", ls="--", alpha=0.4)
-ax[0].hist(np.log10(data2[:,5]), bins=bins, label="top-left optimized", histtype="step", color="C1", linewidth=3)
+ax[0].hist(np.log10(data2[:,5]), bins=bins, label="Optimized", histtype="step", color="C1", linewidth=3)
 ax[0].axvline(x=np.log10(np.median(data2[:,5])), color="C1", ls="--", alpha=0.4)
 ax[0].set_ylabel('Count')
 ax[0].legend(loc='upper left')
 
-ax[1].hist(np.log10(data4[:,4]), bins=bins, alpha=0.3, color="C0")
+ax[1].hist(np.log10(data4[:,4]), bins=bins, label='Original', alpha=0.3, color="C0")
 ax[1].axvline(x=np.log10(np.median(data4[:,4])), color="C0", ls="--", alpha=0.4)
-ax[1].hist(np.log10(data4[:,5]), bins=bins, label="bottom-right optimized", histtype="step", color="C1", linewidth=3)
+ax[1].hist(np.log10(data4[:,5]), bins=bins, label="Optimized", histtype="step", color="C1", linewidth=3)
 ax[1].axvline(x=np.log10(np.median(data4[:,5])), color="C1", ls="--", alpha=0.4)
 ax[1].set_xlabel('$\log(\mathcal{M})$')
 ax[1].set_ylabel('Count')

### 2023-05-15 src/scripts/loss.py

@@ -41,10 +41,11 @@ fig, ax1 = plt.subplots()
 
 ax2 = ax1.twinx()
 ax1.plot(validation_data[:,0], validation_data[:,1], color="C1", linewidth=2, label='validation')
-ax2.plot(train_data[:,0], train_data[:,1], color="C0", linewidth=2, label='train')
+ax2.plot(train_data[:,0], train_data[:,1], color="C0", linewidth=2, label='training')
 
-ax1.set_xlabel('N')
-ax1.set_ylabel('validation loss (dashed)')
-ax2.set_ylabel('test loss')
+ax1.set_xlabel('Iteration')
+ax1.set_ylabel('Validation Loss')
+ax2.set_ylabel('Training Loss')
+fig.legend(loc='upper right', bbox_to_anchor=(0.89, 0.875))
 
 plt.savefig(paths.figures / "loss.pdf", bbox_inches="tight")

### 2023-05-15 src/scripts/ps_q148_chi1chi2.py

@@ -40,9 +40,9 @@ logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-cbar = plt.colorbar(sc)
+cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
 cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
-cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
+cbar.ax.set_yticklabels(['$<-1$', '$-0.5$', '$0$', '$0.5$', '$>1$'])
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 

### 2023-05-15 src/scripts/ps_q148_qchi.py

@@ -35,14 +35,17 @@ params = {
 mpl.rcParams.update(params)
 
 data = np.loadtxt(paths.data / "q148.dat")
+q = data[:,0]/data[:,1]
+eta = data[:,0]*data[:,1]/(data[:,0]+data[:,1])**2
+chi_PN = (data[:,0]*data[:,2]+data[:,1]*data[:,3])/(data[:,0]+data[:,1])-38*eta*(data[:,2]+data[:,3])/113
 
 logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
-sc = plt.scatter(data[:,0], data[:,1], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-cbar = plt.colorbar(sc)
+sc = plt.scatter(q, chi_PN, c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
 cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
-cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
+cbar.ax.set_yticklabels(['$<-1$', '$-0.5$', '$0$', '$0.5$', '$>1$'])
 plt.xlabel('$q$')
 plt.ylabel('$\chi_{\mathrm{PN}}$')
 

### 2023-05-15 src/scripts/ps_q148_quadrants.py

@@ -34,15 +34,15 @@ params = {
 
 mpl.rcParams.update(params)
 
-data = np.loadtxt(paths.data / "q148_quadrants.dat")
+data = np.loadtxt(paths.data / "all_quadrants.dat")
 
 logdiff = np.log10(data[:,5] / data[:,4])
 cm = plt.cm.get_cmap('RdYlBu')
 colors = logdiff
 sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
-cbar = plt.colorbar(sc)
+cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
 cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
-cbar.ax.set_yticklabels(['$< -1$', '$-0.75$', '$-0.5$', '$-0.25$', '$0$', '$0.25$', '$0.5$', '$0.75$', '$> 1$'])
+cbar.ax.set_yticklabels(['$<-1$', '$-0.5$', '$0$', '$0.5$', '$>1$'])
 plt.xlabel('$\chi_1$')
 plt.ylabel('$\chi_2$')
 

### 2023-05-15 src/scripts/q148_q1248_compare.py

@@ -41,8 +41,10 @@ bins = np.arange(-2, 2, 0.2)
 q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
 q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
 
-plt.hist(q148_logdiff, bins=bins, alpha=0.3, color="C0")
+plt.hist(q148_logdiff, bins=bins, label='q148', alpha=0.3, color="C0")
+plt.axvline(x=np.median(q148_logdiff), color="C0", zorder=-20, alpha=0.4, ls=":")
 plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1", linewidth=3)
+plt.axvline(x=np.median(q1248_logdiff), color="C1", zorder=-20, alpha=0.4, ls=":")
 
 plt.legend()
 plt.ylabel('Count')

### 2023-05-17 src/tex/ms.tex
intrin_space fix

@@
 regions respond to optimization differently. In the top-left 
-region, waveforms generally improve with along optimization. However, mismatches 
-in the bottom-right region do not improve after optimization. Most waveforms even 
+region, waveforms generally improve slightly with along optimization. However, waveforms 
+in the bottom-right region do not improve after optimization. Some waveforms even 
 turned worse after optimization.
@@
 		SXS:BBH:1426 & 8.0 & 0.48     & 0.75     \\
-		SXS:BBH:0167 & 8.0 & 0.00     & 0.00     \\ \midrule 
+		SXS:BBH:0063 & 8.0 & 0.00     & 0.00     \\ \midrule 
 		SXS:BBH:0370 & 1.0 & -0.20    & 0.40     \\
@@
 We divided the parameter space into four regions to analyze the effect of the
-recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148}). The
+recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). The
 training waveforms used for fitting are listed in Table~\ref{tab:quadrants}. The
 top-left and bottom-right regions have limited data for $q>4$, hence the results are
-only valid up to $q\leq4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe
-improvements in waveform mismatch for all waveforms, except those in the bottom-right 
-region. For equal-spin waveforms lying in the proximity of the diagonal axis,
-the mismatch has additional improvement due to most training waveforms being
-equal-spin. In the top-left region, the improvement is significant, except for a
+only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
+waveforms lying in the lower diagonal have mismatches worsened while other waveforms 
+generally improve.
+For the top-right and bottom-left regions, we see that the optimization prefers waveforms with $\chi_1<\chi_2$.
+This further  In the top-left region, the improvement is significant, except for a
 few defects caused by some testing waveforms having $q>4$. This can be seen clearer in 
-Fig.~\ref{fig:all_quadrants}, which we see that the optimized histogram shifts downward uniformly. 
-The optimized waveforms in the bottom-right region have a higher mismatch than the original ones,
-indicating poor fitting. These waveforms correspond to the waveforms in the high mismatch tail 
-in Fig.~\ref{fig:q148}, which we can see in Fig.~\ref{fig:all_quadrants}. Thus, we show that 
-further optimization in a smaller subspace does not improve the results. \kw{I am a bit confused about why you can get worse result.}
+Fig.~\ref{fig:all_quadrants}, which we see that the optimized histogram shifts downward 
+rather uniformly. The optimized waveforms in the bottom-right region have two distinct 
+features. Within this region, waveforms in the bottom-right corner improves significantly 
+while others worsened. This phenomenon can also be seen in Fig.~\ref{fig:all_quadrants}, 
+where the mismatch distribution shifted upward. Hence, we can deduce that the loss function 
+descended into a different minimum compared to Fig.~\ref{fig:ps_q148}, making the optimized 
+coefficients not useful in GW analysis. THus, further optimizing in a smaller region does 
+not generally improve the results.
+%\kw{I am a bit confused about why you can get worse result.}
 
@@
 optimization procedure should be able to improve the accuracy of IMRPhenomD on a
-similar scale regardless of the differences. Here, this result serves as a
+similar scale regardless of the differences. Here, the result serves as a
 demonstration of the general method used.  
@@
 space with negative slopes that depend on the mass ratio, as shown in
-Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top left
-region than the bottom right along a degeneracy line. In an attempt to address
+Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
+region than the bottom-right region. In an attempt to address
 this issue, we partitioned the parameter space into four regions, as described
-in Sec. \ref{sec:result}. However, even with separate optimizations for each
-regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the bottom-right region
-exhibits similar mismatches as before, while the top-left region's performance
-has improved. This observation suggests that the ansatz is specific to certain
-regions of the parameter space, with a preference for BBH events featuring
-$\chi_1<0$ and $\chi_2>0$.
+in Sec. \ref{sec:result}. With separate optimizations for each
+regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
+performance has improved, while the bottom-right region exhibits two distinct 
+features. This is because the loss function has descended into a different 
+minimum which highly favors events with $\chi_1\approx1$ and $\chi_2\approx-1$. 
+This observation suggests the ansatz is specific to certain regions of the 
+parameter space, with a preference for BBH events lying in the upper diagonal 
+region, and it hardly fits with waveforms lying in the lower diagonal region.
 
@@
 versatile approach can be applied to other differentiable GW models, such as the
-IMRPhenomP \citep{hannam2014simple} or IMRPhenomX \citep{pratten2020setting} 
+IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
 models within the same family. By jointly optimizing a new set of coefficients,
@@
 approximants as part of the ansatz in the inspiral segment. It will be interesting 
-to recalibratethe IMRPhenomX model \citep{pratten2020setting}. Because it is
+to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
 parameterized by an additional anti-symmetric spin parameter, it is expected not
-to exhibit the degeneracy previously described. A more detailed investigation
+to exhibit the degeneracy previously described. With the currently developing 
+{\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
 may provide valuable insights into the systematics of the Phenom models.
 Furthermore, this approach is applicable to other GW model families, such as NR
-surrogate models or EOB models introduced previously. Such an approach could 
+surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
 simplify NR waveform calibration procedures and lead to the improvement of 
@@
 certain preference for waveforms lying in the top-left region while the bottom-right 
-region cannot be improved. To further test this result, we recalibrate
+region hardly improved. To further test this result, we recalibrate
 waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
 we can see that this recalibration process gives further
-improvement to the top-left region while the bottom-right region shows similar
-result. This indicates that the ansatz does not fit waveforms in this region. 
-This phenomenon is due to the reduced spin approximation used in parameterizing the 
+improvement to the top-left region while the bottom-right region falls into a different 
+loss minimum. This indicates that the ansatz does not fit certain waveforms in this 
+region, hence introduces bias to downstream GW analysis. This phenomenon is mainly 
+due to the reduced spin approximation used in parameterizing the 
 ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.

### 2023-05-17 src/scripts/intrin_space.py

@@ -37,9 +37,13 @@ mpl.rcParams.update(params)
 train_data = np.loadtxt(paths.data / "ps_train.dat")
 test_data = np.loadtxt(paths.data / "q148.dat")
 
+q = test_data[:,0]/test_data[:,1]
+eta = test_data[:,0]*test_data[:,1]/(test_data[:,0]+test_data[:,1])**2
+chi_PN = (test_data[:,0]*test_data[:,2]+test_data[:,1]*test_data[:,3])/(test_data[:,0]+test_data[:,1])-38*eta*(test_data[:,2]+test_data[:,3])/113
+
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
-plt.scatter(test_data[:,0], test_data[:,1], s=30, label='testing')
+plt.scatter(q, chi_PN, s=30, label='testing')
 plt.scatter(train_data[:,0], train_data[:,1], s=100, label='training')
 plt.legend()
 plt.xlabel('$q$')

### 2023-05-18 src/scripts/0154.py
Updating and unifying plots

@@ -6,7 +6,7 @@ import paths
 params = {
     "font.size": 15,
     "legend.fontsize": 12,
-    "legend.frameon": True,
+    "legend.frameon": False,
     "axes.labelsize": 15,
     "axes.titlesize": 15,
     "xtick.labelsize": 15,
@@ -35,57 +35,96 @@ params = {
 MSUN = 1.988409902147041637325262574352366540e30  # kg
 G = 6.67430e-11  # m^3 / kg / s^2
 C = 299792458.0  # m / s
-gt = G * MSUN / (C ** 3.0)
+gt = G * MSUN / (C**3.0)
 
 f_ins = 0.018
-f_rd =  149.1018563859192 * (25. + 25.) * gt
+f_rd = 149.1018563859192 * (25.0 + 25.0) * gt
 
 mpl.rcParams.update(params)
 
 data = np.loadtxt(paths.data / "0154_compare.dat")
 
-f_uniform = data[:,0]
-NR_amp = data[:,1]
-NR_angle = data[:,2]
-IMR_amp = data[:,3]
-IMR_angle = data[:,4]
-IMR_opt_amp = data[:,5]
-IMR_opt_angle = data[:,6]
+f_uniform = data[:, 0]
+NR_amp = data[:, 1]
+NR_angle = data[:, 2]
+IMR_amp = data[:, 3]
+IMR_angle = data[:, 4]
+IMR_opt_amp = data[:, 5]
+IMR_opt_angle = data[:, 6]
 
-fig, ax = plt.subplots(2, 2, figsize=(15, 10))
+fig, ax = plt.subplots(2, 2, figsize=(12, 7), sharex=True)
+# green = "#406a3d"
+# red = "#b14a49"
+# yellow = "#b0882f"
+c1 = "#102F68"  # blue
+c2 = "#B02423"  # red
 
-ax[0, 0].loglog(f_uniform, NR_amp, label='NR')
-ax[0, 0].loglog(f_uniform, IMR_amp, label='IMR original')
-ax[0, 0].loglog(f_uniform, IMR_opt_amp, label='IMR optimized')
-ax[0, 0].axvline(x=f_ins, color="k", alpha=0.5, ls="--")
-ax[0, 0].axvline(x=f_rd, color="k", alpha=0.4, ls="--")
-ax[0, 0].set_xlabel(r'$Mf$')
[…]

### 2023-05-18 src/scripts/intrin_space.py

@@ -4,15 +4,15 @@ import matplotlib as mpl
 import paths
 
 params = {
-    "font.size": 18,
-    "legend.fontsize": 12,
+    "font.size": 17,
+    "legend.fontsize": 14,
     "legend.frameon": True,
-    "axes.labelsize": 15,
-    "axes.titlesize": 15,
-    "xtick.labelsize": 15,
-    "ytick.labelsize": 15,
+    "axes.labelsize": 17,
+    "axes.titlesize": 17,
+    "xtick.labelsize": 17,
+    "ytick.labelsize": 17,
     "axes.unicode_minus": False,
-    "figure.figsize": (7, 5),
+    "figure.figsize": (6, 5),
     "xtick.top": True,
     "ytick.right": True,
     "xtick.bottom": True,
@@ -33,15 +33,26 @@ params = {
 }
 
 mpl.rcParams.update(params)
+# blue = "#102F68"
+# red = "#B02423"
+c1 = "#648FFF"
+c2 = "#FE6100"
+# c2 = "#FFB000"
 
 train_data = np.loadtxt(paths.data / "ps_train.dat")
 test_data = np.loadtxt(paths.data / "q148.dat")
 
+q = test_data[:, 0] / test_data[:, 1]
+eta = test_data[:, 0] * test_data[:, 1] / (test_data[:, 0] + test_data[:, 1]) ** 2
+chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3]) / (
+    test_data[:, 0] + test_data[:, 1]
+) - 38 * eta * (test_data[:, 2] + test_data[:, 3]) / 113
+
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
-plt.scatter(test_data[:,0], test_data[:,1], s=30, label='testing')
-plt.scatter(train_data[:,0], train_data[:,1], s=100, label='training')
+plt.scatter(train_data[:, 0], train_data[:, 1], s=80, color=c1, label="Training")
+plt.scatter(q, chi_PN, s=40, color=c2, label="Validation", marker="*", zorder=-20)
 plt.legend()
-plt.xlabel('$q$')
[…]

### 2023-05-18 src/scripts/loss.py

@@ -4,13 +4,13 @@ import matplotlib as mpl
 import paths
 
 params = {
-    "font.size": 18,
-    "legend.fontsize": 12,
-    "legend.frameon": True,
-    "axes.labelsize": 16,
-    "axes.titlesize": 16,
-    "xtick.labelsize": 14,
-    "ytick.labelsize": 14,
+    "font.size": 17,
+    "legend.fontsize": 14,
+    "legend.frameon": False,
+    "axes.labelsize": 17,
+    "axes.titlesize": 17,
+    "xtick.labelsize": 17,
+    "ytick.labelsize": 17,
     "axes.unicode_minus": False,
     "figure.figsize": (7, 5),
     "xtick.top": True,
@@ -33,19 +33,37 @@ params = {
 }
 
 mpl.rcParams.update(params)
-
+# blue = "#102F68"
+# red = "#B02423"
+# c1 = "#4e7ab1"  # blue Nash
+# c1 = "#b14a49"  # red Nash
+# c2 = "#406a3d"  # green Nash
+c1 = "#648FFF"
+c2 = "#FE6100"
+# 16, 47, 104)
+# \definecolor{darkgreen}{rgb}{0.075,0.302,0.047}
+# % \definecolor{dullred}{rgb}{0.706,0.208,0.192}
+# \definecolor{dullred}{RGB}{176, 36, 35}
 train_data = np.loadtxt(paths.data / "train_waveform_loss.txt")
 validation_data = np.loadtxt(paths.data / "validation_waveform_loss.txt")
 
-fig, ax1 = plt.subplots()
+fig, ax1 = plt.subplots(figsize=(6, 5))
 
 ax2 = ax1.twinx()
-ax1.plot(validation_data[:,0], validation_data[:,1], color="C1", linewidth=2, label='validation')
-ax2.plot(train_data[:,0], train_data[:,1], color="C0", linewidth=2, label='training')
+ax1.plot(train_data[:, 0], train_data[:, 1], color=c1, linewidth=2, label="Training")
+ax2.plot(
+    validation_data[:, 0],
+    validation_data[:, 1],
+    color=c2,
+    linewidth=2,
+    label="Validation",
+    linestyle=(0, (3, 1)),
+)
+ax1.axvline(x=12000, alpha=0.5, color="k", linestyle=(0, (1, 1.05)))
 
-ax1.set_xlabel('Iteration')
-ax1.set_ylabel('Validation Loss')
[…]

### 2023-05-18 src/scripts/ps_q148_chi1chi2.py

@@ -6,11 +6,11 @@ import paths
 params = {
     "font.size": 18,
     "legend.fontsize": 12,
-    "legend.frameon": True,
-    "axes.labelsize": 15,
-    "axes.titlesize": 15,
-    "xtick.labelsize": 15,
-    "ytick.labelsize": 15,
+    "legend.frameon": False,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
     "axes.unicode_minus": False,
     "figure.figsize": (7, 5),
     "xtick.top": True,
@@ -36,14 +36,16 @@ mpl.rcParams.update(params)
 
 data = np.loadtxt(paths.data / "q148.dat")
 
-logdiff = np.log10(data[:,5] / data[:,4])
-cm = plt.cm.get_cmap('RdYlBu')
+logdiff = np.log10(data[:, 5] / data[:, 4])
+cm = plt.cm.get_cmap("RdYlBu")
 colors = logdiff
-sc = plt.scatter(data[:,2], data[:,3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+sc = plt.scatter(data[:, 2], data[:, 3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm)
 cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
-cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
-cbar.ax.set_yticklabels(['$<-1$', '$-0.5$', '$0$', '$0.5$', '$>1$'])
-plt.xlabel('$\chi_1$')
-plt.ylabel('$\chi_2$')
+cbar.ax.set_ylabel(
+    "$\log_{10}(\mathcal{M}_f/\mathcal{M}_i)$", rotation=270, labelpad=20
+)
+cbar.ax.set_yticklabels(["$<-1$", "$-0.5$", "$0$", "$0.5$", "$>1$"])
+plt.xlabel("$\chi_1$")
+plt.ylabel("$\chi_2$")
 
 plt.savefig(paths.figures / "ps_q148_chi1chi2.pdf", bbox_inches="tight")

### 2023-05-18 src/scripts/ps_q148_qchi.py

@@ -6,11 +6,11 @@ import paths
 params = {
     "font.size": 18,
     "legend.fontsize": 12,
-    "legend.frameon": True,
-    "axes.labelsize": 15,
-    "axes.titlesize": 15,
-    "xtick.labelsize": 15,
-    "ytick.labelsize": 15,
+    "legend.frameon": False,
+    "axes.labelsize": 18,
+    "axes.titlesize": 18,
+    "xtick.labelsize": 18,
+    "ytick.labelsize": 18,
     "axes.unicode_minus": False,
     "figure.figsize": (7, 5),
     "xtick.top": True,
@@ -35,18 +35,22 @@ params = {
 mpl.rcParams.update(params)
 
 data = np.loadtxt(paths.data / "q148.dat")
-q = data[:,0]/data[:,1]
-eta = data[:,0]*data[:,1]/(data[:,0]+data[:,1])**2
-chi_PN = (data[:,0]*data[:,2]+data[:,1]*data[:,3])/(data[:,0]+data[:,1])-38*eta*(data[:,2]+data[:,3])/113
+q = data[:, 0] / data[:, 1]
+eta = data[:, 0] * data[:, 1] / (data[:, 0] + data[:, 1]) ** 2
+chi_PN = (data[:, 0] * data[:, 2] + data[:, 1] * data[:, 3]) / (
+    data[:, 0] + data[:, 1]
+) - 38 * eta * (data[:, 2] + data[:, 3]) / 113
 
-logdiff = np.log10(data[:,5] / data[:,4])
-cm = plt.cm.get_cmap('RdYlBu')
+logdiff = np.log10(data[:, 5] / data[:, 4])
+cm = plt.cm.get_cmap("RdYlBu")
 colors = logdiff
-sc = plt.scatter(q, chi_PN, c=colors, vmin=-1, vmax=1, s=30, cmap=cm, alpha=0.8)
+sc = plt.scatter(q, chi_PN, c=colors, vmin=-1, vmax=1, s=30, cmap=cm)
 cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
-cbar.ax.set_ylabel('$\log(\mathcal{M}/\mathcal{M}_i)$', rotation=270, labelpad=20)
-cbar.ax.set_yticklabels(['$<-1$', '$-0.5$', '$0$', '$0.5$', '$>1$'])
[…]

### 2023-05-18 src/scripts/q148.py

@@ -4,15 +4,15 @@ import matplotlib as mpl
 import paths
 
 params = {
-    "font.size": 18,
+    "font.size": 15,
     "legend.fontsize": 12,
-    "legend.frameon": True,
-    "axes.labelsize": 18,
-    "axes.titlesize": 18,
-    "xtick.labelsize": 18,
-    "ytick.labelsize": 18,
+    "legend.frameon": False,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
     "axes.unicode_minus": False,
-    "figure.figsize": (7, 5),
+    "figure.figsize": (6, 5),
     "xtick.top": True,
     "ytick.right": True,
     "xtick.bottom": True,
@@ -34,16 +34,60 @@ params = {
 
 mpl.rcParams.update(params)
 
+# c2 = "#4e7ab1"  # blue Nash
+
+# c2 = "#406a3d"
+# c1 = "#b14a49"
+# c1 = "#b0882f"  # yellow nash
+c1 = "#102F68"
+c2 = "#B02423"
+
+
 q148 = np.loadtxt(paths.data / "q148.dat")
 bins = np.linspace(-5, -1, 15)
 
-plt.hist(np.log10(q148[:,4]), bins=bins, alpha=0.2, color="C0", label="Original")
-plt.axvline(x=np.log10(np.median(q148[:,4])), color="C0", zorder=-20, alpha=0.4, ls=":")
+plt.hist(np.log10(q148[:, 4]), bins=bins, alpha=0.2, color=c1)
+plt.hist(
+    np.log10(q148[:, 4]),
+    bins=bins,
+    color=c1,
+    histtype="step",
+    linewidth=2,
+    label="Original",
+)
+plt.axvline(
+    x=np.log10(np.median(q148[:, 4])),
+    color=c1,
+    zorder=-20,
+    alpha=1.0,
+    lw=2,
+    linestyle=(0, (1, 1.3)),
+)
 
-plt.hist(np.log10(q148[:,5]), bins=bins, label="$\mathcal{L}_{\mathrm{mean}}$", histtype="step", color="C1", linewidth=3)
-plt.axvline(x=np.log10(np.median(q148[:,5])), color="C1", zorder=-20, alpha=0.4, ls=":")
+plt.hist(
+    np.log10(q148[:, 5]),
[…]

### 2023-05-18 src/scripts/q148_q1248_compare.py

@@ -4,15 +4,15 @@ import matplotlib as mpl
 import paths
 
 params = {
-    "font.size": 18,
+    "font.size": 15,
     "legend.fontsize": 12,
-    "legend.frameon": True,
-    "axes.labelsize": 18,
-    "axes.titlesize": 18,
-    "xtick.labelsize": 18,
-    "ytick.labelsize": 18,
+    "legend.frameon": False,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
     "axes.unicode_minus": False,
-    "figure.figsize": (7, 5),
+    "figure.figsize": (6, 5),
     "xtick.top": True,
     "ytick.right": True,
     "xtick.bottom": True,
@@ -34,19 +34,55 @@ params = {
 
 mpl.rcParams.update(params)
 
+c1 = "#B02423"
+c2 = "#266D1F"
+# c2 = "#785EF0"
+
+
 q148_data = np.loadtxt(paths.data / "q148.dat")
 q1248_data = np.loadtxt(paths.data / "q1248.dat")
 
-bins = np.arange(-2, 2, 0.2)
-q148_logdiff = np.log10(q148_data[:,5]/q148_data[:,4])
-q1248_logdiff = np.log10(q1248_data[:,5]/q1248_data[:,4])
+# bins = np.arange(-2, 2, 0.2)
+bins = np.linspace(-2, 2, 15)
+q148_logdiff = np.log10(q148_data[:, 5] / q148_data[:, 4])
+q1248_logdiff = np.log10(q1248_data[:, 5] / q1248_data[:, 4])
 
-plt.hist(q148_logdiff, bins=bins, label='q148', alpha=0.3, color="C0")
-plt.axvline(x=np.median(q148_logdiff), color="C0", zorder=-20, alpha=0.4, ls=":")
-plt.hist(q1248_logdiff, bins=bins, label="q1248", histtype="step", color="C1", linewidth=3)
-plt.axvline(x=np.median(q1248_logdiff), color="C1", zorder=-20, alpha=0.4, ls=":")
+plt.hist(q148_logdiff, bins=bins, alpha=0.2, color=c1)
+plt.hist(q148_logdiff, bins=bins, label="Optimized", color=c1, histtype="step", lw=2)
+plt.axvline(
[…]

### 2023-05-25 src/tex/ms.tex
update plots

@@
 We divided the parameter space into four regions to analyze the effect of the
-recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). The
-training waveforms used for fitting are listed in Table~\ref{tab:quadrants}. The
+recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). 
+The training waveforms used for fitting are listed in Tab.~\ref{tab:quadrants}. The
 top-left and bottom-right regions have limited data for $q>4$, hence the results are
 only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
-waveforms lying in the lower diagonal have mismatches worsened while other waveforms 
-generally improve.
-For the top-right and bottom-left regions, we see that the optimization prefers waveforms with $\chi_1<\chi_2$.
-This further  In the top-left region, the improvement is significant, except for a
-few defects caused by some testing waveforms having $q>4$. This can be seen clearer in 
-Fig.~\ref{fig:all_quadrants}, which we see that the optimized histogram shifts downward 
-rather uniformly. The optimized waveforms in the bottom-right region have two distinct 
-features. Within this region, waveforms in the bottom-right corner improves significantly 
-while others worsened. This phenomenon can also be seen in Fig.~\ref{fig:all_quadrants}, 
-where the mismatch distribution shifted upward. Hence, we can deduce that the loss function 
-descended into a different minimum compared to Fig.~\ref{fig:ps_q148}, making the optimized 
-coefficients not useful in GW analysis. THus, further optimizing in a smaller region does 
-not generally improve the results.
-%\kw{I am a bit confused about why you can get worse result.}
+equal-spin waveforms lying on the diagonal has great improvement. Above the diagonal, 
+the improvement is significant, except for a few defects caused by some testing waveforms 
+having $q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see 
+that the optimized histogram shifts downward uniformly. Below the diagonal, waveforms 
+have negligible improvement (Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}), 
+indicating the original set of coefficients is the optimal set of coefficients, and 
+cannot be further improved. Hence, we can deduce IMRPhenomD has better match with the 
+top-left region and does not fit well with waveforms lying in the bottom-right 
+region. Thus, the ansatz restricts additional improvements and further optimizing in a 
+smaller region does not generally improve the results.
 
@@
 demonstration of the general method used.
-% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for
-% both amplitude and phase. Similarly, for the merger-ringdown region, the
-% optimized waveforms also show great improvement.
 
@@
 unequal spin events. This degeneracy produces straight lines in the parameter
-space with negative slopes that depend on the mass ratio, as shown in
+space with negative slopes that depend on the mass ratio, which can be seen in 
 Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
@@
 regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
-performance has improved, while the bottom-right region exhibits two distinct 
-features. This is because the loss function has descended into a different 
-minimum which highly favors events with $\chi_1\approx1$ and $\chi_2\approx-1$. 
+performance has improved, while the bottom-right region hardly improves. 
 This observation suggests the ansatz is specific to certain regions of the 
-parameter space, with a preference for BBH events lying in the upper diagonal 
-region, and it hardly fits with waveforms lying in the lower diagonal region.
+parameter space, with a preference for BBH events lying above the diagonal, 
+and it has limited enhancement for events lying below the diagonal.
 
@@
 parameter space, we can perform optimization with fewer restrictions and select
-regions more systematically.
+regions more systematically. Other than how to divide regions of interest, the 
+choice of training waveforms also affects the final results greatly. Note that 
+the choice of training waveforms listed in Tab.~\ref{tab:quadrants} were taken 
+arbitrarily to test the effectiveness of separate fitting. Hence, if one takes 
+a different set of training waveforms over the parameter space, the result might 
+give additional features that can test and explain IMRPhenomD better.
 
@@
 we can see that this recalibration process gives further
-improvement to the top-left region while the bottom-right region falls into a different 
-loss minimum. This indicates that the ansatz does not fit certain waveforms in this 
-region, hence introduces bias to downstream GW analysis. This phenomenon is mainly 
-due to the reduced spin approximation used in parameterizing the 
-ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
+improvement to the top-left region while the bottom-right region only have little 
+improvement. This indicates that the ansatz has limited match in this region and 
+does not fit most waveforms in this region, hence introduces bias to downstream GW 
+analysis. This phenomenon is mainly due to the reduced spin approximation used in 
+parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
 
@@
 construct better GW Phenom models in the future.
-
-% From analyzing mismatches in parameter space, we investigated the accuracy of
-% generated waveforms based on their mass ratio and spin. Similar studies can be
-% performed on other GW models to give better criteria to
-
-% Improved the IMRPhenomD model. Found regions that can be improved. Checked
-% ansatz.
 

### 2023-05-25 src/scripts/all_quadrants.py

@@ -37,9 +37,8 @@ mpl.rcParams.update(params)
 c1 = "#102F68"
 c2 = "#B02423"
 
-q148 = np.loadtxt(paths.data / "all_quadrants.dat")
-data2 = np.array([i for i in q148 if (i[2] < 0 and i[3] > 0)])
-data4 = np.array([i for i in q148 if (i[2] > 0 and i[3] < 0)])
+data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
+data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
 
 bins = np.linspace(-5, -1, 15)
 

### 2023-05-25 src/scripts/ps_q148_quadrants.py

@@ -34,7 +34,11 @@ params = {
 
 mpl.rcParams.update(params)
 
-data = np.loadtxt(paths.data / "all_quadrants.dat")
+data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
+data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
+data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
+data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
+data = np.concat((data1, data2, data3, data4))
 
 logdiff = np.log10(data[:, 5] / data[:, 4])
 cm = plt.cm.get_cmap("RdYlBu")

### 2023-05-25 src/scripts/ps_q148_quadrants.py
showyourwork plot fix

@@ -38,7 +38,7 @@ data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
 data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
 data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
-data = np.concat((data1, data2, data3, data4))
+data = np.concatenate((data1, data2, data3, data4))
 
 logdiff = np.log10(data[:, 5] / data[:, 4])
 cm = plt.cm.get_cmap("RdYlBu")

### 2023-05-26 src/tex/ms.tex
update

@@
 waveform coefficients because they lie in the neighborhood of the minimum that we wish 
-to find. Then, by taking $\alpha=10^{-6}$, we see in Fig. \ref{fig:loss} plateau at 
-around $N=30000$, hence we end the optimization here.  
+to find. Then, by taking $\alpha=10^{-6}$, we see the validation loss in 
+Fig.~\ref{fig:loss} plateau at around $N=12000$, hence we end the optimization here.  
 %\te{I think we need a bit more discussion about the stopping criterion here.
@@
 
-To evaluate the effectiveness of the optimization, we analyze approximately 530
-NR waveforms \kw{Just say how many you have} from the SXS catalog in this study.
-We specifically select waveforms with negligible eccentricity
-(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
-consistent with the constraints of the waveform model. As shown in
-Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
-waveforms fall within the parameter space defined by the training waveforms.
-This suggests that these test waveforms can be used for direct comparison with
-the original model.
+To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
+from the SXS catalog in this study. We specifically select waveforms with 
+negligible eccentricity (${e<2\times10^{-3}}$) and precession 
+(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
+the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
+parameters of the chosen test waveforms fall within the parameter space defined 
+by the training waveforms. This suggests that these test waveforms can be used 
+for direct comparison with the original model.
 
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches in the top-left and bottom-right regions.
+	\caption{Distributions of mismatches in the top-left (Top) and bottom-right (Bottom) regions.
 	We use a constant noise spectrum to calculate mismatch and
 	$\mathcal{L}_{\mathrm{mean}}$ as the loss function.
-Waveforms in the top-left region generally improves while waveforms in the bottom-right region worsened, as indicated by the median (dashed lines).}
+Waveforms in the top-left region generally improves while waveforms in the bottom-right region has very little improvement, as indicated by the median (dashed lines).}
 	\label{fig:all_quadrants}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Parameter space of testing waveforms.
-Each region is fitted independently with waveforms listed in Tab.~\ref{tab:quadrants}.
-Colorbar represents the $\log_{10}$ difference of mismatches before and after optimization.}
+	\caption{Parameter space of testing waveforms. Each region is fitted independently with waveforms listed in Tab.~\ref{tab:quadrants}.}
 	\label{fig:ps_q148_quadrant}
@@
 observations suggest that the parameterized ansatz employed in our study may not
-be suitable for certain regions in the parameter space, leading to mismatches
+be suitable for certain regions in the parameter space, leading to low mismatches
 for some waveforms while other waveforms remain at the high mismatch tail with
@@
 observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
-parameter, $\chi_{\mathrm{PN}}$, as described in Sec. \ref{sec:method}. The
+parameter, $\chi_{\mathrm{PN}}$, as described in Sec.~\ref{sec:method}. The
 parameterization of BBH mergers using a single spin parameter results in a
@@
 space with negative slopes that depend on the mass ratio, which can be seen in 
-Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
+Fig.~\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
 region than the bottom-right region. In an attempt to address
 this issue, we partitioned the parameter space into four regions, as described
-in Sec. \ref{sec:result}. With separate optimizations for each
-regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
+in Sec.~\ref{sec:result}. With separate optimizations for each
+regions, Fig.~\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
 performance has improved, while the bottom-right region hardly improves. 

### 2023-06-02 src/tex/ms.tex
First edits to 2.1

@@
 
-\section{Method} \label{sec:method}
+\section{Optimization Method} \label{sec:method}
+
+In this section, we first briefly review the contruction of the IMRPhenomD model and discuss
+how the calibration parameters enter the waveform.
+We then outline the mismatch and how it can be used as a loss fucntion. 
+Finally we discuss the gradient descent algorithm and our stopping criterion.
 
@@
 
-In order to recalibrate the model, we have to understand what parameters the model has.
-Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
-For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
+% Paragraphs we need:
+% 1. What is the model
+% 2. How do the parameters enter the model
+We start by giving a succinct summary of the IMRPhenomD model and the relevant parameters.
+Interested readers should refer to \citep{khan2016frequency} for more details on construction of the model.
 
-The IMRPhenomD model is constructed by combining three individually fitted parts
-into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
-\begin{align}\label{eq:joint_waveform}
-	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
+Aligned-spin, frequency domain waveform models (such as IMRPhenomD) can be written as a
+combination of amplitude and phase functions ($A$ and $\phi$ respectively):
+\begin{align}\label{eq:}
+	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,.
 \end{align}
-Instead of fitting the strain, which is a highly oscillatory function that is
-difficult to fit, the amplitude and phase are fitted since they are smoother
-functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
-amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
-\begin{equation}\label{eq:amplitude}
-\begin{aligned}
-	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
-	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
-	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
-	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
-\end{aligned}
-\end{equation}
-where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
-The ansatzes for phase can be found in \citep{khan2016frequency}.
-These simple analytic functions
-consists of parameters $\Lambda^i=\{\rho_i,\delta_i,\gamma_i\}$, which are defined as follows. 
+\te{Check minus sign}
+Here $f$ is the frequnecy, $\theta$ are of intrinsic parameters of the binary, and $\Lambda$ is a set of additional
+parameters which will be discussed below. 
+The phase and amplitude functions are then split into three sections which represent the
+inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
+During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
+PhenomD uses the TaylorF2 model \citep{boyle2007high} which is known up to 3.5PN order.
+To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
+uses a series of parameterizations\footnote{
+	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
+} 
+which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
+The three sections are then \textit{stictched} together using step functions.
+Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
+boundary between each section.
+
+In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
+Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
 \begin{align} \label{eq:Lambda}
@@
 	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
-	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2),
+	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
 \end{align}
-where $\lambda$ are fitting coefficients obtained during calibration, $\eta$ is
+where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
 the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
@@
 \begin{align}
-	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
+	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
 \end{align}
 Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
-respectively. Finally, the individual segments are first connected directly using 
-step functions. Then, by fixing coefficients in the intermediate segment, one can make 
-the final waveform is continuous in its first derivative.
+respectively.
+
+% More precisely, $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+% the correlations between different segments are omitted, which could limit the 
+% accuracy of the model. Also, since fitting was performed before connecting 
+% individual segments, the final waveform does not guarantee to achieve the optimal 
+% waveform.
+% Finally, the individual segments are first connected directly using 
+% step functions. Then, by fixing coefficients in the intermediate segment, one can make 
+% the final waveform is continuous in its first derivative.
+
+% In order to recalibrate the model, we have to understand what parameters the model has.
+% Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
+% For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
+
+% The IMRPhenomD model is constructed by combining three individually fitted parts
+% into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
+% \begin{align}\label{eq:joint_waveform}
+% 	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
+% \end{align}
+% Instead of fitting the strain, which is a highly oscillatory function that is
+% difficult to fit, the amplitude and phase are fitted since they are smoother
+% functions. In each part, the amplitude and phase are made using simple functions
+% of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
+% amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
+% \begin{equation}\label{eq:amplitude}
+% \begin{aligned}
+% 	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
+% 	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
+% 	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
+% 	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
+% \end{aligned}
+% \end{equation}
+% where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
 
 %\kw{Need to work on this paragraph}
-Combining Eq.~\ref{eq:joint_waveform},~\ref{eq:amplitude}~and~\ref{eq:Lambda}, 
-we can see that the entire waveform is non-linear in $\lambda$. A slightly 
-inaccurate set of $\lambda$ can significantly affect the shape of the generated 
-waveforms. Thus, having a set of accurate waveform coefficients is important 
-and fundamental to having accurate GW models. Generally, waveform coefficients 
-are obtained by calibrating with NR waveforms, which are waveforms computed 
-using NR simulations. In the case of \citep{khan2016frequency}, they first 
-obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
-and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
-sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
-Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
-the correlations between different segments are omitted, which could limit the 
-accuracy of the model. Also, since fitting was performed before connecting 
-individual segments, the final waveform does not guarantee to achieve the optimal 
-waveform. The connecting procedure alters the previously fitted waveform. 
-Hence, the model generated waveforms contains additional inaccuracies.
+Although initially independent, the stiching procedure means that each section
+of the waveform intrinsically depends on the full set of $\lambda$s. 
+A slightly inaccurate set of $\lambda$s can therefore lead to inaccuracies in
+the generated waveforms. 
+Thus, having a set of accurate waveform coefficients is crucial to the accuracy
+of the GW models. 
+Importantly, since the fitting was performed on the individual segments,
+the final waveform is not guarenteed to have $\lambda$s close to global minima.
 
-Instead, we recalibrate coefficients jointly, which we can remove inaccuracies
-and biases discussed above, and can improve model accuracy. In the past, due to
-the complex nature of GW strains and piece-wise formalism of IMRPhenomD,
-non-linear fitting was difficult to be performed in optimizing coefficients.
-Hence, piece-wise optimization was done to obtain coefficients. However, with
-{\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus
-allowing the use of gradient-based algorithms for us to recalibrate the model.
+% Generally, waveform coefficients 
+% are obtained by calibrating with NR waveforms, which are waveforms computed 
+% using NR simulations. In the case of \citep{khan2016frequency}, they first 
+% obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+% the correlations between different segments are omitted, which could limit the 
+% accuracy of the model. Also, since fitting was performed before connecting 
+% individual segments, the final waveform does not guarantee to achieve the optimal 
+% waveform. The connecting procedure alters the previously fitted waveform. 
+% Hence, the model generated waveforms contains additional inaccuracies.
+
+At the time of construction this piece-wise fitting was necessary since
+$\lambda$ has 209 components making the optimization computationally prohibitive.
+Here, we recalibrate the $\lambda$ coefficients jointly. 
+This is made possible by the use of gradient-based optimization algorithms which are
+enabled by AD from \jax and {\ripple}.
+% , gradients of IMRPhenomD can be easily obtained, thus
+% allowing the use of gradient-based algorithms for us to recalibrate the model.
 

### 2023-06-02 src/tex/ms.tex
Additional 2.1 edits

@@
 
-Aligned-spin, frequency domain waveform models (such as IMRPhenomD) can be written as a
+Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
 combination of amplitude and phase functions ($A$ and $\phi$ respectively):
 \begin{align}\label{eq:}
-	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,.
+	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
 \end{align}
-\te{Check minus sign}
-Here $f$ is the frequnecy, $\theta$ are of intrinsic parameters of the binary, and $\Lambda$ is a set of additional
+where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
 parameters which will be discussed below. 
@@
 During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
-PhenomD uses the TaylorF2 model \citep{boyle2007high} which is known up to 3.5PN order.
+IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
 To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
@@
 
-At the time of construction this piece-wise fitting was necessary since
-$\lambda$ has 209 components making the optimization computationally prohibitive.
-Here, we recalibrate the $\lambda$ coefficients jointly. 
-This is made possible by the use of gradient-based optimization algorithms which are
-enabled by AD from \jax and {\ripple}.
+At the time of construction this piece-wise approach was necessary since
+$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
+Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
+This is made possible by the use of gradient-based optimization algorithms,
+enabled by AD from \jax and {\ripple}, which are significatnly moe efficient in high dimensions.
 % , gradients of IMRPhenomD can be easily obtained, thus

### 2023-06-18 src/tex/ms.tex
Update

--- 2023-05-26 src/tex/ms.tex
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Parameter space with mass ratio $q$ against post-newtonian spin parameter $\chi_{\mathrm{PN}}$. Orange: Training waveforms; Blue: Testing waveforms}
+	\caption{Distribution of training and testing waveforms in $q-\chi_{PN}$
+	space. The blue dots represent the waveforms used in training and the orange
+	stars represent the waveforms used in testing.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
 	\label{fig:intrin_space}
@@
 
-To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
-from the SXS catalog in this study. We specifically select waveforms with 
-negligible eccentricity (${e<2\times10^{-3}}$) and precession 
-(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
-the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
-parameters of the chosen test waveforms fall within the parameter space defined 
-by the training waveforms. This suggests that these test waveforms can be used 
-for direct comparison with the original model.
+To evaluate how well the optimization procedure generalize to waveforms that are
+not in the training set, we evaluate the mismatch between the fine-tuned model
+and the data provided in the SXS catalog for 536 NR waveforms. We select
+waveforms that share the same part of the parameter space with the training
+waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
+and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
+shows how the training and testing waveforms are distributed in the
+$q-\chi_{PN}$ space.
 
-To illustrate the improvement on an individual waveform level, we compare a
-particular waveform before and after optimization in Fig. \ref{fig:0154}.
-Compared to the original IMRPhenomD waveform, we can see the optimized waveform
-has smaller residual from NR waveform both in amplitude and phase, particularly
-in the inspiral region, where the amplitude displays a $50\%$ reduction in
-error. For a fair comparison, we selected one of the testing waveforms from the
-catalog presented in \citep{khan2016frequency}.
+To illustrate the effect of optimiziation on an individual waveform level, we
+compare the mismatch of a particular waveform before and after optimization with
+respect to the NR waveform taken directly from the SXS catalog in Fig.
+\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
+optimized waveform has smaller residual from NR waveform both in amplitude and
+phase, particularly in the inspiral region, where the amplitude displays a
+$50\%$ reduction in error. For a fair comparison, we selected one of the testing
+waveforms from the catalog presented in \citep{khan2016frequency}.
 
-To quantify the effect of optimization over the entire dataset, We evaluate the
-mismatch of all testing waveforms using a constant PSD weighted loss function,
-$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
-\ref{fig:q148}. The peak of the distribution has shifted towards a lower
-mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
-in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
-improvement with a 22.9\% decrease in the median. Even though the distribution
-lacks a clear peak due to the problem discussed in Section
-\ref{subsec:optimization}, both distributions show observable improvement.
+With the purpose of improving downstream tasks such as parameter estimation in
+mind, the more relevant metric of improvement is the distribution of improvement
+in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
+distribution of log-mismatch for the testing waveforms before and after the
+optimization procedure. For generality, we use a constant PSD in our loss
+function. One can see the distribution have been skewed toward lower mismatch,
+where the peak of the distribution is shifted by approximately an order of
+magnitude, and a median mismatch is reduced by 50\%. When using
+$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
+decrease in the median.
 
@@
 
-%\kw{Which plot is this paragraph referring to?}
-Applying the same methods, we will see that the distributions of mismatches
-calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
-unweighted mismatch.
-The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
-This result is expected, as the IMRPhenomD model was initially constructed and
-fitted using the {\zdethp} weighted mismatch. Consequently, the model is
-anticipated to closely fit the NR waveforms while incorporating the influence of
-the {\zdethp} PSD, rather than a constant PSD.
+Note that the IMRPhenomD model was initially constructed and fitted using the
+{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
+methods using the {\zdethp} PSD exhibit superior improvement compared to the
+unweighted mismatch. We find no significant qualitative and quantitative
+difference between the two PSDs in the resulting distribution of mismatches.
 

### 2023-06-18 src/tex/ms.tex
Finishing result section

@@
 Furthermore, by comparing the accuracy of optimized and original waveforms, we
-find that model-generated waveforms are very sensitive to their intrinsic
+find that model-generated waveforms are very sensitive to their source
 parameters. Specifically, IMRPhenomD favors certain parts of the parameter
@@
 of optimized waveforms with original waveforms. We also show how the
-optimization result differs with waveforms of different intrinsic parameters. In
+optimization result differs with waveforms of different source parameters. In
 Sec.~\ref{sec:discussion}, we address the difference between our calibrating
@@
 where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the power spectral density(PSD) of noise from the instrument, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the relevant maximum and minimum frequencies for the integration. $t_0$ and $\phi_0$ are maximized by performing regression between $h_1$ and $h_2$. The mismatch
-is a quantity that is closely related to the mean square error (MSE) between the
-two waveforms, as it is often considered a weighted version of the MSE.
+$S_n(f)$ is the power spectral density(PSD) of noise from the instrument,
+$f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the relevant maximum and minimum
+frequencies for the integration. $t_0$ and $\phi_0$ are maximized by performing
+regression between $h_1$ and $h_2$. The mismatch is a quantity that is closely
+related to the mean square error (MSE) between the two waveforms, as it is often
+considered a weighted version of the MSE.
 
-Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
-However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
-To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
-We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
+Since we wish to optimize the model over the whole parameter space, we need to
+compare multiple model generated waveforms with NR waveforms. However, mismatch
+only quantifies the difference between IMRPhenomD and NR waveform for one
+particular set of intrinsic parameters. To take into account of various
+different waveforms in the parameter space, we pick waveforms from the different
+parts of the parameter space. We define the loss function as an average of
+training waveforms in two ways, the simple average of mismatches and the
+normalize average of mismatches,  
 \begin{align}
@@
 
-Motivated by the successful improvement of the waveforms, we expand the training
-dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. We 
-specifically choose to use $q=2$ events since we have abundant $q=2$ NR waveforms 
-to validate the final result. The new set of coefficients generated from this
-optimization process yields only marginal improvements in the newly produced
-waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
-of the optimized distribution remains comparable in length and endpoint to the
-original distribution, indicating that our procedure is incapable of improving
-these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
-function with additional waveforms leads to similar improvement in the resulting
-distribution.
+To understand whether additional training data can further improve the
+performance of the model, we include waveforms that are not presence in the
+original IMRPhenomD calibration in our training dataset with parameters
+tabulated in Tab. \ref{tab:q1248}. We specifically choose to use $q=2$ events
+since we have abundant $q=2$ NR waveforms to validate the final result. The new
+set of coefficients generated from this optimization process yields only
+marginal improvements in the newly produced waveforms, as seen in Fig.
+\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
+distribution remains comparable to the original distribution, indicating that
+the original dataset is sufficient for this task. Similarly, utilizing the
+{\zdethp} PSD to optimize the loss function with additional waveforms results in 
+similar level of improvement.
 
-% We believe that either the ansatz does not suit waveforms within certain
-% regions of parameter space or the set of optimized coefficients falls to a
-% minimum where they do not fit some testing waveforms. If in some region of
-% parameter space, the ansatz does not fit well with NR waveforms, it would
-% never be able to have consistent improvement under optimization. Instead, it
-% would surf around the loss manifold with random fluctuations in mismatches. If
-% the set of coefficients does not fit, then dividing the parameter space into
-% regions and fitting separate sets of coefficients should give better results.
-Although most of the waveforms show improvement in figure \ref{fig:q148}, the
-high mismatch tail of the distribution remains unaffected. Given that the
-waveform model's ansatz may not be entirely compatible with NR, and the
-optimization procedure is carried out over a distribution of waveforms with
-varying intrinsic parameters, it is conceivable that some trade-offs in accuracy
-exist between different parts of the parameter space. If this is the cause of
-the lack of improvement in the high mismatch tail of the distribution,
-segmenting the parameter space into smaller subspaces should alleviate this
-problem. On the other hand, if the ansatz lacks the correct parameterized form
-to capture the NR waveforms' behavior as a function of the intrinsic parameters,
-the results will always be biased, and we should not expect any improvement,
-even if we segment the parameter space during training.
+To investigate the performance of recalibration over the source parameter space,
+we plot the improvement of mismatch in log projected over the parameter space of
+$q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
+axis, we can see the waveforms with $q\lq4$ show most consistent average
+improvement. This is because that part of the parameter space is better covered
+by the training waveform, as compared to testing waveforms from $q=4$ to $q=8$,
+where some testing waveforms lie outside the source parameter space covered by
+the training set of waveforms. On the spin axis, we can see the waveform with
+$\chi_{\textrm{PN}}$ close to zero show the most consistent improvement, and we
+start deviate from 0, there could be fluctuations in the average improvement. On
+top of that, in the $q\lq4$ region, there seems to be more consistent
+improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as
+since the original waveform was developed for aligned spin system, the waveforms
+with $\chi_{\textrm{PN}}$ is less well-fitted \kw{Varify this.}, hence there
+could be more room for improvement. To show this, we plot the parameter space of
+$\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal
+axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements as discussed
+above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
+($\chi_1>0>\chi_2$) regions respond to optimization differently. In the top-left
+region, waveforms generally improve slightly with along optimization. However,
+waveforms in the bottom-right region do not improve after optimization. Some
+waveforms even turned worse after optimization.
 
-Since we know intrinsic parameters play an important role in the ansatz, we
-would like to investigate how intrinsic parameters affect the recalibration
-process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
-Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
-more consistent improvement, since the ansatz are developed base on non-spinning 
-waveforms. Also, the original coefficients were fitted using NR waveforms with
-equal or similar spin, hence the model prefers waveforms with similar spin. On 
-the other hand, spinning waveforms can either improve or worsen in terms of 
-mismatch base on their intrinsic parameters. To further discuss the behavior of 
-non-spinning waveforms, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
-Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. 
-$\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
-Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right ($\chi_1>0>\chi_2$) 
-regions respond to optimization differently. In the top-left 
-region, waveforms generally improve slightly with along optimization. However, waveforms 
-in the bottom-right region do not improve after optimization. Some waveforms even 
-turned worse after optimization.
+ Given that the waveform model's ansatz may not be entirely compatible with NR,
+and the optimization procedure is carried out over a distribution of waveforms
+with varying source parameters, it is conceivable that different parts of the
+source parameter space may not share the same set of optimal IMRPhenomD
+parameters, mean there could be some trade-offs in accuracy between different
+parts of the parameter space. If this is the cause of the lack of improvement in
+the high mismatch tail of the distribution, segmenting the parameter space into
+smaller subspaces should alleviate this problem. On the other hand, if the
+ansatz lacks the correct parameterized form to capture the NR waveforms'
+behavior as a function of the source parameters, the results will always be
+biased, and we should not expect any improvement, even if we segment the
+parameter space during training.
 
@@
 We divided the parameter space into four regions to analyze the effect of the
-recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). 
-The training waveforms used for fitting are listed in Tab.~\ref{tab:quadrants}. The
-top-left and bottom-right regions have limited data for $q>4$, hence the results are
-only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
-equal-spin waveforms lying on the diagonal has great improvement. Above the diagonal, 
-the improvement is significant, except for a few defects caused by some testing waveforms 
-having $q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see 
-that the optimized histogram shifts downward uniformly. Below the diagonal, waveforms 
-have negligible improvement (Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}), 
-indicating the original set of coefficients is the optimal set of coefficients, and 
-cannot be further improved. Hence, we can deduce IMRPhenomD has better match with the 
-top-left region and does not fit well with waveforms lying in the bottom-right 
-region. Thus, the ansatz restricts additional improvements and further optimizing in a 
-smaller region does not generally improve the results.
+recalibration procedure on each region separately
+(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
+this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
+bottom-right regions have limited data for $q>4$, hence the results are only
+valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
+equal-spin waveforms lying on the diagonal has great improvement. The
+improvement above the diagonal is also substantial compared to recalibrate using
+all waveforms at once, barring a few caused by some testing waveforms having
+$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
+that the optimized histogram shifts downward uniformly. On the other hand, the
+improvement below the diagonal is negligible, indicating our method struggles to
+further improve on top of the original IMRPhenomD waveform. Interestingly, we
+can see in the top-right and bottom-left quadrant, the lower right corner shows
+systematic worsening similar to the case of all waveforms. This hinted in
+general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
+anti-aligned spin where the primary spin is positive.
 
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches in the top-left (Top) and bottom-right (Bottom) regions.
-	We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ as the loss function.
-Waveforms in the top-left region generally improves while waveforms in the bottom-right region has very little improvement, as indicated by the median (dashed lines).}
+	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
+	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
+	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
+	top-left region generally improves while waveforms in the bottom-right
+	region has very little improvement, as indicated by the median (dashed
+	lines). \kw{Is the original referring to original IMRPhenomD or optimized
+	using everything? Can we have a reference for optimizing using everything?}}
 	\label{fig:all_quadrants}
@@
 
-Furthermore, we investigated how the intrinsic parameters affect the
+Furthermore, we investigated how the source parameters affect the
 improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a

### 2023-06-18 src/tex/ms.tex
Update

@@
 
-In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
-will finish its maintenance and start observing new gravitational wave (GW)
-results. This new O4 run is expected to double the rate of current binary black
+% Brief intro of waveform models
+
+Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
+as match filtering search and parameter estimation, rely on accurate waveform
+models. Because using numerical relativity (NR) waveforms in these tasks is
+prohibitively expensive, the community has constructed waveform approximants
+that can be evaluated much faster. There are three families of commonly used GW
+approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
+cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
+surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
+and phenomenological (Phenom) models \citep{husa2016frequency,
+khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
+detail of construction of these models are different, they all have a set of
+internal parameters that can be calibrated to NR waveforms. The quality of a
+waveform model is determined by the ansatz used and the accuracy of the
+calibrated parameters.
+
+% Why there is a need
+
+The LIGO-VIRGO-KAGRA (LVK) collaboration
+(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
+VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
+run on May 26, 2023. This new run is expected to double the rate of current binary black
 hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
@@
 
-% Waveform families and slight motivation of why we take Phenom model
-Currently, three families of GW models are commonly used. They are the
-effective-one-body (EOB) \citep{ossokine2020multipolar, cotesta2020frequency, taracchini2014effective}, Numerical Relativity
-(NR) surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2} and phenomenological (Phenom) models
-\citep{husa2016frequency, khan2016frequency, garcia2020multimode, pratten2021computationally}. 
-EOB models are constructed by mapping two masses onto an
-effective body under an effective metric; NR surrogate models construct
-waveforms using combinations of NR waveforms; Phenom models are formulated using
-specific ansatz and inspiral approximations. While EOB and NR surrogate models
-give better waveform approximants, Phenom waveforms can be produced much faster,
-hence it is used mostly in data analysis tasks that requires many waveform
-generations. This advantage scales up in data analysis tasks such as matched
-filtering and parameter estimation, where many waveforms are required in each
-run. This motivates us to improve upon the current framework of Phenom models,
-thus can retain the advantage of fast waveform generation while improving the
-model's accuracy.
+% AD waveforms
 

### 2023-06-19 src/tex/ms.tex
Update

@@
 Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
-as match filtering search and parameter estimation, rely on accurate waveform
+as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
 models. Because using numerical relativity (NR) waveforms in these tasks is
@@
 VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
-run on May 26, 2023. This new run is expected to double the rate of current binary black
-hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
-sensitivity of interferometers will be increased to capture more details of GW.
-Having instruments with higher sensitivity, GW models of equal or higher
-accuracy than observations should be used to extract GW information. Otherwise,
-the extracted information would be affected more by GW models instead of
-interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
-GW models are accurate enough for current analyses, the accuracy of current
-models will no longer suffice for future data analyses
-\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
-improve GW models.
+run on May 26, 2023. Because of the sensitivity improvement, the new run is
+expected to double the size of current binary black hole (BBH) observations
+\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
+detect individual events with a higher signal-to-noise ratio (SNR). This means
+we can resolve more features in the signal, at the same time put a more
+stringent requirement on the accuracy of our waveform model\citep{fill}.
 
@@
 
-Automatic differentiation (AD) is a method to calculate derivatives of functions
-up to machine precision. In traditional numerical calculations, derivatives are
-usually obtained through numerical derivatives. Symbolic derivatives were
-available but it was less efficient. Both methods were not viable in machine
-learning, where back-propagation requires precise and rapid derivative
-calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch},
-\texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train
-machine learning models. AD's algorithm is intuitive in nature. Functions
-defined are decomposed into tree structures of primitives, such as addition or
-function evaluations. Since these operations are fundamental, they were saved as
-pairs internally. Differentiation proceeds forward following the tree structure,
-with the application of the chain rule in each step to evaluate its derivative.
-Analytic derivatives of such operations are applied in each step and the desired
-derivative can then be obtained by composing back the original function
-according to the original structure. {\ripple} \citep{ripple} was a new
-implementation of IMRPhenomD, one of the models within the Phenom family. It was first implemented
-in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
-rewritten using \jax, a {\python} package that supports AD. Using
-{\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
-allowing one to freely use derivative-based algorithms to perform data analyses. 
-% Need to cite references for AD
+Due to the number of calibration parameters (often in a couple hundreds if not
+more), waveform models is usually calibrated in pieces \citep{fill}. This
+ignores the correlation between different parts of the waveform model and limits
+it quality. Recently, there has been efforts in rebuilding waveform models
+\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
+a technique used to compute derivatives of functions down to machine precision
+without issues of scaling up to high dimension or expression swelling. In
+particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
+user. This allows us to make use of infrastructures that are heavily used in
+machine learning, such as gradient descent and back propagation\cite{fill}, to
+improve the calibration of the waveform models.
 
@@
 In this paper, we investigate the possibility of further improving the accuracy
-of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
-waveforms, and what constraints one may face when trying to further improve. We
-find that simply by applying gradient descent algorithm, one can obtain a better
-set of waveform coefficients, thus improving the accuracy of the model.
-Furthermore, by comparing the accuracy of optimized and original waveforms, we
-find that model-generated waveforms are very sensitive to their source
-parameters. Specifically, IMRPhenomD favors certain parts of the parameter
-space. This means IMRPhenomD introduces systematic bias to other GW analysis
-tasks. This showcases the flaws of the ansatz and allows us to have a deeper
-understanding of Phenom models.
+of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
+calibration coefficients given a set of NR waveforms. Training on a small number
+of NR waveforms, we demonstrate one can improves the match between IMRPhenomD
+and NR waveforms over a decently sized parameter space, up to mass ratio $q=8$
+We also explore how different parts of the source parameter space (e.g. the
+primary and second spins) respond to the optimization procedure, by optimizing
+the waveform separately for different regions of the parameter space. This gives
+us hints on whether the waveform model performs equally well in different
+regions of the parameter space.
 
@@
 
+% Moved down from intro
+% In particular, we find it is easier IMRPhenomD
+% favors certain parts of the parameter space. This means IMRPhenomD introduces
+% systematic bias to other GW analysis tasks. This showcases the flaws of the
+% ansatz and allows us to have a deeper understanding of Phenom models.
+
 \section{Conclusion} \label{sec:conclusion}

### 2023-06-19 src/tex/ms.tex
Update ms.tex
Started on 2.2

--- 2023-06-18 src/tex/ms.tex
@@
 Furthermore, by comparing the accuracy of optimized and original waveforms, we
-find that model-generated waveforms are very sensitive to their source
+find that model-generated waveforms are very sensitive to their intrinsic
 parameters. Specifically, IMRPhenomD favors certain parts of the parameter
@@
 of optimized waveforms with original waveforms. We also show how the
-optimization result differs with waveforms of different source parameters. In
+optimization result differs with waveforms of different intrinsic parameters. In
 Sec.~\ref{sec:discussion}, we address the difference between our calibrating
@@
 
-\section{Method} \label{sec:method}
+\section{Optimization Method} \label{sec:method}
+
+In this section, we first briefly review the construction of the IMRPhenomD model and discuss
+how the calibration parameters enter the waveform.
+We then outline the mismatch and how it can be used as a loss fucntion. 
+Finally we discuss the gradient descent algorithm and our stopping criterion.
 
@@
 
-In order to recalibrate the model, we have to understand what parameters the model has.
-Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
-For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
+% Paragraphs we need:
+% 1. What is the model
+% 2. How do the parameters enter the model
+We start by giving a succinct summary of the IMRPhenomD model and the relevant parameters.
+Interested readers should refer to \citep{khan2016frequency} for more details.
 
-The IMRPhenomD model is constructed by combining three individually fitted parts
-into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
-\begin{align}\label{eq:joint_waveform}
-	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
+Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
+combination of amplitude and phase functions ($A$ and $\phi$ respectively):
+\begin{align}\label{eq:}
+	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
 \end{align}
-Instead of fitting the strain, which is a highly oscillatory function that is
-difficult to fit, the amplitude and phase are fitted since they are smoother
-functions. In each part, the amplitude and phase are made using simple functions
-of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
-amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
-\begin{equation}\label{eq:amplitude}
-\begin{aligned}
-	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
-	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
-	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
-	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
-\end{aligned}
-\end{equation}
-where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
-The ansatzes for phase can be found in \citep{khan2016frequency}.
-These simple analytic functions
-consists of parameters $\Lambda^i=\{\rho_i,\delta_i,\gamma_i\}$, which are defined as follows. 
+where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
+parameters which will be discussed below. 
+The phase and amplitude functions are then split into three sections which represent the
+inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
+During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
+IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
+To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
+uses a series of parameterizations\footnote{
+	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
+} 
+which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
+The three sections are then \textit{stictched} together using step functions.
+Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
+boundary between each section.
+
+In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
+Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
 \begin{align} \label{eq:Lambda}
@@
 \begin{align}
-	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
+	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
 \end{align}
 Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
-respectively. Finally, the individual segments are first connected directly using 
-step functions. Then, by fixing coefficients in the intermediate segment, one can make 
-the final waveform is continuous in its first derivative.
+respectively.
+
+% More precisely, $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+% the correlations between different segments are omitted, which could limit the 
+% accuracy of the model. Also, since fitting was performed before connecting 
+% individual segments, the final waveform does not guarantee to achieve the optimal 
+% waveform.
+% Finally, the individual segments are first connected directly using 
+% step functions. Then, by fixing coefficients in the intermediate segment, one can make 
+% the final waveform is continuous in its first derivative.
+
+% In order to recalibrate the model, we have to understand what parameters the model has.
+% Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
+% For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
+
+% The IMRPhenomD model is constructed by combining three individually fitted parts
+% into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
+% \begin{align}\label{eq:joint_waveform}
+% 	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
+% \end{align}
+% Instead of fitting the strain, which is a highly oscillatory function that is
+% difficult to fit, the amplitude and phase are fitted since they are smoother
+% functions. In each part, the amplitude and phase are made using simple functions
+% of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
+% amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
+% \begin{equation}\label{eq:amplitude}
+% \begin{aligned}
+% 	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
+% 	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
+% 	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
+% 	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
+% \end{aligned}
+% \end{equation}
+% where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
 
 %\kw{Need to work on this paragraph}
-Combining Eq.~\ref{eq:joint_waveform},~\ref{eq:amplitude}~and~\ref{eq:Lambda}, 
-we can see that the entire waveform is non-linear in $\lambda$. A slightly 
-inaccurate set of $\lambda$ can significantly affect the shape of the generated 
-waveforms. Thus, having a set of accurate waveform coefficients is important 
-and fundamental to having accurate GW models. Generally, waveform coefficients 
-are obtained by calibrating with NR waveforms, which are waveforms computed 
-using NR simulations. In the case of \citep{khan2016frequency}, they first 
-obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
-and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
-sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
-Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
-the correlations between different segments are omitted, which could limit the 
-accuracy of the model. Also, since fitting was performed before connecting 
-individual segments, the final waveform does not guarantee to achieve the optimal 
-waveform. The connecting procedure alters the previously fitted waveform. 
-Hence, the model generated waveforms contains additional inaccuracies.
+Although initially independent, the stiching procedure means that each section
+of the waveform intrinsically depends on the full set of $\lambda$'s. 
+A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
+the generated waveforms. 
+Thus, the calibration of these coefficients is crucial to the accuracy
+of IMRPhenom GW models. 
+Importantly, since the fitting was performed on the individual segments,
+the final waveform is not guarenteed to have $\lambda$'s close to global minima.
 
-Instead, we recalibrate coefficients jointly, which we can remove inaccuracies
-and biases discussed above, and can improve model accuracy. In the past, due to
-the complex nature of GW strains and piece-wise formalism of IMRPhenomD,
-non-linear fitting was difficult to be performed in optimizing coefficients.
-Hence, piece-wise optimization was done to obtain coefficients. However, with
-{\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus
-allowing the use of gradient-based algorithms for us to recalibrate the model.
+% Generally, waveform coefficients 
+% are obtained by calibrating with NR waveforms, which are waveforms computed 
+% using NR simulations. In the case of \citep{khan2016frequency}, they first 
+% obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
+% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
+% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
+% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
+% the correlations between different segments are omitted, which could limit the 
+% accuracy of the model. Also, since fitting was performed before connecting 
+% individual segments, the final waveform does not guarantee to achieve the optimal 
+% waveform. The connecting procedure alters the previously fitted waveform. 
+% Hence, the model generated waveforms contains additional inaccuracies.
+
+At the time of construction this piece-wise approach was necessary since
+$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
+Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
+This is made possible by the use of gradient-based optimization algorithms,
+enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
+% , gradients of IMRPhenomD can be easily obtained, thus
+% allowing the use of gradient-based algorithms for us to recalibrate the model.
 
@@
 
-In order to recalibrate the model, we need to define an objective function that
-quantify the performance of the model. A common choice for such metric is the
-\textit{mismatch} function \citep{husa2016frequency}. It is defined as
+In order to optimize the coefficients, we need to define a loss function that
+quantifies the difference between waveform model and the target NR simulations which we want to match.
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{husa2016frequency}. 
+It is defined as
 \begin{align} \label{eq:mismatch}
@@
 \end{align}
-where $h_{1,2}$ are the two GW waveform we are comparing, and $t_0$ and $\phi_0$
-are time shift and phase shift respectively. The $\langle h_1, h_2 \rangle$ is
-commonly referred as the inner product, which is defined as 
+where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
+are a relative time and phase shift respectively. 
+The inner product, $\langle h_1, h_2 \rangle$, is defined as 
 \begin{align}\label{eq:inner_prod}
@@
 where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the power spectral density(PSD) of noise from the instrument,
-$f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the relevant maximum and minimum
-frequencies for the integration. $t_0$ and $\phi_0$ are maximized by performing
-regression between $h_1$ and $h_2$. The mismatch is a quantity that is closely
-related to the mean square error (MSE) between the two waveforms, as it is often
-considered a weighted version of the MSE.
+$S_n(f)$ is the power spectral density (PSD), $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
+the relevant maximum and minimum frequencies for the integration.
+We note here that the mismatch can be viewed as the mean square error (MSE) between the
+two waveforms.
 
-Since we wish to optimize the model over the whole parameter space, we need to
-compare multiple model generated waveforms with NR waveforms. However, mismatch
-only quantifies the difference between IMRPhenomD and NR waveform for one
-particular set of intrinsic parameters. To take into account of various
-different waveforms in the parameter space, we pick waveforms from the different
-parts of the parameter space. We define the loss function as an average of
-training waveforms in two ways, the simple average of mismatches and the
-normalize average of mismatches,  
+\te{Up to here}
+
+Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
+However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
+To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
+We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
 \begin{align}
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Distribution of training and testing waveforms in $q-\chi_{PN}$
-	space. The blue dots represent the waveforms used in training and the orange
-	stars represent the waveforms used in testing.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
+	\caption{Parameter space with mass ratio $q$ against post-newtonian spin parameter $\chi_{\mathrm{PN}}$. Orange: Training waveforms; Blue: Testing waveforms}
 	\label{fig:intrin_space}
@@
 
-To evaluate how well the optimization procedure generalize to waveforms that are
-not in the training set, we evaluate the mismatch between the fine-tuned model
-and the data provided in the SXS catalog for 536 NR waveforms. We select
-waveforms that share the same part of the parameter space with the training
-waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
-and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
-shows how the training and testing waveforms are distributed in the
-$q-\chi_{PN}$ space.
+To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
+from the SXS catalog in this study. We specifically select waveforms with 
+negligible eccentricity (${e<2\times10^{-3}}$) and precession 
+(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
+the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
+parameters of the chosen test waveforms fall within the parameter space defined 
+by the training waveforms. This suggests that these test waveforms can be used 
+for direct comparison with the original model.
 
-To illustrate the effect of optimiziation on an individual waveform level, we
-compare the mismatch of a particular waveform before and after optimization with
-respect to the NR waveform taken directly from the SXS catalog in Fig.
-\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
-optimized waveform has smaller residual from NR waveform both in amplitude and
-phase, particularly in the inspiral region, where the amplitude displays a
-$50\%$ reduction in error. For a fair comparison, we selected one of the testing
-waveforms from the catalog presented in \citep{khan2016frequency}.
+To illustrate the improvement on an individual waveform level, we compare a
+particular waveform before and after optimization in Fig. \ref{fig:0154}.
+Compared to the original IMRPhenomD waveform, we can see the optimized waveform
+has smaller residual from NR waveform both in amplitude and phase, particularly
+in the inspiral region, where the amplitude displays a $50\%$ reduction in
+error. For a fair comparison, we selected one of the testing waveforms from the
+catalog presented in \citep{khan2016frequency}.
 
-With the purpose of improving downstream tasks such as parameter estimation in
-mind, the more relevant metric of improvement is the distribution of improvement
-in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
-distribution of log-mismatch for the testing waveforms before and after the
-optimization procedure. For generality, we use a constant PSD in our loss
-function. One can see the distribution have been skewed toward lower mismatch,
-where the peak of the distribution is shifted by approximately an order of
-magnitude, and a median mismatch is reduced by 50\%. When using
-$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
-decrease in the median.
+To quantify the effect of optimization over the entire dataset, We evaluate the
+mismatch of all testing waveforms using a constant PSD weighted loss function,
+$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
+\ref{fig:q148}. The peak of the distribution has shifted towards a lower
+mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
+in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
+improvement with a 22.9\% decrease in the median. Even though the distribution
+lacks a clear peak due to the problem discussed in Section
+\ref{subsec:optimization}, both distributions show observable improvement.
 
@@
 
-Note that the IMRPhenomD model was initially constructed and fitted using the
-{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
-methods using the {\zdethp} PSD exhibit superior improvement compared to the
-unweighted mismatch. We find no significant qualitative and quantitative
-difference between the two PSDs in the resulting distribution of mismatches.
+%\kw{Which plot is this paragraph referring to?}
+Applying the same methods, we will see that the distributions of mismatches
+calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
+unweighted mismatch.
+The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
+This result is expected, as the IMRPhenomD model was initially constructed and
+fitted using the {\zdethp} weighted mismatch. Consequently, the model is
+anticipated to closely fit the NR waveforms while incorporating the influence of
+the {\zdethp} PSD, rather than a constant PSD.
 
-To understand whether additional training data can further improve the
-performance of the model, we include waveforms that are not presence in the
-original IMRPhenomD calibration in our training dataset with parameters
-tabulated in Tab. \ref{tab:q1248}. We specifically choose to use $q=2$ events
-since we have abundant $q=2$ NR waveforms to validate the final result. The new
-set of coefficients generated from this optimization process yields only
-marginal improvements in the newly produced waveforms, as seen in Fig.
-\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
-distribution remains comparable to the original distribution, indicating that
-the original dataset is sufficient for this task. Similarly, utilizing the
-{\zdethp} PSD to optimize the loss function with additional waveforms results in 
-similar level of improvement.
+Motivated by the successful improvement of the waveforms, we expand the training
+dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. We 
+specifically choose to use $q=2$ events since we have abundant $q=2$ NR waveforms 
+to validate the final result. The new set of coefficients generated from this
+optimization process yields only marginal improvements in the newly produced
+waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
+of the optimized distribution remains comparable in length and endpoint to the
+original distribution, indicating that our procedure is incapable of improving
+these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
+function with additional waveforms leads to similar improvement in the resulting
+distribution.
 
-To investigate the performance of recalibration over the source parameter space,
-we plot the improvement of mismatch in log projected over the parameter space of
-$q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
-axis, we can see the waveforms with $q\lq4$ show most consistent average
-improvement. This is because that part of the parameter space is better covered
-by the training waveform, as compared to testing waveforms from $q=4$ to $q=8$,
-where some testing waveforms lie outside the source parameter space covered by
-the training set of waveforms. On the spin axis, we can see the waveform with
-$\chi_{\textrm{PN}}$ close to zero show the most consistent improvement, and we
-start deviate from 0, there could be fluctuations in the average improvement. On
-top of that, in the $q\lq4$ region, there seems to be more consistent
-improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as
-since the original waveform was developed for aligned spin system, the waveforms
-with $\chi_{\textrm{PN}}$ is less well-fitted \kw{Varify this.}, hence there
-could be more room for improvement. To show this, we plot the parameter space of
-$\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal
-axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements as discussed
-above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
-($\chi_1>0>\chi_2$) regions respond to optimization differently. In the top-left
-region, waveforms generally improve slightly with along optimization. However,
-waveforms in the bottom-right region do not improve after optimization. Some
-waveforms even turned worse after optimization.
+% We believe that either the ansatz does not suit waveforms within certain
+% regions of parameter space or the set of optimized coefficients falls to a
+% minimum where they do not fit some testing waveforms. If in some region of
+% parameter space, the ansatz does not fit well with NR waveforms, it would
+% never be able to have consistent improvement under optimization. Instead, it
+% would surf around the loss manifold with random fluctuations in mismatches. If
+% the set of coefficients does not fit, then dividing the parameter space into
+% regions and fitting separate sets of coefficients should give better results.
+Although most of the waveforms show improvement in figure \ref{fig:q148}, the
+high mismatch tail of the distribution remains unaffected. Given that the
+waveform model's ansatz may not be entirely compatible with NR, and the
+optimization procedure is carried out over a distribution of waveforms with
+varying intrinsic parameters, it is conceivable that some trade-offs in accuracy
+exist between different parts of the parameter space. If this is the cause of
+the lack of improvement in the high mismatch tail of the distribution,
+segmenting the parameter space into smaller subspaces should alleviate this
+problem. On the other hand, if the ansatz lacks the correct parameterized form
+to capture the NR waveforms' behavior as a function of the intrinsic parameters,
+the results will always be biased, and we should not expect any improvement,
+even if we segment the parameter space during training.
 
- Given that the waveform model's ansatz may not be entirely compatible with NR,
-and the optimization procedure is carried out over a distribution of waveforms
-with varying source parameters, it is conceivable that different parts of the
-source parameter space may not share the same set of optimal IMRPhenomD
-parameters, mean there could be some trade-offs in accuracy between different
-parts of the parameter space. If this is the cause of the lack of improvement in
-the high mismatch tail of the distribution, segmenting the parameter space into
-smaller subspaces should alleviate this problem. On the other hand, if the
-ansatz lacks the correct parameterized form to capture the NR waveforms'
-behavior as a function of the source parameters, the results will always be
-biased, and we should not expect any improvement, even if we segment the
-parameter space during training.
+Since we know intrinsic parameters play an important role in the ansatz, we
+would like to investigate how intrinsic parameters affect the recalibration
+process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
+Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
+more consistent improvement, since the ansatz are developed base on non-spinning 
+waveforms. Also, the original coefficients were fitted using NR waveforms with
+equal or similar spin, hence the model prefers waveforms with similar spin. On 
+the other hand, spinning waveforms can either improve or worsen in terms of 
+mismatch base on their intrinsic parameters. To further discuss the behavior of 
+non-spinning waveforms, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
+Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. 
+$\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
+Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right ($\chi_1>0>\chi_2$) 
+regions respond to optimization differently. In the top-left 
+region, waveforms generally improve slightly with along optimization. However, waveforms 
+in the bottom-right region do not improve after optimization. Some waveforms even 
+turned worse after optimization.
 
@@
 We divided the parameter space into four regions to analyze the effect of the
-recalibration procedure on each region separately
-(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
-this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
-bottom-right regions have limited data for $q>4$, hence the results are only
-valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
-equal-spin waveforms lying on the diagonal has great improvement. The
-improvement above the diagonal is also substantial compared to recalibrate using
-all waveforms at once, barring a few caused by some testing waveforms having
-$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
-that the optimized histogram shifts downward uniformly. On the other hand, the
-improvement below the diagonal is negligible, indicating our method struggles to
-further improve on top of the original IMRPhenomD waveform. Interestingly, we
-can see in the top-right and bottom-left quadrant, the lower right corner shows
-systematic worsening similar to the case of all waveforms. This hinted in
-general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
-anti-aligned spin where the primary spin is positive.
+recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). 
+The training waveforms used for fitting are listed in Tab.~\ref{tab:quadrants}. The
+top-left and bottom-right regions have limited data for $q>4$, hence the results are
+only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
+equal-spin waveforms lying on the diagonal has great improvement. Above the diagonal, 
+the improvement is significant, except for a few defects caused by some testing waveforms 
+having $q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see 
+that the optimized histogram shifts downward uniformly. Below the diagonal, waveforms 
+have negligible improvement (Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}), 
+indicating the original set of coefficients is the optimal set of coefficients, and 
+cannot be further improved. Hence, we can deduce IMRPhenomD has better match with the 
+top-left region and does not fit well with waveforms lying in the bottom-right 
+region. Thus, the ansatz restricts additional improvements and further optimizing in a 
+smaller region does not generally improve the results.
 
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
-	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
-	top-left region generally improves while waveforms in the bottom-right
-	region has very little improvement, as indicated by the median (dashed
-	lines). \kw{Is the original referring to original IMRPhenomD or optimized
-	using everything? Can we have a reference for optimizing using everything?}}
+	\caption{Distributions of mismatches in the top-left (Top) and bottom-right (Bottom) regions.
+	We use a constant noise spectrum to calculate mismatch and
+	$\mathcal{L}_{\mathrm{mean}}$ as the loss function.
+Waveforms in the top-left region generally improves while waveforms in the bottom-right region has very little improvement, as indicated by the median (dashed lines).}
 	\label{fig:all_quadrants}
@@
 
-Furthermore, we investigated how the source parameters affect the
+Furthermore, we investigated how the intrinsic parameters affect the
 improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a

### 2023-06-20 src/tex/ms.tex
Update tesxt

--- 2023-06-19 src/tex/ms.tex
@@
 
-Due to the number of calibration parameters (often in a couple hundreds if not
+Because of the large number of calibration parameters (often in a couple hundreds if not
 more), waveform models is usually calibrated in pieces \citep{fill}. This
@@
 of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
-calibration coefficients given a set of NR waveforms. Training on a small number
-of NR waveforms, we demonstrate one can improves the match between IMRPhenomD
-and NR waveforms over a decently sized parameter space, up to mass ratio $q=8$
-We also explore how different parts of the source parameter space (e.g. the
-primary and second spins) respond to the optimization procedure, by optimizing
-the waveform separately for different regions of the parameter space. This gives
-us hints on whether the waveform model performs equally well in different
-regions of the parameter space.
+calibration coefficients given a set of NR waveforms. Training on a few NR
+waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
+waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
+explore how different parts of the source parameter space (e.g. the primary and
+second spins) respond to the optimization procedure, by optimizing the waveform
+separately for different regions of the parameter space. This gives us hints on
+whether the waveform model performs equally well in different regions of the
+parameter space.
 

### 2023-06-21 src/tex/ms.tex
update abstract

@@
 \begin{abstract}
-    We present a simple and general method of recalibrating gravitational wave
-    (GW) phenomenological waveform models jointly. By using {\jax} and
-    {\ripple}, we can perform automatic differentiation to functions, which
-    allows us to use gradient-based optimization methods to recalibrate waveform
-    coefficients in IMRPhenomD model. This method reduces systematic bias
-    previously introduced to the model and generally can improve waveform
-    accuracy. With recalibrated coefficients, we found that the typical
-    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
-    accuracy base on the waveform's intrinsic parameters. We found that waveform
-    accuracy has significant dependence on black hole spin. Reduced spin
-    approximation introduces degeneracy in spin, which prevented further
-    improvement. We isolated regions in the parameter space that does not fit
-    the waveform ansatz. These results allow us to understand more about how to
-    develop newer phenomenological models. 
+	We investigate the possibility of improving the accuracy of a
+	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
+	calibration coefficients at once, given a set of numerical relativity (NR)
+	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
+	parts of the waveform were calibrated separately. Using \ripple, a library
+	of waveform models compatible with automatic differentiation, we can perform
+	gradients-based optimization to all the waveform coefficients the same time,
+	which should improve the quality of waveform by capturing correlations
+	between previous separately fitted parts. We found that after recalibration,
+	the median mismatch between the model and NR waveforms decrease by $50\%$.
+	We further explore how different parts of the source parameter space respond
+	to the optimization procedure. We find that the degree of improvement
+	correlates with the spins of the source. This work shows a promising avenue
+	to help understand and treat systematic error in waveform models.
 \end{abstract}

### 2023-06-21 src/tex/ms.tex
Correct for less than sign

@@
 $q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
-axis, we can see the waveforms with $q\lq4$ show most consistent average
+axis, we can see the waveforms with $q\le4$ show most consistent average
 improvement. This is because that part of the parameter space is better covered
@@
 start deviate from 0, there could be fluctuations in the average improvement. On
-top of that, in the $q\lq4$ region, there seems to be more consistent
+top of that, in the $q\le4$ region, there seems to be more consistent
 improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as

### 2023-06-22 src/tex/ms.tex
Start working on discussion
added citation and fixed some comment

@@
 Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
-as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
+as match filtering search \citep{owen1996search, owen1999matched} and parameter 
+estimation\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative}, rely on accurate waveform
 models. Because using numerical relativity (NR) waveforms in these tasks is
@@
 we can resolve more features in the signal, at the same time put a more
-stringent requirement on the accuracy of our waveform model\citep{fill}.
+stringent requirement on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.
 
@@
 Because of the large number of calibration parameters (often in a couple hundreds if not
-more), waveform models is usually calibrated in pieces \citep{fill}. This
+more), waveform models is usually calibrated in pieces \citep{khan2016frequency, 
+santamaria2010matching, pratten2021computationally}. This
 ignores the correlation between different parts of the waveform model and limits
 it quality. Recently, there has been efforts in rebuilding waveform models
-\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
-a technique used to compute derivatives of functions down to machine precision
+\citep{khan2016frequency} that supports automatic differentiation (AD) 
+\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}, which is a technique 
+used to compute derivatives of functions down to machine precision
 without issues of scaling up to high dimension or expression swelling. In
@@
 user. This allows us to make use of infrastructures that are heavily used in
-machine learning, such as gradient descent and back propagation\cite{fill}, to
-improve the calibration of the waveform models.
+machine learning, such as gradient descent and back propagation \cite{jax2018github, 
+pytorch, tensorflow2015}, to improve the calibration of the waveform models.
 
@@
 In this paper, we investigate the possibility of further improving the accuracy
-of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
+of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
 calibration coefficients given a set of NR waveforms. Training on a few NR
@@
 In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
-Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
+Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
 \begin{align} \label{eq:Lambda}
@@
 %\kw{Need to work on this paragraph}
-Although initially independent, the stiching procedure means that each section
+Although initially independent, the stitching procedure means that each section
 of the waveform intrinsically depends on the full set of $\lambda$'s. 
@@
 Importantly, since the fitting was performed on the individual segments,
-the final waveform is not guarenteed to have $\lambda$'s close to global minima.
+the final waveform is not guaranteed to have $\lambda$'s close to global minima.
 
@@
 This is made possible by the use of gradient-based optimization algorithms,
-enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
+enabled by AD from \jax\, and {\ripple}, which are significantly more efficient in high dimensions.
 % , gradients of IMRPhenomD can be easily obtained, thus
@@
 \begin{align}
-	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
+	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
 \end{align}	
@@
 average eliminates the aforementioned issue. Nevertheless, it excludes the
-information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
+information on initial mismatches. $\mathcal{L}_{\mathrm{nl}}$ restricts every
 training waveform to decrease at similar rates, hence it is hard to obtain
@@
 their ratios in mismatches would remain approximately the same. Conversely,
-$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+$\mathcal{L}_{\mathrm{ave}}$ allows the loss function to automatically adjust
 and individual mismatches would be in a similar order of magnitude after 
@@
 IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
+frequency \citep{santamaria2010matching}. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
 which is sufficient to capture all features of GW strain.
@@
 	here, 9 of them are also used in the original IMRPhenomD calibration.
-	\citep{khan2016frequency} The two remaining waveforms were from BAM
-	simulation, to which we do not have access.}
+	\citep{khan2016frequency}}
 	\label{tab:q148}
@@
 magnitude, and a median mismatch is reduced by 50\%. When using
-$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
+$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
 decrease in the median.
@@
 	\caption{Distributions of waveform mismatches calculated using 
-	$\mathcal{L}_{\mathrm{mean}}$. We use training waveforms listed in 
+	$\mathcal{L}_{\mathrm{ave}}$. We use training waveforms listed in 
 	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
@@
 	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$. 
+	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
 	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
@@
 	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
-	We show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with constant PSD 
+	We show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with constant PSD 
 	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
@@
 	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
-	show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with the
+	show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with the
 	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
@@
 (Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
-this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
+this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
+are calculated using simple average of the mismatches. The top-left and
 bottom-right regions have limited data for $q>4$, hence the results are only
@@
 	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
+	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. Waveforms in the
 	top-left region generally improves while waveforms in the bottom-right
@@
 loss function result in significantly different final mismatch distributions. As 
-seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
-$\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
+seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{ave}}$ outperforms 
+$\mathcal{L}_{\mathrm{nl}}$. By increasing the number of training waveforms, we see a 
 slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.

### 2023-06-22 src/scripts/all_quadrants.py

@@ -36,18 +36,33 @@ mpl.rcParams.update(params)
 
 c1 = "#102F68"
 c2 = "#B02423"
+c3 = "#FFB300"
 
 data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
 data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
+
+data = np.loadtxt(paths.data / "q148.dat")
+data2_q148 = np.array([data[i, 5] for i in range(len(data[:,0])) if data[i, 2] < 0 and data[i, 3] > 0])
+data4_q148 = np.array([data[i, 5] for i in range(len(data[:,0])) if data[i, 2] > 0 and data[i, 3] < 0])
 
 bins = np.linspace(-5, -1, 15)
 
 fig, ax = plt.subplots(2, figsize=(6, 10), sharex=True)
 plt.subplots_adjust(hspace=0.05)
 
-ax[0].hist(np.log10(data2[:, 4]), bins=bins, alpha=0.2, color=c1)
 ax[0].hist(
-    np.log10(data2[:, 4]), bins=bins, label="Original", color=c1, lw=2, histtype="step"
+    np.log10(data2[:, 4]), 
+    bins=bins, 
+    alpha=0.2, 
+    color=c1
+)
+ax[0].hist(
+    np.log10(data2[:, 4]), 
+    bins=bins, 
+    label="Original", 
+    color=c1, 
+    lw=2, 
+    histtype="step"
 )
 ax[0].axvline(
     x=np.log10(np.median(data2[:, 4])),
@@ -62,17 +77,38 @@ ax[0].hist(
     bins=bins,
     label="Optimized",
     histtype="step",
-    color=c2,
+    color=c3,
     linewidth=2,
 )
 ax[0].hist(
     np.log10(data2[:, 5]),
     bins=bins,
-    color=c2,
+    color=c3,
     alpha=0.2,
 )
 ax[0].axvline(
     x=np.log10(np.median(data2[:, 5])),
+    color=c3,
+    lw=2,
+    alpha=1.0,
+    linestyle=(0, (1, 1.3)),
+)
+ax[0].hist(
+    np.log10(data2_q148), 
+    bins=bins, 
+    alpha=0.2, 
+    color=c2
+)
+ax[0].hist(
+    np.log10(data2_q148), 
+    bins=bins, 
[…]

### 2023-06-22 src/tex/ms.tex
Finished section 2

--- 2023-06-21 src/tex/ms.tex
@@
 \begin{abstract}
-	We investigate the possibility of improving the accuracy of a
-	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
-	calibration coefficients at once, given a set of numerical relativity (NR)
-	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
-	parts of the waveform were calibrated separately. Using \ripple, a library
-	of waveform models compatible with automatic differentiation, we can perform
-	gradients-based optimization to all the waveform coefficients the same time,
-	which should improve the quality of waveform by capturing correlations
-	between previous separately fitted parts. We found that after recalibration,
-	the median mismatch between the model and NR waveforms decrease by $50\%$.
-	We further explore how different parts of the source parameter space respond
-	to the optimization procedure. We find that the degree of improvement
-	correlates with the spins of the source. This work shows a promising avenue
-	to help understand and treat systematic error in waveform models.
+    We present a simple and general method of recalibrating gravitational wave
+    (GW) phenomenological waveform models jointly. By using {\jax} and
+    {\ripple}, we can perform automatic differentiation to functions, which
+    allows us to use gradient-based optimization methods to recalibrate waveform
+    coefficients in IMRPhenomD model. This method reduces systematic bias
+    previously introduced to the model and generally can improve waveform
+    accuracy. With recalibrated coefficients, we found that the typical
+    \textit{mismatch} has a $50\%$ decrease. Furthermore, we analyze the
+    accuracy base on the waveform's intrinsic parameters. We found that waveform
+    accuracy has significant dependence on black hole spin. Reduced spin
+    approximation introduces degeneracy in spin, which prevented further
+    improvement. We isolated regions in the parameter space that does not fit
+    the waveform ansatz. These results allow us to understand more about how to
+    develop newer phenomenological models. 
 \end{abstract}
@@
 
-% Brief intro of waveform models
+In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
+will finish its maintenance and start observing new gravitational wave (GW)
+results. This new O4 run is expected to double the rate of current binary black
+hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
+sensitivity of interferometers will be increased to capture more details of GW.
+Having instruments with higher sensitivity, GW models of equal or higher
+accuracy than observations should be used to extract GW information. Otherwise,
+the extracted information would be affected more by GW models instead of
+interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
+GW models are accurate enough for current analyses, the accuracy of current
+models will no longer suffice for future data analyses
+\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
+improve GW models.
 
-Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
-as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
-models. Because using numerical relativity (NR) waveforms in these tasks is
-prohibitively expensive, the community has constructed waveform approximants
-that can be evaluated much faster. There are three families of commonly used GW
-approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
-cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
-surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
-and phenomenological (Phenom) models \citep{husa2016frequency,
-khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
-detail of construction of these models are different, they all have a set of
-internal parameters that can be calibrated to NR waveforms. The quality of a
-waveform model is determined by the ansatz used and the accuracy of the
-calibrated parameters.
+% Waveform families and slight motivation of why we take Phenom model
+Currently, three families of GW models are commonly used. They are the
+effective-one-body (EOB) \citep{ossokine2020multipolar, cotesta2020frequency, taracchini2014effective}, Numerical Relativity
+(NR) surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2} and phenomenological (Phenom) models
+\citep{husa2016frequency, khan2016frequency, garcia2020multimode, pratten2021computationally}. 
+EOB models are constructed by mapping two masses onto an
+effective body under an effective metric; NR surrogate models construct
+waveforms using combinations of NR waveforms; Phenom models are formulated using
+specific ansatz and inspiral approximations. While EOB and NR surrogate models
+give better waveform approximants, Phenom waveforms can be produced much faster,
+hence it is used mostly in data analysis tasks that requires many waveform
+generations. This advantage scales up in data analysis tasks such as matched
+filtering and parameter estimation, where many waveforms are required in each
+run. This motivates us to improve upon the current framework of Phenom models,
+thus can retain the advantage of fast waveform generation while improving the
+model's accuracy.
 
-% Why there is a need
-
-The LIGO-VIRGO-KAGRA (LVK) collaboration
-(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
-VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
-run on May 26, 2023. Because of the sensitivity improvement, the new run is
-expected to double the size of current binary black hole (BBH) observations
-\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
-detect individual events with a higher signal-to-noise ratio (SNR). This means
-we can resolve more features in the signal, at the same time put a more
-stringent requirement on the accuracy of our waveform model\citep{fill}.
-
-% AD waveforms
-
-Because of the large number of calibration parameters (often in a couple hundreds if not
-more), waveform models is usually calibrated in pieces \citep{fill}. This
-ignores the correlation between different parts of the waveform model and limits
-it quality. Recently, there has been efforts in rebuilding waveform models
-\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
-a technique used to compute derivatives of functions down to machine precision
-without issues of scaling up to high dimension or expression swelling. In
-particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
-user. This allows us to make use of infrastructures that are heavily used in
-machine learning, such as gradient descent and back propagation\cite{fill}, to
-improve the calibration of the waveform models.
+Automatic differentiation (AD) is a method to calculate derivatives of functions
+up to machine precision. In traditional numerical calculations, derivatives are
+usually obtained through numerical derivatives. Symbolic derivatives were
+available but it was less efficient. Both methods were not viable in machine
+learning, where back-propagation requires precise and rapid derivative
+calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch},
+\texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train
+machine learning models. AD's algorithm is intuitive in nature. Functions
+defined are decomposed into tree structures of primitives, such as addition or
+function evaluations. Since these operations are fundamental, they were saved as
+pairs internally. Differentiation proceeds forward following the tree structure,
+with the application of the chain rule in each step to evaluate its derivative.
+Analytic derivatives of such operations are applied in each step and the desired
+derivative can then be obtained by composing back the original function
+according to the original structure. {\ripple} \citep{ripple} was a new
+implementation of IMRPhenomD, one of the models within the Phenom family. It was first implemented
+in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was
+rewritten using \jax, a {\python} package that supports AD. Using
+{\ripple}, one can apply AD to GW models to obtain precise derivatives, thus
+allowing one to freely use derivative-based algorithms to perform data analyses. 
+% Need to cite references for AD
 
@@
 In this paper, we investigate the possibility of further improving the accuracy
-of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
-calibration coefficients given a set of NR waveforms. Training on a few NR
-waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
-waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
-explore how different parts of the source parameter space (e.g. the primary and
-second spins) respond to the optimization procedure, by optimizing the waveform
-separately for different regions of the parameter space. This gives us hints on
-whether the waveform model performs equally well in different regions of the
-parameter space.
+of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
+waveforms, and what constraints one may face when trying to further improve. We
+find that simply by applying gradient descent algorithm, one can obtain a better
+set of waveform coefficients, thus improving the accuracy of the model.
+Furthermore, by comparing the accuracy of optimized and original waveforms, we
+find that model-generated waveforms are very sensitive to their intrinsic
+parameters. Specifically, IMRPhenomD favors certain parts of the parameter
+space. This means IMRPhenomD introduces systematic bias to other GW analysis
+tasks. This showcases the flaws of the ansatz and allows us to have a deeper
+understanding of Phenom models.
 
@@
 In order to optimize the coefficients, we need to define a loss function that
-quantifies the difference between waveform model and the target NR simulations which we want to match.
-Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{husa2016frequency}. 
+quantifies the difference between the waveform model and the target NR simulations which we want to match.
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
 It is defined as
@@
 where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the power spectral density (PSD), $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
-the relevant maximum and minimum frequencies for the integration.
+$S_n(f)$ is the power spectral density (PSD), and $f_{\mathrm{max}}$ ($f_{\mathrm{min}}$) is
+the maximum (minimum) frequency for the integration.
 We note here that the mismatch can be viewed as the mean square error (MSE) between the
@@
 
-\te{Up to here}
-
-Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
-However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
-To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
-We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
+Since we wish to optimize the model over the whole parameter space,
+we need to compare multiple model generated waveforms with NR waveforms.
+However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
+We therefore define the loss function as an average of training waveforms in two ways,
+the simple average of mismatches and the normalized average of mismatches,  
 \begin{align}
-	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \,,\\
+	\mathcal{L}_{\mathrm{norm}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}}\,,
 \end{align}	
@@
 $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of individual training waveforms.
-Note that we choose to use two different averages, since they have different 
-preferences in optimization base on waveform mismatches. For the first choice, 
-simple average serves as the simplest choice of loss function, but is prone to 
-be dominated by a single waveform with a large mismatch. Other waveforms with 
-smaller mismatches would be insignificant comparatively, and might not be able 
-to improve under such optimization. Alternatively, the second choice, normalized 
-average eliminates the aforementioned issue. Nevertheless, it excludes the
-information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
-training waveform to decrease at similar rates, hence it is hard to obtain
-optimized waveforms with mismatches in the same order of magnitude. Instead,
-their ratios in mismatches would remain approximately the same. Conversely,
-$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and individual mismatches would be in a similar order of magnitude after 
-optimization. In this paper, we showcase the results of using both loss functions 
+training waveform, and $N$ is the total number of training waveforms.
+
+The two loss functions are chosen to give different 
+behaviour during the optimization process.
+The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
+but is prone to be dominated by a single point in parameter space with a large mismatch.
+Other points with smaller mismatches would be insignificant comparatively, and might not be able 
+to improve under such a loss function.
+Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
+by encouraging the waveform to improve at each training point at a similar rate.
+The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
+Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
+In this paper, we show the results of using both loss functions 
 and examine the differences between them.
@@
 
-To compute the loss functions, we have to take NR waveforms for calculating the mismatch. We choose 11-16 NR waveforms from the set of waveforms used in the original
-calibration process as training waveforms. Originally, 19 waveforms are taken from 
-NR simulations for calibrating IMRPhenomD \citep{khan2016frequency}, 
-which are waveforms from the SXS catalog \citep{boyle2019sxs} or BAM simulation. 
-As BAM waveforms are not publicly available, we cannot take the identical training 
-set as them.
-Instead, we take the available waveforms from the SXS catalog to construct our loss function.
-Training waveforms used are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
-The training waveforms chosen has maximum mass ratio to be 8.
-This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
+Originally, 19 NR simulations  were used to calibrate IMRPhenomD \citep{khan2016frequency}; 9 of these 
+are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
+As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
+Instead, we take the same nine waveforms from the SXS catalog plus two additional waveforms which closely mimic
+two of the low mass ratio BAM simulations. 
+The remaining BAM simulations have no close SXS counterparts, and are therefore not included in the training set.
+Our main results therefore uses 11 NR waveforms for calibration which are listed in Tab.~\ref{tab:q148}.
+Additional NR waveforms which are used for further calibration are listed in Tab.~\ref{tab:q1248}.
+
+The training set has a maximum mass ratio of eight due to the lack 
+of high mass ratio simulations in the SXS catalog.
 In fact, the SXS catalog only has NR waveforms with $q\leq10$.
-Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed from LIGO have $q\leq8$.
-Hence, we calibrate IMRPhenomD with waveforms of $q\leq8$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$,
+as most BBH events observed by LIGO and Virgo have $q\leq8$.
 
-In the SXS catalog, NR waveforms are in the form of time-series strain.
-Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
-Hence, we transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
-same intrinsic parameters. We taper the time-series using Tukey window.
-\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
+The SXS NR waveforms are given as time-series strain.
+Since IMRPhenomD is modeled in the frequency domain, we need to Fourier transform the simulation results
+in order to compute the mismatch, \eqref{eq:mismatch}.
+For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
+	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
 	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-	entire GW strain. } 
-Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series.
+	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
+} before using standard FFT routines to compute the Fourier transform.
 
-%\te{I think this discussion needs to be much more precise.
-Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
-So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
-This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
-Other than NR waveforms, one need to choose a relevant noise PSD for 
-mismatch.
-We have opted to use a flat PSD for this purpose, as it provides results that are independent of the detector sensitivity and mass scale.
-The use of a flat PSD ensures that the improvement in accuracy is due mainly to the difference in high-dimensional fitting and piece-wise fitting, but not due to the use of a different mass scale.
-Furthermore, we are interested in examining the effect of introducing a detector PSD on the optimization process.
-For this, we have chosen the zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
-Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD.
-We selected an arbitrary mass scale of $M=50M_{\odot}$ for all waveforms for demonstration, as this is a commonly observed mass scale in LIGO observations.
+In addition to choosing the NR waveforms for the training set,
+one needs to choose a relevant PSD for mismatch.
+We have opted to use a flat PSD for the mismatch calculation,
+as it provides results that are independent of the detector sensitivity and mass scale.
+The use of a flat PSD ensures that the improvement in accuracy is due mainly
+to the difference in high-dimensional fitting.
+Additionally, we are interested in examining the effect of introducing a detector PSD
+on the optimization process. For this, we have chosen the
+zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
+Since the total mass of the system scales with the frequency of the waveform,
+we must choose a corresponding mass scale to match the frequency range of our noise PSD.
+To demonstrate the effect of introducing a detector specific PSD,
+we selected an arbitrary mass scale of $M=50M_{\odot}$, as binaries
+of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.
 %\te{This choice seems pretty arbitrary.
@@
 
-We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
-In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
-(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
-waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models.
-Thus, for simplicity sake, we ignore this procedure.
-On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
-which they are long enough to contain part of the inspiral segment and all
-merger and ringdown frequency information. We take the frequency limits as
+We point out that our treatment of NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
+In the original calibration process, the training waveforms are hybrids of NR and SpinAlignedEOB (SEOB) waveforms.
+The low frequency inspiral part is taken from the SEOB
+waveforms while the rest is taken from NR simulations.
+Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and testing)
+have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, to contain part of the inspiral segment and all
+merger and ringdown frequency information. We use the frequency limits
 $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
 where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set higher
-than that in the original calibration due to NR length. When compared with
-IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
-which is sufficient to capture all features of GW strain.
+the IMRPhenomD's frequency range, except the minimum frequency is set slightly higher
+than in the original calibration due to the NR simulation length. 
+We use the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
+which is sufficient to capture all features of the GW strain.
 
@@
 coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
-original coefficients given in \citep{khan2016frequency}. We take them as the initial 
-waveform coefficients because they lie in the neighborhood of the minimum that we wish 
-to find. Then, by taking $\alpha=10^{-6}$, we see the validation loss in 
-Fig.~\ref{fig:loss} plateau at around $N=12000$, hence we end the optimization here.  
-%\te{I think we need a bit more discussion about the stopping criterion here.
-What is N, and why did we choose it to be this number?
-Ideally we would actually have a plot of the loss function during training.
-I don't have a good intuition for if its noisy.
-Does it plateau?
-If we wanted to ensure that people don't complain about overfitting we could also plot the loss of the validation set to show its not going up.}
+original coefficients given in \citep{khan2016frequency} because they likely lie in the
+neighborhood of the minimum that we wish to find. 
+We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
+don't move far from the minimum.
+Finally, we stop the optimization when the validation loss stops decreasing.
+This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.
 
@@
 
-To evaluate how well the optimization procedure generalize to waveforms that are
-not in the training set, we evaluate the mismatch between the fine-tuned model
-and the data provided in the SXS catalog for 536 NR waveforms. We select
-waveforms that share the same part of the parameter space with the training
-waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
-and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
-shows how the training and testing waveforms are distributed in the
-$q-\chi_{PN}$ space.
+To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
+from the SXS catalog in this study. We specifically select waveforms with 
+negligible eccentricity (${e<2\times10^{-3}}$) and precession 
+(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
+the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
+parameters of the chosen test waveforms fall within the parameter space defined 
+by the training waveforms. This suggests that these test waveforms can be used 
+for direct comparison with the original model.
 
-To illustrate the effect of optimiziation on an individual waveform level, we
-compare the mismatch of a particular waveform before and after optimization with
-respect to the NR waveform taken directly from the SXS catalog in Fig.
-\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
-optimized waveform has smaller residual from NR waveform both in amplitude and
-phase, particularly in the inspiral region, where the amplitude displays a
-$50\%$ reduction in error. For a fair comparison, we selected one of the testing
-waveforms from the catalog presented in \citep{khan2016frequency}.
+To illustrate the improvement on an individual waveform level, we compare a
+particular waveform before and after optimization in Fig. \ref{fig:0154}.
+Compared to the original IMRPhenomD waveform, we can see the optimized waveform
+has smaller residual from NR waveform both in amplitude and phase, particularly
+in the inspiral region, where the amplitude displays a $50\%$ reduction in
+error. For a fair comparison, we selected one of the testing waveforms from the
+catalog presented in \citep{khan2016frequency}.
 
-With the purpose of improving downstream tasks such as parameter estimation in
-mind, the more relevant metric of improvement is the distribution of improvement
-in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
-distribution of log-mismatch for the testing waveforms before and after the
-optimization procedure. For generality, we use a constant PSD in our loss
-function. One can see the distribution have been skewed toward lower mismatch,
-where the peak of the distribution is shifted by approximately an order of
-magnitude, and a median mismatch is reduced by 50\%. When using
-$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
-decrease in the median.
+To quantify the effect of optimization over the entire dataset, We evaluate the
+mismatch of all testing waveforms using a constant PSD weighted loss function,
+$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
+\ref{fig:q148}. The peak of the distribution has shifted towards a lower
+mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
+in the median. When using $\mathcal{L}_{\mathrm{norm}}$, we observe a similar
+improvement with a 22.9\% decrease in the median. Even though the distribution
+lacks a clear peak due to the problem discussed in Section
+\ref{subsec:optimization}, both distributions show observable improvement.
 
@@
 
-Note that the IMRPhenomD model was initially constructed and fitted using the
-{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
-methods using the {\zdethp} PSD exhibit superior improvement compared to the
-unweighted mismatch. We find no significant qualitative and quantitative
-difference between the two PSDs in the resulting distribution of mismatches.
+%\kw{Which plot is this paragraph referring to?}
+Applying the same methods, we will see that the distributions of mismatches
+calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
+unweighted mismatch.
+The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
+This result is expected, as the IMRPhenomD model was initially constructed and
+fitted using the {\zdethp} weighted mismatch. Consequently, the model is
+anticipated to closely fit the NR waveforms while incorporating the influence of
+the {\zdethp} PSD, rather than a constant PSD.
 
-To understand whether additional training data can further improve the
-performance of the model, we include waveforms that are not presence in the
-original IMRPhenomD calibration in our training dataset with parameters
-tabulated in Tab. \ref{tab:q1248}. We specifically choose to use $q=2$ events
-since we have abundant $q=2$ NR waveforms to validate the final result. The new
-set of coefficients generated from this optimization process yields only
-marginal improvements in the newly produced waveforms, as seen in Fig.
-\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
-distribution remains comparable to the original distribution, indicating that
-the original dataset is sufficient for this task. Similarly, utilizing the
-{\zdethp} PSD to optimize the loss function with additional waveforms results in 
-similar level of improvement.
+Motivated by the successful improvement of the waveforms, we expand the training
+dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. We 
+specifically choose to use $q=2$ events since we have abundant $q=2$ NR waveforms 
+to validate the final result. The new set of coefficients generated from this
+optimization process yields only marginal improvements in the newly produced
+waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
+of the optimized distribution remains comparable in length and endpoint to the
+original distribution, indicating that our procedure is incapable of improving
+these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
+function with additional waveforms leads to similar improvement in the resulting
+distribution.
 
-To investigate the performance of recalibration over the source parameter space,
-we plot the improvement of mismatch in log projected over the parameter space of
-$q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
-axis, we can see the waveforms with $q\le4$ show most consistent average
-improvement. This is because that part of the parameter space is better covered
-by the training waveform, as compared to testing waveforms from $q=4$ to $q=8$,
-where some testing waveforms lie outside the source parameter space covered by
-the training set of waveforms. On the spin axis, we can see the waveform with
-$\chi_{\textrm{PN}}$ close to zero show the most consistent improvement, and we
-start deviate from 0, there could be fluctuations in the average improvement. On
-top of that, in the $q\le4$ region, there seems to be more consistent
-improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as
-since the original waveform was developed for aligned spin system, the waveforms
-with $\chi_{\textrm{PN}}$ is less well-fitted \kw{Varify this.}, hence there
-could be more room for improvement. To show this, we plot the parameter space of
-$\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal
-axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements as discussed
-above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
-($\chi_1>0>\chi_2$) regions respond to optimization differently. In the top-left
-region, waveforms generally improve slightly with along optimization. However,
-waveforms in the bottom-right region do not improve after optimization. Some
-waveforms even turned worse after optimization.
+% We believe that either the ansatz does not suit waveforms within certain
+% regions of parameter space or the set of optimized coefficients falls to a
+% minimum where they do not fit some testing waveforms. If in some region of
+% parameter space, the ansatz does not fit well with NR waveforms, it would
+% never be able to have consistent improvement under optimization. Instead, it
+% would surf around the loss manifold with random fluctuations in mismatches. If
+% the set of coefficients does not fit, then dividing the parameter space into
+% regions and fitting separate sets of coefficients should give better results.
+Although most of the waveforms show improvement in figure \ref{fig:q148}, the
+high mismatch tail of the distribution remains unaffected. Given that the
+waveform model's ansatz may not be entirely compatible with NR, and the
+optimization procedure is carried out over a distribution of waveforms with
+varying intrinsic parameters, it is conceivable that some trade-offs in accuracy
+exist between different parts of the parameter space. If this is the cause of
+the lack of improvement in the high mismatch tail of the distribution,
+segmenting the parameter space into smaller subspaces should alleviate this
+problem. On the other hand, if the ansatz lacks the correct parameterized form
+to capture the NR waveforms' behavior as a function of the intrinsic parameters,
+the results will always be biased, and we should not expect any improvement,
+even if we segment the parameter space during training.
 
- Given that the waveform model's ansatz may not be entirely compatible with NR,
-and the optimization procedure is carried out over a distribution of waveforms
-with varying source parameters, it is conceivable that different parts of the
-source parameter space may not share the same set of optimal IMRPhenomD
-parameters, mean there could be some trade-offs in accuracy between different
-parts of the parameter space. If this is the cause of the lack of improvement in
-the high mismatch tail of the distribution, segmenting the parameter space into
-smaller subspaces should alleviate this problem. On the other hand, if the
-ansatz lacks the correct parameterized form to capture the NR waveforms'
-behavior as a function of the source parameters, the results will always be
-biased, and we should not expect any improvement, even if we segment the
-parameter space during training.
+Since we know intrinsic parameters play an important role in the ansatz, we
+would like to investigate how intrinsic parameters affect the recalibration
+process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
+Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
+more consistent improvement, since the ansatz are developed base on non-spinning 
+waveforms. Also, the original coefficients were fitted using NR waveforms with
+equal or similar spin, hence the model prefers waveforms with similar spin. On 
+the other hand, spinning waveforms can either improve or worsen in terms of 
+mismatch base on their intrinsic parameters. To further discuss the behavior of 
+non-spinning waveforms, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
+Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. 
+$\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
+Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right ($\chi_1>0>\chi_2$) 
+regions respond to optimization differently. In the top-left 
+region, waveforms generally improve slightly with along optimization. However, waveforms 
+in the bottom-right region do not improve after optimization. Some waveforms even 
+turned worse after optimization.
 
@@
 
-% Moved down from intro
-% In particular, we find it is easier IMRPhenomD
-% favors certain parts of the parameter space. This means IMRPhenomD introduces
-% systematic bias to other GW analysis tasks. This showcases the flaws of the
-% ansatz and allows us to have a deeper understanding of Phenom models.
-
 \section{Conclusion} \label{sec:conclusion}
@@
 seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
-$\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
+$\mathcal{L}_{\mathrm{norm}}$. By increasing the number of training waveforms, we see a 
 slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
 
-Furthermore, we investigated how the source parameters affect the
+Furthermore, we investigated how the intrinsic parameters affect the
 improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a

### 2023-06-22 src/tex/ms.tex

--- 2023-06-22 src/tex/ms.tex
@@
 
-\te{Up to here}
-
-Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
-However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
-To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
-We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
+Since we wish to optimize the model over the whole parameter space,
+we need to compare multiple model generated waveforms with NR waveforms.
+However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
+We therefore define the loss function as an average of training waveforms in two ways,
+the simple average of mismatches and the normalized average of mismatches,  
 \begin{align}
@@
 $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of individual training waveforms.
-Note that we choose to use two different averages, since they have different 
-preferences in optimization base on waveform mismatches. For the first choice, 
-simple average serves as the simplest choice of loss function, but is prone to 
-be dominated by a single waveform with a large mismatch. Other waveforms with 
-smaller mismatches would be insignificant comparatively, and might not be able 
-to improve under such optimization. Alternatively, the second choice, normalized 
-average eliminates the aforementioned issue. Nevertheless, it excludes the
-information on initial mismatches. $\mathcal{L}_{\mathrm{nl}}$ restricts every
-training waveform to decrease at similar rates, hence it is hard to obtain
-optimized waveforms with mismatches in the same order of magnitude. Instead,
-their ratios in mismatches would remain approximately the same. Conversely,
-$\mathcal{L}_{\mathrm{ave}}$ allows the loss function to automatically adjust
-and individual mismatches would be in a similar order of magnitude after 
-optimization. In this paper, we showcase the results of using both loss functions 
+training waveform, and $N$ is the total number of training waveforms.
+
+The two loss functions are chosen to give different 
+behaviour during the optimization process.
+The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
+but is prone to be dominated by a single point in parameter space with a large mismatch.
+Other points with smaller mismatches would be insignificant comparatively, and might not be able 
+to improve under such a loss function.
+Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
+by encouraging the waveform to improve at each training point at a similar rate.
+The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
+Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
+In this paper, we show the results of using both loss functions 
 and examine the differences between them.
@@
 
-We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
-In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
-(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
-waveforms while the rest of the waveforms are taken from NR simulations.
-Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models.
-Thus, for simplicity sake, we ignore this procedure.
-On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
-which they are long enough to contain part of the inspiral segment and all
-merger and ringdown frequency information. We take the frequency limits as
+We point out that our treatment of NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
+In the original calibration process, the training waveforms are hybrids of NR and SpinAlignedEOB (SEOB) waveforms.
+The low frequency inspiral part is taken from the SEOB
+waveforms while the rest is taken from NR simulations.
+Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and testing)
+have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, to contain part of the inspiral segment and all
+merger and ringdown frequency information. We use the frequency limits
 $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
 where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set higher
-than that in the original calibration due to NR length. When compared with
-IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency \citep{santamaria2010matching}. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
-which is sufficient to capture all features of GW strain.
+the IMRPhenomD's frequency range, except the minimum frequency is set slightly higher
+than in the original calibration due to the NR simulation length.
+When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum
+frequency \citep{santamaria2010matching}.
+We use the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
+which is sufficient to capture all features of the GW strain.
 

### 2023-06-22 src/tex/ms.tex
Finished abstract and section 1

@@
 \begin{abstract}
-	We investigate the possibility of improving the accuracy of a
+	We investigate the possibility of improving the accuracy of the
 	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
@@
 	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
-	parts of the waveform were calibrated separately. Using \ripple, a library
-	of waveform models compatible with automatic differentiation, we can perform
-	gradients-based optimization to all the waveform coefficients the same time,
-	which should improve the quality of waveform by capturing correlations
-	between previous separately fitted parts. We found that after recalibration,
-	the median mismatch between the model and NR waveforms decrease by $50\%$.
-	We further explore how different parts of the source parameter space respond
+	parts (i.e., the inspiral, merger, and ringdown) of the waveform were calibrated separately. 
+	Using \ripple, a library of waveform models compatible with automatic differentiation, we can, for the first time,
+	perform gradient-based optimization on all the waveform coefficients at the same time.
+	This joint optimization process allows us to capture previously ignored correlations between separate parts of the waveform.
+	We found that after recalibration, the median mismatch between the model and NR waveforms decreases by $50\%$.
+	We further explore how different regions of the source parameter space respond
 	to the optimization procedure. We find that the degree of improvement
@@
 
-Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
-as match filtering search \citep{owen1996search, owen1999matched} and parameter 
-estimation\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative}, rely on accurate waveform
-models. Because using numerical relativity (NR) waveforms in these tasks is
+Many data analysis tasks in gravitational wave (GW) astrophysics, such
+as match filtering~\citep{owen1996search, owen1999matched} and parameter 
+estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
+models. Since the generation of numerical relativity (NR) waveforms is
 prohibitively expensive, the community has constructed waveform approximants
 that can be evaluated much faster. There are three families of commonly used GW
-approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
-cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
-surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
+approximants: the effective-one-body (EOB) \citep{ossokine2020multipolar,
+cotesta2020frequency, taracchini2014effective}, NR
+surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2},
 and phenomenological (Phenom) models \citep{husa2016frequency,
 khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
-detail of construction of these models are different, they all have a set of
-internal parameters that can be calibrated to NR waveforms. The quality of a
-waveform model is determined by the ansatz used and the accuracy of the
+detailed construction of each model is different, they all have a set of
+internal parameters that can be calibrated to NR waveforms. The quality of the
+waveform model is therefore determined by the ansatz used and the accuracy of the
 calibrated parameters.
@@
 The LIGO-VIRGO-KAGRA (LVK) collaboration
-(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
-VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
-run on May 26, 2023. Because of the sensitivity improvement, the new run is
-expected to double the size of current binary black hole (BBH) observations
-\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
-detect individual events with a higher signal-to-noise ratio (SNR). This means
-we can resolve more features in the signal, at the same time put a more
-stringent requirement on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.
+~\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
+VIRGO:2014yos,KAGRA:2020tym} recently started their fourth observational
+run on May 26, 2023. Impressively, they are
+expected to double the total number of observed binary black holes (BBHs)~\citep{abbott2020prospects}.
+Morover, the improved sensitivity also implies that we expect to
+detect individual events with a higher signal-to-noise ratio (SNR) than ever before. This means
+we can resolve more features in the signal, therefore putting more
+stringent requirements on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.
 
@@
 
-Because of the large number of calibration parameters (often in a couple hundreds if not
-more), waveform models is usually calibrated in pieces \citep{khan2016frequency, 
+Because of the large number of calibration parameters (often a few hundred if not
+more), waveform models are usually calibrated separately for the inspiral, merger,
+and ringdown parts of the waveform \citep{khan2016frequency, 
 santamaria2010matching, pratten2021computationally}. This
 ignores the correlation between different parts of the waveform model and limits
-it quality. Recently, there has been efforts in rebuilding waveform models
-\citep{khan2016frequency} that supports automatic differentiation (AD) 
-\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}, which is a technique 
-used to compute derivatives of functions down to machine precision
-without issues of scaling up to high dimension or expression swelling. In
-particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
-user. This allows us to make use of infrastructures that are heavily used in
-machine learning, such as gradient descent and back propagation \cite{jax2018github, 
+it quality. Recently, there has been an effort to rebuild waveform models
+\citep{khan2016frequency} using programming languages that support automatic differentiation (AD) 
+\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}; AD is a technique 
+used to compute machine precision derivatives of functions 
+without the issues of scaling up to high dimension or expression swelling. In
+particular, \ripple~\citep{ripple} exposes the calibration parameters to the
+user. This allows us to make use of common techniques from
+machine learning, such as gradient descent and back propagation \citep{jax2018github, 
 pytorch, tensorflow2015}, to improve the calibration of the waveform models.
@@
 of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
-calibration coefficients given a set of NR waveforms. Training on a few NR
-waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
-waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
+calibration coefficients for given a set of NR waveforms. Using a similar set of NR
+waveforms to those used in \citep{khan2016frequency, husa2016frequency},
+we demonstrate that one can improve the match between IMRPhenomD and NR
+waveforms over a decently sized parameter space, up to a mass ratio $q=8$. We additionally
 explore how different parts of the source parameter space (e.g. the primary and
-second spins) respond to the optimization procedure, by optimizing the waveform
-separately for different regions of the parameter space. This gives us hints on
-whether the waveform model performs equally well in different regions of the
+second spins) respond to the optimization procedure by optimizing the waveform
+separately for different regions. This can help in understanding
+whether the waveform model ansatz performs equally well in different regions of the
 parameter space.
@@
 review the parameterization of the IMRPhenomD model and the mismatch function
-that is used as an objective function for the calibration, followed by outlining the
+that is used as an objective/loss function for the calibration, followed by outlining the
 specific optimization scheme used for recalibration. In
 Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
-of optimized waveforms with original waveforms. We also show how the
-optimization result differs with waveforms of different source parameters. In
-Sec.~\ref{sec:discussion}, we address the difference between our calibrating
-procedure with \citep{khan2016frequency}. We also explain how reduced spin
+of the optimized waveforms with the original waveforms. We also show how the
+optimization result differs as a function of the source parameters. In
+Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
+procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
 parameterization affects the accuracy of the model.

### 2023-06-23 src/tex/ms.tex
Working on sec 3

@@
 don't move far from the minimum.
-Finally, we stop the optimization when the validation loss stops decreasing.
+Finally, we stop the optimization when the validation loss stops decreasing 
+(see Sec.~\ref{sec:result} for description of the validation set).
 This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.
@@
 
-To evaluate how well the optimization procedure generalize to waveforms that are
+To evaluate how well the optimization procedure generalizes to waveforms that are
 not in the training set, we evaluate the mismatch between the fine-tuned model
-and the data provided in the SXS catalog for 536 NR waveforms. We select
-waveforms that share the same part of the parameter space with the training
-waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
-and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
-shows how the training and testing waveforms are distributed in the
-$q-\chi_{PN}$ space.
+and an additional 536 NR waveforms in the SXS catalog i.e., the validation set.
+We select waveforms that share the same part of the parameter space with the training
+set i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
+and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
+shows how the training and validation waveforms are distributed in the
+$q-\chi_{\rm PN}$ space.
 
-To illustrate the effect of optimiziation on an individual waveform level, we
-compare the mismatch of a particular waveform before and after optimization with
-respect to the NR waveform taken directly from the SXS catalog in Fig.
-\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
-optimized waveform has smaller residual from NR waveform both in amplitude and
-phase, particularly in the inspiral region, where the amplitude displays a
+To illustrate the effect of optimiziation on an individual waveform level, in Fig.
+\ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
+the NR waveform taken directly from the SXS catalog. In the bottom panel one can see that,
+compared to the original IMRPhenomD waveform, the optimized waveform has smaller residuals both in amplitude and
+phase, particularly in the inspiral region where the amplitude displays a
 $50\%$ reduction in error. For a fair comparison, we selected one of the testing
-waveforms from the catalog presented in \citep{khan2016frequency}.
+waveforms from the catalog presented in \citep{khan2016frequency}.\te{Which one?}
 
@@
 mind, the more relevant metric of improvement is the distribution of improvement
-in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
+in mismatch over the entire training dataset. Figure~\ref{fig:q148} shows the
 distribution of log-mismatch for the testing waveforms before and after the

### 2023-06-23 src/tex/ms.tex
Finished section 3

@@
 the simple average of mismatches and the normalized average of mismatches,  
-\begin{align}
+\begin{align}\label{eqn:loss}
 	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
@@
 $50\%$ reduction in error. For a fair comparison, we selected one of the testing
-waveforms from the catalog presented in \citep{khan2016frequency}.\te{Which one?}
+waveforms from the catalog presented in \citep{khan2016frequency}.\te{Which one? Please write the name from the table.}
 
@@
 mind, the more relevant metric of improvement is the distribution of improvement
-in mismatch over the entire training dataset. Figure~\ref{fig:q148} shows the
-distribution of log-mismatch for the testing waveforms before and after the
-optimization procedure. For generality, we use a constant PSD in our loss
-function. One can see the distribution have been skewed toward lower mismatch,
-where the peak of the distribution is shifted by approximately an order of
-magnitude, and a median mismatch is reduced by 50\%. When using
-$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
+in mismatch over the entire validation dataset. Figure~\ref{fig:q148} shows the
+distribution of log-mismatches for the validation waveforms before and after the
+optimization procedure. Here we show results using a constant PSD in our loss
+function. One can see the distribution of the optimized waveform is skewed toward lower mismatches,
+with the peak of the distribution being shifted by approximately an order of
+magnitude, and a median mismatch being reduced by 50\% (see vertical dotted lines). When using
+$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
 decrease in the median.
@@
 
-Note that the IMRPhenomD model was initially constructed and fitted using the
-{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
-methods using the {\zdethp} PSD exhibit superior improvement compared to the
-unweighted mismatch. We find no significant qualitative and quantitative
-difference between the two PSDs in the resulting distribution of mismatches.
+Note that the IMRPhenomD model was initally tested using the
+{\zdethp} weighted mismatch. We would therefore like to examine whether using the {\zdethp} PSD in our loss function
+could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
+between the results using the two PSDs in the distribution of mismatches.
 
 To understand whether additional training data can further improve the
-performance of the model, we include waveforms that are not presence in the
-original IMRPhenomD calibration in our training dataset with parameters
-tabulated in Tab. \ref{tab:q1248}. We specifically choose to use $q=2$ events
+performance of the model, we include waveforms that are not present in the
+original IMRPhenomD calibration in our training dataset; the parameters
+can be found in Tab.~\ref{tab:q1248}. We specifically choose to use $q=2$ events
 since we have abundant $q=2$ NR waveforms to validate the final result. The new
@@
 marginal improvements in the newly produced waveforms, as seen in Fig.
-\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
-distribution remains comparable to the original distribution, indicating that
+\ref{fig:q148_q1248_compare}. The high mismatch tail of the newly optimized
+distribution remains comparable to the distribution from the first optimization, indicating that
 the original dataset is sufficient for this task. Similarly, utilizing the
-{\zdethp} PSD to optimize the loss function with additional waveforms results in 
+{\zdethp} PSD in the loss function together with additional waveforms results in a
 similar level of improvement.
@@
 To investigate the performance of recalibration over the source parameter space,
-we plot the improvement of mismatch in log projected over the parameter space of
-$q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
-axis, we can see the waveforms with $q\le4$ show most consistent average
-improvement. This is because that part of the parameter space is better covered
-by the training waveform, as compared to testing waveforms from $q=4$ to $q=8$,
-where some testing waveforms lie outside the source parameter space covered by
-the training set of waveforms. On the spin axis, we can see the waveform with
-$\chi_{\textrm{PN}}$ close to zero show the most consistent improvement, and we
-start deviate from 0, there could be fluctuations in the average improvement. On
-top of that, in the $q\le4$ region, there seems to be more consistent
-improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as
-since the original waveform was developed for aligned spin system, the waveforms
-with $\chi_{\textrm{PN}}$ is less well-fitted \kw{Varify this.}, hence there
-could be more room for improvement. To show this, we plot the parameter space of
-$\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal
-axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements as discussed
+we plot the improvement of the log-mismatch as a function of the parameter space
+$q-\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. Red points indicate that the waveform
+is improved by the optimization procedure, while blue points indicate that the waveform mismatch
+\textit{increases} during optimization.
+We can see that waveforms with $q\le4$ show the most consistent average improvement.
+This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
+On the spin axis, we can see the waveforms with
+$\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
+When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
+This is particularly true in the $q\le4$ region, where we see a consistent
+improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
+Since the original waveform was developed for aligned-spin systems, the waveforms
+with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
+To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
+Waveforms along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
 above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
-($\chi_1>0>\chi_2$) regions respond to optimization differently. In the top-left
-region, waveforms generally improve slightly with along optimization. However,
-waveforms in the bottom-right region do not improve after optimization. Some
-waveforms even turned worse after optimization.
+($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
+region, waveforms generally improve slightly after optimization. However,
+waveforms in the bottom-right region do not improve after optimization.
 
@@
 source parameter space may not share the same set of optimal IMRPhenomD
-parameters, mean there could be some trade-offs in accuracy between different
+parameters.
+This would mean that there are trade-offs in accuracy between different
 parts of the parameter space. If this is the cause of the lack of improvement in
@@
 this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
-are calculated using simple average of the mismatches. The top-left and
+are calculated using simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
 bottom-right regions have limited data for $q>4$, hence the results are only
 valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
-equal-spin waveforms lying on the diagonal has great improvement. The
-improvement above the diagonal is also substantial compared to recalibrate using
-all waveforms at once, barring a few caused by some testing waveforms having
-$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
-that the optimized histogram shifts downward uniformly. On the other hand, the
-improvement below the diagonal is negligible, indicating our method struggles to
-further improve on top of the original IMRPhenomD waveform. Interestingly, we
-can see in the top-right and bottom-left quadrant, the lower right corner shows
-systematic worsening similar to the case of all waveforms. This hinted in
-general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
-anti-aligned spin where the primary spin is positive.
+equal-spin waveforms lying on the diagonal improve the most. The
+improvement above the diagonal is also significantly more than when optimizing using
+all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
+that the optimized histogram shifts to lower mismatches.\footnote{
+	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
+	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
+} On the other hand, the
+improvement below the diagonal is negligible, indicating that our method struggles to
+further improve on top of the original IMRPhenomD waveform. 
+Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
+to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
+This suggests that the IMRPhenomD waveform ansatz fits waveforms with
+$\chi_1>0$ and $\chi_2<0$ less well.
 
@@
 	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
-	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
+	(Bottom) regions of Fig.~\ref{fig:ps_q148_quadrant}. We use a constant noise spectrum to calculate mismatch and
 	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. Waveforms in the

### 2023-06-23 src/tex/ms.tex
Update

--- 2023-06-22 src/tex/ms.tex
@@
 \begin{abstract}
-	We investigate the possibility of improving the accuracy of the
+	We investigate the possibility of improving the accuracy of a
 	phenomenological waveform model, IMRPhenomD, by jointly optimizing all the
@@
 	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
-	parts (i.e., the inspiral, merger, and ringdown) of the waveform were calibrated separately. 
-	Using \ripple, a library of waveform models compatible with automatic differentiation, we can, for the first time,
-	perform gradient-based optimization on all the waveform coefficients at the same time.
-	This joint optimization process allows us to capture previously ignored correlations between separate parts of the waveform.
-	We found that after recalibration, the median mismatch between the model and NR waveforms decreases by $50\%$.
-	We further explore how different regions of the source parameter space respond
+	parts of the waveform were calibrated separately. Using \ripple, a library
+	of waveform models compatible with automatic differentiation, we can perform
+	gradients-based optimization to all the waveform coefficients the same time,
+	which should improve the quality of waveform by capturing correlations
+	between previous separately fitted parts. We found that after recalibration,
+	the median mismatch between the model and NR waveforms decrease by $50\%$.
+	We further explore how different parts of the source parameter space respond
 	to the optimization procedure. We find that the degree of improvement
@@
 
-Many data analysis tasks in gravitational wave (GW) astrophysics, such
-as match filtering~\citep{owen1996search, owen1999matched} and parameter 
-estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
-models. Since the generation of numerical relativity (NR) waveforms is
+Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
+as match filtering search \citep{fill} and parameter estimation\citep{fill}, rely on accurate waveform
+models. Because using numerical relativity (NR) waveforms in these tasks is
 prohibitively expensive, the community has constructed waveform approximants
 that can be evaluated much faster. There are three families of commonly used GW
-approximants: the effective-one-body (EOB) \citep{ossokine2020multipolar,
-cotesta2020frequency, taracchini2014effective}, NR
-surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2},
+approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
+cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
+surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
 and phenomenological (Phenom) models \citep{husa2016frequency,
 khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
-detailed construction of each model is different, they all have a set of
-internal parameters that can be calibrated to NR waveforms. The quality of the
-waveform model is therefore determined by the ansatz used and the accuracy of the
+detail of construction of these models are different, they all have a set of
+internal parameters that can be calibrated to NR waveforms. The quality of a
+waveform model is determined by the ansatz used and the accuracy of the
 calibrated parameters.
@@
 The LIGO-VIRGO-KAGRA (LVK) collaboration
-~\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
-VIRGO:2014yos,KAGRA:2020tym} recently started their fourth observational
-run on May 26, 2023. Impressively, they are
-expected to double the total number of observed binary black holes (BBHs)~\citep{abbott2020prospects}.
-Morover, the improved sensitivity also implies that we expect to
-detect individual events with a higher signal-to-noise ratio (SNR) than ever before. This means
-we can resolve more features in the signal, therefore putting more
-stringent requirements on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.
+(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
+VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
+run on May 26, 2023. Because of the sensitivity improvement, the new run is
+expected to double the size of current binary black hole (BBH) observations
+\citep{abbott2020prospects}. The improved sensitivity also implies we expect to
+detect individual events with a higher signal-to-noise ratio (SNR). This means
+we can resolve more features in the signal, at the same time put a more
+stringent requirement on the accuracy of our waveform model\citep{fill}.
 
@@
 
-Because of the large number of calibration parameters (often a few hundred if not
-more), waveform models are usually calibrated separately for the inspiral, merger,
-and ringdown parts of the waveform \citep{khan2016frequency, 
-santamaria2010matching, pratten2021computationally}. This
+Because of the large number of calibration parameters (often in a couple hundreds if not
+more), waveform models is usually calibrated in pieces \citep{fill}. This
 ignores the correlation between different parts of the waveform model and limits
-it quality. Recently, there has been an effort to rebuild waveform models
-\citep{khan2016frequency} using programming languages that support automatic differentiation (AD) 
-\citep{ripple, Iacovelli:2022bbs, Iacovelli:2022mbg, Coogan:2022qxs}; AD is a technique 
-used to compute machine precision derivatives of functions 
-without the issues of scaling up to high dimension or expression swelling. In
-particular, \ripple~\citep{ripple} exposes the calibration parameters to the
-user. This allows us to make use of common techniques from
-machine learning, such as gradient descent and back propagation \citep{jax2018github, 
-pytorch, tensorflow2015}, to improve the calibration of the waveform models.
+it quality. Recently, there has been efforts in rebuilding waveform models
+\citep{fill} that supports automatic differentiation (AD)\citep{fill}, which is
+a technique used to compute derivatives of functions down to machine precision
+without issues of scaling up to high dimension or expression swelling. In
+particular, {\ripple} \citep{ripple} exposes the calibration parameters to the
+user. This allows us to make use of infrastructures that are heavily used in
+machine learning, such as gradient descent and back propagation\cite{fill}, to
+improve the calibration of the waveform models.
 
@@
 In this paper, we investigate the possibility of further improving the accuracy
-of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
-calibration coefficients for given a set of NR waveforms. Using a similar set of NR
-waveforms to those used in \citep{khan2016frequency, husa2016frequency},
-we demonstrate that one can improve the match between IMRPhenomD and NR
-waveforms over a decently sized parameter space, up to a mass ratio $q=8$. We additionally
+of a waveform model, IMRPhenomD\citep{fill}, by jointly optimizing all the
+calibration coefficients given a set of NR waveforms. Training on a few NR
+waveforms, we demonstrate one can improve the match between IMRPhenomD and NR
+waveforms over a decently sized parameter space, up to mass ratio $q=8$ We also
 explore how different parts of the source parameter space (e.g. the primary and
-second spins) respond to the optimization procedure by optimizing the waveform
-separately for different regions. This can help in understanding
-whether the waveform model ansatz performs equally well in different regions of the
+second spins) respond to the optimization procedure, by optimizing the waveform
+separately for different regions of the parameter space. This gives us hints on
+whether the waveform model performs equally well in different regions of the
 parameter space.
@@
 review the parameterization of the IMRPhenomD model and the mismatch function
-that is used as an objective/loss function for the calibration, followed by outlining the
+that is used as an objective function for the calibration, followed by outlining the
 specific optimization scheme used for recalibration. In
 Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
-of the optimized waveforms with the original waveforms. We also show how the
-optimization result differs as a function of the source parameters. In
-Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
-procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
+of optimized waveforms with original waveforms. We also show how the
+optimization result differs with waveforms of different source parameters. In
+Sec.~\ref{sec:discussion}, we address the difference between our calibrating
+procedure with \citep{khan2016frequency}. We also explain how reduced spin
 parameterization affects the accuracy of the model.
@@
 In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
-Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
+Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
 \begin{align} \label{eq:Lambda}
@@
 %\kw{Need to work on this paragraph}
-Although initially independent, the stitching procedure means that each section
+Although initially independent, the stiching procedure means that each section
 of the waveform intrinsically depends on the full set of $\lambda$'s. 
@@
 Importantly, since the fitting was performed on the individual segments,
-the final waveform is not guaranteed to have $\lambda$'s close to global minima.
+the final waveform is not guarenteed to have $\lambda$'s close to global minima.
 
@@
 This is made possible by the use of gradient-based optimization algorithms,
-enabled by AD from \jax\, and {\ripple}, which are significantly more efficient in high dimensions.
+enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
 % , gradients of IMRPhenomD can be easily obtained, thus
@@
 In order to optimize the coefficients, we need to define a loss function that
-quantifies the difference between the waveform model and the target NR simulations which we want to match.
-Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
+quantifies the difference between waveform model and the target NR simulations which we want to match.
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{husa2016frequency}. 
 It is defined as
@@
 where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
-$S_n(f)$ is the power spectral density (PSD), and $f_{\mathrm{max}}$ ($f_{\mathrm{min}}$) is
-the maximum (minimum) frequency for the integration.
+$S_n(f)$ is the power spectral density (PSD), $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
+the relevant maximum and minimum frequencies for the integration.
 We note here that the mismatch can be viewed as the mean square error (MSE) between the
@@
 
-Since we wish to optimize the model over the whole parameter space,
-we need to compare multiple model generated waveforms with NR waveforms.
-However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
-We therefore define the loss function as an average of training waveforms in two ways,
-the simple average of mismatches and the normalized average of mismatches,  
+\te{Up to here}
+
+Since we wish to optimize the model over the whole parameter space, we need to compare multiple model generated waveforms with NR waveforms.
+However, mismatch only quantifies the difference between IMRPhenomD and NR waveform for one particular set of intrinsic parameters.
+To take into account of various different waveforms in the parameter space, we pick waveforms from the different parts of the parameter space.
+We define the loss function as an average of training waveforms in two ways, the simple average of mismatches and the normalize average of mismatches,
 \begin{align}
-	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+	\mathcal{L}_{\mathrm{mean}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
+	\mathcal{L}_{\mathrm{fl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
 \end{align}	
@@
 $\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of training waveforms.
-
-The two loss functions are chosen to give different 
-behaviour during the optimization process.
-The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
-but is prone to be dominated by a single point in parameter space with a large mismatch.
-Other points with smaller mismatches would be insignificant comparatively, and might not be able 
-to improve under such a loss function.
-Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
-by encouraging the waveform to improve at each training point at a similar rate.
-The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
-Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
-In this paper, we show the results of using both loss functions 
+training waveform, and $N$ is the total number of individual training waveforms.
+Note that we choose to use two different averages, since they have different 
+preferences in optimization base on waveform mismatches. For the first choice, 
+simple average serves as the simplest choice of loss function, but is prone to 
+be dominated by a single waveform with a large mismatch. Other waveforms with 
+smaller mismatches would be insignificant comparatively, and might not be able 
+to improve under such optimization. Alternatively, the second choice, normalized 
+average eliminates the aforementioned issue. Nevertheless, it excludes the
+information on initial mismatches. $\mathcal{L}_{\mathrm{fl}}$ restricts every
+training waveform to decrease at similar rates, hence it is hard to obtain
+optimized waveforms with mismatches in the same order of magnitude. Instead,
+their ratios in mismatches would remain approximately the same. Conversely,
+$\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+and individual mismatches would be in a similar order of magnitude after 
+optimization. In this paper, we showcase the results of using both loss functions 
 and examine the differences between them.
@@
 
-To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
-Originally, 19 NR simulations  were used to calibrate IMRPhenomD \citep{khan2016frequency}; 9 of these 
-are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
-As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
-Instead, we take the same nine waveforms from the SXS catalog plus two additional waveforms which closely mimic
-two of the low mass ratio BAM simulations. 
-The remaining BAM simulations have no close SXS counterparts, and are therefore not included in the training set.
-Our main results therefore uses 11 NR waveforms for calibration which are listed in Tab.~\ref{tab:q148}.
-Additional NR waveforms which are used for further calibration are listed in Tab.~\ref{tab:q1248}.
+To compute the loss functions, we have to take NR waveforms for calculating the mismatch. We choose 11-16 NR waveforms from the set of waveforms used in the original
+calibration process as training waveforms. Originally, 19 waveforms are taken from 
+NR simulations for calibrating IMRPhenomD \citep{khan2016frequency}, 
+which are waveforms from the SXS catalog \citep{boyle2019sxs} or BAM simulation. 
+As BAM waveforms are not publicly available, we cannot take the identical training 
+set as them.
+Instead, we take the available waveforms from the SXS catalog to construct our loss function.
+Training waveforms used are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}.
+The training waveforms chosen has maximum mass ratio to be 8.
+This is because SXS catalog does not have NR waveforms with extremely high mass ratio.
+In fact, the SXS catalog only has NR waveforms with $q\leq10$.
+Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed from LIGO have $q\leq8$.
+Hence, we calibrate IMRPhenomD with waveforms of $q\leq8$.
 
-The training set has a maximum mass ratio of eight due to the lack 
-of high mass ratio simulations in the SXS catalog.
-In fact, the SXS catalog only has NR waveforms with $q\leq10$.
-Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$,
-as most BBH events observed by LIGO and Virgo have $q\leq8$.
+In the SXS catalog, NR waveforms are in the form of time-series strain.
+Since time-series data is oscillatory, performing optimization in the time-domain is not ideal.
+Hence, we transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the
+same intrinsic parameters. We taper the time-series using Tukey window.
+\footnote{ Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
+	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
+	entire GW strain. } 
+Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series.
 
-The SXS NR waveforms are given as time-series strain.
-Since IMRPhenomD is modeled in the frequency domain, we need to Fourier transform the simulation results
-in order to compute the mismatch, \eqref{eq:mismatch}.
-For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
-	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
-	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
-} before using standard FFT routines to compute the Fourier transform.
-
-In addition to choosing the NR waveforms for the training set,
-one needs to choose a relevant PSD for mismatch.
-We have opted to use a flat PSD for the mismatch calculation,
-as it provides results that are independent of the detector sensitivity and mass scale.
-The use of a flat PSD ensures that the improvement in accuracy is due mainly
-to the difference in high-dimensional fitting.
-Additionally, we are interested in examining the effect of introducing a detector PSD
-on the optimization process. For this, we have chosen the
-zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
-Since the total mass of the system scales with the frequency of the waveform,
-we must choose a corresponding mass scale to match the frequency range of our noise PSD.
-To demonstrate the effect of introducing a detector specific PSD,
-we selected an arbitrary mass scale of $M=50M_{\odot}$, as binaries
-of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.
+%\te{I think this discussion needs to be much more precise.
+Having a waveform model calibrated to particular noise curve could make sense, but its inherently different to just fitting the phase and the amplitude separately like they do in the original paper.
+So I would recommend that our default is to use a flat PSD and then in addition discuss the one with the PSD.
+This way, I think we demonstrate more clearly that its the high dimensional fitting that helps, not the just changing the metric used when fitting.}
+Other than NR waveforms, one need to choose a relevant noise PSD for 
+mismatch.
+We have opted to use a flat PSD for this purpose, as it provides results that are independent of the detector sensitivity and mass scale.
+The use of a flat PSD ensures that the improvement in accuracy is due mainly to the difference in high-dimensional fitting and piece-wise fitting, but not due to the use of a different mass scale.
+Furthermore, we are interested in examining the effect of introducing a detector PSD on the optimization process.
+For this, we have chosen the zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
+Since the total mass of the system scales with the frequency of the waveform, we must choose a corresponding mass scale to match the frequency range of our noise PSD.
+We selected an arbitrary mass scale of $M=50M_{\odot}$ for all waveforms for demonstration, as this is a commonly observed mass scale in LIGO observations.
 %\te{This choice seems pretty arbitrary.
@@
 
-We point out that our treatment of NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
-In the original calibration process, the training waveforms are hybrids of NR and SpinAlignedEOB (SEOB) waveforms.
-The low frequency inspiral part is taken from the SEOB
-waveforms while the rest is taken from NR simulations.
-Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and testing)
-have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, to contain part of the inspiral segment and all
-merger and ringdown frequency information. We use the frequency limits
+We point out that our treatment to NR waveforms is different from that of \citep{husa2016frequency, khan2016frequency}.
+In the original calibration process, training waveforms are hybrid waveforms of NR and SpinAlignedEOB
+(SEOB) waveforms. The low frequency inspiral part is taken from the SEOB
+waveforms while the rest of the waveforms are taken from NR simulations.
+Instead, we solely use NR waveforms for comparison, since we are only exploring the possibility of optimizing waveform models.
+Thus, for simplicity sake, we ignore this procedure.
+On the other hand, most NR waveforms used (for both training and testing) have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in
+which they are long enough to contain part of the inspiral segment and all
+merger and ringdown frequency information. We take the frequency limits as
 $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$,	
 where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of
-the IMRPhenomD's frequency range, except the minimum frequency is set slightly higher
-than in the original calibration due to the NR simulation length.
-When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum
-frequency \citep{santamaria2010matching}.
-We use the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
-which is sufficient to capture all features of the GW strain.
+the IMRPhenomD's frequency range, except the minimum frequency is set higher
+than that in the original calibration due to NR length. When compared with
+IMRPhenomC, the frequency range is slightly extended to have a higher maximum
+frequency. We have the dimensionless frequency spacing $M\Delta f=2.5\times10^{-6}$, 
+which is sufficient to capture all features of GW strain.
 
@@
 coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
-original coefficients given in \citep{khan2016frequency} because they likely lie in the
-neighborhood of the minimum that we wish to find. 
-We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
-don't move far from the minimum.
-Finally, we stop the optimization when the validation loss stops decreasing.
-This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.
+original coefficients given in \citep{khan2016frequency}. We take them as the initial 
+waveform coefficients because they lie in the neighborhood of the minimum that we wish 
+to find. Then, by taking $\alpha=10^{-6}$, we see the validation loss in 
+Fig.~\ref{fig:loss} plateau at around $N=12000$, hence we end the optimization here.  
+%\te{I think we need a bit more discussion about the stopping criterion here.
+What is N, and why did we choose it to be this number?
+Ideally we would actually have a plot of the loss function during training.
+I don't have a good intuition for if its noisy.
+Does it plateau?
+If we wanted to ensure that people don't complain about overfitting we could also plot the loss of the validation set to show its not going up.}
 
@@
 	here, 9 of them are also used in the original IMRPhenomD calibration.
-	\citep{khan2016frequency}}
+	\citep{khan2016frequency} The two remaining waveforms were from BAM
+	simulation, to which we do not have access.}
 	\label{tab:q148}
@@
 magnitude, and a median mismatch is reduced by 50\%. When using
-$\mathcal{L}_{\mathrm{nl}}$, we observe a similar improvement with a 22.9\%
+$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
 decrease in the median.
@@
 	\caption{Distributions of waveform mismatches calculated using 
-	$\mathcal{L}_{\mathrm{ave}}$. We use training waveforms listed in 
+	$\mathcal{L}_{\mathrm{mean}}$. We use training waveforms listed in 
 	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
@@
 	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
+	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{mean}}$. 
 	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
@@
 	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
-	We show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with constant PSD 
+	We show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with constant PSD 
 	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
@@
 	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
-	show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with the
+	show the result from optimizing $\mathcal{L}_{\mathrm{mean}}$ with the
 	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
@@
 (Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
-this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
-are calculated using simple average of the mismatches. The top-left and
+this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
 bottom-right regions have limited data for $q>4$, hence the results are only
@@
 	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. Waveforms in the
+	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
 	top-left region generally improves while waveforms in the bottom-right
@@
 \section{Discussion} \label{sec:discussion}
+
+The results in this work show a promising way to understand and improve the
+accuracy of phenomenological waveform models by jointly optimizing all
+coefficients at once. There are a number of caveats multiple ways to improve the
+current work.
+
+While this study focuses on the IMRPhenomD model because its availability as an
+automatically differentiable waveform, other more sophisticate waveforms such as
+IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} and IMRPhenomXAS
+\citep{pratten2020setting,pratten2021computationally} can potentially benefit
+from recalibration even more, because of the larger number of calibration
+parameters. % Extend this a bit?
+
+The correlation we observe between the mismatches and the waveform's parameters
+can also be used to design more physical ansatz. For example, we see the
+recalibration process struggles to improve in the regime where $\chi_{1}$ is
+positive and $\chi_{2}$ is negative, this means there is a dependency between
+the two spins that is not captured by the ansatz. A systematic scheme that
+incorporate this recalibration scheme with waveform design will be handy in
+constructing new waveform models in the future. In making figure
+\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions
+because of simplicity. These simple cuts demonstrate the accuracy of the
+waveform can be further improved, but they are almost certainly not the optimal
+way to incorporate the extra information we have about the waveform. A more
+general way such as adding new functional dependency between the parameters
+should be explored in the future.
+
+We use mismatch to quantify the accuracy of the waveform model in this work. It
+is a natural extension to investigate how would the recalibration affects
+downstream analysis such as parameter estimation and population analysis.
+Whether it is possible to reduce the systematic error in parameter estimation
+using recalibrated waveforms is an interesting question we will investigate in
+the sooner future. On top of this, recently relatively more differentiable
+sampler such as \citep{Wong:2022xvh} has been introduced to the community, one
+can also optimize the waveform directly using metrics from parameter estimation,
+such as bias in an injection-recovery run. This will be an example of
+more direct and efficient optimization of waveform models for downstream analysis.
+
+\kw{End here.}
 
@@
 
-Although our study primarily focused on the IMRPhenomD model, this simple yet
-versatile approach can be applied to other differentiable GW models, such as the
-IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
-models within the same family. By jointly optimizing a new set of coefficients,
-it is expected that both models can be enhanced since they share similar
-construction principles to the IMRPhenomD model. For instance, they also use PN
-approximants as part of the ansatz in the inspiral segment. It will be interesting 
-to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
-parameterized by an additional anti-symmetric spin parameter, it is expected not
-to exhibit the degeneracy previously described. With the currently developing 
-{\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
-may provide valuable insights into the systematics of the Phenom models.
-Furthermore, this approach is applicable to other GW model families, such as NR
-surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
-simplify NR waveform calibration procedures and lead to the improvement of 
-existing models.
+% \section{Conclusion} \label{sec:conclusion}
 
-% Moved down from intro
-% In particular, we find it is easier IMRPhenomD
-% favors certain parts of the parameter space. This means IMRPhenomD introduces
-% systematic bias to other GW analysis tasks. This showcases the flaws of the
-% ansatz and allows us to have a deeper understanding of Phenom models.
+% In this paper, we have presented a systematic method to recalibrate GW models.
+% This method utilizes {\jax}'s automatic differentiation to apply
+% derivative-based optimization to recalibrate GW models jointly. Using the new
+% implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
+% conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
+% coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
+% improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
+% mismatches have a slightly better improvement. In contrast, different types of
+% loss function result in significantly different final mismatch distributions. As 
+% seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
+% $\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
+% slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
 
-\section{Conclusion} \label{sec:conclusion}
+% Furthermore, we investigated how the source parameters affect the
+% improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
+% certain preference for waveforms lying in the top-left region while the bottom-right 
+% region hardly improved. To further test this result, we recalibrate
+% waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
+% we can see that this recalibration process gives further
+% improvement to the top-left region while the bottom-right region only have little 
+% improvement. This indicates that the ansatz has limited match in this region and 
+% does not fit most waveforms in this region, hence introduces bias to downstream GW 
+% analysis. This phenomenon is mainly due to the reduced spin approximation used in 
+% parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
 
-In this paper, we have presented a systematic method to recalibrate GW models.
-This method utilizes {\jax}'s automatic differentiation to apply
-derivative-based optimization to recalibrate GW models jointly. Using the new
-implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
-conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
-coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
-improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
-mismatches have a slightly better improvement. In contrast, different types of
-loss function result in significantly different final mismatch distributions. As 
-seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{ave}}$ outperforms 
-$\mathcal{L}_{\mathrm{nl}}$. By increasing the number of training waveforms, we see a 
-slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
-
-Furthermore, we investigated how the source parameters affect the
-improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-certain preference for waveforms lying in the top-left region while the bottom-right 
-region hardly improved. To further test this result, we recalibrate
-waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
-we can see that this recalibration process gives further
-improvement to the top-left region while the bottom-right region only have little 
-improvement. This indicates that the ansatz has limited match in this region and 
-does not fit most waveforms in this region, hence introduces bias to downstream GW 
-analysis. This phenomenon is mainly due to the reduced spin approximation used in 
-parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
-
-While we naively separate the optimization process into 4 regions, one can
-perform systematic region-selection. In principle, we can apply this general
-method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
-models. Then, we can perform all the above analyses to understand how to
-construct better GW Phenom models in the future.
+% While we naively separate the optimization process into 4 regions, one can
+% perform systematic region-selection. In principle, we can apply this general
+% method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
+% models. Then, we can perform all the above analyses to understand how to
+% construct better GW Phenom models in the future.
 

### 2023-06-23 src/tex/ms.tex
Finished section 3

--- 2023-06-23 src/tex/ms.tex
@@
 with the peak of the distribution being shifted by approximately an order of
-magnitude, and a median mismatch being reduced by 50\% (see vertical dotted lines). When using
+magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
 $\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
@@
 
-Note that the IMRPhenomD model was initally tested using the
-{\zdethp} weighted mismatch. We would therefore like to examine whether using the {\zdethp} PSD in our loss function
+Note that the performace of the IMRPhenomD model was initally tested using the
+{\zdethp} weighted mismatch.\footnote{
+	Note, however, that the mismatch was never directly used during the calibration process~\citep{khan2016frequency, khan2019phenomenological}.
+} We would therefore like to examine whether using the {\zdethp} PSD in our loss function
 could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
@@
 This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
-On the spin axis, we can see the waveforms with
+On the spin axis, we can see that the waveforms with
 $\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
@@
 improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
-Since the original waveform was developed for aligned-spin systems, the waveforms
+Since the original model was developed for aligned-spin systems, NR waveforms
 with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
 To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
-Waveforms along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
+Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
 above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
 ($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
-region, waveforms generally improve slightly after optimization. However,
-waveforms in the bottom-right region do not improve after optimization.
+region, the waveform generally improves after optimization. However,
+in the bottom-right region, the waveform does not improve after optimization.
 
@@
 this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
-are calculated using simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
+are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
 bottom-right regions have limited data for $q>4$, hence the results are only
@@
 improvement below the diagonal is negligible, indicating that our method struggles to
-further improve on top of the original IMRPhenomD waveform. 
+further improve over the original IMRPhenomD waveform. 
 Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change

### 2023-06-23 src/tex/ms.tex
Finished with captions

@@
 
-% More precisely, $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
-% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
-% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
-% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
-% the correlations between different segments are omitted, which could limit the 
-% accuracy of the model. Also, since fitting was performed before connecting 
-% individual segments, the final waveform does not guarantee to achieve the optimal 
-% waveform.
-% Finally, the individual segments are first connected directly using 
-% step functions. Then, by fixing coefficients in the intermediate segment, one can make 
-% the final waveform is continuous in its first derivative.
-
-% In order to recalibrate the model, we have to understand what parameters the model has.
-% Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
-% For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.
-
-% The IMRPhenomD model is constructed by combining three individually fitted parts
-% into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
-% \begin{align}\label{eq:joint_waveform}
-% 	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
-% \end{align}
-% Instead of fitting the strain, which is a highly oscillatory function that is
-% difficult to fit, the amplitude and phase are fitted since they are smoother
-% functions. In each part, the amplitude and phase are made using simple functions
-% of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
-% amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
-% \begin{equation}\label{eq:amplitude}
-% \begin{aligned}
-% 	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
-% 	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
-% 	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
-% 	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],
-% \end{aligned}
-% \end{equation}
-% where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.
-
 %\kw{Need to work on this paragraph}
@@
 
-% Generally, waveform coefficients 
-% are obtained by calibrating with NR waveforms, which are waveforms computed 
-% using NR simulations. In the case of \citep{khan2016frequency}, they first 
-% obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
-% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
-% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
-% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
-% the correlations between different segments are omitted, which could limit the 
-% accuracy of the model. Also, since fitting was performed before connecting 
-% individual segments, the final waveform does not guarantee to achieve the optimal 
-% waveform. The connecting procedure alters the previously fitted waveform. 
-% Hence, the model generated waveforms contains additional inaccuracies.
-
 At the time of construction this piece-wise approach was necessary since
@@
 enabled by AD from \jax\, and {\ripple}, which are significantly more efficient in high dimensions.
-% , gradients of IMRPhenomD can be easily obtained, thus
-% allowing the use of gradient-based algorithms for us to recalibrate the model.
 
@@
 of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.
-%\te{This choice seems pretty arbitrary.
-Looking at Fig 8 of 2111.03606, they seem to cluster closer to 60-70.
-This choice is important and should be more informed.}We will then examine the differences in optimization.
-
-%Other than NR waveforms, one need to choose a relevant noise PSD for 
-%mismatch. Since a common goal for creating waveform approximant is to
-%deploy the waveform model in downstream data analysis tasks, a reasonable choice
-%for the noise PSD should be closely related to the instrument of interest. We
-%choose the zero-detuned high-power (\zdethp) noise PSD as our
-%first choice of noise PSD \citep{aasi2015advanced}. Because the total mass of the system simple scales
-%the frequency of the waveform, we need to choose a corresponding mass scale to
-%match the frequency range of our noise PSD. Here, we choose $M=50M_{\odot}$ for
-%all waveforms, since this mass scale is common in LIGO observations. In addition
-%to a detector-dependent PSD, we are also interested in a more
-%instrument-agnostic case such that we can evaluate the effectiveness of the
-%recalibration without considering detector effect. In this case, we set the
-%noise PSD to be a constant value.
 
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{Loss functions against number of iterations. We take the set of coefficients at the minimum of the validation loss.}
+	\caption{Loss functions against the number of iterations.
+	The vertical dotted line indicates minimum of the validation loss,
+	where we stop the optimization and obtain a new set of waveform coefficients, $\lambda$.}
 	\label{fig:loss}
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Distribution of training and testing waveforms in $q-\chi_{PN}$
-	space. The blue dots represent the waveforms used in training and the orange
-	stars represent the waveforms used in testing.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
+	\caption{Distribution of training and validation NR waveforms in the $q-\chi_{PN}$
+	space. The blue stars represent the waveforms used to compute the training loss and the orange
+	dots represent the waveforms used for validation.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
 	\label{fig:intrin_space}
@@
 	\end{tabularx}
-	\caption{List of waveforms used to recalibrate the model. The mass ratio
-	$q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed
-	here, 9 of them are also used in the original IMRPhenomD calibration.
-	\citep{khan2016frequency}}
+	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
+	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
+	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
+	are SXS analogues to two of the BAM NR simulations used in the original calibration
+	\citep{khan2016frequency}.}
 	\label{tab:q148}
@@
 	\end{tabularx}
-	\caption{Additional waveforms used in further recalibration.}
+	\caption{Additional NR waveforms used in further recalibration.}
 	\label{tab:q1248}
@@
 	\caption{Comparison between original and optimized IMRPhenomD waveforms.
-	Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and
-	$\chi_1=\chi_2=-0.8$. The original mismatch is around $2.8\times10^{-4}$ and
-	the optimized mismatch is around $5.3\times10^{-5}$. Top: It shows the
-	amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD
-	waveform. Bottom: It shows the relative error of amplitudes between NR and
-	IMRPhenomD waveforms, and the absolute error of phases between NR and
-	IMRPhenomD waveforms}
+	Here we show the SXS:BBH:0154 NR waveform, which has a mass ratio of $q=1$ and spins
+	$\chi_1=\chi_2=-0.8$. The original mismatch is $2.8\times10^{-4}$ and
+	the optimized mismatch is $5.3\times10^{-5}$. \textit{Top}: Here we show the
+	amplitude (left) and phase (right) of the NR, original IMRPhenomD, and optimized IMRPhenomD
+	waveform. \textit{Bottom}: Here we show the relative error between the NR and
+	IMRPhenomD waveform ampltiudes (left) as well as the absolute error of the phases between the NR and
+	IMRPhenomD waveforms (right).}
 	\label{fig:0154}
@@
 phase, particularly in the inspiral region where the amplitude displays a
-$50\%$ reduction in error. For a fair comparison, we selected one of the testing
-waveforms from the catalog presented in \citep{khan2016frequency}.\te{Which one? Please write the name from the table.}
+$50\%$ reduction in error. For a fair comparison, we selected the SXS:BBH:0154 NR waveform
+which was also used in \citep{khan2016frequency} to validate the original waveform.
 
@@
 	\caption{Distributions of waveform mismatches calculated using 
-	$\mathcal{L}_{\mathrm{ave}}$. We use training waveforms listed in 
-	Tab.~\ref{tab:q148} and	mismatches are weighted with the constant PSD. 
-	Dashed lines represent the median of the distributions, which has decreased by 45.3\%.}
+	$\mathcal{L}_{\mathrm{ave}}$. We use the training waveforms listed in 
+	Tab.~\ref{tab:q148} and	mismatches are weighted with a constant PSD. 
+	The dashed lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
 	\label{fig:q148}
@@
 	\includegraphics[width=\columnwidth]{figures/q148_q1248_compare.pdf}
-	\caption{Distributions of $\log_{10}$ difference in mismatch. The
-	distribution labeled \texttt{q148} uses training waveforms listed in
-	Tab.~\ref{tab:q148} while the \texttt{q1248} distribution uses waveforms
+	\caption{Distributions of $\log_{10}$ difference in mismatch. The red
+	distribution uses the training waveforms listed in
+	Tab.~\ref{tab:q148} while the green distribution uses waveforms
 	listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Mismatches are calculated
-	using the constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
-	Dashed lines represent the median of the distributions, which has decreased by 10.4\%.}
+	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
+	The dashed lines represent the median of the distributions, which decreased by 10.4\% during optimization.}
 	\label{fig:q148_q1248_compare}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
-	\caption{Parameter space of testing waveforms with $q$ against $\chi_{\mathrm{PN}}$. 
-	We show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with constant PSD 
+	\caption{Fractional mismatch change of the validation waveforms in the $q-\chi_{\mathrm{PN}}$ plane. 
+	We show results for the $\mathcal{L}_{\mathrm{ave}}$ loss function with a constant PSD 
 	and training waveforms in Tab.~\ref{tab:q148}. Here, the colorbar represents the 
-	$\log_{10}$ difference between optimized and original mismatches.}
+	$\log_{10}$ difference between optimized and original mismatches. Red points indicate that the waveform
+	is improved by the optimization procedure, while blue points indicate that the waveform mismatch
+	\textit{increases} during optimization.}
 	\label{fig:ps_q148_qchi}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_chi1chi2.pdf}
-	\caption{Parameter space of testing waveforms with $\chi_1$ against $\chi_2$. We
-	show the result from optimizing $\mathcal{L}_{\mathrm{ave}}$ with the
+	\caption{Fractional mismatch change of the validation waveforms in the $\chi_1-\chi_2$ plane. We
+	show results for the $\mathcal{L}_{\mathrm{ave}}$ loss function with a
 	constant PSD and training waveforms listed in Tab.~\ref{tab:q148}.}
@@
 	\end{tabularx}
-	\caption{List of waveforms used in recalibrating coefficients in 4 regions.
-From top to down are the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$), bottom-left ($\chi_1,\chi_2<0$) and bottom-right ($\chi_1>0>\chi_2$) regions.
-Note that for the top-right and bottom-left regions, waveforms are chosen to have $\chi_1\approx\chi_2$, while the training waveforms for the other two regions are chosen to have $\chi_1\approx-\chi_2$.}
+	\caption{List of NR waveforms used in recalibrating the coefficients in the four $\chi_1-\chi_2$ quadrants.
+	From top to bottom the lines denote the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$),
+	bottom-left ($\chi_1,\chi_2<0$), and bottom-right ($\chi_1>0>\chi_2$) regions.
+	Note that for the top-right and bottom-left regions, waveforms are chosen to have $\chi_1\approx\chi_2$,
+	while the training waveforms for the other two regions are chosen to have $\chi_1\approx-\chi_2$.}
 	\label{tab:quadrants}
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
-	(Bottom) regions of Fig.~\ref{fig:ps_q148_quadrant}. We use a constant noise spectrum to calculate mismatch and
-	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. Waveforms in the
-	top-left region generally improves while waveforms in the bottom-right
-	region has very little improvement, as indicated by the median (dashed
+	\caption{Distributions of mismatches in the top-left (top) and bottom-right
+	(bottom) regions of Fig.~\ref{fig:ps_q148_quadrant}. We use a constant PSD to calculate the mismatch and
+	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The waveform generally improves in the
+	top-left region while it shows very little improvement in the bottom-right,
+	as indicated by the change of the medians of each distribution (dashed
 	lines). \kw{Is the original referring to original IMRPhenomD or optimized
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Parameter space of testing waveforms. Each region is fitted independently with waveforms listed in Tab.~\ref{tab:quadrants}.}
+	\caption{Fractional mismatch change of the validation waveforms.
+	Each quadrant is fitted independently with to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
 	\label{fig:ps_q148_quadrant}

### 2023-06-24 src/tex/ms.tex

@@
 
+The results in this work show a promising way to understand and improve the
+accuracy of phenomenological waveform models by jointly optimizing all
+coefficients at once. There are a number of caveats multiple ways to improve the
+current work.
+
+While this study focuses on the IMRPhenomD model because its availability as an
+automatically differentiable waveform, other more sophisticate waveforms such as
+IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} and IMRPhenomXAS
+\citep{pratten2020setting,pratten2021computationally} can potentially benefit
+from recalibration even more, because of the larger number of calibration
+parameters. % Extend this a bit?
+
+The correlation we observe between the mismatches and the waveform's parameters
+can also be used to design more physical ansatz. For example, we see the
+recalibration process struggles to improve in the regime where $\chi_{1}$ is
+positive and $\chi_{2}$ is negative, this means there is a dependency between
+the two spins that is not captured by the ansatz. A systematic scheme that
+incorporate this recalibration scheme with waveform design will be handy in
+constructing new waveform models in the future. In making figure
+\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions
+because of simplicity. These simple cuts demonstrate the accuracy of the
+waveform can be further improved, but they are almost certainly not the optimal
+way to incorporate the extra information we have about the waveform. A more
+general way such as adding new functional dependency between the parameters
+should be explored in the future.
+
+We use mismatch to quantify the accuracy of the waveform model in this work. It
+is a natural extension to investigate how would the recalibration affects
+downstream analysis such as parameter estimation and population analysis.
+Whether it is possible to reduce the systematic error in parameter estimation
+using recalibrated waveforms is an interesting question we will investigate in
+the sooner future. On top of this, recently relatively more differentiable
+sampler such as \citep{Wong:2022xvh} has been introduced to the community, one
+can also optimize the waveform directly using metrics from parameter estimation,
+such as bias in an injection-recovery run. This will be an example of
+more direct and efficient optimization of waveform models for downstream analysis.
+
+\kw{End here.}
+
 We have shown the result of recalibrating waveform coefficients. One thing to
@@
 
-Although our study primarily focused on the IMRPhenomD model, this simple yet
-versatile approach can be applied to other differentiable GW models, such as the
-IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
-models within the same family. By jointly optimizing a new set of coefficients,
-it is expected that both models can be enhanced since they share similar
-construction principles to the IMRPhenomD model. For instance, they also use PN
-approximants as part of the ansatz in the inspiral segment. It will be interesting 
-to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
-parameterized by an additional anti-symmetric spin parameter, it is expected not
-to exhibit the degeneracy previously described. With the currently developing 
-{\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
-may provide valuable insights into the systematics of the Phenom models.
-Furthermore, this approach is applicable to other GW model families, such as NR
-surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
-simplify NR waveform calibration procedures and lead to the improvement of 
+% \section{Conclusion} \label{sec:conclusion}
+
+% In this paper, we have presented a systematic method to recalibrate GW models.
+% This method utilizes {\jax}'s automatic differentiation to apply
+% derivative-based optimization to recalibrate GW models jointly. Using the new
+% implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
+% conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
+% coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
+% improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
+% mismatches have a slightly better improvement. In contrast, different types of
+% loss function result in significantly different final mismatch distributions. As 
+% seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
+% $\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
+% slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
+
+% Furthermore, we investigated how the source parameters affect the
+% improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
+% certain preference for waveforms lying in the top-left region while the bottom-right 
+% region hardly improved. To further test this result, we recalibrate
+% waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
+% we can see that this recalibration process gives further
+% improvement to the top-left region while the bottom-right region only have little 
+% improvement. This indicates that the ansatz has limited match in this region and 
+% does not fit most waveforms in this region, hence introduces bias to downstream GW 
+% analysis. This phenomenon is mainly due to the reduced spin approximation used in 
+% parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
+
+% While we naively separate the optimization process into 4 regions, one can
+% perform systematic region-selection. In principle, we can apply this general
+% method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
+% models. Then, we can perform all the above analyses to understand how to
+% construct better GW Phenom models in the future.
+
+% Although our study primarily focused on the IMRPhenomD model, this simple yet
+% versatile approach can be applied to other differentiable GW models, such as the
+% IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
+% models within the same family. By jointly optimizing a new set of coefficients,
+% it is expected that both models can be enhanced since they share similar
+% construction principles to the IMRPhenomD model. For instance, they also use PN
+% approximants as part of the ansatz in the inspiral segment. It will be interesting 
+% to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
+% parameterized by an additional anti-symmetric spin parameter, it is expected not
+% to exhibit the degeneracy previously described. With the currently developing 
+% {\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
+% may provide valuable insights into the systematics of the Phenom models.
+% Furthermore, this approach is applicable to other GW model families, such as NR
+% surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
+% simplify NR waveform calibration procedures and lead to the improvement of 
 existing models.
-
-% Moved down from intro
-% In particular, we find it is easier IMRPhenomD
-% favors certain parts of the parameter space. This means IMRPhenomD introduces
-% systematic bias to other GW analysis tasks. This showcases the flaws of the
-% ansatz and allows us to have a deeper understanding of Phenom models.
-
-\section{Conclusion} \label{sec:conclusion}
-
-In this paper, we have presented a systematic method to recalibrate GW models.
-This method utilizes {\jax}'s automatic differentiation to apply
-derivative-based optimization to recalibrate GW models jointly. Using the new
-implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
-conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
-coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
-improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
-mismatches have a slightly better improvement. In contrast, different types of
-loss function result in significantly different final mismatch distributions. As 
-seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{ave}}$ outperforms 
-$\mathcal{L}_{\mathrm{nl}}$. By increasing the number of training waveforms, we see a 
-slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
-
-Furthermore, we investigated how the source parameters affect the
-improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-certain preference for waveforms lying in the top-left region while the bottom-right 
-region hardly improved. To further test this result, we recalibrate
-waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
-we can see that this recalibration process gives further
-improvement to the top-left region while the bottom-right region only have little 
-improvement. This indicates that the ansatz has limited match in this region and 
-does not fit most waveforms in this region, hence introduces bias to downstream GW 
-analysis. This phenomenon is mainly due to the reduced spin approximation used in 
-parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
-
-While we naively separate the optimization process into 4 regions, one can
-perform systematic region-selection. In principle, we can apply this general
-method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
-models. Then, we can perform all the above analyses to understand how to
-construct better GW Phenom models in the future.
 

### 2023-06-24 src/tex/ms.tex
Finished editing sec 4

@@
 
-\section{Discussion} \label{sec:discussion}
+\section{Discussion and Conclusion} \label{sec:discussion}
 
@@
 accuracy of phenomenological waveform models by jointly optimizing all
-coefficients at once. There are a number of caveats multiple ways to improve the
-current work.
+coefficients at once. However, there are a number of caveats as well
+as multiple ways to improve upon the current work.
 
-While this study focuses on the IMRPhenomD model because its availability as an
-automatically differentiable waveform, other more sophisticate waveforms such as
+While this study focuses on the IMRPhenomD model due to its availability as an
+automatically differentiable waveform in \ripple, other more sophisticated waveforms such as
 IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} and IMRPhenomXAS
-\citep{pratten2020setting,pratten2021computationally} can potentially benefit
-from recalibration even more, because of the larger number of calibration
-parameters. % Extend this a bit?
+\citep{pratten2020setting,pratten2021computationally} can also potentially benefit
+from recalibration. In fact, since these waveforms have a larger number of calibration
+parameters and are fit to a larger number of NR simulations, it is possible that the 
+improvement will be more dramatic than for IMRPhenomD. {\ripple} is being regularly
+updated with new waveforms, and future work should carefully examine whether these more
+modern waveforms can be improved. % Extend this a bit?
 
@@
 recalibration process struggles to improve in the regime where $\chi_{1}$ is
-positive and $\chi_{2}$ is negative, this means there is a dependency between
-the two spins that is not captured by the ansatz. A systematic scheme that
-incorporate this recalibration scheme with waveform design will be handy in
-constructing new waveform models in the future. In making figure
-\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions
-because of simplicity. These simple cuts demonstrate the accuracy of the
-waveform can be further improved, but they are almost certainly not the optimal
-way to incorporate the extra information we have about the waveform. A more
-general way such as adding new functional dependency between the parameters
-should be explored in the future.
+positive and $\chi_{2}$ is negative. This means there is a dependency between
+the two spins that is not captured by the current ansatz. A systematic scheme that
+incorporates our calibration method with waveform design will be handy in
+constructing new waveform models in the future.
 
-We use mismatch to quantify the accuracy of the waveform model in this work. It
-is a natural extension to investigate how would the recalibration affects
-downstream analysis such as parameter estimation and population analysis.
-Whether it is possible to reduce the systematic error in parameter estimation
-using recalibrated waveforms is an interesting question we will investigate in
-the sooner future. On top of this, recently relatively more differentiable
-sampler such as \citep{Wong:2022xvh} has been introduced to the community, one
-can also optimize the waveform directly using metrics from parameter estimation,
-such as bias in an injection-recovery run. This will be an example of
-more direct and efficient optimization of waveform models for downstream analysis.
+In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions purely for simplicity.
+These simple cuts demonstrate the accuracy of the waveform can be further improved,
+but they are almost certainly not the optimal way to incorporate the extra information
+we have about the waveform. A more general way such as adding new functional dependency
+between the parameters should be explored in the future.
+
+In this work, we use the mismatch to quantify the accuracy of the waveform model. 
+A natural extension of this work to investigate how the recalibration process affects
+downstream analyses such as parameter estimation and population modeling.
+Differentiable samplers such as~\citep{Wong:2022xvh}, which have been recently
+introduced to the community, potentially allow one to optimize the waveform directly
+using metrics from parameter estimation. For example, minimizing the bias in an injection-recovery run
+could be used as a loss function. Overall, this approach could help reduce systematic waveform error
+in parameter estimation simply through recalibrated waveforms. We plan to investigate these avenues in
+the future.
 
@@
 
-We have shown the result of recalibrating waveform coefficients. One thing to
-note is that our recalibration procedure is not exactly the same as the original
-calibration. For instance, we use a different set of NR waveforms, frequency
-range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
-optimization procedure should be able to improve the accuracy of IMRPhenomD on a
-similar scale regardless of the differences. Here, the result serves as a
-demonstration of the general method used.
+% As discussed in Sec.~\ref{sec:method} our recalibration procedure is not exactly the same as the original
+% calibration. For instance, we use a different set of NR waveforms, frequency
+% ranges, etc. Nonetheless, since the decrease in mismatch is rather significant, we believe that
+% the high-dimensional optimization procedure is primarily responsible to the improved performance.
 
-The results presented in Fig.~\ref{fig:q148_q1248_compare} demonstrate that
-increasing the number of training waveforms used in waveform optimization yields
-only a marginal increase in accuracy. Our analysis suggests that this marginal
-improvement is a consequence of over-determination of the waveform coefficients.
-Consequently, increasing the number of calibration NR waveforms is unlikely to
-result in any significant improvement of the model's accuracy. These
-observations suggest that the parameterized ansatz employed in our study may not
-be suitable for certain regions in the parameter space, leading to low mismatches
-for some waveforms while other waveforms remain at the high mismatch tail with
-negligible changes. This highlights the constraints of the model's flexibility
-that ultimately limit its performance.
+The results presented in Fig.~\ref{fig:q148_q1248_compare} indicate that
+increasing the number of training NR waveforms used in the waveform optimization yields
+only a marginal increase in accuracy. This observation suggests that the parameterized ansatz employed 
+in IMRPhenomD struggles to capture the full complexity of the NR waveforms. 
+However, as the accuracy requirements of waveform models increases with more sensitive detectors,
+more NR waveforms will be required to train more flexible Phenom models. 
+Accurate NR simulations must therefore be developed in parallel with waveform models
+to ensure we meet future detector accuracy requirements.
 
-The reduced spin approximation is a major contributor to the inaccuracies
-observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
-parameter, $\chi_{\mathrm{PN}}$, as described in Sec.~\ref{sec:method}. The
-parameterization of BBH mergers using a single spin parameter results in a
-degeneracy within the parameter space. Specifically, black hole events with
-different spins may generate the same waveform due to identical values of
-$\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
-unequal spin events. This degeneracy produces straight lines in the parameter
-space with negative slopes that depend on the mass ratio, which can be seen in 
-Fig.~\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
-region than the bottom-right region. In an attempt to address
-this issue, we partitioned the parameter space into four regions, as described
-in Sec.~\ref{sec:result}. With separate optimizations for each
-regions, Fig.~\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
-performance has improved, while the bottom-right region hardly improves. 
-This observation suggests the ansatz is specific to certain regions of the 
-parameter space, with a preference for BBH events lying above the diagonal, 
-and it has limited enhancement for events lying below the diagonal.
+Overall, the development of accurate waveform models is crucial for the success of GW astronomy.
+In this work, we have explored how modern computational tools (automatic differentiation and gradient descent)
+can help with this task. 
+Our method is general and can be applied to any waveform model that is differentiable.
+We therefore encourage the waveform development community to utlize these tools
+and hope that this work can effetively contribute to the development of accurate waveform models.
+
+% The reduced spin approximation is a major contributor to the inaccuracies
+% observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
+% parameter, $\chi_{\mathrm{PN}}$, as described in Sec.~\ref{sec:method}. The
+% parameterization of BBH mergers using a single spin parameter results in a
+% degeneracy within the parameter space. Specifically, black hole events with
+% different spins may generate the same waveform due to identical values of
+% $\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
+% unequal spin events. This degeneracy produces straight lines in the parameter
+% space with negative slopes that depend on the mass ratio, which can be seen in 
+% Fig.~\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
+% region than the bottom-right region. In an attempt to address
+% this issue, we partitioned the parameter space into four regions, as described
+% in Sec.~\ref{sec:result}. With separate optimizations for each
+% regions, Fig.~\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
+% performance has improved, while the bottom-right region hardly improves. 
+% This observation suggests the ansatz is specific to certain regions of the 
+% parameter space, with a preference for BBH events lying above the diagonal, 
+% and it has limited enhancement for events lying below the diagonal.
+
+\te{I suggest we end here}
 
@@
 a different set of training waveforms over the parameter space, the result might 
-give additional features that can test and explain IMRPhenomD better.
+give additional features that can test and explain IMRPhenomD better. \te{I don't really
+understand this paragraph. Should we remove it?}
 
@@
 % simplify NR waveform calibration procedures and lead to the improvement of 
-existing models.
+% existing models.
 

### 2023-06-24 src/tex/ms.tex
minor updates

@@
 as match filtering~\citep{owen1996search, owen1999matched} and parameter 
-estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
+estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
+Romero-Shaw:2020owr, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
 models. Since the generation of numerical relativity (NR) waveforms is
@@
 explore how different parts of the source parameter space (e.g. the primary and
-second spins) respond to the optimization procedure by optimizing the waveform
+secondary spins) respond to the optimization procedure by optimizing the waveform
 separately for different regions. This can help in understanding
@@
 how the calibration parameters enter the waveform.
-We then outline the mismatch and how it can be used as a loss fucntion. 
+We then outline the mismatch and how it can be used as a loss function. 
 Finally we discuss the gradient descent algorithm and our stopping criterion.
@@
 \end{align}
-where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
+where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
 parameters which will be discussed below. 
@@
 uses a series of parameterizations\footnote{
-	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
+	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
 } 
 which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
-The three sections are then \textit{stictched} together using step functions.
+The three sections are then \textit{stitched} together using step functions.
 Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
@@
 
-In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
+In practice the $\Lambda$ parameters are fit for each section independently, i.e., intermediate coefficients are fit whilst ignoring the MR region.
 Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
@@
 quantifies the difference between the waveform model and the target NR simulations which we want to match.
-Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm, husa2016frequency}. 
 It is defined as
@@
 	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+	\mathcal{L}_{\mathrm{norm}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
 \end{align}	
@@
 Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
+and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
 In this paper, we show the results of using both loss functions 
@@
 	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
+	$t_{\mathrm{RD}}$ is the duration of ringdown (maximum amplitude to the end of the strain) and $T$ is the duration of the
 	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
@@
 	space. The blue stars represent the waveforms used to compute the training loss and the orange
-	dots represent the waveforms used for validation.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
+	dots represent the waveforms used for validation.}
 	\label{fig:intrin_space}
@@
 We select waveforms that share the same part of the parameter space with the training
-set i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
+set, i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
 and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
@@
 
-To illustrate the effect of optimiziation on an individual waveform level, in Fig.
+To illustrate the effect of optimization on an individual waveform level, in Fig.
 \ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
@@
 magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
-$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
+$\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
 decrease in the median.
@@
 
-Note that the performace of the IMRPhenomD model was initally tested using the
+Note that the performance of the IMRPhenomD model was initially tested using the
 {\zdethp} weighted mismatch.\footnote{
@@
 } We would therefore like to examine whether using the {\zdethp} PSD in our loss function
-could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
+could lead to an improved to mismatch. Performing the optimization again, we find no significant difference
 between the results using the two PSDs in the distribution of mismatches.
@@
 \textit{increases} during optimization.
-We can see that waveforms with $q\le4$ show the most consistent average improvement.
+We can see that waveforms with $q\leq4$ show the most consistent average improvement.
 This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
@@
 When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
-This is particularly true in the $q\le4$ region, where we see a consistent
+This is particularly true in the $q\leq4$ region, where we see a consistent
 improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
-Since the original model was developed for aligned-spin systems, NR waveforms
-with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
-To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
+We also plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
 Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
-above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
-($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
+above. Meanwhile, the top-left and bottom-right regions respond to the optimization differently. In the top-left
 region, the waveform generally improves after optimization. However,
@@
 	\end{tabularx}
-	\caption{List of NR waveforms used in recalibrating the coefficients in the four $\chi_1-\chi_2$ quadrants.
+	\caption{List of NR waveforms used in recalibrating the coefficients in the four $\chi_1-\chi_2$ regions.
 	From top to bottom the lines denote the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$),
-	bottom-left ($\chi_1,\chi_2<0$), and bottom-right ($\chi_1>0>\chi_2$) regions.
+	bottom-left ($\chi_1,\chi_2<0$), and bottom-right ($\chi_1>0>\chi_2$) regions respectively.
 	Note that for the top-right and bottom-left regions, waveforms are chosen to have $\chi_1\approx\chi_2$,
@@
 valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
-equal-spin waveforms lying on the diagonal improve the most. The
-improvement above the diagonal is also significantly more than when optimizing using
-all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
-that the optimized histogram shifts to lower mismatches.\footnote{
+equal-spin waveforms lying on the diagonal improve the most. 
+\kl{The improvement above the diagonal is also significant comparing with the original waveform model.
+However, the improvement are not as profound as optimizing all waveform at once.
+This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.
+}
+%The improvement above the diagonal is also significantly more than when optimizing using
+%all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
+%that the optimized histogram shifts to lower mismatches.
+\footnote{
 	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
@@
 	as indicated by the change of the medians of each distribution (dashed
-	lines). \kw{Is the original referring to original IMRPhenomD or optimized
-	using everything? Can we have a reference for optimizing using everything?}}
+	lines).}
 	\label{fig:all_quadrants}
@@
 Our method is general and can be applied to any waveform model that is differentiable.
-We therefore encourage the waveform development community to utlize these tools
-and hope that this work can effetively contribute to the development of accurate waveform models.
+We therefore encourage the waveform development community to utilize these tools
+and hope that this work can effectively contribute to the development of accurate waveform models.
 

### 2023-06-24 src/scripts/all_quadrants.py

@@ -73,6 +73,27 @@ ax[0].axvline(
     linestyle=(0, (1, 1.3)),
 )
 ax[0].hist(
+    np.log10(data2_q148), 
+    bins=bins, 
+    alpha=0.2, 
+    color=c2
+)
+ax[0].hist(
+    np.log10(data2_q148), 
+    bins=bins, 
+    label="Optimized (all))", 
+    color=c2, 
+    lw=2, 
+    histtype="step"
+)
+ax[0].axvline(
+    x=np.log10(np.median(data2_q148)),
+    color=c2,
+    lw=2,
+    alpha=1.0,
+    linestyle=(0, (1, 1.3)),
+)
+ax[0].hist(
     np.log10(data2[:, 5]),
     bins=bins,
     label="Optimized",
@@ -93,27 +114,7 @@ ax[0].axvline(
     alpha=1.0,
     linestyle=(0, (1, 1.3)),
 )
-ax[0].hist(
-    np.log10(data2_q148), 
-    bins=bins, 
-    alpha=0.2, 
-    color=c2
-)
-ax[0].hist(
-    np.log10(data2_q148), 
-    bins=bins, 
-    label="Optimized $(\mathcal{L}_{\mathrm{ave}})$", 
-    color=c2, 
-    lw=2, 
-    histtype="step"
-)
-ax[0].axvline(
-    x=np.log10(np.median(data2_q148)),
-    color=c2,
-    lw=2,
-    alpha=1.0,
-    linestyle=(0, (1, 1.3)),
-)
+
 ax[0].set_ylabel("Count")
 ax[0].legend(loc="upper left")
 
@@ -139,26 +140,6 @@ ax[1].axvline(
     linestyle=(0, (1, 1.3)),
 )
 ax[1].hist(
-    np.log10(data4[:, 5]), 
-    bins=bins, 
-    histtype="step", 
-    color=c3, 
-    lw=2
-)
-ax[1].hist(
-    np.log10(data4[:, 5]), 
-    bins=bins, 
-    color=c3, 
-    alpha=0.2
-)
-ax[1].axvline(
-    x=np.log10(np.median(data4[:, 5])),
-    color=c3,
-    lw=2,
-    alpha=1.0,
-    linestyle=(0, (1, 1.3)),
-)
-ax[1].hist(
     np.log10(data4_q148), 
     bins=bins, 
     alpha=0.2, 
[…]

### 2023-06-24 src/scripts/intrin_space.py

@@ -50,8 +50,8 @@ chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3])
 
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
-plt.scatter(train_data[:, 0], train_data[:, 1], s=80, color=c1, label="Training", marker='*')
-plt.scatter(q, chi_PN, s=40, color=c2, label="Validation", zorder=-20)
+plt.scatter(train_data[:, 0], train_data[:, 1], s=160, color=c1, label="Training", marker='*')
+plt.scatter(q, chi_PN, s=20, color=c2, label="Validation", zorder=-20)
 plt.legend()
 plt.xlabel("$q$")
 plt.ylabel("$\chi_{\mathrm{PN}}$")

### 2023-06-25 src/tex/ms.tex
minor changes

@@
 
-% The reduced spin approximation is a major contributor to the inaccuracies
-% observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
-% parameter, $\chi_{\mathrm{PN}}$, as described in Sec.~\ref{sec:method}. The
-% parameterization of BBH mergers using a single spin parameter results in a
-% degeneracy within the parameter space. Specifically, black hole events with
-% different spins may generate the same waveform due to identical values of
-% $\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
-% unequal spin events. This degeneracy produces straight lines in the parameter
-% space with negative slopes that depend on the mass ratio, which can be seen in 
-% Fig.~\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
-% region than the bottom-right region. In an attempt to address
-% this issue, we partitioned the parameter space into four regions, as described
-% in Sec.~\ref{sec:result}. With separate optimizations for each
-% regions, Fig.~\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
-% performance has improved, while the bottom-right region hardly improves. 
-% This observation suggests the ansatz is specific to certain regions of the 
-% parameter space, with a preference for BBH events lying above the diagonal, 
-% and it has limited enhancement for events lying below the diagonal.
-
-\te{I suggest we end here}
-
-The division of the parameter space into four regions was a simple approach
-taken for practical reasons. A more systematic approach would involve the use of
-level set estimation algorithms to identify regions of interest within the
-parameter space. Such an algorithm can reveal additional degeneracies or issues
-that may exist within the ansatz. One possible strategy is to recalibrate
-individual regions of interest to achieve better results. An alternative
-approach is to select regions based on degeneracy lines. However, due to the
-limited number of NR waveforms available, we were unable to implement this
-approach. With more NR waveforms available in the future that cover the entire
-parameter space, we can perform optimization with fewer restrictions and select
-regions more systematically. Other than how to divide regions of interest, the 
-choice of training waveforms also affects the final results greatly. Note that 
-the choice of training waveforms listed in Tab.~\ref{tab:quadrants} were taken 
-arbitrarily to test the effectiveness of separate fitting. Hence, if one takes 
-a different set of training waveforms over the parameter space, the result might 
-give additional features that can test and explain IMRPhenomD better. \te{I don't really
-understand this paragraph. Should we remove it?}
-
-% \section{Conclusion} \label{sec:conclusion}
-
-% In this paper, we have presented a systematic method to recalibrate GW models.
-% This method utilizes {\jax}'s automatic differentiation to apply
-% derivative-based optimization to recalibrate GW models jointly. Using the new
-% implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
-% conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
-% coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
-% improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
-% mismatches have a slightly better improvement. In contrast, different types of
-% loss function result in significantly different final mismatch distributions. As 
-% seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
-% $\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
-% slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
-
-% Furthermore, we investigated how the source parameters affect the
-% improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
-% certain preference for waveforms lying in the top-left region while the bottom-right 
-% region hardly improved. To further test this result, we recalibrate
-% waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
-% we can see that this recalibration process gives further
-% improvement to the top-left region while the bottom-right region only have little 
-% improvement. This indicates that the ansatz has limited match in this region and 
-% does not fit most waveforms in this region, hence introduces bias to downstream GW 
-% analysis. This phenomenon is mainly due to the reduced spin approximation used in 
-% parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
-
-% While we naively separate the optimization process into 4 regions, one can
-% perform systematic region-selection. In principle, we can apply this general
-% method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
-% models. Then, we can perform all the above analyses to understand how to
-% construct better GW Phenom models in the future.
-
-% Although our study primarily focused on the IMRPhenomD model, this simple yet
-% versatile approach can be applied to other differentiable GW models, such as the
-% IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
-% models within the same family. By jointly optimizing a new set of coefficients,
-% it is expected that both models can be enhanced since they share similar
-% construction principles to the IMRPhenomD model. For instance, they also use PN
-% approximants as part of the ansatz in the inspiral segment. It will be interesting 
-% to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
-% parameterized by an additional anti-symmetric spin parameter, it is expected not
-% to exhibit the degeneracy previously described. With the currently developing 
-% {\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
-% may provide valuable insights into the systematics of the Phenom models.
-% Furthermore, this approach is applicable to other GW model families, such as NR
-% surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
-% simplify NR waveform calibration procedures and lead to the improvement of 
-% existing models.
-
 \section{ACKNOWLEDGMENTS}

### 2023-06-25 src/tex/ms.tex
Add comments

@@
 The two loss functions are chosen to give different 
-behaviour during the optimization process.
+behavior during the optimization process.
 The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{Loss functions against the number of iterations.
-	The vertical dotted line indicates minimum of the validation loss,
-	where we stop the optimization and obtain a new set of waveform coefficients, $\lambda$.}
+	\caption{Loss functions against the number of iterations. The vertical
+	dotted line indicates minimum of the validation loss, which is where we stop
+	the optimization.}
 	\label{fig:loss}
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Distribution of training and validation NR waveforms in the $q-\chi_{PN}$
-	space. The blue stars represent the waveforms used to compute the training loss and the orange
-	dots represent the waveforms used for validation.}
+	\caption{Distribution of training and validation NR waveforms in the
+	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
+	the training loss and the orange dots represent the waveforms used for
+	validation. \kw{It is hard to see the blue stars on top of the orange. Play
+	around with the size of the dots and change the colors. Also add the extra
+	NR waveforms in table 2 with a different color but use the stars.}}}
 	\label{fig:intrin_space}
@@
 	are SXS analogues to two of the BAM NR simulations used in the original calibration
-	\citep{khan2016frequency}.}
+	\citep{khan2016frequency}. \kw{BE EXPLICIT ABOUT WHAT DO YOU MEAN THE REMAINING TWO!}}
 	\label{tab:q148}
@@
 	\caption{Fractional mismatch change of the validation waveforms.
-	Each quadrant is fitted independently with to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
+	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
 	\label{fig:ps_q148_quadrant}
@@
 
-\kw{End here.}
-
-% As discussed in Sec.~\ref{sec:method} our recalibration procedure is not exactly the same as the original
-% calibration. For instance, we use a different set of NR waveforms, frequency
-% ranges, etc. Nonetheless, since the decrease in mismatch is rather significant, we believe that
-% the high-dimensional optimization procedure is primarily responsible to the improved performance.
-
 The results presented in Fig.~\ref{fig:q148_q1248_compare} indicate that

### 2023-06-26 src/tex/ms.tex
update

@@
 	the training loss and the orange dots represent the waveforms used for
-	validation. \kw{It is hard to see the blue stars on top of the orange. Play
-	around with the size of the dots and change the colors. Also add the extra
-	NR waveforms in table 2 with a different color but use the stars.}}}
+	validation.}
 	\label{fig:intrin_space}
@@
 	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
-	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
-	are SXS analogues to two of the BAM NR simulations used in the original calibration
-	\citep{khan2016frequency}. \kw{BE EXPLICIT ABOUT WHAT DO YOU MEAN THE REMAINING TWO!}}
+	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}).
+The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
 	\label{tab:q148}
@@
 are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
-bottom-right regions have limited data for $q>4$, hence the results are only
-valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
+bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions.
+From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
 equal-spin waveforms lying on the diagonal improve the most. 
-\kl{The improvement above the diagonal is also significant comparing with the original waveform model.
-However, the improvement are not as profound as optimizing all waveform at once.
-This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.
-}
+\kl{The improvement above the diagonal is also significant comparing with the original waveform model and it has a similar improvement as optimizing all waveform at once.
+Notice that optimizing over all waveforms gives a slightly better improvement.
+This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
 %The improvement above the diagonal is also significantly more than when optimizing using
@@
 	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
-} On the other hand, the
-improvement below the diagonal is negligible, indicating that our method struggles to
+} On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
 further improve over the original IMRPhenomD waveform. 
-Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
+Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
 to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 

### 2023-06-26 src/scripts/all_quadrants.py

@@ -38,12 +38,11 @@ c1 = "#102F68"
 c2 = "#B02423"
 c3 = "#FFB300"
 
-data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
-data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
+data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
+data4 = np.loadtxt(paths.data / "q148_1pos2neg_region_f.dat")
 
-data = np.loadtxt(paths.data / "q148.dat")
-data2_q148 = np.array([data[i, 5] for i in range(len(data[:,0])) if data[i, 2] < 0 and data[i, 3] > 0])
-data4_q148 = np.array([data[i, 5] for i in range(len(data[:,0])) if data[i, 2] > 0 and data[i, 3] < 0])
+data2_q148 = np.loadtxt(paths.data / "q148_1neg2pos_all_f.dat")
+data4_q148 = np.loadtxt(paths.data / "q148_1pos2neg_all_f.dat")
 
 bins = np.linspace(-5, -1, 15)
 
@@ -73,27 +72,6 @@ ax[0].axvline(
     linestyle=(0, (1, 1.3)),
 )
 ax[0].hist(
-    np.log10(data2_q148), 
-    bins=bins, 
-    alpha=0.2, 
-    color=c2
-)
-ax[0].hist(
-    np.log10(data2_q148), 
-    bins=bins, 
-    label="Optimized (all))", 
-    color=c2, 
-    lw=2, 
-    histtype="step"
-)
-ax[0].axvline(
-    x=np.log10(np.median(data2_q148)),
-    color=c2,
-    lw=2,
-    alpha=1.0,
-    linestyle=(0, (1, 1.3)),
-)
-ax[0].hist(
     np.log10(data2[:, 5]),
     bins=bins,
     label="Optimized",
@@ -114,7 +92,27 @@ ax[0].axvline(
     alpha=1.0,
     linestyle=(0, (1, 1.3)),
 )
-
+ax[0].hist(
+    np.log10(data2_q148[:, 5]), 
+    bins=bins, 
+    alpha=0.2, 
+    color=c2
+)
+ax[0].hist(
+    np.log10(data2_q148[:, 5]), 
+    bins=bins, 
+    label="Optimized $(\mathcal{L}_{\mathrm{ave}})$", 
+    color=c2, 
[…]

### 2023-06-26 src/scripts/ps_q148_quadrants.py

@@ -35,9 +35,9 @@ params = {
 mpl.rcParams.update(params)
 
 data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
-data2 = np.loadtxt(paths.data / "1neg2pos_mismatch.dat")
+data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
 data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
-data4 = np.loadtxt(paths.data / "1pos2neg_mismatch.dat")
+data4 = np.loadtxt(paths.data / "q148_1pos2neg_region_f.dat")
 data = np.concatenate((data1, data2, data3, data4))
 
 logdiff = np.log10(data[:, 5] / data[:, 4])

### 2023-06-26 src/tex/ms.tex
minor update

@@
 quantifies the difference between the waveform model and the target NR simulations which we want to match.
-Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm, husa2016frequency}. 
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{owen1996search, husa2016frequency}. 
 It is defined as
@@
 in order to compute the mismatch, \eqref{eq:mismatch}.
-For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
+For this, we taper the time-series using Tukey window~\citep{usman2016pycbc}\footnote{
 	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where

### 2023-06-26 src/scripts/intrin_space.py

@@ -35,7 +35,7 @@ params = {
 mpl.rcParams.update(params)
 # blue = "#102F68"
 # red = "#B02423"
-c1 = "#648FFF"
+c1 = "#0562ED"
 c2 = "#FE6100"
 # c2 = "#FFB000"
 

### 2023-06-26 src/scripts/intrin_space.py
update

@@ -37,9 +37,11 @@ mpl.rcParams.update(params)
 # red = "#B02423"
 c1 = "#0562ED"
 c2 = "#FE6100"
-# c2 = "#FFB000"
+c3 = '#266D1F'
 
-train_data = np.loadtxt(paths.data / "ps_train.dat")
+
+train_data = np.loadtxt(paths.data / "ps_train.dat")[:11]
+add_train_data = np.loadtxt(paths.data / "ps_train.dat")[11:]
 test_data = np.loadtxt(paths.data / "q148.dat")
 
 q = test_data[:, 0] / test_data[:, 1]
@@ -50,8 +52,9 @@ chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3])
 
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
-plt.scatter(train_data[:, 0], train_data[:, 1], s=160, color=c1, label="Training", marker='*')
-plt.scatter(q, chi_PN, s=20, color=c2, label="Validation", zorder=-20)
+plt.scatter(train_data[:, 0], train_data[:, 1], s=180, color=c1, label="Training", marker='*')
+plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Add. Training', marker='*')
+plt.scatter(q, chi_PN, s=10, color=c2, label="Validation", zorder=-20)
 plt.legend()
 plt.xlabel("$q$")
 plt.ylabel("$\chi_{\mathrm{PN}}$")

### 2023-06-26 src/scripts/loss.py

@@ -33,17 +33,10 @@ params = {
 }
 
 mpl.rcParams.update(params)
-# blue = "#102F68"
-# red = "#B02423"
-# c1 = "#4e7ab1"  # blue Nash
-# c1 = "#b14a49"  # red Nash
-# c2 = "#406a3d"  # green Nash
-c1 = "#648FFF"
+
+c1 = "#0562ED"
 c2 = "#FE6100"
-# 16, 47, 104)
-# \definecolor{darkgreen}{rgb}{0.075,0.302,0.047}
-# % \definecolor{dullred}{rgb}{0.706,0.208,0.192}
-# \definecolor{dullred}{RGB}{176, 36, 35}
+
 train_data = np.loadtxt(paths.data / "train_waveform_loss.txt")
 validation_data = np.loadtxt(paths.data / "validation_waveform_loss.txt")
 

### 2023-06-26 src/tex/ms.tex
update

@@
 	$t_{\mathrm{RD}}$ is the duration of ringdown (maximum amplitude to the end of the strain) and $T$ is the duration of the
-	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
+	entire GW strain. 
+%	\te{What is $t_{\mathrm{RD}}$? and how is it computed?}
 } before using standard FFT routines to compute the Fourier transform.
@@
 waveforms while the rest is taken from NR simulations.
-Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and testing)
+Instead, we solely use NR waveforms for calibration since most NR waveforms used (for both training and validation)
 have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, to contain part of the inspiral segment and all
@@
 recalibration procedure on each region separately
-(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
+(Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
 this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
@@
 bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions.
-From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
+From top panel of Fig.~\ref{fig:ps_q148_quadrant}, we observe that
 equal-spin waveforms lying on the diagonal improve the most. 
-\kl{The improvement above the diagonal is also significant comparing with the original waveform model and it has a similar improvement as optimizing all waveform at once.
-Notice that optimizing over all waveforms gives a slightly better improvement.
+\kl{Both top-right and bottom-left region have good performance comparing with the original model.
+Especially for the bottom-right region, where the improvement is significantly better than optimizing with all waveforms.
+This suggests the ansatz are well-fitted to this part of the parameter space.
+Waveforms above the diagonal also give good improvements when comparing with the original model and it has a similar improvement as optimizing all waveforms at once.
+Notice optimizing over all waveforms gives a slightly better improvement.
 This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
@@
 \footnote{
-	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
+	Note that the bottom-left panel of Fig.~\ref{fig:all_quadrants} specifically shows the
 	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
-} On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
+} 
+On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
 further improve over the original IMRPhenomD waveform. 
 Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
-to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
+to the overall mismatch (see bottom-right panel of Fig.~\ref{fig:all_quadrants}). 
 This suggests that the IMRPhenomD waveform ansatz fits waveforms with
@@
 	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches in the top-left (top) and bottom-right
-	(bottom) regions of Fig.~\ref{fig:ps_q148_quadrant}. We use a constant PSD to calculate the mismatch and
-	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The waveform generally improves in the
-	top-left region while it shows very little improvement in the bottom-right,
-	as indicated by the change of the medians of each distribution (dashed
-	lines).}
+	\caption{Distributions of mismatches for both optimization schemes in all 4 regions of Fig.~\ref{fig:ps_q148_quadrant}. 
+	We use a constant PSD to calculate the mismatch and
+	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distributions are indicated by dashed
+	lines.}
 	\label{fig:all_quadrants}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Fractional mismatch change of the validation waveforms.
-	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
+	\caption{\textit{Top}: Fractional mismatch change of the validation waveforms.
+	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.
+	\textit{Bottom}: Combined distribution of both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
 	\label{fig:ps_q148_quadrant}

### 2023-06-26 src/scripts/all_quadrants.py

@@ -38,147 +38,113 @@ c1 = "#102F68"
 c2 = "#B02423"
 c3 = "#FFB300"
 
+data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
+data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
 data4 = np.loadtxt(paths.data / "q148_1pos2neg_region_f.dat")
 
+data1_q148_ = np.loadtxt(paths.data / "q148.dat")
+data1_q148 = np.array([data1_q148_[i] for i in range(len(data1_q148_[:,0])) if data1_q148_[i,2] > 0 and data1_q148_[i,3] > 0])
 data2_q148 = np.loadtxt(paths.data / "q148_1neg2pos_all_f.dat")
+data3_q148_ = np.loadtxt(paths.data / "q148.dat")
+data3_q148 = np.array([data3_q148_[i] for i in range(len(data3_q148_[:,0])) if data3_q148_[i,2] < 0 and data3_q148_[i,3] < 0])
 data4_q148 = np.loadtxt(paths.data / "q148_1pos2neg_all_f.dat")
 
 bins = np.linspace(-5, -1, 15)
 
-fig, ax = plt.subplots(2, figsize=(6, 10), sharex=True)
+fig, ax = plt.subplots(2, 2, figsize=(12, 10), sharex=True)
 plt.subplots_adjust(hspace=0.05)
+plt.subplots_adjust(wspace=0.04)
 
-ax[0].hist(
-    np.log10(data2[:, 4]), 
-    bins=bins, 
-    alpha=0.2, 
-    color=c1
-)
-ax[0].hist(
-    np.log10(data2[:, 4]), 
-    bins=bins, 
-    label="Original", 
-    color=c1, 
-    lw=2, 
-    histtype="step"
-)
-ax[0].axvline(
-    x=np.log10(np.median(data2[:, 4])),
-    color=c1,
-    zorder=-20,
-    lw=2,
-    alpha=1.0,
-    linestyle=(0, (1, 1.3)),
-)
-ax[0].hist(
-    np.log10(data2[:, 5]),
-    bins=bins,
-    label="Optimized",
-    histtype="step",
-    color=c3,
-    linewidth=2,
-)
[…]

### 2023-06-26 src/scripts/ps_q148_quadrants.py

@@ -38,18 +38,38 @@ data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
 data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
 data4 = np.loadtxt(paths.data / "q148_1pos2neg_region_f.dat")
-data = np.concatenate((data1, data2, data3, data4))
+data_region = np.concatenate((data1, data2, data3, data4))
 
-logdiff = np.log10(data[:, 5] / data[:, 4])
+data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
+data2 = np.loadtxt(paths.data / "q148_1neg2pos_all_f.dat")
+data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
+data4 = np.loadtxt(paths.data / "q148_1pos2neg_all_f.dat")
+data_q148 = np.concatenate((data1, data2, data3, data4))
+
+logdiff = np.log10(data_region[:, 5] / data_region[:, 4])
+
+fig, ax = plt.subplots(2, 1, figsize=(7, 12))
+
 cm = plt.cm.get_cmap("RdYlBu")
 colors = logdiff
-sc = plt.scatter(data[:, 2], data[:, 3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm)
-cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1])
+sc = ax[0].scatter(data_region[:, 2], data_region[:, 3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm)
+cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1], ax=ax[0])
 cbar.ax.set_ylabel(
     "$\log_{10}(\mathcal{M}_f/\mathcal{M}_i)$", rotation=270, labelpad=20
 )
 cbar.ax.set_yticklabels(["$<-1$", "$-0.5$", "$0$", "$0.5$", "$>1$"])
-plt.xlabel("$\chi_1$")
-plt.ylabel("$\chi_2$")
+ax[0].set_xlabel("$\chi_1$")
+ax[0].set_ylabel("$\chi_2$")
+
+bins = np.linspace(-5, -1, 15)
+
+ax[1].hist(np.log10(data_q148[:,5]), alpha=0.2, bins=bins, color=c2)
[…]

### 2023-06-26 src/scripts/ps_q148_quadrants.py
minor fix to plot

@@ -34,6 +34,9 @@ params = {
 
 mpl.rcParams.update(params)
 
+c2 = "#B02423"
+c3 = "#FFB300"
+
 data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
 data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")

### 2023-06-26 src/tex/ms.tex
update

@@
 $\chi_1>0$ and $\chi_2<0$ less well. 
+\kl{Overall, the improvement for both optimization schemes are similar (bottom panel of Fig.~\ref{fig:ps_q148_quadrant}).
+Although split region optimization uses more training waveforms, some of the waveforms, e.g.
+opposite spins, hinders the optimization procedure and hence gives a slightly worse result when comparing with optimization over all waveforms.}
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\script{all_quadrants.py}
@@
 	\label{fig:all_quadrants}
-\end{figure}
+\end{figure*}
 
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{\textit{Top}: Fractional mismatch change of the validation waveforms.
+	\caption{\textit{Top}: Fractional mismatch change of the validation waveforms ().
 	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.

### 2023-06-27 src/scripts/all_quadrants.py
fix formatting issue

@@ -130,6 +130,7 @@ ax[1, 1].set_ylim(0, 21)
 ax[1, 1].yaxis.set_ticks([0, 5, 10, 15, 20])
 plt.setp(ax[1, 1].get_yticklabels(), visible=False)
 
+
 ax[0, 0].text(-1.5, 42.8, r'$\chi_1,\chi_2>0$',
         horizontalalignment='left',
         verticalalignment='top',

### 2023-06-27 src/scripts/intrin_space.py

@@ -53,7 +53,7 @@ chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3])
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
 plt.scatter(train_data[:, 0], train_data[:, 1], s=180, color=c1, label="Training", marker='*')
-plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Add. Training', marker='*')
+plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Extra Training', marker='*')
 plt.scatter(q, chi_PN, s=10, color=c2, label="Validation", zorder=-20)
 plt.legend()
 plt.xlabel("$q$")

### 2023-06-27 src/scripts/ps_q148_quadrants.py

@@ -67,12 +67,13 @@ ax[0].set_ylabel("$\chi_2$")
 bins = np.linspace(-5, -1, 15)
 
 ax[1].hist(np.log10(data_q148[:,5]), alpha=0.2, bins=bins, color=c2)
-ax[1].hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2)
+ax[1].hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2, label='Optimized (All Regions)')
 ax[1].hist(np.log10(data_region[:,5]), alpha=0.2, bins=bins, color=c3)
-ax[1].hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3)
+ax[1].hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Optimized (Split Regions)')
 ax[1].axvline(x=np.log10(np.median(data_q148[:,5])), color=c2, ls='--', lw=2)
 ax[1].axvline(x=np.log10(np.median(data_region[:,5])), color=c3, ls='--', lw=2)
 ax[1].set_xlabel("$\log_{10}(\mathcal{M})$")
 ax[1].set_ylabel("Count")
+ax[1].legend(loc='upper left')
 
 plt.savefig(paths.figures / "ps_q148_quadrants.pdf", bbox_inches="tight")

### 2023-06-27 src/scripts/q148_q1248_compare.py

@@ -48,7 +48,7 @@ q148_logdiff = np.log10(q148_data[:, 5] / q148_data[:, 4])
 q1248_logdiff = np.log10(q1248_data[:, 5] / q1248_data[:, 4])
 
 plt.hist(q148_logdiff, bins=bins, alpha=0.2, color=c1)
-plt.hist(q148_logdiff, bins=bins, label="Optimized", color=c1, histtype="step", lw=2)
+plt.hist(q148_logdiff, bins=bins, label="Without Extra Data", color=c1, histtype="step", lw=2)
 plt.axvline(
     x=np.median(q148_logdiff),
     color=c1,
@@ -66,7 +66,7 @@ plt.hist(
 plt.hist(
     q1248_logdiff,
     bins=bins,
-    label=f"Add. NR data",
+    label=f"With Extra Data",
     histtype="step",
     color=c2,
     linewidth=2,

### 2023-06-27 src/scripts/intrin_space.py
Update intrin_space.py
Kevin forgot to change figure one god damn it

@@ -54,7 +54,7 @@ chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3])
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
 plt.scatter(train_data[:, 0], train_data[:, 1], s=180, color=c1, label="Training", marker='*')
 plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Extra Training', marker='*')
-plt.scatter(q, chi_PN, s=10, color=c2, label="Validation", zorder=-20)
+plt.scatter(q, chi_PN, s=3, color=c2, label="Validation", zorder=-20)
 plt.legend()
 plt.xlabel("$q$")
 plt.ylabel("$\chi_{\mathrm{PN}}$")

### 2023-06-27 src/tex/ms.tex
fixing minor typos

--- 2023-06-26 src/tex/ms.tex
@@
 as match filtering~\citep{owen1996search, owen1999matched} and parameter 
-estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
-Romero-Shaw:2020owr, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
+estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
 models. Since the generation of numerical relativity (NR) waveforms is
@@
 explore how different parts of the source parameter space (e.g. the primary and
-secondary spins) respond to the optimization procedure by optimizing the waveform
+second spins) respond to the optimization procedure by optimizing the waveform
 separately for different regions. This can help in understanding
@@
 how the calibration parameters enter the waveform.
-We then outline the mismatch and how it can be used as a loss function. 
+We then outline the mismatch and how it can be used as a loss fucntion. 
 Finally we discuss the gradient descent algorithm and our stopping criterion.
@@
 \end{align}
-where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
+where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
 parameters which will be discussed below. 
@@
 uses a series of parameterizations\footnote{
-	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
+	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
 } 
 which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
-The three sections are then \textit{stitched} together using step functions.
+The three sections are then \textit{stictched} together using step functions.
 Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
@@
 
-In practice the $\Lambda$ parameters are fit for each section independently, i.e., intermediate coefficients are fit whilst ignoring the MR region.
+In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
 Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
@@
 quantifies the difference between the waveform model and the target NR simulations which we want to match.
-Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{owen1996search, husa2016frequency}. 
+Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
 It is defined as
@@
 	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
-	\mathcal{L}_{\mathrm{norm}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
+	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
 \end{align}	
@@
 The two loss functions are chosen to give different 
-behavior during the optimization process.
+behaviour during the optimization process.
 The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
@@
 Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
-and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
+and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
 In this paper, we show the results of using both loss functions 
@@
 To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
-Originally, 19 NR simulations were used to calibrate IMRPhenomD \citep{khan2016frequency}; 9 of these 
+Originally, 19 NR simulations were used to calibrate IMRPhenomD \citep{khan2016frequency}; nine of these 
 are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
@@
 In fact, the SXS catalog only has NR waveforms with $q\leq10$.
-Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$,
+Nevertheless, we are interested in the behavior of the IMRPhenomD model with small $q$,
 as most BBH events observed by LIGO and Virgo have $q\leq8$.
@@
 in order to compute the mismatch, \eqref{eq:mismatch}.
-For this, we taper the time-series using Tukey window~\citep{usman2016pycbc}\footnote{
+For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
 	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
-	$t_{\mathrm{RD}}$ is the duration of ringdown (maximum amplitude to the end of the strain) and $T$ is the duration of the
+	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
 	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
-} before using standard FFT routines to compute the Fourier transform.
+} before using standard routines to compute the fast Fourier transform.
 
 In addition to choosing the NR waveforms for the training set,
-one needs to choose a relevant PSD for mismatch.
+one needs to choose a relevant PSD for the mismatch.
 We have opted to use a flat PSD for the mismatch calculation,
@@
 neighborhood of the minimum that we wish to find. 
-We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
+We fix our learning rate, $\alpha$, to be $10^{-6}$, which is small enough to ensure we 
 don't move far from the minimum.
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{Loss functions against the number of iterations. The vertical
-	dotted line indicates minimum of the validation loss, which is where we stop
-	the optimization.}
+	\caption{Loss functions against the number of iterations.
+	The vertical dotted line indicates minimum of the validation loss,
+	where we stop the optimization and obtain a new set of waveform coefficients, $\lambda$.}
 	\label{fig:loss}
@@
 	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Distribution of training and validation NR waveforms in the
-	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
-	the training loss and the orange dots represent the waveforms used for
-	validation.}
+	\caption{Distribution of training and validation NR waveforms in the $q-\chi_{PN}$
+	space. The blue stars represent the waveforms used to compute the training loss and the orange
+	dots represent the waveforms used for validation.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
 	\label{fig:intrin_space}
@@
 	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
-	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}).
-The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
+	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
+	are SXS analogues to two of the BAM NR simulations used in the original calibration
+	\citep{khan2016frequency}.}
 	\label{tab:q148}
@@
 We select waveforms that share the same part of the parameter space with the training
-set, i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
+set i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
 and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
@@
 
-To illustrate the effect of optimization on an individual waveform level, in Fig.
+To illustrate the effect of optimiziation on an individual waveform level, in Fig.
 \ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
@@
 magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
-$\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
+$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
 decrease in the median.
@@
 
-Note that the performance of the IMRPhenomD model was initially tested using the
+Note that the performace of the IMRPhenomD model was initally tested using the
 {\zdethp} weighted mismatch.\footnote{
@@
 } We would therefore like to examine whether using the {\zdethp} PSD in our loss function
-could lead to an improved to mismatch. Performing the optimization again, we find no significant difference
+could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
 between the results using the two PSDs in the distribution of mismatches.
@@
 \textit{increases} during optimization.
-We can see that waveforms with $q\leq4$ show the most consistent average improvement.
+We can see that waveforms with $q\le4$ show the most consistent average improvement.
 This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
@@
 When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
-This is particularly true in the $q\leq4$ region, where we see a consistent
+This is particularly true in the $q\le4$ region, where we see a consistent
 improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
-We also plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
+Since the original model was developed for aligned-spin systems, NR waveforms
+with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
+To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
 Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
-above. Meanwhile, the top-left and bottom-right regions respond to the optimization differently. In the top-left
+above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
+($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
 region, the waveform generally improves after optimization. However,
@@
 	\end{tabularx}
-	\caption{List of NR waveforms used in recalibrating the coefficients in the four $\chi_1-\chi_2$ regions.
+	\caption{List of NR waveforms used in recalibrating the coefficients in the four $\chi_1-\chi_2$ quadrants.
 	From top to bottom the lines denote the top-right ($\chi_1,\chi_2>0$), top-left ($\chi_1<0<\chi_2$),
-	bottom-left ($\chi_1,\chi_2<0$), and bottom-right ($\chi_1>0>\chi_2$) regions respectively.
+	bottom-left ($\chi_1,\chi_2<0$), and bottom-right ($\chi_1>0>\chi_2$) regions.
 	Note that for the top-right and bottom-left regions, waveforms are chosen to have $\chi_1\approx\chi_2$,
@@
 are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
-bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions.
-From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
-equal-spin waveforms lying on the diagonal improve the most. 
-\kl{The improvement above the diagonal is also significant comparing with the original waveform model and it has a similar improvement as optimizing all waveform at once.
-Notice that optimizing over all waveforms gives a slightly better improvement.
-This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
-%The improvement above the diagonal is also significantly more than when optimizing using
-%all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
-%that the optimized histogram shifts to lower mismatches.
-\footnote{
+bottom-right regions have limited data for $q>4$, hence the results are only
+valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
+equal-spin waveforms lying on the diagonal improve the most. The
+improvement above the diagonal is also significantly more than when optimizing using
+all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
+that the optimized histogram shifts to lower mismatches.\footnote{
 	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
 	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
-} On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
+} On the other hand, the
+improvement below the diagonal is negligible, indicating that our method struggles to
 further improve over the original IMRPhenomD waveform. 
-Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
+Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
 to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
@@
 	as indicated by the change of the medians of each distribution (dashed
-	lines).}
+	lines). \kw{Is the original referring to original IMRPhenomD or optimized
+	using everything? Can we have a reference for optimizing using everything?}}
 	\label{fig:all_quadrants}
@@
 	\caption{Fractional mismatch change of the validation waveforms.
-	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
+	Each quadrant is fitted independently with to the NR waveforms listed in Tab.~\ref{tab:quadrants}.}
 	\label{fig:ps_q148_quadrant}
@@
 
-\section{Discussion and Conclusion} \label{sec:discussion}
+\section{Discussion} \label{sec:discussion}
 
@@
 accuracy of phenomenological waveform models by jointly optimizing all
-coefficients at once. However, there are a number of caveats as well
-as multiple ways to improve upon the current work.
+coefficients at once. There are a number of caveats multiple ways to improve the
+current work.
 
-While this study focuses on the IMRPhenomD model due to its availability as an
-automatically differentiable waveform in \ripple, other more sophisticated waveforms such as
+While this study focuses on the IMRPhenomD model because its availability as an
+automatically differentiable waveform, other more sophisticate waveforms such as
 IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} and IMRPhenomXAS
-\citep{pratten2020setting,pratten2021computationally} can also potentially benefit
-from recalibration. In fact, since these waveforms have a larger number of calibration
-parameters and are fit to a larger number of NR simulations, it is possible that the 
-improvement will be more dramatic than for IMRPhenomD. {\ripple} is being regularly
-updated with new waveforms, and future work should carefully examine whether these more
-modern waveforms can be improved. % Extend this a bit?
+\citep{pratten2020setting,pratten2021computationally} can potentially benefit
+from recalibration even more, because of the larger number of calibration
+parameters. % Extend this a bit?
 
@@
 recalibration process struggles to improve in the regime where $\chi_{1}$ is
-positive and $\chi_{2}$ is negative. This means there is a dependency between
-the two spins that is not captured by the current ansatz. A systematic scheme that
-incorporates our calibration method with waveform design will be handy in
-constructing new waveform models in the future.
+positive and $\chi_{2}$ is negative, this means there is a dependency between
+the two spins that is not captured by the ansatz. A systematic scheme that
+incorporate this recalibration scheme with waveform design will be handy in
+constructing new waveform models in the future. In making figure
+\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions
+because of simplicity. These simple cuts demonstrate the accuracy of the
+waveform can be further improved, but they are almost certainly not the optimal
+way to incorporate the extra information we have about the waveform. A more
+general way such as adding new functional dependency between the parameters
+should be explored in the future.
 
-In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions purely for simplicity.
-These simple cuts demonstrate the accuracy of the waveform can be further improved,
-but they are almost certainly not the optimal way to incorporate the extra information
-we have about the waveform. A more general way such as adding new functional dependency
-between the parameters should be explored in the future.
+We use mismatch to quantify the accuracy of the waveform model in this work. It
+is a natural extension to investigate how would the recalibration affects
+downstream analysis such as parameter estimation and population analysis.
+Whether it is possible to reduce the systematic error in parameter estimation
+using recalibrated waveforms is an interesting question we will investigate in
+the sooner future. On top of this, recently relatively more differentiable
+sampler such as \citep{Wong:2022xvh} has been introduced to the community, one
+can also optimize the waveform directly using metrics from parameter estimation,
+such as bias in an injection-recovery run. This will be an example of
+more direct and efficient optimization of waveform models for downstream analysis.
 
-In this work, we use the mismatch to quantify the accuracy of the waveform model. 
-A natural extension of this work to investigate how the recalibration process affects
-downstream analyses such as parameter estimation and population modeling.
-Differentiable samplers such as~\citep{Wong:2022xvh}, which have been recently
-introduced to the community, potentially allow one to optimize the waveform directly
-using metrics from parameter estimation. For example, minimizing the bias in an injection-recovery run
-could be used as a loss function. Overall, this approach could help reduce systematic waveform error
-in parameter estimation simply through recalibrated waveforms. We plan to investigate these avenues in
-the future.
+\kw{End here.}
 
-The results presented in Fig.~\ref{fig:q148_q1248_compare} indicate that
-increasing the number of training NR waveforms used in the waveform optimization yields
-only a marginal increase in accuracy. This observation suggests that the parameterized ansatz employed 
-in IMRPhenomD struggles to capture the full complexity of the NR waveforms. 
-However, as the accuracy requirements of waveform models increases with more sensitive detectors,
-more NR waveforms will be required to train more flexible Phenom models. 
-Accurate NR simulations must therefore be developed in parallel with waveform models
-to ensure we meet future detector accuracy requirements.
+We have shown the result of recalibrating waveform coefficients. One thing to
+note is that our recalibration procedure is not exactly the same as the original
+calibration. For instance, we use a different set of NR waveforms, frequency
+range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
+optimization procedure should be able to improve the accuracy of IMRPhenomD on a
+similar scale regardless of the differences. Here, the result serves as a
+demonstration of the general method used.
 
-Overall, the development of accurate waveform models is crucial for the success of GW astronomy.
-In this work, we have explored how modern computational tools (automatic differentiation and gradient descent)
-can help with this task. 
-Our method is general and can be applied to any waveform model that is differentiable.
-We therefore encourage the waveform development community to utilize these tools
-and hope that this work can effectively contribute to the development of accurate waveform models.
+The results presented in Fig.~\ref{fig:q148_q1248_compare} demonstrate that
+increasing the number of training waveforms used in waveform optimization yields
+only a marginal increase in accuracy. Our analysis suggests that this marginal
+improvement is a consequence of over-determination of the waveform coefficients.
+Consequently, increasing the number of calibration NR waveforms is unlikely to
+result in any significant improvement of the model's accuracy. These
+observations suggest that the parameterized ansatz employed in our study may not
+be suitable for certain regions in the parameter space, leading to low mismatches
+for some waveforms while other waveforms remain at the high mismatch tail with
+negligible changes. This highlights the constraints of the model's flexibility
+that ultimately limit its performance.
+
+The reduced spin approximation is a major contributor to the inaccuracies
+observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
+parameter, $\chi_{\mathrm{PN}}$, as described in Sec.~\ref{sec:method}. The
+parameterization of BBH mergers using a single spin parameter results in a
+degeneracy within the parameter space. Specifically, black hole events with
+different spins may generate the same waveform due to identical values of
+$\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
+unequal spin events. This degeneracy produces straight lines in the parameter
+space with negative slopes that depend on the mass ratio, which can be seen in 
+Fig.~\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
+region than the bottom-right region. In an attempt to address
+this issue, we partitioned the parameter space into four regions, as described
+in Sec.~\ref{sec:result}. With separate optimizations for each
+regions, Fig.~\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
+performance has improved, while the bottom-right region hardly improves. 
+This observation suggests the ansatz is specific to certain regions of the 
+parameter space, with a preference for BBH events lying above the diagonal, 
+and it has limited enhancement for events lying below the diagonal.
+
+The division of the parameter space into four regions was a simple approach
+taken for practical reasons. A more systematic approach would involve the use of
+level set estimation algorithms to identify regions of interest within the
+parameter space. Such an algorithm can reveal additional degeneracies or issues
+that may exist within the ansatz. One possible strategy is to recalibrate
+individual regions of interest to achieve better results. An alternative
+approach is to select regions based on degeneracy lines. However, due to the
+limited number of NR waveforms available, we were unable to implement this
+approach. With more NR waveforms available in the future that cover the entire
+parameter space, we can perform optimization with fewer restrictions and select
+regions more systematically. Other than how to divide regions of interest, the 
+choice of training waveforms also affects the final results greatly. Note that 
+the choice of training waveforms listed in Tab.~\ref{tab:quadrants} were taken 
+arbitrarily to test the effectiveness of separate fitting. Hence, if one takes 
+a different set of training waveforms over the parameter space, the result might 
+give additional features that can test and explain IMRPhenomD better.
+
+% \section{Conclusion} \label{sec:conclusion}
+
+% In this paper, we have presented a systematic method to recalibrate GW models.
+% This method utilizes {\jax}'s automatic differentiation to apply
+% derivative-based optimization to recalibrate GW models jointly. Using the new
+% implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in
+% conjunction with NR waveforms from the SXS catalog, we recalibrate waveform
+% coefficients of the IMRPhenomD model. In general, the waveform accuracy can be
+% improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted
+% mismatches have a slightly better improvement. In contrast, different types of
+% loss function result in significantly different final mismatch distributions. As 
+% seen in Fig.~\ref{fig:q148}, $\mathcal{L}_{\mathrm{mean}}$ outperforms 
+% $\mathcal{L}_{\mathrm{fl}}$. By increasing the number of training waveforms, we see a 
+% slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
+
+% Furthermore, we investigated how the source parameters affect the
+% improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
+% certain preference for waveforms lying in the top-left region while the bottom-right 
+% region hardly improved. To further test this result, we recalibrate
+% waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
+% we can see that this recalibration process gives further
+% improvement to the top-left region while the bottom-right region only have little 
+% improvement. This indicates that the ansatz has limited match in this region and 
+% does not fit most waveforms in this region, hence introduces bias to downstream GW 
+% analysis. This phenomenon is mainly due to the reduced spin approximation used in 
+% parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
+
+% While we naively separate the optimization process into 4 regions, one can
+% perform systematic region-selection. In principle, we can apply this general
+% method to other newer and more accurate models such as IMRPhenomX or IMRPhenomP
+% models. Then, we can perform all the above analyses to understand how to
+% construct better GW Phenom models in the future.
+
+% Although our study primarily focused on the IMRPhenomD model, this simple yet
+% versatile approach can be applied to other differentiable GW models, such as the
+% IMRPhenomP \citep{hannam2014simple, khan2019phenomenological} or IMRPhenomX \citep{pratten2020setting,pratten2021computationally} 
+% models within the same family. By jointly optimizing a new set of coefficients,
+% it is expected that both models can be enhanced since they share similar
+% construction principles to the IMRPhenomD model. For instance, they also use PN
+% approximants as part of the ansatz in the inspiral segment. It will be interesting 
+% to recalibratethe IMRPhenomXAS model \citep{pratten2020setting}. Because it is
+% parameterized by an additional anti-symmetric spin parameter, it is expected not
+% to exhibit the degeneracy previously described. With the currently developing 
+% {\jax} IMRPhenomXAS model in {\ripple}, A more detailed investigation
+% may provide valuable insights into the systematics of the Phenom models.
+% Furthermore, this approach is applicable to other GW model families, such as NR
+% surrogate models or EOB models introduced in Sec. \ref{sec:intro}. Such an approach could 
+% simplify NR waveform calibration procedures and lead to the improvement of 
+existing models.
 

### 2023-06-27 src/tex/ms.tex
merging

--- 2023-06-26 src/tex/ms.tex
@@
 In addition to choosing the NR waveforms for the training set,
-one needs to choose a relevant PSD for mismatch.
+one needs to choose a relevant PSD for the mismatch.
 We have opted to use a flat PSD for the mismatch calculation,

### 2023-06-27 src/tex/ms.tex
merging correctly

@@
 In this work, we use the mismatch to quantify the accuracy of the waveform model. 
-A natural extension of this work to investigate how the recalibration process affects
+A natural extension of this work is to investigate how the recalibration process affects
 downstream analyses such as parameter estimation and population modeling.

### 2023-06-27 src/scripts/all_quadrants.py
Update

@@ -36,7 +36,7 @@ mpl.rcParams.update(params)
 
 c1 = "#102F68"
 c2 = "#B02423"
-c3 = "#FFB300"
+c3 = "#B58500"
 
 data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")

### 2023-06-27 src/scripts/intrin_space.py

@@ -53,7 +53,7 @@ chi_PN = (test_data[:, 0] * test_data[:, 2] + test_data[:, 1] * test_data[:, 3])
 # train_data = np.loadtxt('./plotting_data/ps_train.dat', delimiter=' ') # [q, chi_PN, chi1, chi2]
 # test_data = np.loadtxt('./plotting_data/q148.dat', delimiter=' ') # [q, chi_PN, chi1, chi2, ori_loss, opt_loss]
 plt.scatter(train_data[:, 0], train_data[:, 1], s=180, color=c1, label="Training", marker='*')
-plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Extra Training', marker='*')
+plt.scatter(add_train_data[:, 0], add_train_data[:, 1], s=180, color=c3, label='Extra Training', marker='^')
 plt.scatter(q, chi_PN, s=3, color=c2, label="Validation", zorder=-20)
 plt.legend()
 plt.xlabel("$q$")

### 2023-06-27 src/scripts/ps_q148_quadrants.py

@@ -35,7 +35,7 @@ params = {
 mpl.rcParams.update(params)
 
 c2 = "#B02423"
-c3 = "#FFB300"
+c3 = "#B58500"
 
 data1 = np.loadtxt(paths.data / "1pos2pos_mismatch.dat")
 data2 = np.loadtxt(paths.data / "q148_1neg2pos_region_f.dat")
@@ -51,29 +51,17 @@ data_q148 = np.concatenate((data1, data2, data3, data4))
 
 logdiff = np.log10(data_region[:, 5] / data_region[:, 4])
 
-fig, ax = plt.subplots(2, 1, figsize=(7, 12))
-
-cm = plt.cm.get_cmap("RdYlBu")
-colors = logdiff
-sc = ax[0].scatter(data_region[:, 2], data_region[:, 3], c=colors, vmin=-1, vmax=1, s=30, cmap=cm)
-cbar = plt.colorbar(sc, ticks=[-1, -0.5, 0, 0.5, 1], ax=ax[0])
-cbar.ax.set_ylabel(
-    "$\log_{10}(\mathcal{M}_f/\mathcal{M}_i)$", rotation=270, labelpad=20
-)
-cbar.ax.set_yticklabels(["$<-1$", "$-0.5$", "$0$", "$0.5$", "$>1$"])
-ax[0].set_xlabel("$\chi_1$")
-ax[0].set_ylabel("$\chi_2$")
-
 bins = np.linspace(-5, -1, 15)
 
-ax[1].hist(np.log10(data_q148[:,5]), alpha=0.2, bins=bins, color=c2)
-ax[1].hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2, label='Optimized (All Regions)')
-ax[1].hist(np.log10(data_region[:,5]), alpha=0.2, bins=bins, color=c3)
-ax[1].hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Optimized (Split Regions)')
-ax[1].axvline(x=np.log10(np.median(data_q148[:,5])), color=c2, ls='--', lw=2)
-ax[1].axvline(x=np.log10(np.median(data_region[:,5])), color=c3, ls='--', lw=2)
-ax[1].set_xlabel("$\log_{10}(\mathcal{M})$")
-ax[1].set_ylabel("Count")
-ax[1].legend(loc='upper left')
+plt.figure(figsize=(6, 5))
[…]

### 2023-06-27 src/scripts/ps_q148_quadrants.py
fix

@@ -60,8 +60,8 @@ plt.hist(np.log10(data_region[:,5]), alpha=0.2, bins=bins, color=c3)
 plt.hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Optimized (Split Regions)')
 plt.axvline(x=np.log10(np.median(data_q148[:,5])), color=c2, ls='--', lw=2)
 plt.axvline(x=np.log10(np.median(data_region[:,5])), color=c3, ls='--', lw=2)
-plt.set_xlabel("$\log_{10}(\mathcal{M})$")
-plt.set_ylabel("Count")
+plt.xlabel("$\log_{10}(\mathcal{M})$")
+plt.ylabel("Count")
 plt.legend(loc='upper left')
 
 plt.savefig(paths.figures / "ps_q148_quadrants.pdf", bbox_inches="tight")

### 2023-06-27 src/tex/ms.tex
Update
make Fig. 8 larger
change caption and legend

@@
 	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
-	the training loss and the orange dots represent the waveforms used for
-	validation.}
+	the training loss, green triangles represent extra training data, and the 
+	orange dots represent the waveforms used for validation.}
 	\label{fig:intrin_space}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
-	\caption{Distributions of mismatches for both optimization schemes in all 4 regions of Fig.~\ref{fig:ps_q148_quadrant}. 
-	We use a constant PSD to calculate the mismatch and
+	\includegraphics[width=\textwidth]{figures/all_quadrants.pdf}
+	\caption{Distributions of mismatches for both split region optimization and all region optimization in all 4 
+	regions. We use a constant PSD to calculate the mismatch and
 	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distributions are indicated by dashed
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{\textit{Top}: Fractional mismatch change of the validation waveforms ().
-	Each quadrant is fitted independently to the NR waveforms listed in Tab.~\ref{tab:quadrants}.
-	\textit{Bottom}: Combined distribution of both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
+	\caption{Combined distribution of both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
 	\label{fig:ps_q148_quadrant}

### 2023-06-27 src/scripts/ps_q148_quadrants.py

@@ -55,9 +55,9 @@ bins = np.linspace(-5, -1, 15)
 
 plt.figure(figsize=(6, 5))
 plt.hist(np.log10(data_q148[:,5]), alpha=0.2, bins=bins, color=c2)
-plt.hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2, label='Optimized (All Regions)')
+plt.hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2, label='Opt. (All)')
 plt.hist(np.log10(data_region[:,5]), alpha=0.2, bins=bins, color=c3)
-plt.hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Optimized (Split Regions)')
+plt.hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Opt. (Split)')
 plt.axvline(x=np.log10(np.median(data_q148[:,5])), color=c2, ls='--', lw=2)
 plt.axvline(x=np.log10(np.median(data_region[:,5])), color=c3, ls='--', lw=2)
 plt.xlabel("$\log_{10}(\mathcal{M})$")

### 2023-06-27 src/scripts/all_quadrants.py
fix

@@ -60,12 +60,12 @@ ax[0, 0].hist(np.log10(data1[:, 4]), bins=bins, alpha=0.2, color=c1)
 ax[0, 0].hist(np.log10(data1[:, 4]), bins=bins, label="Original", color=c1, lw=2, histtype="step")
 ax[0,0].axvline(x=np.log10(np.median(data1[:, 4])), color=c1,zorder=-20,lw=2,alpha=1.0,linestyle=(0, (1, 1.3)))
 
-ax[0, 0].hist(np.log10(data1[:, 5]),bins=bins,label="Optimized (Split Regions)",histtype="step",color=c3,linewidth=2)
+ax[0, 0].hist(np.log10(data1[:, 5]),bins=bins,label="Opt. (Split)",histtype="step",color=c3,linewidth=2)
 ax[0, 0].hist(np.log10(data1[:, 5]),bins=bins,color=c3,alpha=0.2)
 ax[0, 0].axvline(x=np.log10(np.median(data1[:, 5])),color=c3,lw=2,alpha=1.0,linestyle=(0, (1, 1.3)))
 
 ax[0, 0].hist(np.log10(data1_q148[:, 5]), bins=bins, alpha=0.2, color=c2)
-ax[0, 0].hist(np.log10(data1_q148[:, 5]), bins=bins, label="Optimized (All Regions)", color=c2, lw=2, histtype="step")
+ax[0, 0].hist(np.log10(data1_q148[:, 5]), bins=bins, label="Opt. (All)", color=c2, lw=2, histtype="step")
 ax[0, 0].axvline(x=np.log10(np.median(data1_q148[:, 5])),color=c2,lw=2,alpha=1.0,linestyle=(0, (1, 1.3)))
 
 ax[0, 0].set_ylabel("Count")

### 2023-06-28 src/tex/ms.tex
Final typos

@@
 of the optimized waveforms with the original waveforms. We also show how the
-optimization result differs as a function of the source parameters. In
+optimization result differs as a function of the source parameters. Finally, in
 Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
@@
 parameterization affects the accuracy of the model. 
+Note that throughout this paper we use the terms recalibration and optimization interchangeably.
 
@@
 boundary between each section.
+
+\begin{figure}
+	\script{loss.py}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/loss.pdf}
+	\caption{Loss functions against the number of iterations. The vertical
+	dotted line indicates the minimum of the validation loss, which is where we stop
+	the optimization.}
+	\label{fig:loss}
+\end{figure}
 
@@
 two waveforms.
+
+\begin{figure}[t]
+	\script{intrin_space.py}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
+	\caption{Distribution of training and validation NR waveforms in the
+	$q-\chi_{\rm PN}$ space. The blue stars represent the waveforms used to compute
+	the training loss, green triangles represent extra training data, and the 
+	orange dots represent the waveforms used for validation.}
+	\label{fig:intrin_space}
+\end{figure}
 
@@
 
-\begin{figure}
-	\script{loss.py}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{Loss functions against the number of iterations. The vertical
-	dotted line indicates minimum of the validation loss, which is where we stop
-	the optimization.}
-	\label{fig:loss}
-\end{figure}
-
 \begin{algorithm}[t]
@@
 \end{algorithm}
-
-\begin{figure}[t]
-	\script{intrin_space.py}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
-	\caption{Distribution of training and validation NR waveforms in the
-	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
-	the training loss, green triangles represent extra training data, and the 
-	orange dots represent the waveforms used for validation.}
-	\label{fig:intrin_space}
-\end{figure}
 
@@
 	\end{tabularx}
-	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
+	\caption{List of NR waveforms used to recalibrate the model. The mass ratio is defined as
 	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
@@
 	amplitude (left) and phase (right) of the NR, original IMRPhenomD, and optimized IMRPhenomD
-	waveform. \textit{Bottom}: Here we show the relative error between the NR and
+	waveforms. \textit{Bottom}: Here we show the relative error between the NR and
 	IMRPhenomD waveform ampltiudes (left) as well as the absolute error of the phases between the NR and
@@
 	\includegraphics[width=\columnwidth]{figures/q148.pdf}
-	\caption{Distributions of waveform mismatches calculated using 
-	$\mathcal{L}_{\mathrm{ave}}$. We use the training waveforms listed in 
-	Tab.~\ref{tab:q148} and	mismatches are weighted with a constant PSD. 
+	\caption{Distributions of mismatches before and after optimization using the 
+	$\mathcal{L}_{\mathrm{ave}}$ loss function. Mismatches are calculated using the training waveforms listed in 
+	Tab.~\ref{tab:q148} and	are weighted with a constant PSD. 
 	The dashed lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
@@
 	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
-	The dashed lines represent the median of the distributions, which decreased by 10.4\% during optimization.}
+	The dashed lines represent the median of the distributions, which decreased \te{by an additional 10.4\% during optimization with the additional data.}}
 	\label{fig:q148_q1248_compare}
@@
 	regions. We use a constant PSD to calculate the mismatch and
-	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distributions are indicated by dashed
+	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distribution are indicated by dashed
 	lines.}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Combined distribution of both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
+	\caption{Combined mismatch distributions for both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
 	\label{fig:ps_q148_quadrant}

### 2023-06-28 src/scripts/ps_q148_quadrants.py

@@ -54,14 +54,40 @@ logdiff = np.log10(data_region[:, 5] / data_region[:, 4])
 bins = np.linspace(-5, -1, 15)
 
 plt.figure(figsize=(6, 5))
-plt.hist(np.log10(data_q148[:,5]), alpha=0.2, bins=bins, color=c2)
-plt.hist(np.log10(data_q148[:,5]), histtype='step', lw=2, bins=bins, color=c2, label='Opt. (All)')
-plt.hist(np.log10(data_region[:,5]), alpha=0.2, bins=bins, color=c3)
-plt.hist(np.log10(data_region[:,5]), histtype='step', lw=2, bins=bins, color=c3, label='Opt. (Split)')
-plt.axvline(x=np.log10(np.median(data_q148[:,5])), color=c2, ls='--', lw=2)
-plt.axvline(x=np.log10(np.median(data_region[:,5])), color=c3, ls='--', lw=2)
+plt.hist(np.log10(data_q148[:, 5]), alpha=0.2, bins=bins, color=c2)
+plt.hist(
+    np.log10(data_q148[:, 5]),
+    histtype="step",
+    lw=2,
+    bins=bins,
+    color=c2,
+    label="Opt. (All)",
+)
+plt.hist(np.log10(data_region[:, 5]), alpha=0.2, bins=bins, color=c3)
+plt.hist(
+    np.log10(data_region[:, 5]),
+    histtype="step",
+    lw=2,
+    bins=bins,
+    color=c3,
+    label="Opt. (Split)",
+)
+plt.axvline(
+    x=np.log10(np.median(data_q148[:, 5])),
+    color=c2,
+    zorder=-20,
+    lw=2,
+    linestyle=(0, (1, 1.3)),
+)
+plt.axvline(
+    x=np.log10(np.median(data_region[:, 5])),
+    color=c3,
+    zorder=-20,
+    lw=2,
+    linestyle=(0, (1, 1.3)),
+)
 plt.xlabel("$\log_{10}(\mathcal{M})$")
 plt.ylabel("Count")
-plt.legend(loc='upper left')
+plt.legend(loc="upper left")
 
 plt.savefig(paths.figures / "ps_q148_quadrants.pdf", bbox_inches="tight")

### 2023-06-28 src/scripts/all_quadrants.py
Fixing fig 8

@@ -44,10 +44,22 @@ data3 = np.loadtxt(paths.data / "1neg2neg_mismatch.dat")
 data4 = np.loadtxt(paths.data / "q148_1pos2neg_region_f.dat")
 
 data1_q148_ = np.loadtxt(paths.data / "q148.dat")
-data1_q148 = np.array([data1_q148_[i] for i in range(len(data1_q148_[:,0])) if data1_q148_[i,2] > 0 and data1_q148_[i,3] > 0])
+data1_q148 = np.array(
+    [
+        data1_q148_[i]
+        for i in range(len(data1_q148_[:, 0]))
+        if data1_q148_[i, 2] > 0 and data1_q148_[i, 3] > 0
+    ]
+)
 data2_q148 = np.loadtxt(paths.data / "q148_1neg2pos_all_f.dat")
 data3_q148_ = np.loadtxt(paths.data / "q148.dat")
-data3_q148 = np.array([data3_q148_[i] for i in range(len(data3_q148_[:,0])) if data3_q148_[i,2] < 0 and data3_q148_[i,3] < 0])
+data3_q148 = np.array(
+    [
+        data3_q148_[i]
+        for i in range(len(data3_q148_[:, 0]))
+        if data3_q148_[i, 2] < 0 and data3_q148_[i, 3] < 0
+    ]
+)
 data4_q148 = np.loadtxt(paths.data / "q148_1pos2neg_all_f.dat")
 
 bins = np.linspace(-5, -1, 15)
@@ -56,96 +68,250 @@ fig, ax = plt.subplots(2, 2, figsize=(12, 10), sharex=True)
 plt.subplots_adjust(hspace=0.05)
 plt.subplots_adjust(wspace=0.04)
 
-ax[0, 0].hist(np.log10(data1[:, 4]), bins=bins, alpha=0.2, color=c1)
-ax[0, 0].hist(np.log10(data1[:, 4]), bins=bins, label="Original", color=c1, lw=2, histtype="step")
-ax[0,0].axvline(x=np.log10(np.median(data1[:, 4])), color=c1,zorder=-20,lw=2,alpha=1.0,linestyle=(0, (1, 1.3)))
+ax[0, 1].hist(np.log10(data1[:, 4]), bins=bins, alpha=0.2, color=c1, density=True)
+ax[0, 1].hist(
[…]

### 2023-06-28 src/tex/ms.tex
Update

@@
 expected to double the total number of observed binary black holes (BBHs)~\citep{abbott2020prospects}.
-Morover, the improved sensitivity also implies that we expect to
+Moreover, the improved sensitivity also implies that we expect to
 detect individual events with a higher signal-to-noise ratio (SNR) than ever before. This means
@@
 %	\te{What is $t_{\mathrm{RD}}$? and how is it computed?}
-} before using standard FFT routines to compute the Fourier transform.
+} before using standard FFT routines to compute the fast Fourier transform.
 
@@
 not in the training set, we evaluate the mismatch between the fine-tuned model
-and an additional 536 NR waveforms in the SXS catalog i.e., the validation set.
+and an additional 526 NR waveforms in the SXS catalog i.e., the validation set.
 We select waveforms that share the same part of the parameter space with the training
@@
 	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
-	The dashed lines represent the median of the distributions, which decreased \te{by an additional 10.4\% during optimization with the additional data.}}
+	The dashed lines represent the median of the distributions, which decreased by an additional 10.8\% during optimization with the additional data.}
 	\label{fig:q148_q1248_compare}
@@
 } We would therefore like to examine whether using the {\zdethp} PSD in our loss function
-could lead to an improved to mismatch. Performing the optimization again, we find no significant difference
+could lead to an improved mismatch. Performing the optimization again, we find no significant difference
 between the results using the two PSDs in the distribution of mismatches.
@@
 \kl{Both top-right and bottom-left region have good performance comparing with the original model.
-Especially for the bottom-right region, where the improvement is significantly better than optimizing with all waveforms.
+Especially for the bottom-left region, where the improvement is significantly better than optimizing all regions simultaneously.
 This suggests the ansatz are well-fitted to this part of the parameter space.
 Waveforms above the diagonal also give good improvements when comparing with the original model and it has a similar improvement as optimizing all waveforms at once.
-Notice optimizing over all waveforms gives a slightly better improvement.
+Notice optimizing all regions gives a slightly better improvement.
 This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
@@
 Although split region optimization uses more training waveforms, some of the waveforms, e.g.
-opposite spins, hinders the optimization procedure and hence gives a slightly worse result when comparing with optimization over all waveforms.}
+opposite spins, hinders the optimization procedure and hence gives a slightly worse result when comparing with all region optimization.}
 

### 2023-06-28 src/tex/ms.tex
final check

@@
 bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions. 
-From top panel of Fig.~\ref{fig:ps_q148_quadrant}, we observe that
-equal-spin waveforms lying on the diagonal improve the most. 
-\kl{Both top-right and bottom-left region have good performance comparing with the original model.
+\kl{From Fig.~\ref{fig:all_quadrants}, we observe that both top-right and bottom-left region have good performance comparing with the original model.
 Especially for the bottom-left region, where the improvement is significantly better than optimizing all regions simultaneously.
 This suggests the ansatz are well-fitted to this part of the parameter space.
-Waveforms above the diagonal also give good improvements when comparing with the original model and it has a similar improvement as optimizing all waveforms at once.
+Waveforms lying in the top-left region also give good improvements when comparing with the original model and it has a similar improvement as optimizing all waveforms at once.
 Notice optimizing all regions gives a slightly better improvement.
 This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
-%The improvement above the diagonal is also significantly more than when optimizing using
-%all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
-%that the optimized histogram shifts to lower mismatches.
-\footnote{
-	Note that the bottom-left panel of Fig.~\ref{fig:all_quadrants} specifically shows the
-	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
-} 
-On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
+On the other hand, the improvement in the bottom-right region is negligible, indicating that our method struggles to
 further improve over the original IMRPhenomD waveform. 
@@
 $\chi_1>0$ and $\chi_2<0$ less well. 
-\kl{Overall, the improvement for both optimization schemes are similar (bottom panel of Fig.~\ref{fig:ps_q148_quadrant}).
+\kl{Overall, the improvement for both optimization schemes are similar (Fig.~\ref{fig:ps_q148_quadrant}).
 Although split region optimization uses more training waveforms, some of the waveforms, e.g.

### 2023-06-29 src/tex/ms.tex
Editing final paragraph of section 3

@@
 are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
-bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions. 
-\kl{From Fig.~\ref{fig:all_quadrants}, we observe that both top-right and bottom-left region have good performance comparing with the original model.
-Especially for the bottom-left region, where the improvement is significantly better than optimizing all regions simultaneously.
-This suggests the ansatz are well-fitted to this part of the parameter space.
-Waveforms lying in the top-left region also give good improvements when comparing with the original model and it has a similar improvement as optimizing all waveforms at once.
-Notice optimizing all regions gives a slightly better improvement.
-This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
-On the other hand, the improvement in the bottom-right region is negligible, indicating that our method struggles to
-further improve over the original IMRPhenomD waveform. 
-Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
-to the overall mismatch (see bottom-right panel of Fig.~\ref{fig:all_quadrants}). 
-This suggests that the IMRPhenomD waveform ansatz fits waveforms with
-$\chi_1>0$ and $\chi_2<0$ less well. 
-\kl{Overall, the improvement for both optimization schemes are similar (Fig.~\ref{fig:ps_q148_quadrant}).
-Although split region optimization uses more training waveforms, some of the waveforms, e.g.
-opposite spins, hinders the optimization procedure and hence gives a slightly worse result when comparing with all region optimization.}
+bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we only use waveforms with $q\leq4$ as validation waveforms for these two regions. 
+From Fig.~\ref{fig:all_quadrants}, we observe that both the top-right and bottom-left regions improve significantly over the original model.
+This is especially pronounced for the bottom-left region, where the improvement is significantly better than optimizing all regions simultaneously.
+This suggests the ansatz fits this part of the parameter space well.
+The top-left region also improves over the original model although it is similar to when optimizing all waveforms at once.
+% Notice optimizing all regions gives a slightly better improvement.
+% This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
+On the other hand, the bottom-right region does not improve over the original model, indicating that a change to the ansatz is required to fit the NR data better.
+% Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
+% to the overall mismatch (see bottom-right panel of Fig.~\ref{fig:all_quadrants}). 
+% This suggests that the IMRPhenomD waveform ansatz fits waveforms with
+% $\chi_1>0$ and $\chi_2<0$ less well.
+
+Overall, the improvement for both optimization schemes are similar (Fig.~\ref{fig:ps_q148_quadrant}).
+Although split region optimization uses more training waveforms, some of the waveforms,
+e.g. opposite spins, hinders the optimization procedure and hence gives a slightly worse result when comparing with all region optimization.
 

### 2023-09-08 src/tex/ms.tex
fix some minor comments

@@
-% Define document class
 \documentclass[twocolumn]{aastex631}
 \usepackage{showyourwork}
-
 \usepackage{graphicx}
@@
 \newcommand{\zdethp}{\texttt{zdethp}}
+
 \definecolor{rb4}{HTML}{27408B}
@@
 \newcommand{\te}[1]{{\color{rr}[TE: #1 ]}}
+\definecolor{cyan}{HTML}{0097A7}
+\newcommand{\rr}[1]{{\color{cyan}[#1]}}
 
@@
 
-% Begin!
 \begin{document}
 
-% Title
 \title{Recalibrating Gravitational Wave Phenomenological Waveform Model}
 
-% Author list
 \author{Kelvin K.~H.~Lam} 
@@
 
-% Brief intro of waveform models
-
 Many data analysis tasks in gravitational wave (GW) astrophysics, such
-as match filtering~\citep{owen1996search, owen1999matched} and parameter 
+as \rr{matched} filtering~\citep{owen1996search, owen1999matched} and parameter 
 estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
@@
 
-% Why there is a need
-
 The LIGO-VIRGO-KAGRA (LVK) collaboration
@@
 stringent requirements on the accuracy of our waveform model \citep{purrer2020gravitational, hu2022assessing}.
-
-% AD waveforms
 
@@
 
-% Put more focus on the joint optimization instead of having a set of more accurate 
-% coefficients, since the method is the main point
 In this paper, we investigate the possibility of further improving the accuracy
@@
 
-% Paragraphs we need:
-% 1. What is the model
-% 2. How do the parameters enter the model
 We start by giving a succinct summary of the IMRPhenomD model and the relevant parameters.
@@
 \begin{align}\label{eq:}
-	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
+	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda^i)}\,,
 \end{align}
-where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
-parameters which will be discussed below. 
+where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \rr{($i=1,2,\cdots,19$) is a set of 19 additional
+parameters} which will be discussed below. 
 The phase and amplitude functions are then split into three sections which represent the
@@
 } 
-which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
+which depend purely on $\Lambda^i$ and can be calibrated to numerical relativity (NR) simulations.
 The three sections are then \textit{stitched} together using step functions.
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{Loss functions against the number of iterations. The vertical
+	\caption{\rr{Average Loss $\mathcal{L}_{\mathrm{ave}}$} against the number of iterations. The vertical
 	dotted line indicates the minimum of the validation loss, which is where we stop
@@
 
-In practice the $\Lambda$ parameters are fit for each section independently, i.e., intermediate coefficients are fit whilst ignoring the MR region.
-Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
+In practice the $\Lambda^i$ parameters are fit for each section independently, i.e., intermediate coefficients are fit whilst ignoring the MR region.
+Finally, to map the grid of tuned $\Lambda^i$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
 \begin{align} \label{eq:Lambda}
@@
 
-%\kw{Need to work on this paragraph}
 Although initially independent, the stitching procedure means that each section
@@
 the maximum (minimum) frequency for the integration.
-We note here that the mismatch can be viewed as the mean square error (MSE) between the
+We note here that the mismatch can be viewed as \rr{a} mean square error (MSE) between the
 two waveforms.
@@
 where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
-$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
-training waveform, and $N$ is the total number of training waveforms.
+$\mathcal{M}_{i,\mathrm{ini}}$ represents the mismatch \rr{between original IMRPhenomD waveforms and NR waveforms}, and $N$ is the total number of training waveforms.
 
@@
 behavior during the optimization process.
-The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
+The simple average, \rr{$\mathcal{L}_{\mathrm{ave}}$}, serves as the simplest choice of loss function,
 but is prone to be dominated by a single point in parameter space with a large mismatch.
@@
 The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
-Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
+Conversely, \rr{$\mathcal{L}_{\mathrm{ave}}$} allows the loss function to automatically adjust
 and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
@@
 in order to compute the mismatch, \eqref{eq:mismatch}.
-For this, we taper the time-series using Tukey window~\citep{usman2016pycbc}\footnote{
+For this, we taper the time-series using Tukey window \citep{usman2016pycbc}
+\footnote{
 	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
@@
 	entire GW strain. 
-%	\te{What is $t_{\mathrm{RD}}$? and how is it computed?}
-} before using standard FFT routines to compute the fast Fourier transform.
+} 
+before using standard FFT routines to compute the fast Fourier transform.
 
@@
 	\KwIn{initial coefficients $\lambda_i$} \Parameters{number of iterations
-		$N$, learning rate $\alpha$} \Variables{current coefficients $\lambda$,
+		$M$, learning rate $\alpha$} \Variables{current coefficients $\lambda$,
 		mismatch gradient $\nabla\mathcal{L}$} \KwResult{output coefficients
@@
 	\tcc{Gradient Descent}
-	\For{$i<N$}{ $\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
+	\For{$i<M$}{ $\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
 		$\nabla\mathcal{L}\leftarrow AutoDiff(\mathcal{L})$\\
@@
 	Tab.~\ref{tab:q148} and	are weighted with a constant PSD. 
-	The dashed lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
+	The dotted lines represent the median of the distributions, which decreased by 45.3\% during optimization.}
 	\label{fig:q148}
@@
 	using a constant PSD with the loss function $\mathcal{L}_{\mathrm{ave}}$. 
-	The dashed lines represent the median of the distributions, which decreased by an additional 10.8\% during optimization with the additional data.}
+	The dotted lines represent the median of the distributions, which decreased by an additional 10.8\% during optimization with the additional data.}
 	\label{fig:q148_q1248_compare}
@@
 The top-left region also improves over the original model although it is similar to when optimizing all waveforms at once.
-% Notice optimizing all regions gives a slightly better improvement.
-% This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
 On the other hand, the bottom-right region does not improve over the original model, indicating that a change to the ansatz is required to fit the NR data better.
-% Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
-% to the overall mismatch (see bottom-right panel of Fig.~\ref{fig:all_quadrants}). 
-% This suggests that the IMRPhenomD waveform ansatz fits waveforms with
-% $\chi_1>0$ and $\chi_2<0$ less well.
 
@@
 	regions. We use a constant PSD to calculate the mismatch and
-	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distribution are indicated by dashed
+	$\mathcal{L}_{\mathrm{ave}}$ as the loss function. The medians of each distribution are indicated by dotted
 	lines.}
@@
 	\includegraphics[width=\columnwidth]{figures/ps_q148_quadrants.pdf}
-	\caption{Combined mismatch distributions for both optimization schemes shown in Fig.~\ref{fig:all_quadrants}.}
+	\caption{Combined mismatch distributions for both optimization schemes shown in Fig.~\ref{fig:all_quadrants}. The medians of both distributions are indicated by dotted lines. }
 	\label{fig:ps_q148_quadrant}
@@
 updated with new waveforms, and future work should carefully examine whether these more
-modern waveforms can be improved. % Extend this a bit?
+modern waveforms can be improved.
 
@@
 supported by the Horizon Postdoctoral Fellowship.
+
 \bibliography{bib}

### 2023-10-02 src/tex/ms.tex
added responses

@@
 \usepackage{algorithmic}
+
+\usepackage[normalem]{ulem}
 
@@
 Originally, 19 NR simulations were used to calibrate IMRPhenomD \citep{khan2016frequency}; nine of these 
-are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
+are from the SXS catalog 
+\footnote{We took the CoM-corrected waveforms with 4th-order extrapolation.} 
+\citep{boyle2019sxs} and 10 BAM simulations. 
 As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
@@
 recalibration process struggles to improve in the regime where $\chi_{1}$ is
-positive and $\chi_{2}$ is negative. This means there is a dependency between
-the two spins that is not captured by the current ansatz. A systematic scheme that
+positive and $\chi_{2}$ is negative. \kl{\sout{This means there is a dependency between
+the two spins that is not captured by the current ansatz.}} A systematic scheme that
 incorporates our calibration method with waveform design will be handy in
@@
 
-In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions purely for simplicity.
-These simple cuts demonstrate the accuracy of the waveform can be further improved,
-but they are almost certainly not the optimal way to incorporate the extra information
-we have about the waveform. A more general way such as adding new functional dependency
-between the parameters should be explored in the future.
+\kl{We note that in comparison to the original IMRPhenomD calibration, we restricted the 
+parameter space to $q\leq8$ and directly used NR waveform instead of hybridized SEOB-NR 
+waveforms, which clearly put the original IMRPhemonD result in a 
+disadvantage. We have tried to train the model using a set of training waveforms of the same 
+intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. The result 
+shows a marginal improvement, which corresponds to the stitching problem discussed before.
+However, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio regime, 
+and is rarely used in this region of parameter space. Hence, we focus on showing AD with 
+gradient descent, as a general and robust method, can be used to further optimize models 
+independent of the training waveforms used.}
+
+In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions 
+purely for simplicity. These simple cuts demonstrate the accuracy of the waveform can be 
+further improved, but they are almost certainly not the optimal way to incorporate the 
+extra information we have about the waveform. A more general way such as adding new 
+functional dependency between the parameters should be explored in the future.
 

### 2023-10-02 src/tex/ms.tex
add figure

@@
 During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
-IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
+IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up 
+to 3.5PN order \kl{augmented with the next four higher order PN terms}.
 To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
@@
 with the peak of the distribution being shifted by approximately an order of
-magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
+magnitude. The median mismatch is reduced by 50\% (see vertical dotted lines) \kl{while the tail hardly changes}. When using
 $\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
@@
 parameter space to $q\leq8$ and directly used NR waveform instead of hybridized SEOB-NR 
-waveforms, which clearly put the original IMRPhemonD result in a 
-disadvantage. We have tried to train the model using a set of training waveforms of the same 
-intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. The result 
-shows a marginal improvement, which corresponds to the stitching problem discussed before.
-However, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio regime, 
-and is rarely used in this region of parameter space. Hence, we focus on showing AD with 
-gradient descent, as a general and robust method, can be used to further optimize models 
-independent of the training waveforms used.}
+waveforms in training and validation purposes. These choices put the original IMRPhemonD 
+result in a disadvantage, as both the parameter space and frequency range are restricted. 
+We tested our result by training the model using a set of training waveforms of the same 
+intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. We can see 
+the result in Fig.~\ref{fig:q14818}, where the mismatch median changes by around $8\%$, 
+This improvement mainly corresponds to the stitching problem discussed before. 
+Nevertheless, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio 
+regime, and is rarely used in this region of parameter space. Hence, we focus on showing 
+AD with gradient descent, as a general and robust method, can be used to further optimize 
+models independent of the training waveforms used.}
+
+\begin{figure}[t]
+	\script{q14818.py}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/q14818.pdf}
+	\caption{Distributions of mismatches before and after optimization using the 
+		$\mathcal{L}_{\mathrm{ave}}$ loss function.
+Mismatches are calculated using the training waveforms listed in \cite{khan2016frequency}, but BAM waveforms are replaced by SEOBNRv4 waveforms of the same intrinsic parameters.
+		Mismatches are calculated with a constant PSD. The dotted lines represent the median of the distributions, which decreased by $8.2\%$ during optimization.}
+	\label{fig:q14818}
+\end{figure}
 

### 2023-10-02 src/scripts/q14818.py

@@ -1,0 +1,88 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import paths
+
+params = {
+    "font.size": 15,
+    "legend.fontsize": 12,
+    "legend.frameon": False,
+    "axes.labelsize": 15,
+    "axes.titlesize": 15,
+    "xtick.labelsize": 15,
+    "ytick.labelsize": 15,
+    "axes.unicode_minus": False,
+    "figure.figsize": (6, 5),
+    "xtick.top": True,
+    "ytick.right": True,
+    "xtick.bottom": True,
+    "ytick.left": True,
+    "xtick.major.pad": 8,
+    "xtick.major.size": 8,
+    "xtick.minor.size": 4,
+    "ytick.major.size": 8,
+    "ytick.minor.size": 4,
+    "xtick.direction": "in",
+    "ytick.direction": "in",
+    "axes.linewidth": 1.5,
+    "text.usetex": False,
+    "font.family": "serif",
+    "font.serif": "cmr10",
+    "mathtext.fontset": "cm",
+    "axes.formatter.use_mathtext": True,  # needed when using cm=cmr10 for normal text
+}
+
+mpl.rcParams.update(params)
+
+c1 = "#102F68"
+c2 = "#B02423"
+
+
+q14818 = np.loadtxt(paths.data / "q14818.dat")
+bins = np.linspace(-5, -1, 15)
+
+plt.hist(np.log10(q14818[:, 4]), bins=bins, alpha=0.2, color=c1)
+plt.hist(
+    np.log10(q14818[:, 4]),
+    bins=bins,
+    color=c1,
+    histtype="step",
+    linewidth=2,
+    label="Original",
+)
+plt.axvline(
+    x=np.log10(np.median(q14818[:, 4])),
+    color=c1,
+    zorder=-20,
+    alpha=1.0,
+    lw=2,
+    linestyle=(0, (1, 1.3)),
+)
+
+plt.hist(
+    np.log10(q14818[:, 5]),
+    bins=bins,
+    histtype="step",
+    label="Optimized $(\mathcal{L}_{\mathrm{ave}})$",
+    color=c2,
+    linewidth=2,
+)
+plt.hist(
+    np.log10(q14818[:, 5]),
+    bins=bins,
+    color=c2,
+    alpha=0.2,
+)
+plt.axvline(
[…]

### 2024-01-05 src/tex/ms.tex
change color of comments

@@
 \definecolor{cyan}{HTML}{0097A7}
-\newcommand{\rr}[1]{{\color{cyan}[#1]}}
+\newcommand{\corr}[1]{{\color{red}[corr: #1 ]}}
 
@@
 Many data analysis tasks in gravitational wave (GW) astrophysics, such
-as \rr{matched} filtering~\citep{owen1996search, owen1999matched} and parameter 
+as \corr{matched} filtering~\citep{owen1996search, owen1999matched} and parameter 
 estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
@@
 \end{align}
-where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \rr{($i=1,2,\cdots,19$) is a set of 19 additional
+where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \corr{($i=1,2,\cdots,19$) is a set of 19 additional
 parameters} which will be discussed below. 
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{\rr{Average Loss $\mathcal{L}_{\mathrm{ave}}$} against the number of iterations. The vertical
+	\caption{\corr{Average Loss $\mathcal{L}_{\mathrm{ave}}$} against the number of iterations. The vertical
 	dotted line indicates the minimum of the validation loss, which is where we stop
@@
 the maximum (minimum) frequency for the integration.
-We note here that the mismatch can be viewed as \rr{a} mean square error (MSE) between the
+We note here that the mismatch can be viewed as \corr{a} mean square error (MSE) between the
 two waveforms.
@@
 where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
-$\mathcal{M}_{i,\mathrm{ini}}$ represents the mismatch \rr{between original IMRPhenomD waveforms and NR waveforms}, and $N$ is the total number of training waveforms.
+$\mathcal{M}_{i,\mathrm{ini}}$ represents the mismatch \corr{between original IMRPhenomD waveforms and NR waveforms}, and $N$ is the total number of training waveforms.
 
@@
 behavior during the optimization process.
-The simple average, \rr{$\mathcal{L}_{\mathrm{ave}}$}, serves as the simplest choice of loss function,
+The simple average, \corr{$\mathcal{L}_{\mathrm{ave}}$}, serves as the simplest choice of loss function,
 but is prone to be dominated by a single point in parameter space with a large mismatch.
@@
 The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
-Conversely, \rr{$\mathcal{L}_{\mathrm{ave}}$} allows the loss function to automatically adjust
+Conversely, \corr{$\mathcal{L}_{\mathrm{ave}}$} allows the loss function to automatically adjust
 and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.

### 2024-01-11 src/tex/ms.tex
color fix and minor word changes

@@
 IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up 
-to 3.5PN order \kl{augmented with the next four higher order PN terms}.
+to 3.5PN order \corr{augmented with the next four higher order PN terms}.
 To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
@@
 are from the SXS catalog 
-\footnote{We took the CoM-corrected waveforms with 4th-order extrapolation.} 
-\citep{boyle2019sxs} and 10 BAM simulations. 
+\citep{boyle2019sxs} and 10 BAM simulations\footnote{\corr{We took the CoM-corrected waveforms with 4th-order extrapolation.}} . 
 As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
@@
 with the peak of the distribution being shifted by approximately an order of
-magnitude. The median mismatch is reduced by 50\% (see vertical dotted lines) \kl{while the tail hardly changes}. When using
+magnitude. The median mismatch is reduced by 50\% (see vertical dotted lines) \corr{while the tail hardly changes}. When using
 $\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
@@
 recalibration process struggles to improve in the regime where $\chi_{1}$ is
-positive and $\chi_{2}$ is negative. \kl{\sout{This means there is a dependency between
-the two spins that is not captured by the current ansatz.}} A systematic scheme that
+positive and $\chi_{2}$ is negative. \corr{\sout{This means there is a dependency between
+the two spins} \sout{that is not captured by the current ansatz.}} A systematic scheme that
 incorporates our calibration method with waveform design will be handy in
@@
 
-\kl{We note that in comparison to the original IMRPhenomD calibration, we restricted the 
-parameter space to $q\leq8$ and directly used NR waveform instead of hybridized SEOB-NR 
-waveforms in training and validation purposes. These choices put the original IMRPhemonD 
-result in a disadvantage, as both the parameter space and frequency range are restricted. 
-We tested our result by training the model using a set of training waveforms of the same 
-intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. We can see 
-the result in Fig.~\ref{fig:q14818}, where the mismatch median changes by around $8\%$, 
-This improvement mainly corresponds to the stitching problem discussed before. 
-Nevertheless, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio 
-regime, and is rarely used in this region of parameter space. Hence, we focus on showing 
-AD with gradient descent, as a general and robust method, can be used to further optimize 
-models independent of the training waveforms used.}
+\corr{We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
+of $q=18$ are used in the original calibration trials. Also, we have directly used NR 
+waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
+These choices put the original IMRPhenomD result at a disadvantage, as both the parameter 
+space and frequency range are restricted. We tested our result by training the model 
+using a set of training waveforms of the same intrinsic parameters, but with BAM waveforms 
+replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
+mismatch median decreases by about $8\%$, This mismatch improvement mainly concerns the 
+stitching problem discussed before. Nevertheless, it is known that IMRPhenomD has a large 
+inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
+space. Hence, we focus on showing AD with gradient descent, as a general and robust method, 
+can be used to further optimize GW waveform models with arbitrary sets of training 
+waveforms used.}
 

### 2024-01-13 src/tex/ms.tex
added response 4

@@
 \begin{align}\label{eq:}
-	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda^i)}\,,
+	h(f,\theta,\Lambda^i) = A(f,\theta,\Lambda^i)e^{-i\phi(f,\theta,\Lambda^i)}\,,
 \end{align}
@@
 recalibration process struggles to improve in the regime where $\chi_{1}$ is 
-positive and $\chi_{2}$ is negative. \corr{\sout{This means there is a dependency between
-the two spins} \sout{that is not captured by the current ansatz.}} A systematic scheme that
-incorporates our calibration method with waveform design will be handy in
-constructing new waveform models in the future.
+positive and $\chi_{2}$ is negative. This means there is a dependency between the two 
+spins that is not captured in the current ansatz. \corr{In fact, this issue can 
+be explained by PN theory, where the leading spin-orbit coupling term depends on 
+the effective spin while higher order terms depends on both $\chi_1$ and $\chi_2$. 
+Nevertheless, we hope to show a systematic scheme that could reveal hidden degeneracies, 
+as illustrated by the effective spin degeneracy. Hence, such systematic scheme that 
+incorporates our calibration method with waveform design will be handy in constructing 
+new waveform models in the future. }
 
 \corr{We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
-of $q=18$ are used in the original calibration trials. Also, we have directly used NR 
+of $q=18$ are used in the original calibration trial. Also, we have directly used NR 
 waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
@@
 
-In making Fig.~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions 
+In making Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}, we split the parameter space into four regions 
 purely for simplicity. These simple cuts demonstrate the accuracy of the waveform can be 

### 2024-02-01 src/tex/ms.tex
slight text change

@@
 replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
-mismatch median decreases by about $8\%$, This mismatch improvement mainly concerns the 
-stitching problem discussed before. Nevertheless, it is known that IMRPhenomD has a large 
+mismatch median decreases by about $8\%$. Nevertheless, it is known that IMRPhenomD has a large 
 inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
-space. Hence, we focus on showing AD with gradient descent, as a general and robust method, 
+space. Hence, we focus on showing that AD with gradient descent, as a general and robust method, 
 can be used to further optimize GW waveform models with arbitrary sets of training 

### 2024-02-16 src/tex/ms.tex
final cleanup for submission

@@
 \newcommand{\zdethp}{\texttt{zdethp}}
-
-\definecolor{rb4}{HTML}{27408B}
-\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
-\definecolor{cyan}{HTML}{0097A7}
-\newcommand{\kl}[1]{{\color{cyan}[KL: #1 ]}}
-\definecolor{rr}{RGB}{173, 37, 37}
-\newcommand{\te}[1]{{\color{rr}[TE: #1 ]}}
-\definecolor{cyan}{HTML}{0097A7}
-\newcommand{\corr}[1]{{\color{red}[corr: #1 ]}}
 
@@
 	waveforms. When IMRPhenomD was first calibrated to NR waveforms, different
-	parts (i.e., the inspiral, merger, and ringdown) of the waveform were calibrated separately. 
-	Using \ripple, a library of waveform models compatible with automatic differentiation, we can, for the first time,
-	perform gradient-based optimization on all the waveform coefficients at the same time.
-	This joint optimization process allows us to capture previously ignored correlations between separate parts of the waveform.
-	We found that after recalibration, the median mismatch between the model and NR waveforms decreases by $50\%$.
-	We further explore how different regions of the source parameter space respond
-	to the optimization procedure. We find that the degree of improvement
-	correlates with the spins of the source. This work shows a promising avenue
-	to help understand and treat systematic error in waveform models.
+	parts (i.e., the inspiral, merger, and ringdown) of the waveform were 
+	calibrated separately. 	Using \ripple, a library of waveform models 
+	compatible with automatic differentiation, we can, for the first time, 
+	perform gradient-based optimization on all the waveform coefficients at the 
+	same time. This joint optimization process allows us to capture previously 
+	ignored correlations between separate parts of the waveform. We found that 
+	after recalibration, the median mismatch between the model and NR waveforms 
+	decreases by $50\%$. We further explore how different regions of the source 
+	parameter space respond to the optimization procedure. We find that the 
+	degree of improvement correlates with the spins of the source. This work 
+	shows a promising avenue to help understand and treat systematic error 
+	in waveform models.
 \end{abstract}
@@
 Many data analysis tasks in gravitational wave (GW) astrophysics, such
-as \corr{matched} filtering~\citep{owen1996search, owen1999matched} and parameter 
+as matched filtering~\citep{owen1996search, owen1999matched} and parameter 
 estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
@@
 \end{align}
-where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ \corr{($i=1,2,\cdots,19$) is a set of 19 additional
-parameters} which will be discussed below. 
+where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda^i$ ($i=1,2,\cdots,19$) is a set of 19 additional
+parameters which will be discussed below. 
 The phase and amplitude functions are then split into three sections which represent the
@@
 IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up 
-to 3.5PN order \corr{augmented with the next four higher order PN terms}.
+to 3.5PN order augmented with the next four higher order PN terms.
 To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
-uses a series of parameterizations\footnote{
-	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
-} 
+uses a series of parameterizations
+\footnote{The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.} 
 which depend purely on $\Lambda^i$ and can be calibrated to numerical relativity (NR) simulations.
@@
 	\includegraphics[width=\columnwidth]{figures/loss.pdf}
-	\caption{\corr{Average Loss $\mathcal{L}_{\mathrm{ave}}$} against the number of iterations. The vertical
+	\caption{Average Loss $\mathcal{L}_{\mathrm{ave}}$ against the number of iterations. The vertical
 	dotted line indicates the minimum of the validation loss, which is where we stop
@@
 the maximum (minimum) frequency for the integration.
-We note here that the mismatch can be viewed as \corr{a} mean square error (MSE) between the
+We note here that the mismatch can be viewed as a mean square error (MSE) between the
 two waveforms.
@@
 where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
-$\mathcal{M}_{i,\mathrm{ini}}$ represents the mismatch \corr{between original IMRPhenomD waveforms and NR waveforms}, and $N$ is the total number of training waveforms.
+$\mathcal{M}_{i,\mathrm{ini}}$ represents the mismatch between original IMRPhenomD waveforms and NR waveforms, and $N$ is the total number of training waveforms.
 
@@
 behavior during the optimization process.
-The simple average, \corr{$\mathcal{L}_{\mathrm{ave}}$}, serves as the simplest choice of loss function,
+The simple average, $\mathcal{L}_{\mathrm{ave}}$, serves as the simplest choice of loss function,
 but is prone to be dominated by a single point in parameter space with a large mismatch.
@@
 The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
-Conversely, \corr{$\mathcal{L}_{\mathrm{ave}}$} allows the loss function to automatically adjust
+Conversely, $\mathcal{L}_{\mathrm{ave}}$ allows the loss function to automatically adjust
 and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
@@
 are from the SXS catalog 
-\citep{boyle2019sxs} and 10 BAM simulations\footnote{\corr{We took the CoM-corrected waveforms with 4th-order extrapolation.}} . 
+\citep{boyle2019sxs} and 10 BAM simulations. 
+\footnote{We took the CoM-corrected waveforms with 4th-order extrapolation.}
 As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
@@
 with the peak of the distribution being shifted by approximately an order of
-magnitude. The median mismatch is reduced by 50\% (see vertical dotted lines) \corr{while the tail hardly changes}. When using
+magnitude. The median mismatch is reduced by 50\% (see vertical dotted lines) while the tail hardly changes. When using
 $\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
@@
 can also be used to design more physical ansatz. For example, we see the 
-recalibration process struggles to improve in the regime where $\chi_{1}$ is 
-positive and $\chi_{2}$ is negative. This means there is a dependency between the two 
-spins that is not captured in the current ansatz. \corr{In fact, this issue can 
+recalibration process struggles to improve in the regime where $\chi_1$ is 
+positive and $\chi_2$ is negative. This means there is a dependency between the two 
+spins that is not captured in the current ansatz. In fact, this issue can 
 be explained by PN theory, where the leading spin-orbit coupling term depends on 
@@
 incorporates our calibration method with waveform design will be handy in constructing 
-new waveform models in the future. }
+new waveform models in the future.
 
-\corr{We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
+We note that the parameter space we use is restricted to $q\leq8$, while waveforms 
 of $q=18$ are used in the original calibration trial. Also, we have directly used NR 
@@
 can be used to further optimize GW waveform models with arbitrary sets of training 
-waveforms used.}
+waveforms used.
 

### 2024-04-03 src/tex/ms.tex
final minor change
changes for 2nd referee report

@@
 \usepackage[italicdiff]{physics}
+
+\newcommand{\corr}[1]{{\color{red}[corr: #1 ]}}
 
@@
 	same time. This joint optimization process allows us to capture previously 
-	ignored correlations between separate parts of the waveform. We found that 
-	after recalibration, the median mismatch between the model and NR waveforms 
-	decreases by $50\%$. We further explore how different regions of the source 
+	ignored correlations between separate parts of the waveform. \corr{We found 
+	that after recalibrating with a slightly restricted parameter subspace 
+	$(q\leq8)$, despite the tail of the mismatch distribution remains similar, 
+	the median mismatch between the model and NR waveforms decreases by 
+	$50\%$.} We further explore how different regions of the source 
 	parameter space respond to the optimization procedure. We find that the 
@@
 review the parameterization of the IMRPhenomD model and the mismatch function
-that is used as an objective/loss function for the calibration, followed by outlining the
-specific optimization scheme used for recalibration. In
+that is used as an objective/loss function for the calibration, followed by 
+outlining the specific optimization scheme used for recalibration. In
 Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
@@
 optimization result differs as a function of the source parameters. Finally, in
-Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
-procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
-parameterization affects the accuracy of the model. 
-Note that throughout this paper we use the terms recalibration and optimization interchangeably.
+Sec.~\ref{sec:discussion}, we discuss the differences between our calibration
+procedure and the procedure used in \citep{khan2016frequency}. \corr{We point 
+out how some caveats, such as the difference in training waveforms, limit our 
+conclusion to the mismatch improvement.} We also explain how the reduced spin 
+parameterization affects the accuracy of the model. Note that throughout this 
+paper we use the terms recalibration and optimization interchangeably.
 
@@
 waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
-These choices put the original IMRPhenomD result at a disadvantage, as both the parameter 
-space and frequency range are restricted. We tested our result by training the model 
-using a set of training waveforms of the same intrinsic parameters, but with BAM waveforms 
-replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
-mismatch median decreases by about $8\%$. Nevertheless, it is known that IMRPhenomD has a large 
-inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
-space. Hence, we focus on showing that AD with gradient descent, as a general and robust method, 
-can be used to further optimize GW waveform models with arbitrary sets of training 
-waveforms used.
+\corr{These choices may put the original IMRPhenomD result at a disadvantage when 
+assessing accuracy, as both the parameter space and frequency range are restricted.} We 
+tested our result by training the model using a set of training waveforms of the same 
+intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. We can see 
+the result in Fig.~\ref{fig:q14818}, where the mismatch median decreases by about $8\%$. 
+Nevertheless, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio 
+regime, and is rarely used in this region of parameter space. Hence, we focus on showing 
+that AD with gradient descent, as a general and robust method, can be used to further 
+optimize GW waveform models with arbitrary sets of training waveforms used. \corr{With 
+more publicly available data in the future, more systematic studies should be made to 
+precisely assess the improvement one can get out of further optimization. }
 

### 2024-04-10 src/tex/ms.tex
minor change for submission

@@
 Sec.~\ref{sec:discussion}, we discuss the differences between our calibration
-procedure and the procedure used in \citep{khan2016frequency}. \corr{We point 
-out how some caveats, such as the difference in training waveforms, limit our 
-conclusion to the mismatch improvement.} We also explain how the reduced spin 
+procedure and the procedure used in \citep{khan2016frequency}. \corr{Additionally, 
+we discuss some caveats (such as the difference in training waveforms) that limit 
+our conclusions on the mismatch improvement.} We also explain how the reduced spin 
 parameterization affects the accuracy of the model. Note that throughout this 
@@
 
+\newpage
+
 \section{ACKNOWLEDGMENTS}

### 2024-04-18 src/tex/ms.tex
remove highlight

--- 2024-02-16 src/tex/ms.tex
@@
 	same time. This joint optimization process allows us to capture previously 
-	ignored correlations between separate parts of the waveform. We found that 
-	after recalibration, the median mismatch between the model and NR waveforms 
-	decreases by $50\%$. We further explore how different regions of the source 
+	ignored correlations between separate parts of the waveform. We found 
+	that after recalibrating with a slightly restricted parameter subspace 
+	$(q\leq8)$, despite the tail of the mismatch distribution remains similar, 
+	the median mismatch between the model and NR waveforms decreases by 
+	$50\%$. We further explore how different regions of the source 
 	parameter space respond to the optimization procedure. We find that the 
@@
 optimization result differs as a function of the source parameters. Finally, in
-Sec.~\ref{sec:discussion}, we discuss the difference between our calibration
-procedure and the procedure used in \citep{khan2016frequency}. We also explain how the reduced spin
-parameterization affects the accuracy of the model. 
-Note that throughout this paper we use the terms recalibration and optimization interchangeably.
+Sec.~\ref{sec:discussion}, we discuss the differences between our calibration
+procedure and the procedure used in \citep{khan2016frequency}. Additionally, 
+we discuss some caveats (such as the difference in training waveforms) that limit 
+our conclusions on the mismatch improvement. We also explain how the reduced spin 
+parameterization affects the accuracy of the model. Note that throughout this 
+paper we use the terms recalibration and optimization interchangeably.
 
@@
 waveforms instead of hybridized SEOB-NR waveforms for training and validation purposes. 
-These choices put the original IMRPhenomD result at a disadvantage, as both the parameter 
-space and frequency range are restricted. We tested our result by training the model 
-using a set of training waveforms of the same intrinsic parameters, but with BAM waveforms 
-replaced with SEOBNR waveforms. We can see the result in Fig.~\ref{fig:q14818}, where the 
-mismatch median decreases by about $8\%$. Nevertheless, it is known that IMRPhenomD has a large 
-inaccuracy in the large mass-ratio regime, and is rarely used in this region of parameter 
-space. Hence, we focus on showing that AD with gradient descent, as a general and robust method, 
-can be used to further optimize GW waveform models with arbitrary sets of training 
-waveforms used.
+These choices may put the original IMRPhenomD result at a disadvantage when 
+assessing accuracy, as both the parameter space and frequency range are restricted. We 
+tested our result by training the model using a set of training waveforms of the same 
+intrinsic parameters, but with BAM waveforms replaced with SEOBNR waveforms. We can see 
+the result in Fig.~\ref{fig:q14818}, where the mismatch median decreases by about $8\%$. 
+Nevertheless, it is known that IMRPhenomD has a large inaccuracy in the large mass-ratio 
+regime, and is rarely used in this region of parameter space. Hence, we focus on showing 
+that AD with gradient descent, as a general and robust method, can be used to further 
+optimize GW waveform models with arbitrary sets of training waveforms used. With 
+more publicly available data in the future, more systematic studies should be made to 
+precisely assess the improvement one can get out of further optimization.
 
